### 2014-09-04 journal.tex
add notes

%------------------------------------------------------------------------------
% Beginning of journal.tex
%------------------------------------------------------------------------------
%
% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
%
%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
%
% Replace amsart by the documentclass for the target journal, e.g., tran-l.
%
\documentclass{amsart}

%     If your article includes graphics, uncomment this command.
\usepackage{graphicx}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{xca}[theorem]{Exercise}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\numberwithin{equation}{section}

%    Absolute value notation
\newcommand{\abs}[1]{\lvert#1\rvert}

%    Blank box placeholder for figures (to avoid requiring any
%    particular graphics capabilities for printing this document).
\newcommand{\blankbox}[2]{%
  \parbox{\columnwidth}{\centering
%    Set fboxsep to 0 so that the actual size of the box will match the
%    given measurements more closely.
    \setlength{\fboxsep}{0pt}%
    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
  }%
}

\begin{document}

\title{Structural Stability of Random Dynamical Systems}

%    Information for first author
\author{Aviv Bergman}
%    Address of record for the research reported here
\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
%    Current address
\curraddr{}
\email{aviv@einstein.yu.edu}
%    \thanks will become a 1st page footnote.
\thanks{The first author was supported in part by NSF Grant \#000000.}

%    Information for second author
\author{Raymond S. Puzio}
\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
%    Current address
\curraddr{}
\email{rsp@novres.org}
\thanks{Support information for the second author.}

%    Information for third author
\author{Cameron Smith}
\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
%    Current address
\curraddr{}
\email{cameron.smith@med.einstein.yu.edu}
\thanks{Support information for the second author.}

%    General info
\subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}

\date{January 1, 2001 and, in revised form, June 22, 2001.}

\dedicatory{This paper is dedicated to our advisors.}

\keywords{Differential geometry, algebraic geometry}

\begin{abstract}
This paper is a sample prepared to illustrate the use of the American
Mathematical Society's \LaTeX{} document class \texttt{amsart} and
publication-specific variants of that class for AMS-\LaTeX{} version 2.
\end{abstract}

\maketitle

\section{$2 \times 2$ systems}
We call a matrix stable when all its eigenvalues are negative.  For a
$2 \times 2$ matrix, this amounts to the conditions $T < 0$ and $D >
0$ where $T$ and $D$ denote the trace and the determinant.

Suppose we have a stable matrix
$$
\begin{pmatrix}
a & b \\
d & c
\end{pmatrix}
$$
where $a + c < 0$ and $ac > bd$.  We want to know the probability
that, if we resample an element of this matrix, the result will be
stable.  By symmetry, there are two cases to consider; resampling 
$a$ is equivalent to resampling $d$ and resampling $b$ is equivalent
to resampling $c$.

Suppose that we resample $b$.  Since the trace does not involve $b$,
that condition will be satisfied automatically and we only need to
examine the determinant.  We will consider seperately the cases of $d
< 0$ and $d > 0$.  When $d > 0$, the condition becomes $ac/d > b$,
which is satisfiable when $ac > -d$ and results in probability $(ac/d
+ 1)/2$ of remaining stable.  When $d < 0$, the condition becomes
$ac/d < b$, which is satisfiable when $ac > d$, so we have probability
$(1 - ac/d)/2$.

Thus, the probability that a random stable matrix will remain stable
upon resampling $b$ is given as follows:
% \begin{strip}
\begin{align*}
&{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
  P(ac>b'd, a+c < 0) \over
\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \cr
&{{{1 \over 32} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db \,
({ac \over d} + 1) + {1 \over 32} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 2 \atop + 
{1 \over 32} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\, (1
- {ac \over d}) + {1 \over 32} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 2}
\over
{{{1 \over 16} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db\,1
+ {1 \over 16} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 1} \atop
{+ {1 \over 16} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\,1} + {1 \over 16} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 1}}
\end{align*}
% \end{strip}

In order to evaluate the integrals, we begin with the
integrals over $b$, which are particularly easy because none of the
integrands depends upon $b$.  Once we do that and make the change of
variables $d \mapsto -d$ in the second integrals in the numerator and
denominator, we can combine terms and simplify the expression to the
following:
$$
{\int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2 + \int_0^1 dd \int_{d > ac} da\,dc\, 4 \over
{2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)} + 2 \int_0^1 dd \int_{d > ac} da\,dc\, 2}
$$
For $d > 0$ we have
$$
\int_{-d<ac<d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c) 
%\over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
$$
and 
$$
\int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$

% Thus, our expression becomes
% $$
% {\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1)^2 +
% \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1)^2 +
% \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1)^2
% \over
% 2\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1) +
% 2 \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1) +
% 2 \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1) }
% $$

Suppose that we resample $a$.  Then we have to satisfy the conditions
$a + c < 0$ and $ac > bd$ for fixed values of $b,c,d$.  We will
consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
inequalities become $a < -c$ and $a > bd/c$.  For these to be
consistent, we must have $-c > bd/c$, or $bd < -c^2$ and, when we do,
we have a probability $(-c-bd/c)/2$ of reasmpling an $a$ that keeps
our matrix stable.  When $c < 0$, the conditions become $a < -c$ and
$a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
< -c$.  Then the probability of resampling an $a$ which keeps the
matrix stable is $(\min(-c, bd/c) + 1)/2$.

Thus, the probability that a random stable matrix will remain stable upon resampling $a$ is given as follows:
\begin{align*}
&\frac{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
  P(a'c>bd, a'+c < 0)}{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \\
&\frac{\genfrac{}{}{0pt}{}{{1 \over 32} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{\max(bd/c,-1)}^{-c} da \,
(-c - \max({bd \over c},-1)}{+
{1 \over 32} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c,
bd/c)} da\, (\min(-c, bd/c) + 1)}}{{1 \over 16} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{bd/c}^{-c} da\,1
+
{1 \over 16} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c, bd/c)} da\,1}
\end{align*}

For the case $c>0, {bd \over c} < a < -c, bd < -c^2$ we have two possibilities. For $bd < -c$ we have
$$
\int_{-1}^{-c} dd\, \int_{c/d}^1 db\, + \int_c^1 dd\, \int_{-1}^{-c/d} db,
$$
and for $bd > c$
$$
\int_{-1}^{-c} dd\, \int_{-c^2/d}^{-c/d} db\, + \int_{-c}^{-c^2} dd\, \int_{-c^2/d}^{1} db +
\int_{c^2}^{c} dd \int_{-1}^{-c^2/d} db +
\int_{c}^{1}dd\,\int_{-c/d}^{-c^2/d}db.
$$
For the case $c < 0, a < -c, a < bd/c$ we also have two possibilities. For $bd < -c^2$ we have
$$
\int_{-1}^{-c^2} dd\, \int_{-c^2/d}^1 db\, + \int_{c^2}^1 dd\, \int_{-1}^{-c^2/d} db,
$$
and for $bd > -c^2$
$$
\int_{-1}^{-c^2} dd\, \int_{-1}^{-c^2/d} db\, + \int_{-c^2}^{c^2} dd\, \int_{-1}^{1} db + 
\int_{c^2}^{1} dd \int_{-c^2/d}^{1} db.
$$

\section{Arbitrary system size}
We call a matrix $M$ stable if all its eigenvalues have negative real
part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
stable is equivalent to asking that the roots of $p$ all have real
part.  Using some good old fashioned invariant theory, we will
formulate conditions for this to happen in terms of the coefficients of
$p$.

Let us start by fixing some notation.  Suppose that $M$ is of size $n
\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
(x-r_k)$.

Introduce $V$ as the parameter space of all polynomials or degree $n$
with real coefficients.  In other words, $V$ is a real vector space of
dimension $n$ whose coordinates we will identify with the coefficients
$s_k$ of our polynomial.  We now partition this space into regions
corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
subset of $V$ such that $x \in R_{abcd}$ iff the polynomial
corresponding to $x$ has $a$ positive real roots, $b$ negative real
roots, $c$ complex roots with positive real part and $d$ complex roots
with negative real part.  Once we characterize these regions using
invariants, we will have solved the original problem because a matrix
is stable if the point of $V$ corresponding to it lies in a region of
the form $R_{a0c0}$.

Next, we introduce some products:
\begin{align}
D &= \prod_{k=1}^n r_n \cr
\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
\end{align}

Since these quantities are invariant under permutation of the roots
$r_j$, they may be expressed as functions of the coefficients $s_j$.
The explicit expressions may be obtained by such techniques as long
division and determinants.

These quantities are of interest because they describe the boundaries
between the regions introduced above.  Suppose that we have a family
of polynomials whose roots are of one type but turn into another type.
There are three ways this could happen:
\begin{itemize}
\item{1.} A real root could go from being positive to being negative
  or vice versa.  For this to happen, we must pass through an
  intermediate value with a zero root.  For that intermediate value,
  we have $D = 0$.
\item{2.} A pair of real roots could turn into a pair of complex roots
  or vice versa.  For this to happen, we must pass through an
  intermediate value with a repeated root.  For that intermediate
  value, we have $\delta = 0$.
\item{3.} A conjugate pair of complex roots could go from having
  positive real part to having negative real part or vice versa.  For
  this to happen, we must pass through an intermediate value in which
  the real part is zero.  For that intermediate value, the two roots
  are imaginary so, bing conjugates, they sum to zero, hence we have
  $\Delta = 0$.
\end{itemize}
\noindent  Thus, we conclude that the boundaries of our regions
$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
\Delta=0$ and we will characterize them by studying these surfaces.

The simplest case is $n=2$.  There we have the regions $R_{2000},
R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
is summarized in the following table:
\begin{align}
R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0 \cr
R_{1100} \qquad &D < 0, \delta > 0 \cr  
R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0 \cr
R_{0020} \qquad &\delta < 0, \Delta > 0 \cr 
R_{0002} \qquad &\delta < 0, \Delta < 0 \cr
\end{align}

Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
signs to distinguish regions.  For instance, consider the regions
$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
but to tell them apart, we need to note that the region $\delta > 0$
has two connected components and use some auxiliary equalities to
specify which of the components contains our region.  (Looking at the
sign of $\Delta$ will not help here because, while $\Delta > 0$ in the
region $R_{3000}$, it is also positive for some elements of $R_{1200}$
as well.)

% \bibliographystyle{amsplain}
% \begin{thebibliography}{10}

% \bibitem {A} T. Aoki, \textit{Calcul exponentiel des op\'erateurs
% microdifferentiels d'ordre infini.} I, Ann. Inst. Fourier (Grenoble)
% \textbf{33} (1983), 227--250.

% \bibitem {B} R. Brown, \textit{On a conjecture of Dirichlet},
% Amer. Math. Soc., Providence, RI, 1993.

% \bibitem {D} R. A. DeVore, \textit{Approximation of functions},
% Proc. Sympos. Appl. Math., vol. 36,
% Amer. Math. Soc., Providence, RI, 1986, pp. 34--56.

% \end{thebibliography}

\end{document}

%------------------------------------------------------------------------------
% End of journal.tex
%------------------------------------------------------------------------------

### 2014-09-04 journal.tex
clarify 2x2 case and add figures

@@
 that condition will be satisfied automatically and we only need to
-examine the determinant.  We will consider seperately the cases of $d
-< 0$ and $d > 0$.  When $d > 0$, the condition becomes $ac/d > b$,
-which is satisfiable when $ac > -d$ and results in probability $(ac/d
-+ 1)/2$ of remaining stable.  When $d < 0$, the condition becomes
-$ac/d < b$, which is satisfiable when $ac > d$, so we have probability
-$(1 - ac/d)/2$.
+examine the determinant.  Thus, we have the inequalities $ac > bd$ and $-1 < b < 1$.  We will consider seperately the cases of $d
+< 0$ and $d > 0$.  When $d > 0$, the first condition becomes $ac/d > b$.
+Since $b > -1$, this can only be satisfied when $ac/d > -1$, i.e.
+when $ac > -d$.
+In the subcase that $ac < d$, we then have the limits $-1 < b < ac/d$ and in the subcase $ac < d$, we have the limits $-1 < b < 1$.
+When $d < 0$, the condition becomes
+$ac/d < b$, which is satisfiable when $ac > d$.  In the subcase where $ac > -d$, we have $ac/d < b < 1$ and, in the subcase where $ac < -d$, we have $-1 < b <
+1$.
 
@@
 $$
-{\int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2 + \int_0^1 dd \int_{d > ac} da\,dc\, 4 \over
-{2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)} + 2 \int_0^1 dd \int_{d > ac} da\,dc\, 2}
+\frac{\int_0^1 dd \int_{ac > d} da\,dc\, 4 + \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2}{2 \int_0^1 dd \int_{ac > d} da\,dc\, 2 + 2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)}
 $$
-For $d > 0$ we have
-$$
-\int_{-d<ac<d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
-%\over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
-$$
-and 
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-b.pdf}
+\caption{{\bf Resampling b.} (A) The region defined by $ac > d$ (B) The region defined by $d > ac > -d$}
+\label{fig:2x2-resample-b}
+\end{figure}
+
+For $ac > d$ we divide the $a$-axis into two parts indicated in Fig. \ref{fig:2x2-resample-b}A
 $$
 \int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$
+for the numerator $f(a,c) = 4$ and the integral is equal to 2. For $d > ac > -d$ we divide the $a$-axis into three parts indicated in Fig. \ref{fig:2x2-resample-b}B
+$$
+\int_{d > ac > -d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
+% \over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
+$$
+for the numerator $f(a,c) = (\frac{ac}{d} + 1)^2$ and the integral is equal to $\frac{32}{9}$ and for the denominator $f(a,c) = \frac{ac}{d} + 1$ and the integral is equal to $3$. Therefore
+$$P\left(\begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{25}{36} \approx 0.69.$$
 
@@
 
-Suppose that we resample $a$.  Then we have to satisfy the conditions
-$a + c < 0$ and $ac > bd$ for fixed values of $b,c,d$.  We will
+Now suppose we resample $a$.  Then we have to satisfy the conditions
+$a + c < 0$, $ac > bd$, and $-1 < a < 1$ for fixed values of $b,c,d$.  We will
 consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
 inequalities become $a < -c$ and $a > bd/c$.  For these to be
-consistent, we must have $-c > bd/c$, or $bd < -c^2$ and, when we do,
-we have a probability $(-c-bd/c)/2$ of reasmpling an $a$ that keeps
-our matrix stable.  When $c < 0$, the conditions become $a < -c$ and
+consistent, we must have $-c > bd/c$, or $bd < -c^2$.
+In the subcase where $bd < -c$, we have $-1 < a < -c$ and, in the subcase where $-c < bd < -c^2$, we have $bd/c < a < -c$.
+When $c < 0$, the conditions become $a < -c$ and
 $a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
-< -c$.  Then the probability of resampling an $a$ which keeps the
-matrix stable is $(\min(-c, bd/c) + 1)/2$.
+< -c$.  In the subcase where $bd < -c^2$, we have $-1 < a < bd/c$ and, in the subcase where $bd > -c^2$, we have $-1 < a < -c$.
 
@@
 \end{align*}
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-a.pdf}
+\caption{{\bf Resampling a.} (A) The region defined by $bd < -c$ (B) The region defined by $-c < bd < -c^2$ (C) The region defined by $bd < -c^2$ (D) The region defined by $bd > -c^2$}
+\label{fig:2x2-resample-a}
+\end{figure}
 

### 2014-09-05 journal.tex
update text to match computations

--- 2014-09-04 journal.tex
@@
 \section{$2 \times 2$ systems}
-We call a matrix stable when all its eigenvalues are negative.  For a
-$2 \times 2$ matrix, this amounts to the conditions $T < 0$ and $D >
+We call a matrix stable when the real part of all its eigenvalues is negative.  For a $2 \times 2$ matrix, this amounts to the conditions $T < 0$ and $D >
 0$ where $T$ and $D$ denote the trace and the determinant.
@@
 $$
-where $a + c < 0$ and $ac > bd$.  We want to know the probability
-that, if we resample an element of this matrix, the result will be
-stable.  By symmetry, there are two cases to consider; resampling 
-$a$ is equivalent to resampling $d$ and resampling $b$ is equivalent
-to resampling $c$.
+where $a + c < 0$ and $ac > bd$.
+We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
+We want to know the probability that, if we resample an element of this matrix, the result will be stable.
+By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
+to resampling $d$.
 
-Suppose that we resample $b$.  Since the trace does not involve $b$,
-that condition will be satisfied automatically and we only need to
-examine the determinant.  We will consider seperately the cases of $d
-< 0$ and $d > 0$.  When $d > 0$, the condition becomes $ac/d > b$,
-which is satisfiable when $ac > -d$ and results in probability $(ac/d
-+ 1)/2$ of remaining stable.  When $d < 0$, the condition becomes
-$ac/d < b$, which is satisfiable when $ac > d$, so we have probability
-$(1 - ac/d)/2$.
+Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
+% \begin{strip}
+\begin{align}\label{eq:condprob}
+P\left(\begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right)
+& = \frac{P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right)}.
+\end{align}
+The denominator of \ref{eq:condprob} is given by
+\begin{align*}
+P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
+\end{align*}
+Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
+Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
+\begin{align*}
+P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
+\end{align*}
+The analogous equation for resampling $a$ is
+\begin{align*}
+P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+a' & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
+\end{align*}
 
-Thus, the probability that a random stable matrix will remain stable
-upon resampling $b$ is given as follows:
-% \begin{strip}
-\begin{align*}
-&{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
-  P(ac>b'd, a+c < 0) \over
-\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \cr
-&{{{1 \over 32} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db \,
-({ac \over d} + 1) + {1 \over 32} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 2 \atop + 
-{1 \over 32} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\, (1
-- {ac \over d}) + {1 \over 32} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 2}
-\over
-{{{1 \over 16} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db\,1
-+ {1 \over 16} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 1} \atop
-{+ {1 \over 16} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\,1} + {1 \over 16} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 1}}
-\end{align*}
-% \end{strip}
+\newcommand{\specialcell}[2][c]{%
+  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
 
-In order to evaluate the integrals, we begin with the
-integrals over $b$, which are particularly easy because none of the
-integrands depends upon $b$.  Once we do that and make the change of
-variables $d \mapsto -d$ in the second integrals in the numerator and
-denominator, we can combine terms and simplify the expression to the
-following:
-$$
-{\int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2 + \int_0^1 dd \int_{d > ac} da\,dc\, 4 \over
-{2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)} + 2 \int_0^1 dd \int_{d > ac} da\,dc\, 2}
-$$
-For $d > 0$ we have
-$$
-\int_{-d<ac<d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
-%\over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
-$$
-and 
-$$
-\int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$
+\begin{table}[h]
+\begin{tabular}{ c || c | c }
+\hline
+matrix & \specialcell{probability of stability\\to perturbation} & probability of stability\\
+\hline
+  $\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix}$ & 0.62 & 0.25 \\
+  $\begin{pmatrix}
+a & b \\
+d & 0
+\end{pmatrix}$, $\begin{pmatrix}
+0 & b \\
+d & c
+\end{pmatrix}$ & 0.5 & 0.25 \\
+  $\begin{pmatrix}
+a & 0 \\
+d & c
+\end{pmatrix}$, $\begin{pmatrix}
+a & b \\
+0 & c
+\end{pmatrix}$ & 0.67 & 0.25 \\
+$\begin{pmatrix}
+a & 0 \\
+0 & c
+\end{pmatrix}$ & 0.5 & 0.25 \\
+\end{tabular}
+\caption{Probability of matrix stability under resampling and \emph{a priori} stability. All matrices not listed have probability $0$ of stability.}\label{tab:structstabmat}
+\end{table}
 
-% Thus, our expression becomes
+The probability of \emph{a priori} stability and of stability to perturbations for all $2 \times 2$-matrices are given in Table \ref{tab:structstabmat}.
+
+%---------------------------------------------
+
+% We will consider seperately the cases of $d
+% < 0$ and $d > 0$.  When $d > 0$, the first condition becomes $ac/d > b$.
+% Since $b > -1$, this can only be satisfied when $ac/d > -1$, i.e.
+when $ac > -d$.
+In the subcase that $ac < d$, we then have the limits $-1 < b < ac/d$ and in the subcase $ac < d$, we have the limits $-1 < b < 1$.
+When $d < 0$, the condition becomes
+% $ac/d < b$, which is satisfiable when $ac > d$.  In the subcase where $ac > -d$, we have $ac/d < b < 1$ and, in the subcase where $ac < -d$, we have $-1 < b <
+% 1$.
+
+% Thus, the probability that a random stable matrix will remain stable
+% upon resampling $b$ is given as follows:
+% % \begin{strip}
+% \begin{align*}
+% P\left(\begin{pmatrix}
+% a & b' \\
+% d & c
+% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right)
+% & = \frac{P\left(\begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
+% a & b' \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right)}
+% \end{align*}
+% which becomes
+% \begin{align*}
+% &{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
+%   P(ac>b'd, a+c < 0) \over
+% \int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \cr
+% &{{{1 \over 32} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db \,
+% ({ac \over d} + 1) + {1 \over 32} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 2 \atop +
+% {1 \over 32} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\, (1
+% - {ac \over d}) + {1 \over 32} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 2}
+% \over
+% {{{1 \over 16} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db\,1
+% + {1 \over 16} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 1} \atop
+% {+ {1 \over 16} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\,1} + {1 \over 16} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 1}}
+% \end{align*}
+% % \end{strip}
+
+% In order to evaluate the integrals, we begin with the
+% integrals over $b$, which are particularly easy because none of the
+% integrands depends upon $b$.  Once we do that and make the change of
+% variables $d \mapsto -d$ in the second integrals in the numerator and
+% denominator, we can combine terms and simplify the expression to the
+% following:
 % $$
-% {\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1)^2 +
-% \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1)^2 +
-% \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1)^2
-% \over
-% 2\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1) +
-% 2 \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1) +
-% 2 \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1) }
+% \frac{\int_0^1 dd \int_{ac > d} da\,dc\, 4 + \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2}{2 \int_0^1 dd \int_{ac > d} da\,dc\, 2 + 2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)}
 % $$
 
-Suppose that we resample $a$.  Then we have to satisfy the conditions
-$a + c < 0$ and $ac > bd$ for fixed values of $b,c,d$.  We will
-consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
-inequalities become $a < -c$ and $a > bd/c$.  For these to be
-consistent, we must have $-c > bd/c$, or $bd < -c^2$ and, when we do,
-we have a probability $(-c-bd/c)/2$ of reasmpling an $a$ that keeps
-our matrix stable.  When $c < 0$, the conditions become $a < -c$ and
-$a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
-< -c$.  Then the probability of resampling an $a$ which keeps the
-matrix stable is $(\min(-c, bd/c) + 1)/2$.
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-b.pdf}
+% \caption{{\bf Resampling b.} (A) The region defined by $ac > d$ (B) The region defined by $d > ac > -d$}
+% \label{fig:2x2-resample-b}
+% \end{figure}
 
-Thus, the probability that a random stable matrix will remain stable upon resampling $a$ is given as follows:
-\begin{align*}
-&\frac{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
-  P(a'c>bd, a'+c < 0)}{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \\
-&\frac{\genfrac{}{}{0pt}{}{{1 \over 32} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{\max(bd/c,-1)}^{-c} da \,
-(-c - \max({bd \over c},-1)}{+
-{1 \over 32} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c,
-bd/c)} da\, (\min(-c, bd/c) + 1)}}{{1 \over 16} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{bd/c}^{-c} da\,1
-+
-{1 \over 16} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c, bd/c)} da\,1}
-\end{align*}
+% For $ac > d$ we divide the $a$-axis into two parts indicated in Fig. \ref{fig:2x2-resample-b}A
+% $$
+% \int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$
+% for the numerator $f(a,c) = 4$ and the integral is equal to 2. For $d > ac > -d$ we divide the $a$-axis into three parts indicated in Fig. \ref{fig:2x2-resample-b}B
+% $$
+% \int_{d > ac > -d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
+% % \over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
+% $$
+% for the numerator $f(a,c) = (\frac{ac}{d} + 1)^2$ and the integral is equal to $\frac{32}{9}$ and for the denominator $f(a,c) = \frac{ac}{d} + 1$ and the integral is equal to $3$. Therefore
+% $$P\left(\begin{pmatrix}
+% a & b' \\
+% d & c
+% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right) = \frac{25}{36} \approx 0.69.$$
 
-For the case $c>0, {bd \over c} < a < -c, bd < -c^2$ we have two possibilities. For $bd < -c$ we have
-$$
-\int_{-1}^{-c} dd\, \int_{c/d}^1 db\, + \int_c^1 dd\, \int_{-1}^{-c/d} db,
-$$
-and for $bd > c$
-$$
-\int_{-1}^{-c} dd\, \int_{-c^2/d}^{-c/d} db\, + \int_{-c}^{-c^2} dd\, \int_{-c^2/d}^{1} db +
-\int_{c^2}^{c} dd \int_{-1}^{-c^2/d} db +
-\int_{c}^{1}dd\,\int_{-c/d}^{-c^2/d}db.
-$$
-For the case $c < 0, a < -c, a < bd/c$ we also have two possibilities. For $bd < -c^2$ we have
-$$
-\int_{-1}^{-c^2} dd\, \int_{-c^2/d}^1 db\, + \int_{c^2}^1 dd\, \int_{-1}^{-c^2/d} db,
-$$
-and for $bd > -c^2$
-$$
-\int_{-1}^{-c^2} dd\, \int_{-1}^{-c^2/d} db\, + \int_{-c^2}^{c^2} dd\, \int_{-1}^{1} db + 
-\int_{c^2}^{1} dd \int_{-c^2/d}^{1} db.
-$$
+% % Thus, our expression becomes
+% % $$
+% % {\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1)^2 +
+% % \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1)^2 +
+% % \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1)^2
+% % \over
+% % 2\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1) +
+% % 2 \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1) +
+% % 2 \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1) }
+% % $$
+
+% Now suppose we resample $a$.  Then we have to satisfy the conditions
+% $a + c < 0$, $ac > bd$, and $-1 < a < 1$ for fixed values of $b,c,d$.  We will
+% consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
+% inequalities become $a < -c$ and $a > bd/c$.  For these to be
+% consistent, we must have $-c > bd/c$, or $bd < -c^2$.
+In the subcase where $bd < -c$, we have $-1 < a < -c$ and, in the subcase where $-c < bd < -c^2$, we have $bd/c < a < -c$.
+When $c < 0$, the conditions become $a < -c$ and
+% $a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
+% < -c$.  In the subcase where $bd < -c^2$, we have $-1 < a < bd/c$ and, in the subcase where $bd > -c^2$, we have $-1 < a < -c$.
+
+% Thus, the probability that a random stable matrix will remain stable upon resampling $a$ is given as follows:
+% \begin{align*}
+% &\frac{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
+%   P(a'c>bd, a'+c < 0)}{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \\
+% &\frac{\genfrac{}{}{0pt}{}{{1 \over 32} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{\max(bd/c,-1)}^{-c} da \,
+% (-c - \max({bd \over c},-1)}{+
+% {1 \over 32} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c,
+% bd/c)} da\, (\min(-c, bd/c) + 1)}}{{1 \over 16} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{bd/c}^{-c} da\,1
+% +
+% {1 \over 16} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c, bd/c)} da\,1}
+% \end{align*}
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-a.pdf}
+% \caption{{\bf Resampling a.} (A) The region defined by $bd < -c$ (B) The region defined by $-c < bd < -c^2$ (C) The region defined by $bd < -c^2$ (D) The region defined by $bd > -c^2$}
+% \label{fig:2x2-resample-a}
+% \end{figure}
+
+% For the case $c>0, {bd \over c} < a < -c, bd < -c^2$ we have two possibilities. For $bd < -c$ we have
+% $$
+% \int_{-1}^{-c} dd\, \int_{c/d}^1 db\, + \int_c^1 dd\, \int_{-1}^{-c/d} db,
+% $$
+% and for $bd > c$
+% $$
+% \int_{-1}^{-c} dd\, \int_{-c^2/d}^{-c/d} db\, + \int_{-c}^{-c^2} dd\, \int_{-c^2/d}^{1} db +
+% \int_{c^2}^{c} dd \int_{-1}^{-c^2/d} db +
+% \int_{c}^{1}dd\,\int_{-c/d}^{-c^2/d}db.
+% $$
+% For the case $c < 0, a < -c, a < bd/c$ we also have two possibilities. For $bd < -c^2$ we have
+% $$
+% \int_{-1}^{-c^2} dd\, \int_{-c^2/d}^1 db\, + \int_{c^2}^1 dd\, \int_{-1}^{-c^2/d} db,
+% $$
+% and for $bd > -c^2$
+% $$
+% \int_{-1}^{-c^2} dd\, \int_{-1}^{-c^2/d} db\, + \int_{-c^2}^{c^2} dd\, \int_{-1}^{1} db +
+% \int_{c^2}^{1} dd \int_{-c^2/d}^{1} db.
+% $$
+
+%---------------------------------------------
 

### 2014-09-08 journal.tex
draft abstract and intro

@@
 
-\title{Structural Stability of Random Dynamical Systems}
+\title{Structural Stability of Dynamical System Ensembles}
 
@@
 \begin{abstract}
-This paper is a sample prepared to illustrate the use of the American
-Mathematical Society's \LaTeX{} document class \texttt{amsart} and
-publication-specific variants of that class for AMS-\LaTeX{} version 2.
+The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
+When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
+Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
+In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two interacting components.
+We show in this case that structural stability has a non-monotonic relationship with system connectivity.
+These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
+The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
 \end{abstract}
@@
 \maketitle
+
+\section{Stability analysis of arbitrary systems}
+
+Consider an n-dimensional vector
+$$
+x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
+$$
+whose components are solutions to the arbitrary first order system
+$$
+\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
+$$
+where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
+Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
+$$
+\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
+$$
+where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matirx $A$ has components
+$$
+a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
+$$
+The spectral abscissa of the matrix $A$ is defined as
+$$
+\eta(A) = \max_i \{\Re(\lambda_i)\}
+$$
+where $\lambda_i$ are the eigenvalues of $A$. The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
+
+To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
+Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
+
+Another form of stability, referred to as structural stability, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
+This can be formalized as the conditional probability distribution
+$$
+P(A' \, \textrm{stable} | A \, \textrm{stable})
+$$
+where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
+In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
 

### 2014-09-08 journal.tex
draft introduction and add bibliography

@@
 
-\title{Structural Stability of Dynamical System Ensembles}
+\title{Stability of Dynamical System Ensembles}
 
@@
 
+\section{Introduction}
+
+The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
+Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
+In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
+This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
+
+Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
+The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
+For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
+This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
+However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
+
+Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
+We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
+We then proceed to determine the probability of structural stability over the same domain.
+We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
+This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
+
 \section{Stability analysis of arbitrary systems}
@@
 
-Another form of stability, referred to as structural stability, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
+Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
 This can be formalized as the conditional probability distribution
@@
 
-% \bibliographystyle{amsplain}
+\bibliographystyle{amsplain}
 % \begin{thebibliography}{10}
+\bibliography{bib/books,bib/papers}
 

### 2014-09-09 journal.tex
comment out support information
fix typo

@@
 %    \thanks will become a 1st page footnote.
-\thanks{The first author was supported in part by NSF Grant \#000000.}
+% \thanks{The first author was supported in part by NSF Grant \#000000.}
 
@@
 \email{rsp@novres.org}
-\thanks{Support information for the second author.}
+% \thanks{Support information for the second author.}
 
@@
 \email{cameron.smith@med.einstein.yu.edu}
-\thanks{Support information for the second author.}
+% \thanks{Support information for the second author.}
 
 %    General info
-\subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
+% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
 
-\date{January 1, 2001 and, in revised form, June 22, 2001.}
+% \date{January 1, 2001 and, in revised form, June 22, 2001.}
 
-\dedicatory{This paper is dedicated to our advisors.}
+% \dedicatory{This paper is dedicated to our advisors.}
 
-\keywords{Differential geometry, algebraic geometry}
+% \keywords{Differential geometry, algebraic geometry}
 
@@
 
-Introduce $V$ as the parameter space of all polynomials or degree $n$
+Introduce $V$ as the parameter space of all polynomials of degree $n$
 with real coefficients.  In other words, $V$ is a real vector space of

### 2014-09-09 code/rds3x3.jl
add scripts for 3x3 simulations

@@ -1,0 +1,61 @@
+using Permutations
+
+flatten{T}(a::Array{T,1}) = any(map(x->isa(x,Array),a))? flatten(vcat(map(flatten,a)...)): a
+flatten{T}(a::Array{T}) = reshape(a,prod(size(a)))
+flatten(a)=a
+
+function bin2dec(I::Array{Bool,1})
+
+  # convert boolean array to an integer
+  return parseint(join(map(x->convert(Int,x),I)),2)
+
+end
+
+function dec2bin(N::Integer)
+
+  # convert integer to boolean array
+  map(parseint,split(bits(N),""))
+
+end
+
+function genbinmats(N::Integer)
+
+  return [reshape(dec2bin(i)[64-N^2+1:end],N,N) for i in 0:2^(N^2)-1]
+
+end
+
+function findpermclasses(N::Integer)
+    # generate matrix representations of permutations
+    permmats = [matrix(Permutation(i)) for i in permutations([1:N])][2:end]
+    allbinmats = genbinmats(N)
+    conjclassreps = (Array{Int64,2})[]
+
+    for (i,mat) in enumerate(allbinmats)
+        ind = true
+        if bin2dec(bool(flatten(mat'))) < i-1
+            ind = false
+        end
+
+        for p in permmats
+            if bin2dec(bool(flatten((p*mat*inv(p))))) < i-1
+                ind = false
+            end
+
+            if bin2dec(bool(flatten((p*mat*inv(p))'))) < i-1
+                ind = false
+            end
+        end
+
+        if ind
+            push!(conjclassreps,mat)
+        end
+    end
+
+    classbyconnect = Array(Array{Array{Int64,2}},N^2+1)
+    for i in 0:N^2
+        classbyconnect[i+1] = conjclassreps[find(x->sum(x)==i,conjclassreps)]
+    end
+
+    return classbyconnect
+end
+

### 2014-09-10 journal.tex
finish 3x3 simulation

@@
 }
+
+% user added
+\usepackage{longtable}
 
@@
 
-\section{$2 \times 2$ systems}
+\section{Arbitrary system size}
+We call a matrix $M$ stable if all its eigenvalues have negative real
+part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
+roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
+stable is equivalent to asking that the roots of $p$ all have real
+part.  Using some good old fashioned invariant theory, we will
+formulate conditions for this to happen in terms of the coefficients of
+$p$.
+
+Let us start by fixing some notation.  Suppose that $M$ is of size $n
+\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
+$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
+(x-r_k)$.
+
+Introduce $V$ as the parameter space of all polynomials of degree $n$
+with real coefficients.  In other words, $V$ is a real vector space of
+dimension $n$ whose coordinates we will identify with the coefficients
+$s_k$ of our polynomial.  We now partition this space into regions
+corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
+subset of $V$ such that $x \in R_{abcd}$ iff the polynomial
+corresponding to $x$ has $a$ positive real roots, $b$ negative real
+roots, $c$ complex roots with positive real part and $d$ complex roots
+with negative real part.  Once we characterize these regions using
+invariants, we will have solved the original problem because a matrix
+is stable if the point of $V$ corresponding to it lies in a region of
+the form $R_{a0c0}$.
+
+Next, we introduce some products:
+\begin{align}
+D &= \prod_{k=1}^n r_n \cr
+\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
+\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
+\end{align}
+
+Since these quantities are invariant under permutation of the roots
+$r_j$, they may be expressed as functions of the coefficients $s_j$.
+The explicit expressions may be obtained by such techniques as long
+division and determinants.
+
+These quantities are of interest because they describe the boundaries
+between the regions introduced above.  Suppose that we have a family
+of polynomials whose roots are of one type but turn into another type.
+There are three ways this could happen:
+\begin{itemize}
+\item{1.} A real root could go from being positive to being negative
+  or vice versa.  For this to happen, we must pass through an
+  intermediate value with a zero root.  For that intermediate value,
+  we have $D = 0$.
+\item{2.} A pair of real roots could turn into a pair of complex roots
+  or vice versa.  For this to happen, we must pass through an
+  intermediate value with a repeated root.  For that intermediate
+  value, we have $\delta = 0$.
+\item{3.} A conjugate pair of complex roots could go from having
+  positive real part to having negative real part or vice versa.  For
+  this to happen, we must pass through an intermediate value in which
+  the real part is zero.  For that intermediate value, the two roots
+  are imaginary so, bing conjugates, they sum to zero, hence we have
+  $\Delta = 0$.
+\end{itemize}
+\noindent  Thus, we conclude that the boundaries of our regions
+$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
+\Delta=0$ and we will characterize them by studying these surfaces.
+
+The simplest case is $n=2$.  There we have the regions $R_{2000},
+R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
+s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
+is summarized in the following table:
+\begin{align}
+R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0 \cr
+R_{1100} \qquad &D < 0, \delta > 0 \cr
+R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0 \cr
+R_{0020} \qquad &\delta < 0, \Delta > 0 \cr
+R_{0002} \qquad &\delta < 0, \Delta < 0 \cr
+\end{align}
+
+Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
+= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
+\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
+signs to distinguish regions.  For instance, consider the regions
+$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
+but to tell them apart, we need to note that the region $\delta > 0$
+has two connected components and use some auxiliary equalities to
+specify which of the components contains our region.  (Looking at the
+sign of $\Delta$ will not help here because, while $\Delta > 0$ in the
+region $R_{3000}$, it is also positive for some elements of $R_{1200}$
+as well.)
+
+\section{Two component systems}
 We call a matrix stable when the real part of all its eigenvalues is negative.  For a $2 \times 2$ matrix, this amounts to the conditions $T < 0$ and $D >
@@
 \begin{table}[h]
-\begin{tabular}{ c || c | c }
+\begin{tabular}{ c || c | c | c }
 \hline
-matrix & \specialcell{probability of stability\\to perturbation} & probability of stability\\
+matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
 \hline
@@
 d & c
-\end{pmatrix}$ & 0.62 & 0.25 \\
+\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
   $\begin{pmatrix}
@@
 d & c
-\end{pmatrix}$ & 0.5 & 0.25 \\
+\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
   $\begin{pmatrix}
@@
 0 & c
-\end{pmatrix}$ & 0.67 & 0.25 \\
+\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
 $\begin{pmatrix}
@@
 0 & c
-\end{pmatrix}$ & 0.5 & 0.25 \\
+\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
 \end{tabular}
-\caption{Probability of matrix stability under resampling and \emph{a priori} stability. All matrices not listed have probability $0$ of stability.}\label{tab:structstabmat}
+\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
+All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
 \end{table}
 
-The probability of \emph{a priori} stability and of stability to perturbations for all $2 \times 2$-matrices are given in Table \ref{tab:structstabmat}.
+The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
 
@@
 
-\section{Arbitrary system size}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  Using some good old fashioned invariant theory, we will
-formulate conditions for this to happen in terms of the coefficients of
-$p$.
+\section{Three component systems}
 
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
+The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
 
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ iff the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants.
-
-These quantities are of interest because they describe the boundaries
-between the regions introduced above.  Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized in the following table:
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0 \cr
-R_{1100} \qquad &D < 0, \delta > 0 \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0 \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0 \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0 \cr
-\end{align}
-
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use some auxiliary equalities to
-specify which of the components contains our region.  (Looking at the
-sign of $\Delta$ will not help here because, while $\Delta > 0$ in the
-region $R_{3000}$, it is also positive for some elements of $R_{1200}$
-as well.)
+% \begin{table}[h]
+\begin{longtable}{ c || c | c | c }
+\hline
+matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
+\hline
+$\begin{pmatrix}
+1 & 0 & 0\\
+0 & 1 & 0\\
+0 & 0 & 1\\
+\end{pmatrix}$ & 4 & 0.496 & 0.125\\
+$\begin{pmatrix}
+0 & 0 & 1\\
+0 & 0 & 1\\
+1 & 1 & 0\\
+\end{pmatrix}$ & 5 & 0.117 & 0.07\\
+$\begin{pmatrix}
+0 & 0 & 1\\
+0 & 1 & 0\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 5 & 0.49 & 0.123\\
+$\begin{pmatrix}
+1 & 0 & 0\\
+0 & 1 & 1\\
+0 & 0 & 1\\
+\end{pmatrix}$ & 5 & 0.62 & 0.128\\
+$\begin{pmatrix}
+0 & 0 & 1\\
+0 & 0 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 6 & 0.32 & 0.089\\
+$\begin{pmatrix}
+0 & 0 & 1\\
+0 & 1 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 6 & 0.599 & 0.123\\
+$\begin{pmatrix}
+0 & 0 & 1\\
+0 & 1 & 1\\
+1 & 1 & 0\\
+\end{pmatrix}$ & 6 & 0.499 & 0.127\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 0 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 6 & 0.398 & 0.064\\
+$\begin{pmatrix}
+0 & 1 & 0\\
+0 & 1 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 6 & 0.431 & 0.08\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 1 & 0\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 6 & 0.597 & 0.126\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 1 & 1\\
+1 & 0 & 0\\
+\end{pmatrix}$ & 6 & 0.41 & 0.061\\
+$\begin{pmatrix}
+1 & 0 & 1\\
+0 & 1 & 1\\
+0 & 0 & 1\\
+\end{pmatrix}$ & 6 & 0.7 & 0.122\\
+$\begin{pmatrix}
+1 & 0 & 0\\
+0 & 1 & 1\\
+0 & 1 & 1\\
+\end{pmatrix}$ & 6 & 0.595 & 0.123\\
+$\begin{pmatrix}
+1 & 0 & 1\\
+0 & 1 & 0\\
+0 & 1 & 1\\
+\end{pmatrix}$ & 6 & 0.701 & 0.125\\
+$\begin{pmatrix}
+0 & 0 & 1\\
+0 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 7 & 0.574 & 0.117\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 0 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 7 & 0.486 & 0.075\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 1 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 7 & 0.539 & 0.092\\
+$\begin{pmatrix}
+0 & 1 & 0\\
+0 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 7 & 0.497 & 0.088\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 1 & 0\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 7 & 0.664 & 0.126\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 1 & 1\\
+1 & 1 & 0\\
+\end{pmatrix}$ & 7 & 0.482 & 0.086\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+1 & 1 & 0\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 7 & 0.589 & 0.126\\
+$\begin{pmatrix}
+1 & 0 & 1\\
+0 & 1 & 1\\
+0 & 1 & 1\\
+\end{pmatrix}$ & 7 & 0.66 & 0.127\\
+$\begin{pmatrix}
+1 & 1 & 1\\
+0 & 1 & 1\\
+0 & 0 & 1\\
+\end{pmatrix}$ & 7 & 0.749 & 0.121\\
+$\begin{pmatrix}
+1 & 1 & 0\\
+0 & 1 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 7 & 0.61 & 0.097\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+0 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 8 & 0.571 & 0.102\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+1 & 0 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 8 & 0.472 & 0.077\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+1 & 1 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 8 & 0.59 & 0.11\\
+$\begin{pmatrix}
+1 & 1 & 1\\
+0 & 1 & 1\\
+0 & 1 & 1\\
+\end{pmatrix}$ & 8 & 0.709 & 0.125\\
+$\begin{pmatrix}
+1 & 0 & 1\\
+0 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 8 & 0.642 & 0.122\\
+$\begin{pmatrix}
+1 & 1 & 1\\
+0 & 1 & 1\\
+1 & 0 & 1\\
+\end{pmatrix}$ & 8 & 0.63 & 0.108\\
+$\begin{pmatrix}
+0 & 1 & 1\\
+1 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 9 & 0.57 & 0.092\\
+$\begin{pmatrix}
+1 & 1 & 1\\
+0 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 9 & 0.649 & 0.111\\
+$\begin{pmatrix}
+1 & 1 & 1\\
+1 & 1 & 1\\
+1 & 1 & 1\\
+\end{pmatrix}$ & 10 & 0.645 & 0.103\\
+\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3}
+\end{longtable}
+%
+% \end{table}
 

### 2014-09-10 code/rds3x3.jl

@@ -1,5 +1,47 @@
 using Permutations
 
+function pmatrix(M::Array{Int64,2})
+    lines = split(replace(replace(string(M),"[",""),"]",""),"\n")
+    rv = (String)[]
+    push!(rv,"\\begin{pmatrix}")
+    for l in lines
+        push!(rv,string(join(split(l)," & "),"\\\\"))
+    end
+    push!(rv,"\\end{pmatrix}")
+    return join(rv,"\n")
+end
+
+function pmatrix(M::Array{Float64,2})
+    lines = split(replace(replace(string(round(M,2)),"[",""),"]",""),"\n")
+    rv = (String)[]
+    push!(rv,"\\begin{pmatrix}")
+    for l in lines
+        push!(rv,string(join(split(l)," & "),"\\\\"))
+    end
+    push!(rv,"\\end{pmatrix}")
+    return join(rv,"\n")
+end
+
+function savelatex(pcN,sp,rsp)
+    rv = (String)[]
+
+    for (i,plist) in enumerate(sp)
+        for (j,p) in enumerate(plist)
+            if p > 1e-3
+                  pmatstr = pmatrix(pcN[i][j])
+                  push!(rv,
+                        string("\$", pmatstr, "\$", " & ",
+                               "$(i)", " & ",
+                               "$(round(rsp[i][j],3))", " & ",
+                               "$(round(sp[i][j],3))", "\\\\"))
+            end
+        end
+    end
+
+    return join(rv,"\n")
+end
+
+
 flatten{T}(a::Array{T,1}) = any(map(x->isa(x,Array),a))? flatten(vcat(map(flatten,a)...)): a
 flatten{T}(a::Array{T}) = reshape(a,prod(size(a)))
 flatten(a)=a
@@ -58,4 +100,113 @@ function findpermclasses(N::Integer)
 
     return classbyconnect
 end
+
+function sysstab(M::Array{Float64,2})
+
+  # Continuous time system
+  ec,vc = eig(M)
+
+  # Discrete time system L = exp(M)
+  # where exp is the matrix exponential
+  # L = eye(M)+M+M^2/2+M^3/6
+  # ed,vd = eig(L)
+
+  # Print system matrices
+  # println("Continuous system:")
+  # println(M)
+  # println("Discrete system:")
+  # println(L)
+
+  # return maximum(abs(e))/sqrt(N)
+  # return maximum(abs(ed)), maximum(real(ec))
+  return maximum(real(ec))
[…]

### 2014-09-10 journal.tex
add figures with constraints on char poly
remove extra comments
clarify stability condition

@@
 Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-$$
+\begin{equation}\label{eq:lineardynsys}
 \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-$$
+\end{equation}
 where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matirx $A$ has components
@@
 $$
-where $\lambda_i$ are the eigenvalues of $A$. The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
+where $\lambda_i$ are the eigenvalues of $A$.
+The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
+This is because the general solution to equation \ref{eq:lineardynsys} is
+$$
+y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
+$$
+for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
 
@@
 
-%---------------------------------------------
-
-% We will consider seperately the cases of $d
-% < 0$ and $d > 0$.  When $d > 0$, the first condition becomes $ac/d > b$.
-% Since $b > -1$, this can only be satisfied when $ac/d > -1$, i.e.
-when $ac > -d$.
-In the subcase that $ac < d$, we then have the limits $-1 < b < ac/d$ and in the subcase $ac < d$, we have the limits $-1 < b < 1$.
-When $d < 0$, the condition becomes
-% $ac/d < b$, which is satisfiable when $ac > d$.  In the subcase where $ac > -d$, we have $ac/d < b < 1$ and, in the subcase where $ac < -d$, we have $-1 < b <
-% 1$.
-
-% Thus, the probability that a random stable matrix will remain stable
-% upon resampling $b$ is given as follows:
-% % \begin{strip}
-% \begin{align*}
-% P\left(\begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)
-% & = \frac{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}
-% \end{align*}
-% which becomes
-% \begin{align*}
-% &{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
-%   P(ac>b'd, a+c < 0) \over
-% \int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \cr
-% &{{{1 \over 32} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db \,
-% ({ac \over d} + 1) + {1 \over 32} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 2 \atop +
-% {1 \over 32} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\, (1
-% - {ac \over d}) + {1 \over 32} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 2}
-% \over
-% {{{1 \over 16} \int_0^1 dd \int_{d > ac > -d} da\,dc \int_{-1}^{ac/d} db\,1
-% + {1 \over 16} \int_{0}^{1} dd \int_{ac>d} da\,dc \int_{-1}^{1} db 1} \atop
-% {+ {1 \over 16} \int_{-1}^0 dd \int_{-d > ac > d} da\,dc \int_{ac/d}^1 db\,1} + {1 \over 16} \int_{-1}^{0} dd \int_{ac>-d} da\,dc \int_{-1}^{1} db 1}}
-% \end{align*}
-% % \end{strip}
-
-% In order to evaluate the integrals, we begin with the
-% integrals over $b$, which are particularly easy because none of the
-% integrands depends upon $b$.  Once we do that and make the change of
-% variables $d \mapsto -d$ in the second integrals in the numerator and
-% denominator, we can combine terms and simplify the expression to the
-% following:
-% $$
-% \frac{\int_0^1 dd \int_{ac > d} da\,dc\, 4 + \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2}{2 \int_0^1 dd \int_{ac > d} da\,dc\, 2 + 2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)}
-% $$
-
-% \begin{figure}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-b.pdf}
-% \caption{{\bf Resampling b.} (A) The region defined by $ac > d$ (B) The region defined by $d > ac > -d$}
-% \label{fig:2x2-resample-b}
-% \end{figure}
-
-% For $ac > d$ we divide the $a$-axis into two parts indicated in Fig. \ref{fig:2x2-resample-b}A
-% $$
-% \int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$
-% for the numerator $f(a,c) = 4$ and the integral is equal to 2. For $d > ac > -d$ we divide the $a$-axis into three parts indicated in Fig. \ref{fig:2x2-resample-b}B
-% $$
-% \int_{d > ac > -d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
-% % \over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
-% $$
-% for the numerator $f(a,c) = (\frac{ac}{d} + 1)^2$ and the integral is equal to $\frac{32}{9}$ and for the denominator $f(a,c) = \frac{ac}{d} + 1$ and the integral is equal to $3$. Therefore
-% $$P\left(\begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right) = \frac{25}{36} \approx 0.69.$$
-
-% % Thus, our expression becomes
-% % $$
-% % {\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1)^2 +
-% % \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1)^2 +
-% % \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1)^2
-% % \over
-% % 2\int_0^1 dd \int_{-1}^{-d} da \int_{-1}^{-d/a} dc\, ({ac \over d} + 1) +
-% % 2 \int_0^1 dd \int_{-d}^{d} da \int_{-1}^{1} dc\, ({ac \over d} + 1) +
-% % 2 \int_0^1 dd \int_d^1 da \int_{-d/a}^{1} dc\, ({ac \over d} + 1) }
-% % $$
-
-% Now suppose we resample $a$.  Then we have to satisfy the conditions
-% $a + c < 0$, $ac > bd$, and $-1 < a < 1$ for fixed values of $b,c,d$.  We will
-% consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
-% inequalities become $a < -c$ and $a > bd/c$.  For these to be
-% consistent, we must have $-c > bd/c$, or $bd < -c^2$.
-In the subcase where $bd < -c$, we have $-1 < a < -c$ and, in the subcase where $-c < bd < -c^2$, we have $bd/c < a < -c$.
-When $c < 0$, the conditions become $a < -c$ and
-% $a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
-% < -c$.  In the subcase where $bd < -c^2$, we have $-1 < a < bd/c$ and, in the subcase where $bd > -c^2$, we have $-1 < a < -c$.
-
-% Thus, the probability that a random stable matrix will remain stable upon resampling $a$ is given as follows:
-% \begin{align*}
-% &\frac{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,
-%   P(a'c>bd, a'+c < 0)}{\int_{ac>bd \atop a+c<0} da\,db\,dc\,dd\,1} = \\
-% &\frac{\genfrac{}{}{0pt}{}{{1 \over 32} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{\max(bd/c,-1)}^{-c} da \,
-% (-c - \max({bd \over c},-1)}{+
-% {1 \over 32} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c,
-% bd/c)} da\, (\min(-c, bd/c) + 1)}}{{1 \over 16} \int_0^1 dc \int_{bd < -c^2} db\,dd \int_{bd/c}^{-c} da\,1
-% +
-% {1 \over 16} \int_{-1}^0 dc \int_{bd < -c} db\,dd \int_{-1}^{\min(-c, bd/c)} da\,1}
-% \end{align*}
-
-% \begin{figure}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-a.pdf}
-% \caption{{\bf Resampling a.} (A) The region defined by $bd < -c$ (B) The region defined by $-c < bd < -c^2$ (C) The region defined by $bd < -c^2$ (D) The region defined by $bd > -c^2$}
-% \label{fig:2x2-resample-a}
-% \end{figure}
-
-% For the case $c>0, {bd \over c} < a < -c, bd < -c^2$ we have two possibilities. For $bd < -c$ we have
-% $$
-% \int_{-1}^{-c} dd\, \int_{c/d}^1 db\, + \int_c^1 dd\, \int_{-1}^{-c/d} db,
-% $$
-% and for $bd > c$
-% $$
-% \int_{-1}^{-c} dd\, \int_{-c^2/d}^{-c/d} db\, + \int_{-c}^{-c^2} dd\, \int_{-c^2/d}^{1} db +
-% \int_{c^2}^{c} dd \int_{-1}^{-c^2/d} db +
-% \int_{c}^{1}dd\,\int_{-c/d}^{-c^2/d}db.
-% $$
-% For the case $c < 0, a < -c, a < bd/c$ we also have two possibilities. For $bd < -c^2$ we have
-% $$
-% \int_{-1}^{-c^2} dd\, \int_{-c^2/d}^1 db\, + \int_{c^2}^1 dd\, \int_{-1}^{-c^2/d} db,
-% $$
-% and for $bd > -c^2$
-% $$
-% \int_{-1}^{-c^2} dd\, \int_{-1}^{-c^2/d} db\, + \int_{-c^2}^{c^2} dd\, \int_{-1}^{1} db +
-% \int_{c^2}^{1} dd \int_{-c^2/d}^{1} db.
-% $$
-
-%---------------------------------------------
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
+\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} }
+\label{fig:region2x2}
+\end{figure}
 
@@
 
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
+\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} }
+\label{fig:region3x3}
+\end{figure}
+
 \bibliographystyle{amsplain}

### 2014-09-10 code/rds3x3.jl
fix edge case where zero eigenvalues are slightly negative due to numerical error

@@ -136,7 +136,7 @@ function teststab(N::Integer,K::Integer)
         for (i,mat) in enumerate(matlist)
             stabnum = 0
             for j=1:K
-                if sysstab(mat.*randn(N,N)) < 0.0
+                if sysstab(mat.*randn(N,N)) < -1e-4
                     stabnum += 1
                 end
             end
@@ -179,12 +179,12 @@ function teststructstab(N::Integer,K::Integer)
                     end
 
                     candidate = pcN[k][l].*randn(N,N)
-                    if sysstab(candidate) < 0.0
+                    if sysstab(candidate) < -1e-4
                         stabnum += 1
                         nzinds = find(x->x>0,pcN[k][l])
                         ind = nzinds[rand(1:length(nzinds))]
                         candidate[ind] = randn()
-                        if sysstab(candidate) < 0.0
+                        if sysstab(candidate) < -1e-4
                             restabnum += 1
                         end
                     end

### 2014-09-10 code/rds3x3.jl
fix off-by-one error in connectivity value of latex table

@@ -31,7 +31,7 @@ function savelatex(pcN,sp,rsp)
                   pmatstr = pmatrix(pcN[i][j])
                   push!(rv,
                         string("\$", pmatstr, "\$", " & ",
-                               "$(i)", " & ",
+                               "$(i-1)", " & ",
                                "$(round(rsp[i][j],3))", " & ",
                                "$(round(sp[i][j],3))", "\\\\"))
             end

### 2014-09-11 code/rds3x3.jl
plot resampling probabilities for 3x3

@@ -41,6 +41,23 @@ function savelatex(pcN,sp,rsp)
     return join(rv,"\n")
 end
 
+function savedata(filename,sp,rsp)
+    rv = (String)[]
+    push!(rv,"connectivity\tstability\tresampling")
+    for (i,plist) in enumerate(sp)
+        for (j,p) in enumerate(plist)
+            if p > 1e-3
+                  line = "$(i-1)\t$(sp[i][j])\t$(rsp[i][j])"
+                  push!(rv,line)
+            end
+        end
+    end
+    rv = join(rv,"\n")
+    f = open(filename,"w")
+    write(f,rv)
+    close(f)
+    return rv
+end
 
 flatten{T}(a::Array{T,1}) = any(map(x->isa(x,Array),a))? flatten(vcat(map(flatten,a)...)): a
 flatten{T}(a::Array{T}) = reshape(a,prod(size(a)))

### 2014-09-11 data/plot.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+# usage:
+#   ./plot.sh stab3x3
+
+gnuplot $1.p
+rsvg-convert -f pdf -o ../fig/$1.pdf ../fig/$1.svg
+evince ../fig/$1.pdf

### 2014-09-11 journal.tex
modify title
fix edge case where zero eigenvalues are slightly negative due to numerical error
fix off-by-one error in connectivity value of latex table
add plot of stability to perturbations for 3x3 systems
mention case of three component systems in abstract

@@
 
-\title{Stability of Dynamical System Ensembles}
+\title{Stability of Ensembles of Biological Networks}
 
@@
 Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two interacting components.
-We show in this case that structural stability has a non-monotonic relationship with system connectivity.
+In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
+We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
 These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
@@
 The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
+Fig.
+\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems.
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
+\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
+\label{fig:stab3x3}
+\end{figure}
 
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 4 & 0.496 & 0.125\\
-$\begin{pmatrix}
-0 & 0 & 1\\
-0 & 0 & 1\\
-1 & 1 & 0\\
-\end{pmatrix}$ & 5 & 0.117 & 0.07\\
+\end{pmatrix}$ & 3 & 0.499 & 0.126\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.49 & 0.123\\
+\end{pmatrix}$ & 4 & 0.505 & 0.121\\
 $\begin{pmatrix}
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.62 & 0.128\\
-$\begin{pmatrix}
-0 & 0 & 1\\
-0 & 0 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.32 & 0.089\\
+\end{pmatrix}$ & 4 & 0.622 & 0.127\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.599 & 0.123\\
+\end{pmatrix}$ & 5 & 0.595 & 0.121\\
 $\begin{pmatrix}
@@
 1 & 1 & 0\\
-\end{pmatrix}$ & 6 & 0.499 & 0.127\\
+\end{pmatrix}$ & 5 & 0.494 & 0.128\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.398 & 0.064\\
+\end{pmatrix}$ & 5 & 0.41 & 0.061\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.431 & 0.08\\
+\end{pmatrix}$ & 5 & 0.43 & 0.078\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.597 & 0.126\\
+\end{pmatrix}$ & 5 & 0.605 & 0.12\\
 $\begin{pmatrix}
@@
 1 & 0 & 0\\
-\end{pmatrix}$ & 6 & 0.41 & 0.061\\
+\end{pmatrix}$ & 5 & 0.405 & 0.06\\
 $\begin{pmatrix}
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.7 & 0.122\\
+\end{pmatrix}$ & 5 & 0.698 & 0.122\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.595 & 0.123\\
+\end{pmatrix}$ & 5 & 0.587 & 0.128\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.701 & 0.125\\
+\end{pmatrix}$ & 5 & 0.707 & 0.127\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.574 & 0.117\\
+\end{pmatrix}$ & 6 & 0.578 & 0.121\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.486 & 0.075\\
+\end{pmatrix}$ & 6 & 0.487 & 0.081\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 0.539 & 0.092\\
+\end{pmatrix}$ & 6 & 0.543 & 0.098\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.497 & 0.088\\
+\end{pmatrix}$ & 6 & 0.501 & 0.088\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.664 & 0.126\\
+\end{pmatrix}$ & 6 & 0.662 & 0.123\\
 $\begin{pmatrix}
@@
 1 & 1 & 0\\
-\end{pmatrix}$ & 7 & 0.482 & 0.086\\
+\end{pmatrix}$ & 6 & 0.467 & 0.079\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 0.589 & 0.126\\
+\end{pmatrix}$ & 6 & 0.583 & 0.13\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.66 & 0.127\\
+\end{pmatrix}$ & 6 & 0.659 & 0.124\\
 $\begin{pmatrix}
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 7 & 0.749 & 0.121\\
+\end{pmatrix}$ & 6 & 0.751 & 0.124\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 0.61 & 0.097\\
+\end{pmatrix}$ & 6 & 0.604 & 0.097\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 0.571 & 0.102\\
+\end{pmatrix}$ & 7 & 0.564 & 0.103\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 0.472 & 0.077\\
+\end{pmatrix}$ & 7 & 0.475 & 0.068\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 8 & 0.59 & 0.11\\
+\end{pmatrix}$ & 7 & 0.591 & 0.108\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 8 & 0.709 & 0.125\\
+\end{pmatrix}$ & 7 & 0.717 & 0.119\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 0.642 & 0.122\\
+\end{pmatrix}$ & 7 & 0.648 & 0.122\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 8 & 0.63 & 0.108\\
+\end{pmatrix}$ & 7 & 0.627 & 0.105\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 9 & 0.57 & 0.092\\
+\end{pmatrix}$ & 8 & 0.577 & 0.093\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 9 & 0.649 & 0.111\\
+\end{pmatrix}$ & 8 & 0.639 & 0.109\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 10 & 0.645 & 0.103\\
+\end{pmatrix}$ & 9 & 0.638 & 0.106\\
 \caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3}

### 2014-09-11 code/rds3x3.jl
compute cycle number and add to table

@@ -1,4 +1,7 @@
 using Permutations
+using PyCall
+# @pyimport numpy as np
+@pyimport networkx as nx
 
 function pmatrix(M::Array{Int64,2})
     lines = split(replace(replace(string(M),"[",""),"]",""),"\n")
@@ -22,7 +25,37 @@ function pmatrix(M::Array{Float64,2})
     return join(rv,"\n")
 end
 
+function cyclenumber(adjmatlistlist)
+
+    cyclecountlistlist = (Array{Int64,1})[]
+
+    for (i,adjmatlist) in enumerate(adjmatlistlist)
+
+        cyclecountlist = zeros(Int64,length(adjmatlist))
+
+        for (j,adjmat) in enumerate(adjmatlist)
+            g = nx.DiGraph(adjmat)
+            cc = nx.simple_cycles(g)
+            cc = pyeval("list(x)",x=cc)
+            if length(cc)>0
+                c = sum(map(x->length(x)>1,cc))
+            else
+                c = 0
+            end
+            cyclecountlist[j] = c
+        end
+
+        push!(cyclecountlistlist,cyclecountlist)
+
+    end
+
+    return cyclecountlistlist
+
+end
+
 function savelatex(pcN,sp,rsp)
+    cn = cyclenumber(pcN)
+    println(cn)
     rv = (String)[]
 
     for (i,plist) in enumerate(sp)
@@ -32,22 +65,25 @@ function savelatex(pcN,sp,rsp)
                   push!(rv,
                         string("\$", pmatstr, "\$", " & ",
                                "$(i-1)", " & ",
+                               "$(cn[i][j])", " & ",
                                "$(round(rsp[i][j],3))", " & ",
                                "$(round(sp[i][j],3))", "\\\\"))
             end
         end
     end
 
+    savedata("../data/stab3x3.tsv",sp,rsp,cn);
+
     return join(rv,"\n")
 end
 
-function savedata(filename,sp,rsp)
+function savedata(filename,sp,rsp,cn)
     rv = (String)[]
-    push!(rv,"connectivity\tstability\tresampling")
+    push!(rv,"connectivity\tstability\tresampling\tcycle")
     for (i,plist) in enumerate(sp)
         for (j,p) in enumerate(plist)
             if p > 1e-3
-                  line = "$(i-1)\t$(sp[i][j])\t$(rsp[i][j])"
[…]

### 2014-09-11 journal.tex
compute cycle number and add to table
add plots with cycle number

@@
 Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems.
+\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
+Fig.
+\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
+Fig.
+\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
+\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
+\label{fig:connectcycle3D3x3}
+\end{figure}
 
@@
 
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
+\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
+\label{fig:cycle3x3}
+\end{figure}
+
+\pagebreak
+
 % \begin{table}[h]
-\begin{longtable}{ c || c | c | c }
+\begin{longtable}{ c || c | c | c | c }
 \hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
+matrix & connectivity & cycle number & \specialcell{probability of stability\\to perturbation} & probability of stability\\
 \hline
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 3 & 0.499 & 0.126\\
+\end{pmatrix}$ & 3 & 0 & 0.496 & 0.122\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 4 & 0.505 & 0.121\\
+\end{pmatrix}$ & 4 & 1 & 0.496 & 0.13\\
 $\begin{pmatrix}
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 4 & 0.622 & 0.127\\
+\end{pmatrix}$ & 4 & 0 & 0.626 & 0.128\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.595 & 0.121\\
+\end{pmatrix}$ & 5 & 1 & 0.602 & 0.127\\
 $\begin{pmatrix}
@@
 1 & 1 & 0\\
-\end{pmatrix}$ & 5 & 0.494 & 0.128\\
+\end{pmatrix}$ & 5 & 2 & 0.496 & 0.122\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.41 & 0.061\\
+\end{pmatrix}$ & 5 & 2 & 0.405 & 0.061\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.43 & 0.078\\
+\end{pmatrix}$ & 5 & 1 & 0.429 & 0.081\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.605 & 0.12\\
+\end{pmatrix}$ & 5 & 1 & 0.597 & 0.129\\
 $\begin{pmatrix}
@@
 1 & 0 & 0\\
-\end{pmatrix}$ & 5 & 0.405 & 0.06\\
+\end{pmatrix}$ & 5 & 2 & 0.408 & 0.062\\
 $\begin{pmatrix}
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0.698 & 0.122\\
+\end{pmatrix}$ & 5 & 0 & 0.693 & 0.125\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 5 & 0.587 & 0.128\\
+\end{pmatrix}$ & 5 & 1 & 0.606 & 0.124\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 5 & 0.707 & 0.127\\
+\end{pmatrix}$ & 5 & 0 & 0.705 & 0.131\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.578 & 0.121\\
+\end{pmatrix}$ & 6 & 2 & 0.578 & 0.115\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.487 & 0.081\\
+\end{pmatrix}$ & 6 & 3 & 0.489 & 0.085\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.543 & 0.098\\
+\end{pmatrix}$ & 6 & 2 & 0.548 & 0.091\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.501 & 0.088\\
+\end{pmatrix}$ & 6 & 2 & 0.493 & 0.087\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.662 & 0.123\\
+\end{pmatrix}$ & 6 & 1 & 0.669 & 0.128\\
 $\begin{pmatrix}
@@
 1 & 1 & 0\\
-\end{pmatrix}$ & 6 & 0.467 & 0.079\\
+\end{pmatrix}$ & 6 & 3 & 0.474 & 0.086\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.583 & 0.13\\
+\end{pmatrix}$ & 6 & 2 & 0.587 & 0.12\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 6 & 0.659 & 0.124\\
+\end{pmatrix}$ & 6 & 1 & 0.656 & 0.121\\
 $\begin{pmatrix}
@@
 0 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.751 & 0.124\\
+\end{pmatrix}$ & 6 & 0 & 0.741 & 0.125\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0.604 & 0.097\\
+\end{pmatrix}$ & 6 & 1 & 0.596 & 0.098\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.564 & 0.103\\
+\end{pmatrix}$ & 7 & 3 & 0.572 & 0.094\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.475 & 0.068\\
+\end{pmatrix}$ & 7 & 5 & 0.465 & 0.075\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 0.591 & 0.108\\
+\end{pmatrix}$ & 7 & 3 & 0.592 & 0.105\\
 $\begin{pmatrix}
@@
 0 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.717 & 0.119\\
+\end{pmatrix}$ & 7 & 1 & 0.713 & 0.122\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 0.648 & 0.122\\
+\end{pmatrix}$ & 7 & 2 & 0.642 & 0.125\\
 $\begin{pmatrix}
@@
 1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 0.627 & 0.105\\
+\end{pmatrix}$ & 7 & 2 & 0.628 & 0.11\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 0.577 & 0.093\\
+\end{pmatrix}$ & 8 & 5 & 0.578 & 0.092\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 0.639 & 0.109\\
+\end{pmatrix}$ & 8 & 3 & 0.647 & 0.115\\
 $\begin{pmatrix}
@@
 1 & 1 & 1\\
-\end{pmatrix}$ & 9 & 0.638 & 0.106\\
+\end{pmatrix}$ & 9 & 5 & 0.635 & 0.11\\
 \caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3}

### 2014-09-11 journal.tex
add section on decomposition of stability computations in terms of strongly connected components

@@
 \usepackage{longtable}
+
+\def\tr{\mathrm{tr}}
+\def\Path{\mathrm{Path}}
+\def\hier{\mathrm{hier}}
+\def\Vertex{\mathrm{Vertex}}
 
@@
 In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
+
+\section{Decomposition of stability computations into strongly connected components}
+$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$
+where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of
+the matrix $M$ with $z$ a complex variable.  This formula is readily
+verified when $M$ is diagonal and has the nice property of being
+additive over decompositions of the matrix. Making the change of variables $x=\frac{1}{z}$, we have
+$$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p
+\in \Path_n (G)(a,a)} F(p)$$
+where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
+paths of length $n$ from $a$ to $b$.
+$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
+
+We now make use of the additivity.  Let $\hier$ be the functor which
+sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
+in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
+for the fibers of the points in the hierarchy which corresponds to a
+factorization of the characteristic polynomial.
 

### 2014-09-12 journal.tex
improve section on stability criteria

@@
 
-\section{Stability analysis of arbitrary systems}
+\section{Stability analysis of biological networks}
 
@@
 $$
-P(A' \, \textrm{stable} | A \, \textrm{stable})
+P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
 $$
@@
 
-\section{Decomposition of stability computations into strongly connected components}
+\section{System graph and decomposition into strongly connected components}
 $${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$
@@
 
-\section{Arbitrary system size}
+\section{Stability criteria}
 We call a matrix $M$ stable if all its eigenvalues have negative real
@@
 stable is equivalent to asking that the roots of $p$ all have real
-part.  Using some good old fashioned invariant theory, we will
-formulate conditions for this to happen in terms of the coefficients of
-$p$.
+part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
 
@@
 corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ iff the polynomial
+subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
 corresponding to $x$ has $a$ positive real roots, $b$ negative real
@@
 The explicit expressions may be obtained by such techniques as long
-division and determinants.
-
-These quantities are of interest because they describe the boundaries
-between the regions introduced above.  Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
+division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
 
@@
 s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized in the following table:
+is summarized by the conditions
 \begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0 \cr
-R_{1100} \qquad &D < 0, \delta > 0 \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0 \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0 \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0 \cr
+R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
+R_{1100} \qquad &D < 0, \delta > 0, \cr
+R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
+R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
+R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
 \end{align}
+These regions used for the classification of stability of two-component systems are show in Fig.
+\ref{fig:region2x2}.
+The stable regions are given by $R_{2000}$ and $R_{0020}$.
+All of the other regions represent unstable matrices.
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
+\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
+Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
+\label{fig:region2x2}
+\end{figure}
 
@@
 but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use some auxiliary equalities to
-specify which of the components contains our region.  (Looking at the
-sign of $\Delta$ will not help here because, while $\Delta > 0$ in the
-region $R_{3000}$, it is also positive for some elements of $R_{1200}$
-as well.)
+has two connected components and use the auxiliary condition that $s_1 > 0$ to
+specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
+%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
+\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
+This is the plane $s_3=1$.}
+\label{fig:region3x3}
+\end{figure}
 
 \section{Two component systems}
-We call a matrix stable when the real part of all its eigenvalues is negative.  For a $2 \times 2$ matrix, this amounts to the conditions $T < 0$ and $D >
+For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
 0$ where $T$ and $D$ denote the trace and the determinant.
@@
 The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} }
-\label{fig:region2x2}
-\end{figure}
 
@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} }
-\label{fig:region3x3}
-\end{figure}
-
 \bibliographystyle{amsplain}
-% \begin{thebibliography}{10}
 \bibliography{bib/books,bib/papers}
 
-% \bibitem {A} T. Aoki, \textit{Calcul exponentiel des op\'erateurs
-% microdifferentiels d'ordre infini.} I, Ann. Inst. Fourier (Grenoble)
-% \textbf{33} (1983), 227--250.
-
-% \bibitem {B} R. Brown, \textit{On a conjecture of Dirichlet},
-% Amer. Math. Soc., Providence, RI, 1993.
-
-% \bibitem {D} R. A. DeVore, \textit{Approximation of functions},
-% Proc. Sympos. Appl. Math., vol. 36,
-% Amer. Math. Soc., Providence, RI, 1986, pp. 34--56.
-
-% \end{thebibliography}
+\section{Appendix A}\label{sec:appA}
+Suppose that we have a family
+of polynomials whose roots are of one type but turn into another type.
+There are three ways this could happen:
+\begin{itemize}
+\item{1.} A real root could go from being positive to being negative
+  or vice versa.  For this to happen, we must pass through an
+  intermediate value with a zero root.  For that intermediate value,
+  we have $D = 0$.
+\item{2.} A pair of real roots could turn into a pair of complex roots
+  or vice versa.  For this to happen, we must pass through an
+  intermediate value with a repeated root.  For that intermediate
+  value, we have $\delta = 0$.
+\item{3.} A conjugate pair of complex roots could go from having
+  positive real part to having negative real part or vice versa.  For
+  this to happen, we must pass through an intermediate value in which
+  the real part is zero.  For that intermediate value, the two roots
+  are imaginary so, bing conjugates, they sum to zero, hence we have
+  $\Delta = 0$.
+\end{itemize}
+\noindent  Thus, we conclude that the boundaries of our regions
+$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
+\Delta=0$ and we will characterize them by studying these surfaces.
 

### 2014-09-12 journal.tex
add correspondence between dynamical system and Jacobian graph

@@
 \usepackage{longtable}
+\usepackage{dot2texi}
+\usepackage{tikz}
+\usetikzlibrary{automata,shapes,arrows}
 
@@
 \def\Path{\mathrm{Path}}
-\def\hier{\mathrm{hier}}
+\def\hier{\mathrm{Hier}}
 \def\Vertex{\mathrm{Vertex}}
+\def\adj{\mathrm{adj}}
 
@@
 
-\title{Stability of Ensembles of Biological Networks}
+\title{Ensemble Stability of Biological Networks}
 
@@
 We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
+We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
 These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
@@
 where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
+We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
+ \begin{displaymath}
+   \adj(G_X)_{ij} = \left\{
+     \begin{array}{ll}
+       1, & F_i \hbox{ depends on } x_j\\
+       0, & F_i \hbox{ does not depend on } x_j
+     \end{array}
+   \right.
+\end{displaymath}
+For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
+\begin{align*}
+\frac{dx_1}{dt} & = F_1(x_2)\\
+\frac{dx_2}{dt} & = F_2(x_1,x_2)
+\end{align*}
+we have the graph
+\begin{center}
+\begin{tikzpicture}[
+every state/.style={draw=red!50,very thick,fill=red!20}]
+\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
+digraph G {
+d2ttikzedgelabels = true;
+node [style="state"];
+edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
+x_1 -> x_2;
+x_2 -> x_1;
+x_2 -> x_2 [topath="loop above"];
+}
+\end{dot2tex}
+\end{tikzpicture}
+\end{center}
+with adjacency matrix
+$$
+\adj(G_X) = \begin{pmatrix}
+0 & 1 \\
+1 & 1
+\end{pmatrix}
+$$
+
 Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
@@
 \end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matirx $A$ has components
+where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
 $$
@@
 $$
+To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
+ \begin{displaymath}
+   \adj(G_A)_{ij} = \left\{
+     \begin{array}{lr}
+       1, & a_{ij} \neq 0\\
+       0, & a_{ij} = 0
+     \end{array}
+   \right.
+\end{displaymath}
+The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
+
 The spectral abscissa of the matrix $A$ is defined as
@@
 \section{System graph and decomposition into strongly connected components}
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$
-where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of
-the matrix $M$ with $z$ a complex variable.  This formula is readily
-verified when $M$ is diagonal and has the nice property of being
-additive over decompositions of the matrix. Making the change of variables $x=\frac{1}{z}$, we have
-$$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p
-\in \Path_n (G)(a,a)} F(p)$$
+We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
+The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
+$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
+This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
+Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
 where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of

### 2014-09-12 code/rds3x3.jl
add function to generate latex dot2texi code from adjacency matrices

@@ -3,6 +3,10 @@ using PyCall
 # @pyimport numpy as np
 @pyimport networkx as nx
 
+macro R_str(s)
+    s
+end
+
 function pmatrix(M::Array{Int64,2})
     lines = split(replace(replace(string(M),"[",""),"]",""),"\n")
     rv = (String)[]
@@ -22,6 +26,33 @@ function pmatrix(M::Array{Float64,2})
         push!(rv,string(join(split(l)," & "),"\\\\"))
     end
     push!(rv,"\\end{pmatrix}")
+    return join(rv,"\n")
+end
+
+
+function adjtograph(M::Array{Int64,2})
+    # lines = split(replace(replace(string(M),"[",""),"]",""),"\n")
+    rv = (String)[]
+    headstring = R"\begin{tikzpicture}[
+every state/.style={draw=red!50,very thick,fill=red!20}]
+\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
+digraph G {
+d2ttikzedgelabels = true;
+node [style=\"state\"];
+edge [lblstyle=\"auto\",topath=\"bend left\",style=\"line width=1.5pt\"];"
+    tailstring = R"}
+\end{dot2tex}
+\end{tikzpicture}"
+
+    push!(rv,headstring)
+
+    for i in 1:size(M)[1], j in 1:size(M)[2]
+        if M[i,j] != 0
+            push!(rv,string("$j -> $i\;"))
+        end
+    end
+
+    push!(rv,tailstring)
     return join(rv,"\n")
 end
 

### 2014-09-12 tablewithgraphs.tex
add table with graphs

\begin{longtable}{ c || c | c | c | c }
\hline
matrix & connectivity & cycle number & \specialcell{probability of stability\\to perturbation} & probability of stability\\
\hline
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 2 [topath="loop above"];;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 3 & 0 & 0.499 & 0.126\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 4 & 1 & 0.505 & 0.121\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 2 [topath="loop above"];;
3 -> 2;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 4 & 0 & 0.622 & 0.127\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.595 & 0.121\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 2 & 0.494 & 0.128\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 2 & 0.41 & 0.061\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.43 & 0.078\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.605 & 0.12\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 2 & 0.405 & 0.06\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 0 & 0.698 & 0.122\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 2 [topath="loop above"];;
3 -> 2;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.587 & 0.128\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 0 & 0.707 & 0.127\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.578 & 0.121\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 3 & 0.487 & 0.081\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.543 & 0.098\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.501 & 0.088\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 1 & 0.662 & 0.123\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 3 & 0.467 & 0.079\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.583 & 0.13\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 1 & 0.659 & 0.124\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 0 & 0.751 & 0.124\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 1 & 0.604 & 0.097\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 3 & 0.564 & 0.103\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 5 & 0.475 & 0.068\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 3 & 0.591 & 0.108\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 1 & 0.717 & 0.119\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 2 & 0.648 & 0.122\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 2 & 0.627 & 0.105\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 8 & 5 & 0.577 & 0.093\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 8 & 3 & 0.639 & 0.109\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 9 & 5 & 0.638 & 0.106\\
\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3graph}
\end{longtable}

### 2014-09-12 code/rds3x3.jl

@@ -33,7 +33,7 @@ end
 function adjtograph(M::Array{Int64,2})
     # lines = split(replace(replace(string(M),"[",""),"]",""),"\n")
     rv = (String)[]
-    headstring = R"\begin{tikzpicture}[
+    headstring = R"\begin{tikzpicture}[scale=0.6,
 every state/.style={draw=red!50,very thick,fill=red!20}]
 \begin{dot2tex}[styleonly,codeonly,neato,mathmode]
 digraph G {
@@ -48,7 +48,11 @@ edge [lblstyle=\"auto\",topath=\"bend left\",style=\"line width=1.5pt\"];"
 
     for i in 1:size(M)[1], j in 1:size(M)[2]
         if M[i,j] != 0
-            push!(rv,string("$j -> $i\;"))
+            if i != j
+                push!(rv,string("$j -> $i\;"))
+            else
+                push!(rv,string("$j -> $i [topath=\"loop above\"];\;"))
+            end
         end
     end
 
@@ -92,9 +96,11 @@ function savelatex(pcN,sp,rsp)
     for (i,plist) in enumerate(sp)
         for (j,p) in enumerate(plist)
             if p > 1e-3
-                  pmatstr = pmatrix(pcN[i][j])
+                  # pmatstr = pmatrix(pcN[i][j])
+                  pmatstr = adjtograph(pcN[i][j])
+                  # string("\$", pmatstr, "\$", " & ",
                   push!(rv,
-                        string("\$", pmatstr, "\$", " & ",
+                        string(pmatstr, " & ",
                                "$(i-1)", " & ",
                                "$(cn[i][j])", " & ",
                                "$(round(rsp[i][j],3))", " & ",

### 2014-09-12 code/rds3x3.jl
remove extra semicolon in dot2texi generator

@@ -51,7 +51,7 @@ edge [lblstyle=\"auto\",topath=\"bend left\",style=\"line width=1.5pt\"];"
             if i != j
                 push!(rv,string("$j -> $i\;"))
             else
-                push!(rv,string("$j -> $i [topath=\"loop above\"];\;"))
+                push!(rv,string("$j -> $i [topath=\"loop above\"]\;"))
             end
         end
     end

### 2014-09-12 journal.tex
add table with graphs
move table to separate file

@@
 
-\pagebreak
+\newpage
 
-% \begin{table}[h]
-\begin{longtable}{ c || c | c | c | c }
-\hline
-matrix & connectivity & cycle number & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-$\begin{pmatrix}
-1 & 0 & 0\\
-0 & 1 & 0\\
-0 & 0 & 1\\
-\end{pmatrix}$ & 3 & 0 & 0.496 & 0.122\\
-$\begin{pmatrix}
-0 & 0 & 1\\
-0 & 1 & 0\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 4 & 1 & 0.496 & 0.13\\
-$\begin{pmatrix}
-1 & 0 & 0\\
-0 & 1 & 1\\
-0 & 0 & 1\\
-\end{pmatrix}$ & 4 & 0 & 0.626 & 0.128\\
-$\begin{pmatrix}
-0 & 0 & 1\\
-0 & 1 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 1 & 0.602 & 0.127\\
-$\begin{pmatrix}
-0 & 0 & 1\\
-0 & 1 & 1\\
-1 & 1 & 0\\
-\end{pmatrix}$ & 5 & 2 & 0.496 & 0.122\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 0 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 2 & 0.405 & 0.061\\
-$\begin{pmatrix}
-0 & 1 & 0\\
-0 & 1 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 1 & 0.429 & 0.081\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 1 & 0\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 5 & 1 & 0.597 & 0.129\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 1 & 1\\
-1 & 0 & 0\\
-\end{pmatrix}$ & 5 & 2 & 0.408 & 0.062\\
-$\begin{pmatrix}
-1 & 0 & 1\\
-0 & 1 & 1\\
-0 & 0 & 1\\
-\end{pmatrix}$ & 5 & 0 & 0.693 & 0.125\\
-$\begin{pmatrix}
-1 & 0 & 0\\
-0 & 1 & 1\\
-0 & 1 & 1\\
-\end{pmatrix}$ & 5 & 1 & 0.606 & 0.124\\
-$\begin{pmatrix}
-1 & 0 & 1\\
-0 & 1 & 0\\
-0 & 1 & 1\\
-\end{pmatrix}$ & 5 & 0 & 0.705 & 0.131\\
-$\begin{pmatrix}
-0 & 0 & 1\\
-0 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 2 & 0.578 & 0.115\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 0 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 3 & 0.489 & 0.085\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 1 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 2 & 0.548 & 0.091\\
-$\begin{pmatrix}
-0 & 1 & 0\\
-0 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 2 & 0.493 & 0.087\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 1 & 0\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 6 & 1 & 0.669 & 0.128\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 1 & 1\\
-1 & 1 & 0\\
-\end{pmatrix}$ & 6 & 3 & 0.474 & 0.086\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-1 & 1 & 0\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 2 & 0.587 & 0.12\\
-$\begin{pmatrix}
-1 & 0 & 1\\
-0 & 1 & 1\\
-0 & 1 & 1\\
-\end{pmatrix}$ & 6 & 1 & 0.656 & 0.121\\
-$\begin{pmatrix}
-1 & 1 & 1\\
-0 & 1 & 1\\
-0 & 0 & 1\\
-\end{pmatrix}$ & 6 & 0 & 0.741 & 0.125\\
-$\begin{pmatrix}
-1 & 1 & 0\\
-0 & 1 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 6 & 1 & 0.596 & 0.098\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-0 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 3 & 0.572 & 0.094\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-1 & 0 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 5 & 0.465 & 0.075\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-1 & 1 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 3 & 0.592 & 0.105\\
-$\begin{pmatrix}
-1 & 1 & 1\\
-0 & 1 & 1\\
-0 & 1 & 1\\
-\end{pmatrix}$ & 7 & 1 & 0.713 & 0.122\\
-$\begin{pmatrix}
-1 & 0 & 1\\
-0 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 7 & 2 & 0.642 & 0.125\\
-$\begin{pmatrix}
-1 & 1 & 1\\
-0 & 1 & 1\\
-1 & 0 & 1\\
-\end{pmatrix}$ & 7 & 2 & 0.628 & 0.11\\
-$\begin{pmatrix}
-0 & 1 & 1\\
-1 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 5 & 0.578 & 0.092\\
-$\begin{pmatrix}
-1 & 1 & 1\\
-0 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 8 & 3 & 0.647 & 0.115\\
-$\begin{pmatrix}
-1 & 1 & 1\\
-1 & 1 & 1\\
-1 & 1 & 1\\
-\end{pmatrix}$ & 9 & 5 & 0.635 & 0.11\\
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3}
-\end{longtable}
-%
-% \end{table}
+\input{tablenographs}
+
+\newpage
+
+\input{tablewithgraphs}
 

### 2014-09-12 paper.tex
reorganize

--- 2014-09-12 journal.tex
@@
 %------------------------------------------------------------------------------
-% Beginning of journal.tex
+% Beginning of paper.tex
 %------------------------------------------------------------------------------
@@
 
-\section{Stability analysis of biological networks}
+\section{Stability analysis of biological networks}\label{sec:stabanalbn}
 
@@
 \begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
+\input{tex/examplehamdyngraph}
 \end{center}
@@
 The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
+$$\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
 This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
@@
 \section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
+Suppose that $M$ is an $n \times n$ matrix with real coefficients.
+Motivated by the discussion of Section \ref{sec:stabanalbn}, we will call the matrix $M$ stable if all its eigenvalues have negative real part.
+Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
+Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
+Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that $M$ be stable is equivalent to asking that the roots of $\chi$ all have negative real part.
+We will formulate conditions for this to happen in terms of the coefficients of $\chi$.
 
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
+Assume that $M$ is stable and that it has $2m$ complex roots which come in complex conjugate pairs and $n-m$ real roots.
+Let $r_1 \ldots, r_{2m}$ be the complex roots with $r_{2k+1}$ being the complex conjugate of $r_{2k}$ and let $r_{2m+1}, \ldots,  r_{n}$ be the real roots.
+Then we can write the factorization as follows:
+\[
+\chi(M)(z) = \prod_{k=1}^{m} (z^2 - (r_{2k} + r_{2k+1}) z + r_{2k} r_{2k+1})
+             \prod_{k=2m+1}^{n} (z - r_k)
+\]
+Since $r_{2k}$ and $r_{2k+1}$ are complex conjugates, $r_{2k} + r_{2k+1} =
+\Re r_{2k} = \Re r_{2k+1}$ and $r_{k} r_{2k+1} > 0$.   Since all the
+roots have negative real part, all the coefficients of the terms in
+the products above are positive so, when we multiply out the product,
+we will have all positive terms.  Hence, we conclude that a necessary
+condition for $M$ to be stable is that $s_k > 0$ for all $k$.
 
@@
 d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
 \end{align*}
@@
 
-\newpage
-
-\input{tablenographs}
+\bibliographystyle{amsplain}
+\bibliography{bib/books,bib/papers}
 
@@
 
-\input{tablewithgraphs}
+\input{tex/tablenographs}
 
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
+\newpage
+
+\input{tex/tablewithgraphs}
 

### 2014-09-12 paper.tex
add todonotes and autoref labelling

@@
 \usetikzlibrary{automata,shapes,arrows}
+% todo notes see http://www.texample.net/tikz/examples/todo-notes/
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{hyperref}
+\hypersetup{colorlinks=true,
+linkcolor=[rgb]{.67 .27 .27},
+citecolor=[rgb]{.09 .29 .54},
+urlcolor=[rgb]{.09 .29 .54}}
+
+% Set autoref text
+% http://tex.stackexchange.com/a/36576/6784
+\renewcommand*{\figureautorefname}{Fig.}
+\renewcommand*{\equationautorefname}{Eq.}
+\renewcommand*{\tableautorefname}{Table}
+\renewcommand*{\sectionautorefname}{Section}
 
@@
 \begin{document}
+
+% Add Figure, Table prefixes to references
+% http://tex.stackexchange.com/a/6063/6784
+\let\ref\autoref
 
@@
 
+% \section{Todo}
+% \listoftodos
+
 \section{Introduction}
@@
 \section{Stability analysis of biological networks}\label{sec:stabanalbn}
-
+\todo{Describe dynamical systems in terms of vector fields on Manifolds \cite{Deville}}
 Consider an n-dimensional vector
@@
 $$
+\todo{Define strongly connected components and connect to operads \cite{Vagner2014}}
 
@@
 The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
+This is because the general solution to \ref{eq:lineardynsys} is
 $$
@@
 Suppose that $M$ is an $n \times n$ matrix with real coefficients.
-Motivated by the discussion of Section \ref{sec:stabanalbn}, we will call the matrix $M$ stable if all its eigenvalues have negative real part.
+Motivated by the discussion of \ref{sec:stabanalbn}, we will call the matrix $M$ stable if all its eigenvalues have negative real part.
 Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
@@
 \end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
+These regions used for the classification of stability of two-component systems are show in \ref{fig:region2x2}.
 The stable regions are given by $R_{2000}$ and $R_{0020}$.
@@
 has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
+specify which of the components contains our region. The regions analogous to \ref{fig:region2x2} for three component systems where $D=1$ is given in \ref{fig:region3x3}.
 %(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
@@
 The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
 \ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
 \ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
 \ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
@@
 
+\newpage
+
 \bibliographystyle{amsplain}
@@
 
-\input{tex/tablewithgraphs}
+% \input{tex/tablewithgraphs}
 

### 2014-09-13 paper.tex
explain stability analysis

@@
 \section{Stability analysis of biological networks}\label{sec:stabanalbn}
-\todo{Describe dynamical systems in terms of vector fields on Manifolds \cite{Deville}}
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
+In the construction of a class of potential models for biological systems at any level of the biological hierarchy from metabolic to ecosystem-level networks, it is common to first attempt to define a collection of observable phenomena of interest and determine a domain (such as binary numbers, integers, or real numbers) in which each observable can be quantified.
+Next, it is necessary to establish the interdependencies among system components.
+The specific manner in which the components depend upon one another must be clarified, which is often done by determining a particular system of mathematical functions that represents a hypothesis about how the quantified observables evolve in time.
+To the degree to which there is uncertainty about the interactions among system components, parameters are introduced to broaden the class of models under consideration.
+Finally, whatever model class remains may be compared to empirical observations to determine how capable the model is of representing the phenomena of interest.
+
+For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold to each observable, a directed graph to the interdependencies, and a vector field satisfying these interdependencies for each observable over the space determined by taking the product of the manifolds associated to the collection of observables \cite{Deville}.
+For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
+If the system under consideration has the graph
 \begin{center}
@@
 $$
+then for the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
+\begin{align*}
+\frac{dx_1}{dt} & = F_1(x_2)\\
+\frac{dx_2}{dt} & = F_2(x_1,x_2)
+\end{align*}
+
+In the more general case of a system with $n$ components we would have an $n$-dimensional vector of observables
+$$
+x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
+$$
+whose components are solutions to the arbitrary first order system
+$$
+\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
+$$
+where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
+
+This description can be formalized as a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
+ \begin{displaymath}
+   \adj(G_X)_{ij} = \left\{
+     \begin{array}{ll}
+       1, & F_i \hbox{ depends on } x_j\\
+       0, & F_i \hbox{ does not depend on } x_j
+     \end{array}
+   \right.
+\end{displaymath}
 \todo{Define strongly connected components and connect to operads \cite{Vagner2014}}

### 2014-09-15 paper.tex
introduce parameters and work on char poly section

@@
 
+In order to accommodate the possibility of uncertainty in our modelling,
+we will generalize our notion of a system on a network to that of a
+system with random parameters.  Again, this will involve three
+steps. First, we provide a measure space $S$ which represents the
+values over which our parameters can vary. Next, instead of a single
+vector field, we consider a family of vector fields parameterized by
+this space. That is to say, we introduce a mapping $V$ from $S$ to
+sections of the tangent bundle of the product of manifolds. As before,
+for each point $p \in S$, the vector field $V(p)$ must be consistent with
+the system graph. Finally, we select a probability measure $\mu$ on
+$S$ which represents our understanding of which values of the parameters
+are likely. In accord with Bayesian statistics, we might revise this
+distribution as data comes in or use it to estimate parameters and error
+bars from experimental results.
+
+Having done this, we are now in a position to do a probabilistic analysis.
+Given some quantity $q$ characterizing our system (such as, for instance,
+the time it would take a particle to travel between two points), this
+quantity now becomes a random variable.  We may then study the average
+value, variance, and the like.  Another question one can ask is that of
+robustness, which is the conditional probability that, if some property
+holds for a set of parameters, it will still hold if we make some random
+perturbation about those parameter values.
+
 This description can be formalized as a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
@@
 
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$$\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
 \section{Stability criteria}
 Suppose that $M$ is an $n \times n$ matrix with real coefficients.
-Motivated by the discussion of \ref{sec:stabanalbn}, we will call the matrix $M$ stable if all its eigenvalues have negative real part.
+Motivated by the discussion of \ref{sec:stabanalbn}, we will call
+the matrix $M$ stable if all its eigenvalues have negative real part.
 Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
-Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
-Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that $M$ be stable is equivalent to asking that the roots of $\chi$ all have negative real part.
-We will formulate conditions for this to happen in terms of the coefficients of $\chi$.
+Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$,
+writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
+Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
+$M$ be stable is equivalent to asking that the roots of $\chi$ all have
+negative real part.  We will formulate conditions for this to happen in
+terms of the coefficients of $\chi$.
 
-Assume that $M$ is stable and that it has $2m$ complex roots which come in complex conjugate pairs and $n-m$ real roots.
-Let $r_1 \ldots, r_{2m}$ be the complex roots with $r_{2k+1}$ being the complex conjugate of $r_{2k}$ and let $r_{2m+1}, \ldots,  r_{n}$ be the real roots.
-Then we can write the factorization as follows:
+\begin{lemma}
+If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
+\end{lemma}
+
+\begin{proof}
+Assume that $M$ is stable and that it has $2m$ complex roots which come in
+complex conjugate pairs and $n-m$ real roots.  Let $r_1 \ldots, r_{2m}$ be
+the complex roots with $r_{2k+1}$ being the complex conjugate of $r_{2k}$
+and let $r_{2m+1}, \ldots,  r_{n}$ be the real roots. Then we can write the
+factorization as follows:
 \[
@@
 condition for $M$ to be stable is that $s_k > 0$ for all $k$.
+\end{proof}
+
+To state the next criterion, we will require the quantity $\Delta$, which
+is defined as $\prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j)$.  Since
+this expression is invariant under permuting the roots, it may be expressed
+as a function of the coefficients $s_i$; for small values of $n$, the
+resulting expression looks as follows:
+\begin{align*}
+n=2 \qquad &\Delta = s_1 \\
+n=3 \qquad &\Delta = s_3 - s_1 s_2 \\
+n=4 \qquad &\Delta = s_1 s_2 s_3 - s_3^2 - s_1^2 s_4 \\
+n=5 \qquad &\Delta = s_1 s_2 s_3 s_4 - s_3^2 s_4 - s_1^2 s_4^2 -
+                     s_1 s_2^2 s_5 + s_2 s_3 s_5 + 2 s_1 s_4 s_5 - s_5^2
+\end{align*}
+
+\begin{lemma}
+Given a polynomial $p(z) = z^n + \sum_{k=1}^{n-1} s_k z^{n-k}$ and a real number
+$\epsilon > 0$, there exists $\delta \in \mathbb{R}$ such that, if $p'(x) =
+x^n + \sum_{k=1}^{n-1} {s'}_i z^i$ is a polynomial for which $|s_i - {s'}_i|
+< \delta$ then to every root $r$ of $p$ there corresponds a root $r'$ of
+$p'$ such that $|r-r'| < \epsilon$.
+\end{lemma}
+
+\begin{proof}
+Without loss of generality, we may assume that $\epsilon < 1$.
+
+We first consider the case where $p = x^n$.
+
+If we make the rescaling $x = \epsilon y/2$ in the equation $p(x)  = 0$,
+we obtain the equation
+\[
+y^n + \sum_{k=1}^{n-1} \left(\frac{2}{\epsilon}\right)^k s_k y^{n-k} = 0 .
+\]
+Cauchy's bound on the roots of this equation is
+\[
+|y| < 1 + \max_j \left(\frac{2}{\epsilon}\right)^j |s_j| ,
+\]
+which becomes the bound
+\[
+|x| < \frac{\epsilon}{2} + \max_j \left(\frac{2}{\epsilon}\right)^{j-1} |s_j|
+< \frac{\epsilon}{2} + \left(\frac{2}{\epsilon}\right)^{n-1} \max_j |s_j|
+\]
+for the original equation.  Hence, if we choose $\delta = (\epsilon/2)^{n}$,
+we will have $|x| < \epsilon$.
+
+Next, we generalize to the case $p = (x-r)^n$.
+
+To extend this result of the general case, we note that, given two
+polynomials $p = x^m + \sum_{k=1}^n a_k x^{m-k}$ and
+$q = x^n + \sum_{k=1}^n b_k x^{n-k}$, the Jacobian of the map which
+sends the coefficients of the matrices to the coefficients of the
+product $pq$ is the Sylvester matrix, whose determinant is the resultant
+of $p$ and $q$.  Hence, is $p$ and $q$ have no common root, the map
+is locally invertible.
+\end{proof}
+
+\begin{lemma}
+If a point lies on the boundary of the set of stable polynomials, then
+either $s_n = 0$ and $\Delta = 0$.
+\end{lemma}
+
+\begin{proof}
+Since the mapping which sends the roots of a polynomial to its
+coefficients has non-singular Jacobian when the discriminant is not
+zero, if a stable polynomial has distinct roots, it does not lie on
+the boundary.
+
+Suppose that we have a family
+of polynomials whose roots are of one type but turn into another type.
+There are three ways this could happen:
+\begin{itemize}
+\item{1.} A real root could go from being positive to being negative
+  or vice versa.  For this to happen, we must pass through an
+  intermediate value with a zero root.  For that intermediate value,
+  we have $D = 0$.
+\item{2.} A pair of real roots could turn into a pair of complex roots
+  or vice versa.  For this to happen, we must pass through an
+  intermediate value with a repeated root.  For that intermediate
+  value, we have $\delta = 0$.
+\item{3.} A conjugate pair of complex roots could go from having
+  positive real part to having negative real part or vice versa.  For
+  this to happen, we must pass through an intermediate value in which
+  the real part is zero.  For that intermediate value, the two roots
+  are imaginary so, bing conjugates, they sum to zero, hence we have
+  $\Delta = 0$.
+\end{itemize}
+\noindent  Thus, we conclude that the boundaries of our regions
+$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
+\Delta=0$ and we will characterize them by studying these surfaces.
+
+To complete the proof, we show that the condition $\delta = 0$ is not
+needed to describe the boundary.  Suppose that $\delta = 0$.  Then there
+must exist a repeated root.  If the real part of any root, repeated or
+not, is positive, then the polynomial is not stable and, by continuity,
+if one allows the coefficients to vary within a sufficiently small
+ball about that solution, the root will remain negative, so the
+polynomial cannot lie on the boundary of the stable region.  If all
+the roots, including the repeated ones, are negative then, by a similar
+argument the point is in the interior of the stable region.  The only
+other possibility is that some root has zero real part, but then
+$\Delta = 0$ so the condition $\delta = 0$ is superfluous.
+\end{proof}
 
@@
 \end{figure}
+
+\section{System graph and decomposition into strongly connected components}
+Next, we will want to relate prpoperties of our matrix and its eigenvalues
+to  properties of the associated graph $G_A$.  To do this, we will first
+examine the category of paths on $G_A$, and derive a formula expressing
+the chararacteristic polynomial as a sum over cycles.
+
+Given a digraph $G$, let $\Path (a,b)$ denote the set of paths from the
+vertex $a$ to the vertex $b$.  Given a path $p_1$ from $a$ to $b$ and a path
+$p_2$ from $b$ to $c$, we may concatenate them to obtain a path $p_1 \circ
+p_2$ from $a$ to $c$.  Hence, $(\Vertex(G), \Path(G), \circ)$ forms a category.
+
+The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
+$$\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
+This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
+Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
+where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
+paths of length $n$ from $a$ to $b$.  ; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
+
+We now make use of the additivity.  Let $\hier$ be the functor which
+sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
+in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
+for the fibers of the points in the hierarchy which corresponds to a
+factorization of the characteristic polynomial.
 
@@
 \section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
 

### 2014-09-15 paper.tex
add figure to demonstrate system composed of modules

@@
 For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
-If the system under consideration has the graph
-\begin{center}
-\input{tex/examplehamdyngraph}
-\end{center}
+If the system under consideration has the graph shown in \ref{fig:examplesystemmodules}A
+% \begin{center}
+% \input{tex/examplehamdyngraph}
+% \end{center}
 with adjacency matrix
@@
 \end{align*}
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.6\columnwidth]{fig/examplesystemmodules.pdf}
+\caption{{\bf Example of the combination of open system modules to construct closed systems.} (A) Example of combining two open modules to construct a closed system of two components (B) Analogous example for combining three open modules to construct a closed system with three components}
+\label{fig:examplesystemmodules}
+\end{figure}
 

### 2014-09-16 paper.tex
add definition of SCCs

@@
 \end{displaymath}
-\todo{Define strongly connected components and connect to operads \cite{Vagner2014}}
+A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components.
+A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
+The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc}.
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.7\columnwidth]{fig/scc.pdf}
+\caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
+Note that the most hierarchical system in (A) has the highest possible number of connected components, three, whereas the system containing a single cycle and therefore posessing no hierarchy contains only one connected component.
+Systems (B) and (D) represent examples of hierarchical modular systems that posess both modularity and hierarchy.}
+\label{fig:scc}
+\end{figure}
 

### 2014-09-16 paper.tex
move tables/figures to end and refactor into separate files

--- 2014-09-15 paper.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of paper.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-% todo notes see http://www.texample.net/tikz/examples/todo-notes/
-\usepackage[colorinlistoftodos]{todonotes}
-\usepackage{hyperref}
-\hypersetup{colorlinks=true,
-linkcolor=[rgb]{.67 .27 .27},
-citecolor=[rgb]{.09 .29 .54},
-urlcolor=[rgb]{.09 .29 .54}}
-
-% Set autoref text
-% http://tex.stackexchange.com/a/36576/6784
-\renewcommand*{\figureautorefname}{Fig.}
-\renewcommand*{\equationautorefname}{Eq.}
-\renewcommand*{\tableautorefname}{Table}
-\renewcommand*{\sectionautorefname}{Section}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
+\input{ams_header.tex}
 
@@
 
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold to each observable, a directed graph to the interdependencies, and a vector field satisfying these interdependencies for each observable over the space determined by taking the product of the manifolds associated to the collection of observables \cite{Deville}.
+For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph $G$ to the interdependencies, and a vector field $V \in T(\prod_i M_i)$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
 For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
-If the system under consideration has the graph
-\begin{center}
-\input{tex/examplehamdyngraph}
-\end{center}
+If the system under consideration has the graph given in \ref{fig:examplesystemmodules}A
+% \begin{center}
+% \input{tex/examplehamdyngraph}
+% \end{center}
 with adjacency matrix
@@
 system with random parameters.  Again, this will involve three
-steps. First, we provide a measure space $S$ which represents the
+steps. First, we provide a measure space $S, \mathcal{B}(S)$ which represents the
 values over which our parameters can vary. Next, instead of a single
 vector field, we consider a family of vector fields parameterized by
-this space. That is to say, we introduce a mapping $V$ from $S$ to
-sections of the tangent bundle of the product of manifolds. As before,
-for each point $p \in S$, the vector field $V(p)$ must be consistent with
-the system graph. Finally, we select a probability measure $\mu$ on
-$S$ which represents our understanding of which values of the parameters
+this space. That is to say, we introduce a mapping $V \colon S \to \Gamma T(\prod_i M_m)$.
+(Here $\Gamma$ denoted the space of sections of the tangent bundle.)
+As before, for each point $p \in S$, the vector field $V(p)$ must be consistent with
+the system graph. Finally, we select a probability measure $\mu \colon
+\mathcal{B}(S) \to [0,1]$ which represents our understanding of which values of the parameters
 are likely. In accord with Bayesian statistics, we might revise this
@@
 Having done this, we are now in a position to do a probabilistic analysis.
-Given some quantity $q$ characterizing our system (such as, for instance,
-the time it would take a particle to travel between two points), this
-quantity now becomes a random variable.  We may then study the average
-value, variance, and the like.  Another question one can ask is that of
-robustness, which is the conditional probability that, if some property
-holds for a set of parameters, it will still hold if we make some random
-perturbation about those parameter values.
+Given some quantity $q$ characterizing our system, this quantity now
+becomes a random variable.  This variable may be discrete of continuous
+depending on the quantity under consideration.  For example, if the quantity
+is the time it takes for a particle to travel between two points, we have
+a real-valued random variable; if the quantity is the number of fixed
+points, we have an integer-valued random variable; and if the quantity is
+whether or not the system posesses a limit cycle, we have a binary random
+variable.
+
+Another question one can ask is that of robustness, which measures the
+probability that, if some property holds for a set of parameters, it will
+continue to hold if we make a random perturbation about those parameter values.
+To define robustness, we will introduce, in addition to the probability
+distribution $\mu$ described above, a family of probability distributions
+$\mu' \colon S \times \mathcal{B}(S) \to [0,1]$.  Then, given a binary random
+variable $q$, we we define its robustness as the following conditional proability:
+\[
+  \mathrm{Robust} (q) =
+  \frac{\int_S d\mu(x) \int_S d\mu'(y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
+  {\int_S d\mu(x) \mathbf{1}_q(x)}
+\]
 
@@
 \end{displaymath}
-\todo{Define strongly connected components and connect to operads \cite{Vagner2014}}
+A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components.
+A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
+The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc}.
 
@@
 \begin{lemma}
-Given a polynomial $p(z) = z^n + \sum_{k=1}^{n-1} s_k z^{n-k}$ and a real number
-$\epsilon > 0$, there exists $\delta \in \mathbb{R}$ such that, if $p'(x) =
-x^n + \sum_{k=1}^{n-1} {s'}_i z^i$ is a polynomial for which $|s_i - {s'}_i|
-< \delta$ then to every root $r$ of $p$ there corresponds a root $r'$ of
-$p'$ such that $|r-r'| < \epsilon$.
+Given a polynomial $p(x) = z^n + \sum_{k=1}^{n-1} s_k x^{n-k}$ and a real number
+$\epsilon > 0$, there exists $\delta > 0 \in \mathbb{R}$ such that, if $q(x) =
+x^n + \sum_{k=1}^{n-1} \sigma_i x^i$ is a polynomial for which $|s_i - \sigma_i|
+< \delta$ then one can order the roots of $p$ and $q$ such that $p(z) = \prod_{k=1}^n
+(z - r_i)$ and $q(z) = \prod_{k=1}^n (z - \rho_i)$ and $| r_k - \rho_k | < \epsilon$.
 \end{lemma}
@@
 
-We first consider the case where $p = x^n$.
+We first consider the case where $p(x) = x^n$.
 
@@
 \begin{proof}
-Since the mapping which sends the roots of a polynomial to its
-coefficients has non-singular Jacobian when the discriminant is not
-zero, if a stable polynomial has distinct roots, it does not lie on
-the boundary.
+Suppose that none of the real parts of the roots of a polynomial $p$ equal
+zero.  Let $\epsilon$ be the minimum of the absolute values of the real
+parts of the roots of $p$.  Then, by continuity of the map which sends the
+roots of a polynomial to its coefficients, there exists a $\delta > 0$
+such that, if $q$ is any polynomial whose coefficients differ by less than
+$\delta$ from the
 
@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
-Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
-
 Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
@@
 %(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
-This is the plane $s_3=1$.}
-\label{fig:region3x3}
-\end{figure}
 
@@
 
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$$\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
+The characteristic polynomial $\chi(A)$ associated to $A$ can be factored
+according to the strongly connected components of the graph $G$.
+\[
+\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}
+\]
+where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the
+matrix $M$ with $z$ a complex variable.  This formula is readily verified when
+$M$ is diagonal and has the nice property of being additive over decompositions
+of the matrix. Making the change of variables $x=\frac{1}{z}$, we have
+\[
+\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)
+\]
 where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.  ; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
+paths of length $n$ from $a$ to $b$.  ; $F$ is the functor from this category
+to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
+
+As an illustration of this formula, we will use it to compute the characteristic
+polynomial of the matrix associated to the labelled graph in figure ??.  We begin
+by noting that the closed paths of this graph can be described by the following
+regular expressions:
+\begin{align*}
+\Path (1,1) &: (e + bd + bca)^* \\
+\Path (2,2) &: (ae^*(be^*d)^*bc)^* \\
+\Path (3,3) &: ((ca + d)e^*b)^* \\
+\end{align*}
+Since these regular expressions are non-redundant, we can obtain generating
+functions for the paths by considering $a,b,c,d,e$ as non-commutative
+variables and replacing the Kleene star $x^*$ by $(1 - x)^{-1}$:
+\begin{align*}
+\Path (1,1) &: (1 - e - bd - bca)^{-1} \\
+\Path (2,2) &: (1 - a(1-e)^{-1}(1 - b(1-e)^{-1}d)^{-1}bc)^{-1} \\
+\Path (3,3) &: (1 - (ca + d)(1 - e)^{-1}b)^{-1} \\
+\end{align*}
+To obtain the characteristic polynomial, we make the substitutions
+$a \mapsto m_{12}/z, b \mapsto m_{31}/z, c \mapsto m_{23}/z, d \mapsto m_{13}/z,
+e \mapsto m_{11}/z$ and simplify, using the fact that the matrix entries
+commute:
+\begin{align*}
+\Path (1,1) &: \frac{z^3}
+{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} } \\
+\Path (2,2) &: \frac{z(z^2 - m_{11} z - m_{13} m_{31})}
+{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} }  \\
+\Path (3,3) &: \frac{z^2(z^2 - m_{11})}
+{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} } \\
+\end{align*}
+Adding these three quantitites together, we obtain
+\[
+\frac{z(3 z^2 - 2 m_{11} z - m_{13} m_{31})}
+{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} },
+\]
+which we recognize as $\chi'(z)/\chi(z)$, where
+$\chi(z) = z^3−m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31}$.
+finally, we note that this polynomial is indeed the characteristic
+polynomial of the matrix
+\[
+ \begin{matrix}
+   m_{11} & m_{12} & m_{13} \\
+   0      & 0      & m_{23} \\
+   m_{31} & 0      & 0 \\
+ \end{matrix}
+\]
+corresponding to our digraph.
 
@@
 
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\begin{table}[h]
-\begin{tabular}{ c || c | c | c }
-\hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-  $\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
-  $\begin{pmatrix}
-a & b \\
-d & 0
-\end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
-  $\begin{pmatrix}
-a & 0 \\
-d & c
-\end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
-$\begin{pmatrix}
-a & 0 \\
-0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
-\end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
-
 The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
-
 \newpage
@@
 \newpage
+\FloatBarrier
 
-\input{tex/tablenographs}
+\section{Figure Legends}
+\input{tex/figures}
 
 \newpage
+\FloatBarrier
 
+\section{Tables}
+\input{tex/table2x2stab}
+\input{tex/tablenographs}
 % \input{tex/tablewithgraphs}
-
-\section{Appendix A}\label{sec:appA}
 

### 2014-09-16 tex/figures.tex
move tables/figures to end and refactor into separate files

--- 2014-09-12 journal.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of journal.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
+%!TEX root = ../paper.tex
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.6\columnwidth]{fig/examplesystemmodules.pdf}
+\caption{{\bf Example of the combination of open system modules to construct closed systems.} (A) Example of combining two open modules to construct a closed system of two components (B) Analogous example for combining three open modules to construct a closed system with three components}
+\label{fig:examplesystemmodules}
+\end{figure}
 
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
-
-\begin{document}
-
-\title{Ensemble Stability of Biological Networks}
-
-%    Information for first author
-\author{Aviv Bergman}
-%    Address of record for the research reported here
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{aviv@einstein.yu.edu}
-%    \thanks will become a 1st page footnote.
-% \thanks{The first author was supported in part by NSF Grant \#000000.}
-
-%    Information for second author
-\author{Raymond S. Puzio}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{rsp@novres.org}
-% \thanks{Support information for the second author.}
-
-%    Information for third author
-\author{Cameron Smith}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{cameron.smith@med.einstein.yu.edu}
-% \thanks{Support information for the second author.}
-
-%    General info
-% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
-
-% \date{January 1, 2001 and, in revised form, June 22, 2001.}
-
-% \dedicatory{This paper is dedicated to our advisors.}
-
-% \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Stability analysis of biological networks}
-
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
-\begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
-\end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
-
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
-
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
-
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.7\columnwidth]{fig/scc.pdf}
+\caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
+Note that the most hierarchical system in (A) has the highest possible number of connected components, three, whereas the system containing a single cycle and therefore posessing no hierarchy contains only one connected component.
+Systems (B) and (D) represent examples of hierarchical modular systems that posess both modularity and hierarchy.}
+\label{fig:scc}
+\end{figure}
 
@@
 
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
-%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
 \begin{figure}[!ht]
@@
 \end{figure}
-
-\section{Two component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\begin{table}[h]
-\begin{tabular}{ c || c | c | c }
-\hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-  $\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
-  $\begin{pmatrix}
-a & b \\
-d & 0
-\end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
-  $\begin{pmatrix}
-a & 0 \\
-d & c
-\end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
-$\begin{pmatrix}
-a & 0 \\
-0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
-\end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\section{Three component systems}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
 
@@
 \end{figure}
-
-\newpage
-
-\input{tablenographs}
-
-\newpage
-
-\input{tablewithgraphs}
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------

### 2014-09-16 tex/table2x2stab.tex
move tables/figures to end and refactor into separate files

--- 2014-09-12 journal.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of journal.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
-
-\begin{document}
-
-\title{Ensemble Stability of Biological Networks}
-
-%    Information for first author
-\author{Aviv Bergman}
-%    Address of record for the research reported here
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{aviv@einstein.yu.edu}
-%    \thanks will become a 1st page footnote.
-% \thanks{The first author was supported in part by NSF Grant \#000000.}
-
-%    Information for second author
-\author{Raymond S. Puzio}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{rsp@novres.org}
-% \thanks{Support information for the second author.}
-
-%    Information for third author
-\author{Cameron Smith}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{cameron.smith@med.einstein.yu.edu}
-% \thanks{Support information for the second author.}
-
-%    General info
-% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
-
-% \date{January 1, 2001 and, in revised form, June 22, 2001.}
-
-% \dedicatory{This paper is dedicated to our advisors.}
-
-% \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Stability analysis of biological networks}
-
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
-\begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
-\end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
-
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
-
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
-
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
-Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
-
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
-%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
-This is the plane $s_3=1$.}
-\label{fig:region3x3}
-\end{figure}
-
-\section{Two component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
+%!TEX root = ../paper.tex
 \newcommand{\specialcell}[2][c]{%
@@
 \end{table}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\section{Three component systems}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
-
-\newpage
-
-\input{tablenographs}
-
-\newpage
-
-\input{tablewithgraphs}
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------

### 2014-09-16 paper.tex
switch to different style

@@
-\input{ams_header.tex}
+\input{header.tex}
 
@@
 
-%    Information for first author
-\author{Aviv Bergman}
-%    Address of record for the research reported here
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{aviv@einstein.yu.edu}
-%    \thanks will become a 1st page footnote.
-% \thanks{The first author was supported in part by NSF Grant \#000000.}
+\pagenumbering{arabic}
 
-%    Information for second author
-\author{Raymond S. Puzio}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{rsp@novres.org}
-% \thanks{Support information for the second author.}
+% Title must be 150 characters or less
+\begin{center}
+{\Large
+\textbf{Ensemble Stability of Biological Networks}
+}
+% Insert Author names, affiliations and corresponding author email.
+\\[.5cm]
+Aviv Bergman$^{1,2,3,4}$,
+Raymond S. Puzio$^{1}$,
+Cameron Smith$^{1}$,
+\\[.5cm]
+$^1$Department of Systems and Computational Biology,\\
+$^2$Dominick P. Purpura Department of Neuroscience,\\
+$^3$Department of Pathology, Albert Einstein College of Medicine,\\
+1301 Morris Park Ave, Bronx, NY 10461, USA\\
+$^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
+\\[.5cm]
+% $\ast$To whom correspondence should be addressed; E-mail: aviv@einstein.yu.edu.
+\end{center}
 
-%    Information for third author
-\author{Cameron Smith}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{cameron.smith@med.einstein.yu.edu}
-% \thanks{Support information for the second author.}
-
-%    General info
-% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
-
-% \date{January 1, 2001 and, in revised form, June 22, 2001.}
-
-% \dedicatory{This paper is dedicated to our advisors.}
-
-% \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
+{\begin{quote} \bf
+\input{tex/abstract}
+\end{quote}}
 

### 2014-09-16 tex/abstract.tex
switch to different style

--- 2014-09-12 journal.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of journal.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
-
-\begin{document}
-
-\title{Ensemble Stability of Biological Networks}
-
-%    Information for first author
-\author{Aviv Bergman}
-%    Address of record for the research reported here
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{aviv@einstein.yu.edu}
-%    \thanks will become a 1st page footnote.
-% \thanks{The first author was supported in part by NSF Grant \#000000.}
-
-%    Information for second author
-\author{Raymond S. Puzio}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{rsp@novres.org}
-% \thanks{Support information for the second author.}
-
-%    Information for third author
-\author{Cameron Smith}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{cameron.smith@med.einstein.yu.edu}
-% \thanks{Support information for the second author.}
-
-%    General info
-% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
-
-% \date{January 1, 2001 and, in revised form, June 22, 2001.}
-
-% \dedicatory{This paper is dedicated to our advisors.}
-
-% \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
 The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
@@
 The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Stability analysis of biological networks}
-
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
-\begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
-\end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
-
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
-
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
-
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
-Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
-
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
-%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
-This is the plane $s_3=1$.}
-\label{fig:region3x3}
-\end{figure}
-
-\section{Two component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\begin{table}[h]
-\begin{tabular}{ c || c | c | c }
-\hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-  $\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
-  $\begin{pmatrix}
-a & b \\
-d & 0
-\end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
-  $\begin{pmatrix}
-a & 0 \\
-d & c
-\end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
-$\begin{pmatrix}
-a & 0 \\
-0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
-\end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\section{Three component systems}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
-
-\newpage
-
-\input{tablenographs}
-
-\newpage
-
-\input{tablewithgraphs}
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------

### 2014-09-16 tex/ams_author.tex
switch to different style

--- 2014-09-12 journal.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of journal.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
-
-\begin{document}
-
-\title{Ensemble Stability of Biological Networks}
-
 %    Information for first author
@@
 % \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Stability analysis of biological networks}
-
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
-\begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
-\end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
-
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
-
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
-
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
-Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
-
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
-%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
-This is the plane $s_3=1$.}
-\label{fig:region3x3}
-\end{figure}
-
-\section{Two component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\begin{table}[h]
-\begin{tabular}{ c || c | c | c }
-\hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-  $\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
-  $\begin{pmatrix}
-a & b \\
-d & 0
-\end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
-  $\begin{pmatrix}
-a & 0 \\
-d & c
-\end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
-$\begin{pmatrix}
-a & 0 \\
-0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
-\end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\section{Three component systems}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
-
-\newpage
-
-\input{tablenographs}
-
-\newpage
-
-\input{tablewithgraphs}
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------

### 2014-09-16 paper.tex
add results sections

@@
 
-\section{Two component systems}
+\section{Stability and robustness analysis of multi-component systems}
 For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
@@
 The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
+Similarly, the analogous results for all three component systems is given in Table \ref{tab:structstabmat3}.
 
-\section{Three component systems}
+\section{Connectivity is correlated with robustness}
+\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
 
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
+\section{Cycle number is inversely correlated with robustness}
+\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+
+\section{Cycle number and connectivity classify robust systems}
 \ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+
+\section{Hierarchical systems are the most robust}
+\ref{fig:scc} shows systems with varying levels of hierarchy and modularity.
+
+\section{Discussion}
 

### 2014-09-16 paper.tex
simplify characterization of stability

@@
 this space. That is to say, we introduce a mapping $V \colon S \to \Gamma T(\prod_i M_m)$.
-(Here $\Gamma$ denoted the space of sections of the tangent bundle.)
+(Here $\Gamma$ denotes the space of sections of the tangent bundle.)
 As before, for each point $p \in S$, the vector field $V(p)$ must be consistent with
@@
 
-Having done this, we are now in a position to do a probabilistic analysis.
-Given some quantity $q$ characterizing our system, this quantity now
-becomes a random variable.  This variable may be discrete of continuous
-depending on the quantity under consideration.  For example, if the quantity
-is the time it takes for a particle to travel between two points, we have
-a real-valued random variable; if the quantity is the number of fixed
+Having done this, we are now in a position to do a probabilistic analysis
+of our system and its properties.  Given some quantity $q$ characterizing 
+our system, this quantity becomes a random variable which may be discrete or 
+continuous depending on the quantity under consideration.  For example, if 
+the quantity is the time it takes for a particle to travel between two points, 
+we have a real-valued random variable; if the quantity is the number of fixed
 points, we have an integer-valued random variable; and if the quantity is
@@
 \begin{lemma}
-Given a polynomial $p(x) = z^n + \sum_{k=1}^{n-1} s_k x^{n-k}$ and a real number
-$\epsilon > 0$, there exists $\delta > 0 \in \mathbb{R}$ such that, if $q(x) =
-x^n + \sum_{k=1}^{n-1} \sigma_i x^i$ is a polynomial for which $|s_i - \sigma_i|
-< \delta$ then one can order the roots of $p$ and $q$ such that $p(z) = \prod_{k=1}^n
-(z - r_i)$ and $q(z) = \prod_{k=1}^n (z - \rho_i)$ and $| r_k - \rho_k | < \epsilon$.
-\end{lemma}
-
-\begin{proof}
-Without loss of generality, we may assume that $\epsilon < 1$.
-
-We first consider the case where $p(x) = x^n$.
-
-If we make the rescaling $x = \epsilon y/2$ in the equation $p(x)  = 0$,
-we obtain the equation
-\[
-y^n + \sum_{k=1}^{n-1} \left(\frac{2}{\epsilon}\right)^k s_k y^{n-k} = 0 .
-\]
-Cauchy's bound on the roots of this equation is
-\[
-|y| < 1 + \max_j \left(\frac{2}{\epsilon}\right)^j |s_j| ,
-\]
-which becomes the bound
-\[
-|x| < \frac{\epsilon}{2} + \max_j \left(\frac{2}{\epsilon}\right)^{j-1} |s_j|
-< \frac{\epsilon}{2} + \left(\frac{2}{\epsilon}\right)^{n-1} \max_j |s_j|
-\]
-for the original equation.  Hence, if we choose $\delta = (\epsilon/2)^{n}$,
-we will have $|x| < \epsilon$.
-
-Next, we generalize to the case $p = (x-r)^n$.
-
-To extend this result of the general case, we note that, given two
-polynomials $p = x^m + \sum_{k=1}^n a_k x^{m-k}$ and
-$q = x^n + \sum_{k=1}^n b_k x^{n-k}$, the Jacobian of the map which
-sends the coefficients of the matrices to the coefficients of the
-product $pq$ is the Sylvester matrix, whose determinant is the resultant
-of $p$ and $q$.  Hence, is $p$ and $q$ have no common root, the map
-is locally invertible.
-\end{proof}
-
-\begin{lemma}
 If a point lies on the boundary of the set of stable polynomials, then
@@
 \begin{proof}
-Suppose that none of the real parts of the roots of a polynomial $p$ equal
-zero.  Let $\epsilon$ be the minimum of the absolute values of the real
-parts of the roots of $p$.  Then, by continuity of the map which sends the
-roots of a polynomial to its coefficients, there exists a $\delta > 0$
-such that, if $q$ is any polynomial whose coefficients differ by less than
-$\delta$ from the
+Let $S_n$ denote the symmetric group on $n$ objects and let $f \colon
+\mathbb{C}^n/S_n \to \mathbb{C}^n$ be the map that sends a set of roots
+to the n-tuplet of coefficients of a polynomial.  This map is explicitly
+given by the elemetary symmetric polynomials.  By the fundamental theorem
+of algebra, it is invertible and, by ???, its inverse is continuous.
 
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
+Suppose that $p$ is a stable polynomial.  By definition, this means that
+$\Re r > 0$ for every root $r$ of $p$.  Since the set $\{r \mid \Re r > 0\}$
+is an open subset of $\mathbb{C}^n/S_n$, the continuity of $f^{-1}$ implies 
+that the set of stable polynomials is open.  Likewise, the set of unstable
+polynomials none of whose roots have real part equal to zero is open.
 
-To complete the proof, we show that the condition $\delta = 0$ is not
-needed to describe the boundary.  Suppose that $\delta = 0$.  Then there
-must exist a repeated root.  If the real part of any root, repeated or
-not, is positive, then the polynomial is not stable and, by continuity,
-if one allows the coefficients to vary within a sufficiently small
-ball about that solution, the root will remain negative, so the
-polynomial cannot lie on the boundary of the stable region.  If all
-the roots, including the repeated ones, are negative then, by a similar
-argument the point is in the interior of the stable region.  The only
-other possibility is that some root has zero real part, but then
-$\Delta = 0$ so the condition $\delta = 0$ is superfluous.
+Hence, if a point lies on the boundary of the stable region, it must 
+correspond to a polynomial with at least one root whose real part is zero.
+There are two ways this could happen.  If the root is real, then having
+zero real part is the same as the root being zero, hence $s_n = 0$.  If
+the root is complex, then having zero real part means that the root and
+its complex conjugate are purely imaginary and sum to zero, hence $\Delta = 0$.
 \end{proof}
 
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in \ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
-
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to \ref{fig:region2x2} for three component systems where $D=1$ is given in \ref{fig:region3x3}.
-%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
-\section{System graph and decomposition into strongly connected components}
+\section{Paths and eigenvalues}
 Next, we will want to relate prpoperties of our matrix and its eigenvalues
@@
 Given a digraph $G$, let $\Path (a,b)$ denote the set of paths from the
-vertex $a$ to the vertex $b$.  Given a path $p_1$ from $a$ to $b$ and a path
-$p_2$ from $b$ to $c$, we may concatenate them to obtain a path $p_1 \circ
-p_2$ from $a$ to $c$.  Hence, $(\Vertex(G), \Path(G), \circ)$ forms a category.
+vertex $a$ to the vertex $b$ and let $\Path(G) = \bigcup_{a,b \in G} \Path (a,b)$. 
+Given a path $p_1$ from $a$ to $b$ and a path $p_2$ from $b$ to $c$, we may 
+concatenate them to obtain a path $p_1 \circ p_2$ from $a$ to $c$.  Since 
+this operation is associative and has identity elements, 
+$(\Vertex(G), \Path(G), \circ)$ forms a category.  Define 
+$\mathcal{A}(G)$ to be the set of all linear combinations of elements of
+$\Path(G)$ with coefficients in $\mathbb{R}$.  Then $\mathcal{A}(G)$ forms
+an associative algebra under the multiplication induced by $\circ$.
+
+% \def\wprod{\mathrm{wp}}
+Next, we introduce two basic functors from $\Path (G)$.  One of these is
+$\length \colon \Path(G) \to \mathbb{N}$ (regarding the natural numbers
+as a category with a single object) which sends each path to its length.
+The other functor is the weight product functor.  Given an assignement 
+$w \colon \Path(G) \to \mathbb(R)$ of weights to the edges of our digraph,
+define $\wprod \colon \Path(G) \to \mathbb{R}$ to be the function which sends
+each path to the product of the weights associated to the edges which
+that path traverses.  Given the way that graphs are associated to matrices,
+weights on edges correspond to matrix entries and, given a path 
+$v_1, v_2, v_3 \ldots$, we have $\wprod(v_1, v_2, v_3 \ldots) =
+m_{v_1 v_2} m_{v_2 v_3} \cdots$.  Going in the opposite direction, we
+note that any functor from $\Path(G)$ to $\mathbb{R}$ may be regarded as
+a weighted product for a suitable choice of weights.  We extend $\wp$ to
+$\mathcal{A}(G)$ by linearity; the result is an algebra homomorphism.
+
+\todo{Start talking about Abelianizing the category and the algebra, say
+more after decomposing into components.}
+
+\todo{Characterize strongly connected components in terms of 
+fibration over poset.}
 
@@
 \[
-\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)
+\tr (I - xM)^n =
+\sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} \wprod(p) x^{\length(p)}
 \]
@@
 
+\todo{Say a few words about Monte Carlo integration and error bars.}
+
 \section{Connectivity is correlated with robustness}

### 2014-09-17 code/rds3x3.jl
add function to compute orbit size under action of conjugation by permutation-x-transposition

@@ -191,6 +191,50 @@ function findpermclasses(N::Integer)
     return classbyconnect
 end
 
+function countorbitsize(M::Array{Int64,2})
+    N = size(M)[1]
+    # generate matrix representations of permutations
+    permmats = [matrix(Permutation(i)) for i in permutations([1:N])]
+    allbinmats = genbinmats(N)
+    count = 0
+
+    for (i,mat) in enumerate(allbinmats)
+
+        ind = false
+        if M' == mat
+            ind = true
+        end
+
+        for p in permmats
+            if p*M*inv(p) == mat
+                ind = true
+            end
+
+            if (p*M*inv(p))' == mat
+                ind = true
+            end
+        end
+
+        if ind
+            count += 1
+        end
+
+    end
+
+    return count
+end
+
+function weightedavg(VW)
+
+    # input is a tuple (vector, weights)
+    inds = find(x->x>0,VW[1])
+    relw = VW[2][inds]
+    tot = sum(relw)
+
+    return dot(relw/tot,VW[1][inds])
+
+end
+
 function sysstab(M::Array{Float64,2})
 
   # Continuous time system
@@ -299,3 +343,11 @@ function teststructstab(N::Integer,K::Integer)
 
 end
 
+# @elapsed sp,rsp = teststructstab(3,10000)
+# println(savelatex(pc3,sp,rsp))
+# pc3 = findpermclasses(3)
+# cn = cyclenumber(pc3)
+# println(savedata("../data/stab3x3.tsv",sp,rsp,cn))
+# os3 = map(x->map(countorbitsize,x),pc3)
+# rspwa = map(weightedavg,zip(rsp,os3))
+# spwa = map(weightedavg,zip(sp,os3))

### 2014-09-17 paper.tex
add commands to header
add stability figure and combined stability regions figure

@@
 
-% \def\wprod{\mathrm{wp}}
 Next, we introduce two basic functors from $\Path (G)$.  One of these is
@@
 The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-Similarly, the analogous results for all three component systems is given in Table \ref{tab:structstabmat3}.
+Similarly, the analogous results for all three component systems is given in \ref{tab:structstabmat3} and \ref{fig:stab3x3}.
+It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
+For small system sizes such as the two and three component systems, studied here the situation is not so clear cut.
+For two component systems, system stability is constant across the entire range of connectivities.
+For three component systems, the trend is unclear as shown in \ref{fig:apstab3x3}.
 
@@
 \section{Connectivity is correlated with robustness}
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
+The relationship between connectivity and probability of stability to perturbations for two component systems is shown in \ref{tab:structstabmat}.
+Likewise, \ref{fig:stab3x3} and \ref{tab:structstabmat3} show the probability of stability to perturbations for all three component systems as a function of connectivity.
 

### 2014-09-17 tex/figures.tex
add stability figure and combined stability regions figure

@@
 \centering
+\noindent\includegraphics[width=0.5\columnwidth]{fig/region2and3.pdf}
+\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
+\label{fig:region2and3}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
 \noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
@@
 \label{fig:region3x3}
+\end{figure}
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\columnwidth]{fig/apstab3x3.pdf}
+\caption{{\bf Plot of stability versus connectivity for three component systems.} }
+\label{fig:apstab3x3}
 \end{figure}

### 2014-09-17 code/rds3x3.jl
add orbit size to 3x3 table

@@ -88,19 +88,22 @@ function cyclenumber(adjmatlistlist)
 
 end
 
-function savelatex(pcN,sp,rsp)
+function savelatex(pcN,osN,sp,rsp)
     cn = cyclenumber(pcN)
-    println(cn)
+    # osN = map(x->map(countorbitsize,x),pcN)
+
     rv = (String)[]
 
     for (i,plist) in enumerate(sp)
         for (j,p) in enumerate(plist)
             if p > 1e-3
-                  # pmatstr = pmatrix(pcN[i][j])
-                  pmatstr = adjtograph(pcN[i][j])
-                  # string("\$", pmatstr, "\$", " & ",
+                  pmatstr = pmatrix(pcN[i][j])
+                  # pmatstr = adjtograph(pcN[i][j])
+
                   push!(rv,
-                        string(pmatstr, " & ",
+                        string("\$", pmatstr, "\$", " & ",
+                        # string(pmatstr, " & ",
+                               "$(osN[i][j])", " & ",
                                "$(i-1)", " & ",
                                "$(cn[i][j])", " & ",
                                "$(round(rsp[i][j],3))", " & ",
@@ -110,7 +113,7 @@ function savelatex(pcN,sp,rsp)
     end
 
     savedata("../data/stab3x3.tsv",sp,rsp,cn);
-
+    clipboard(join(rv,"\n"))
     return join(rv,"\n")
 end
 
@@ -344,10 +347,10 @@ function teststructstab(N::Integer,K::Integer)
 end
 
 # @elapsed sp,rsp = teststructstab(3,10000)
-# println(savelatex(pc3,sp,rsp))
 # pc3 = findpermclasses(3)
+# os3 = map(x->map(countorbitsize,x),pc3)
+# println(savelatex(pc3,os3,sp,rsp))
 # cn = cyclenumber(pc3)
 # println(savedata("../data/stab3x3.tsv",sp,rsp,cn))
-# os3 = map(x->map(countorbitsize,x),pc3)
 # rspwa = map(weightedavg,zip(rsp,os3))
 # spwa = map(weightedavg,zip(sp,os3))

### 2014-09-18 tex/abstract.tex
update abstract

@@
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
+%!TEX root = ../paper.tex
+The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many degrees of freedom.
+When the details of a system are unknown, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints applying to all such systems.
 Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
+In particular, we precisely determine the probability distribution of structural stability over system connectivity and cycle number, parameters that have significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.
+We show that structural stability is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most stable systems are also the most hierarchical.
+This results in the classification of the dynaical robustness of biological networks based upon a purely topological property.

### 2014-09-18 paper.tex
minor edits
edit abstract and intro

@@
 
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
+Here we build on work aimed at investigating the stability of randomly connected dynamical systems to structural perturbations.
+We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two or three interacting components.
+We then proceed to determine the probability of \emph{structural stability} or \emph{robustness} which is the distribution of stability to perturbation in the system structure for systems which are determined to be stable \cite{Smale1967}.
+This quantity is of interest at all scales of the biological hierarchy, and has been previously addressed in the study of the robustness of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
+We find that stability to structural perturbations of this class of dynamical systems is correlated with connectivity, number of cycles, and the number of links between strongly connected components of the underlying interdependency graph.
+The latter derives from the fact that the permutation of strongly connected components is a symmetry of system robustness.
+This work clarifies a thread of research that is extremely important for the continuing development of systems biology.
 
@@
 \input{tex/tablenographs}
+
+% \newpage
+
 % \input{tex/tablewithgraphs}

### 2014-09-18 paper.tex
integrate changes from sharelatex

@@
 
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph $G$ to the interdependencies, and a vector field $V \in T(\prod_i M_i)$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
+For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph $G$ to the interdependencies, and a vector field $V \in \Gamma T(\prod_i M_i)$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
 For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
@@
   \mathrm{Robust} (q) =
-  \frac{\int_S d\mu(x) \int_S d\mu'(y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
+  \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
   {\int_S d\mu(x) \mathbf{1}_q(x)}
@@
 
+\section{Component decomposition and robustness}
+
+The decomposition of a digraph into strongly connected components 
+corresponds to a block triangular decomposition of matrices
+describle by that digraph.  Say that the graph $G$ has strongly
+connected components $C_1, C_2, \ldots C_n$, which have been
+labelled in such a way that there are no links from vertices in
+component $C_i$ to component $C_j$ when $i < j$.  Label the
+vertices in such a way that $v_1, v_{n_1}$ belong to $C_1$,
+$v_1, v_{n_1}$ belong to $C_2$, etc.  Then if $M$ is a matrix
+consistent with $G$, it must be block triangular.  [explain better]
+
+An important consequence of this is that edges which go between
+different strongly connected components do not contribute to
+determining stability.  Hence, if we resample a matrix entry
+corresponding to such an edge, the conditional probability 
+of the matrix remaining stable upon resampling is 1.  Hence, if
+our graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$
+where $C_i$ has $v_i$ vertices and connectivity $c_1$ and there
+are $\ell_{ij}$ links between components $i$ and $j$, then the
+total robustness to resampling a random entry is
+\[
+R(G) = \frac{\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i R_i}
+            {\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i}
+\]
+
+\[
+R_{\mathrm{max}} = \frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i R_i}
+            {\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i}
+\]
+
 \section{Stability criteria}
+\todo{Since there is no point in publishing proofs of results contemporary
+with Darwin's Origin of Species, this section will be replaced with a
+reference to Gantmacher.}
 Suppose that $M$ is an $n \times n$ matrix with real coefficients.
-Motivated by the discussion of \ref{sec:stabanalbn}, we will call
+As described in \ref{sec:stabanalbn}, we call
 the matrix $M$ stable if all its eigenvalues have negative real part.
@@
 $M$ be stable is equivalent to asking that the roots of $\chi$ all have
-negative real part.  We will formulate conditions for this to happen in
-terms of the coefficients of $\chi$.
+negative real part.  Hence, let us call a polynomial stable if all its
+roots have negative real part.  We will formulate conditions for this to 
+happen in terms of the coefficients of $\chi$.
 
@@
 
+\begin{lemma} 
+If $n < 6$, then a polynomial $p$ is stable if and only
+if $\Delta > 0$ and $s_1 > 0$ for all $i$.
+\end{lemma}
+
+\begin{proof}
+Suppose that, to the contrary, there exists a polynomial which satisfies
+the inequalities $\Delta > 0$ and $s_1 > 0$ but which is not stable.
+Consider a straight line connecting this polynomial to a stable polynomial,
+say $(x - 1)^n$.
+
+Suppose that $p$ is a polynomial for which $\Delta = 0$ and $s_i > 0$ but
+which is not on the boundary of the stable region.  In order to have 
+$\Delta = 0$, we must either have a pair of real roots which sum to zero 
+or a pair of complex conjugate roots with zero real part.  In the former
+case, we would have $x^2 - r^2$ as a factor of our polynomial and in the
+latter, we would have $x^2 + r^2$ as a factor.
+
+Suppose that $x^2 - r^2$ is a factor.  Then we have
+\[
+p = (x^2 - r^2) \left(x^{n-2} + \sum_{k=1}^{n-3} c_k x^{n-k-2} \right) =
+x^n + \sum_{k=3}^{n-3} (c_k - r^2 c_{k-2}) x^{n-k} + 
+\]
+
+Suppose that $x^2 + r^2$ is a factor. Then we have one of the following
+when $2 < n < 6$
+\begin{align*}
+(x^2 + r^2)(x + a_1) &=
+x^3 + a_1 x^3 + r^2 x^2 + a_1 r^2 \\
+(x^2 + r^2)(x^2 + a_1 x + a_2) &=
+x^4 + a_1 x^3 + (a_2 + r^2) x^2 + a_1 r^2 x + a_2 r^2 \\
+(x^2 + r^2)(x^3 + a_1 x^2 + a_2 x + a_3) &=
+x^5 + a_1 x^4 + (a_2 + r^2) x^3 + (a_3 + a_1 r^2) x^2 + a_2 r^2 x + a_3 r^2
+\end{align*}
+Inspecting these, we see that, if the coefficients of $p$ are positive,
+then $a_i > 0$ for all $i$.
+
+$\ldots$
+\end{proof}
+
+For $n = 6$, this argument breaks down because, for instance, we have
+the polynomial $x^6 + x^5 +\frac{1}{2} x^4 + 2x^3 + \frac{1}{2} x^2 +
+x + 1$  which factors as  $(x^2 + 1)(x^4 + x^3 - \frac{1}{2} x^2 + x + 1)$.
+This polynomial has positive cofficients and $\Delta = 0$ but does not
+lie on the boundary of the stable region because the second factor has
+a negative coefficient, hence has a root with positive real part.  Thus,
+for $n = 6$ and higher, we need to supplement our conditions with
+additional inequalities (which may be taken to be linear) to exclude
+such possibilitiies.  However, we will not pursue this topic further
+in this work because we will only be making computations with matrices
+of size less than $6$.
+
 \section{Paths and eigenvalues}
+\todo{I am thinking that we might want to jettison this section from
+the published version and leave it as private notes for future work.
+As a substitute for the published version, I made a section in front 
+of it which takes a more direct route to proving what is needed for
+the purposes of showing that the characteristic polynomial factors and
+making our arguments about total orders being the most stable.  We
+should discuss this point tomorrow.}
+
 Next, we will want to relate prpoperties of our matrix and its eigenvalues
@@
 note that any functor from $\Path(G)$ to $\mathbb{R}$ may be regarded as
-a weighted product for a suitable choice of weights.  We extend $\wp$ to
+a weighted product for a suitable choice of weights.  We extend $\wprod$ to
 $\mathcal{A}(G)$ by linearity; the result is an algebra homomorphism.
@@
 \[
- \begin{matrix}
+ \begin{pmatrix}
    m_{11} & m_{12} & m_{13} \\
@@
    m_{31} & 0      & 0 \\
- \end{matrix}
+ \end{pmatrix}
 \]

### 2014-09-19 paper.tex
update main text

@@
 
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph $G$ to the interdependencies, and a vector field $V \in \Gamma T(\prod_i M_i)$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
+For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V \in \Gamma T(\prod_i M_i)$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
 For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
@@
 $$
-then for the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
+having connectivity $\connectivity$ equal to the number of edges of the graph, in this case $\connectivity = 3$.
+For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
 \begin{align*}
@@
 $\mu' \colon S \times \mathcal{B}(S) \to [0,1]$.  Then, given a binary random
-variable $q$, we we define its robustness as the following conditional proability:
-\[
-  \mathrm{Robust} (q) =
+variable $q$, we define its robustness as the following conditional proability:
+\begin{align}\label{eq:robustness}
+  R (q) =
   \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
   {\int_S d\mu(x) \mathbf{1}_q(x)}
-\]
+\end{align}
 
@@
 
+\section{Stability and robustness analysis of multi-component systems}
+Suppose that $M$ is an $n \times n$ matrix with real coefficients.
+We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
+We want to know the probability that, if we resample the $k^{th}$ element of this matrix, the resulting matrix will correspond to a stable matrix which is given by a special case of \ref{eq:robustness}
+\begin{align}\label{eq:condprobgen}
+ R (stab) =
+  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{-1}^{1} dy \, \mathbf{1}_{stab}(x_1 \cdots x_k \cdots x_d) \mathbf{1}_{stab}(x_1 \cdots y \cdots x_d)}
+  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
+\end{align}
+
+As previously stated, we call the matrix $M$ stable if all its eigenvalues have negative real part.
+Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
+Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
+Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
+$M$ be stable is equivalent to asking that the roots of $\chi$ all have
+negative real part.  Hence, let us call a polynomial stable if all its
+roots have negative real part.  We will formulate conditions for this to
+happen in terms of the coefficients of $\chi$ \cite{Gantmacher1959}.
+
+\begin{lemma}
+If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
+\end{lemma}
+\begin{lemma}
+If a point lies on the boundary of the set of stable polynomials, then
+either $s_n = 0$ and $\Delta = 0$.
+\end{lemma}
+\begin{lemma}
+If $n < 6$, then a polynomial $p$ is stable if and only
+if $\Delta > 0$ and $s_1 > 0$ for all $i$.
+\end{lemma}
+
+For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
+0$ where $T$ and $D$ denote the trace and the determinant. Suppose we have a stable matrix
+$$
+\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix}
+$$
+where $a + c < 0$ and $ac > bd$.
+By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$.
+Suppose that we resample $b$.
+We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is a special case of \ref{eq:condprobgen}
+% \begin{strip}
+\begin{align}\label{eq:condprob}
+P\left(\begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right)
+& = \frac{P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right)}.
+\end{align}
+The denominator of \ref{eq:condprob} is given by
+\begin{align*}
+P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
+\end{align*}
+Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
+Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
+\begin{align*}
+P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+a & b' \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
+\end{align*}
+The analogous equation for resampling $a$ is
+\begin{align*}
+P\left(\begin{pmatrix}
+a & b \\
+d & c
+\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+a' & b \\
+d & c
+\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
+\end{align*}
+
+Using this approach the probability of \emph{a priori} stability and of robustness for all two component systems is given in Table \ref{tab:structstabmat}.
+Similarly, the analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:stab3x3}.
+
+\todo{Say a few words about Monte Carlo integration and error bars.}
+
+\section{Dependency of robustness upon connectivity and cycle number}
+It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
+For small system sizes such as the two and three component systems, studied here the situation is not so clear cut.
+For two component systems, system stability is constant across the entire range of connectivities.
+For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
+The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:stab3x3}.
+If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \ref{fig:stab3x3}.
+% \section{Cycle number is inversely correlated with robustness}
+\ref{fig:cycle3x3} shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+There appears to be a weak negative correlation between robustness and the number of simple cycles.
+
+% \section{Cycle number and connectivity classify robust systems}
+The combination of connectivity and cycle number as shown in \ref{fig:connectcycle3D3x3} provides a better classification of the dependence of robustness upon network topology.
+Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity.
+The network with the highest robustness for three component systems is that of \ref{fig:scc}A.
+This network is the most hierarchical of all three component systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \ref{fig:scc}A has a block triangular structure.
+
 \section{Component decomposition and robustness}
 
-The decomposition of a digraph into strongly connected components 
-corresponds to a block triangular decomposition of matrices
-describle by that digraph.  Say that the graph $G$ has strongly
-connected components $C_1, C_2, \ldots C_n$, which have been
-labelled in such a way that there are no links from vertices in
-component $C_i$ to component $C_j$ when $i < j$.  Label the
-vertices in such a way that $v_1, v_{n_1}$ belong to $C_1$,
-$v_1, v_{n_1}$ belong to $C_2$, etc.  Then if $M$ is a matrix
-consistent with $G$, it must be block triangular.  [explain better]
+The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of matrices whose connectivity is given by that digraph.
+Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
+Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
+Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
+This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
+A simple example of this is given by the graph in \ref{fig:scc}C which has two strongly connected components.
 
-An important consequence of this is that edges which go between
-different strongly connected components do not contribute to
-determining stability.  Hence, if we resample a matrix entry
-corresponding to such an edge, the conditional probability 
-of the matrix remaining stable upon resampling is 1.  Hence, if
-our graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$
-where $C_i$ has $v_i$ vertices and connectivity $c_1$ and there
-are $\ell_{ij}$ links between components $i$ and $j$, then the
-total robustness to resampling a random entry is
-\[
+Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomials factors as the product of the charactericstic polynomials of its diagonal blocks.
+Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.
+Note that this condition does not depend upon the entries off the diagonal (which correspond to links between strongly connected components) and does not depend upon what order the components appear.
+
+An important consequence of this is that stability is invariant to perturbations to the edges that go between different strongly connected components.
+Hence, if we resample a matrix entry corresponding to such an edge, the conditional probability of the matrix remaining stable upon resampling is 1.
+If our graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$ where $C_i$ has $v_i$ vertices and connectivity $c_1$ and there are $\ell_{ij}$ links between components $i$ and $j$, then the total robustness to resampling a random entry is
+\begin{align}
 R(G) = \frac{\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i R_i}
-            {\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i}
-\]
+            {\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i}, \label{eq:sccrobustness}
+\end{align}
+where $R_i$ refers to the independent robustness of component $i$.
+\begin{align}
+R_{\mathrm{max}} = \frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i R_i}{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i}, \label{eq:sccmaxrobustness}
+\end{align}
 
-\[
-R_{\mathrm{max}} = \frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i R_i}
-            {\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i}
-\]
+\section{Strongly connected components, robustness, and hierarchy}
+The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
+Inspection of \ref{eq:sccrobustness} demonstrates that robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
+% \section{Hierarchical systems are the most robust}
+% \ref{fig:scc} shows systems with varying levels of hierarchy and modularity.
+
+\section{Discussion}
+
+\newpage
+
+\bibliographystyle{amsplain}
+\bibliography{bib/books,bib/papers}
+
+\newpage
+\FloatBarrier
+
+\section{Figure Legends}
+\input{tex/figures}
+
+\newpage
+\FloatBarrier
+
+\section{Tables}
+\input{tex/tablenotation}
+\input{tex/table2x2stab}
+\input{tex/tablenographs}
+
+% \newpage
+
+% \input{tex/tablewithgraphs}
 
@@
 
-\section{Stability and robustness analysis of multi-component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-Similarly, the analogous results for all three component systems is given in \ref{tab:structstabmat3} and \ref{fig:stab3x3}.
-It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
-For small system sizes such as the two and three component systems, studied here the situation is not so clear cut.
-For two component systems, system stability is constant across the entire range of connectivities.
-For three component systems, the trend is unclear as shown in \ref{fig:apstab3x3}.
-
-\todo{Say a few words about Monte Carlo integration and error bars.}
-
-\section{Connectivity is correlated with robustness}
-The relationship between connectivity and probability of stability to perturbations for two component systems is shown in \ref{tab:structstabmat}.
-Likewise, \ref{fig:stab3x3} and \ref{tab:structstabmat3} show the probability of stability to perturbations for all three component systems as a function of connectivity.
-
-\section{Cycle number is inversely correlated with robustness}
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-
-\section{Cycle number and connectivity classify robust systems}
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-
-\section{Hierarchical systems are the most robust}
-\ref{fig:scc} shows systems with varying levels of hierarchy and modularity.
-
-\section{Discussion}
-
-\newpage
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\newpage
-\FloatBarrier
-
-\section{Figure Legends}
-\input{tex/figures}
-
-\newpage
-\FloatBarrier
-
-\section{Tables}
-\input{tex/table2x2stab}
-\input{tex/tablenographs}
-
-% \newpage
-
-% \input{tex/tablewithgraphs}
-
 \end{document}

### 2014-09-20 tex/figures.tex
minor edits
update main text
compute graph edit distance and combine figures

@@
 Note that the most hierarchical system in (A) has the highest possible number of connected components, three, whereas the system containing a single cycle and therefore posessing no hierarchy contains only one connected component.
-Systems (B) and (D) represent examples of hierarchical modular systems that posess both modularity and hierarchy.}
+Systems (B) and (C) represent examples of hierarchical modular systems that posess both modularity and hierarchy.}
 \label{fig:scc}
@@
 \centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2and3.pdf}
+\noindent\includegraphics[width=0.8\columnwidth]{fig/region2and3.pdf}
 \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
 \label{fig:region2and3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
-Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
-This is the plane $s_3=1$.}
-\label{fig:region3x3}
 \end{figure}

### 2014-09-20 tex/figuressupp.tex
compute graph edit distance and combine figures

--- 2014-09-12 journal.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of journal.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
-
-\begin{document}
-
-\title{Ensemble Stability of Biological Networks}
-
-%    Information for first author
-\author{Aviv Bergman}
-%    Address of record for the research reported here
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{aviv@einstein.yu.edu}
-%    \thanks will become a 1st page footnote.
-% \thanks{The first author was supported in part by NSF Grant \#000000.}
-
-%    Information for second author
-\author{Raymond S. Puzio}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{rsp@novres.org}
-% \thanks{Support information for the second author.}
-
-%    Information for third author
-\author{Cameron Smith}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{cameron.smith@med.einstein.yu.edu}
-% \thanks{Support information for the second author.}
-
-%    General info
-% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
-
-% \date{January 1, 2001 and, in revised form, June 22, 2001.}
-
-% \dedicatory{This paper is dedicated to our advisors.}
-
-% \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Stability analysis of biological networks}
-
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
-\begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
-\end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
-
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
-
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
-
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
-
 \begin{figure}[!ht]
@@
 \end{figure}
-
-\section{Two component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\begin{table}[h]
-\begin{tabular}{ c || c | c | c }
-\hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-  $\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
-  $\begin{pmatrix}
-a & b \\
-d & 0
-\end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
-  $\begin{pmatrix}
-a & 0 \\
-d & c
-\end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
-$\begin{pmatrix}
-a & 0 \\
-0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
-\end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\section{Three component systems}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
-
-\newpage
-
-\input{tablenographs}
-
-\newpage
-
-\input{tablewithgraphs}
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------

### 2014-09-20 code/rds3x3.jl

@@ -88,8 +88,9 @@ function cyclenumber(adjmatlistlist)
 
 end
 
-function savelatex(pcN,osN,sp,rsp)
+function savelatex(pcN,osN,sp,rsp,ged)
     cn = cyclenumber(pcN)
+    # ged = [map(y->grapheditdistance(pcN[7][11],y),z) for z in pcN]
     # osN = map(x->map(countorbitsize,x),pcN)
 
     rv = (String)[]
@@ -105,6 +106,7 @@ function savelatex(pcN,osN,sp,rsp)
                         # string(pmatstr, " & ",
                                "$(osN[i][j])", " & ",
                                "$(i-1)", " & ",
+                               "$(ged[i][j])", " & ",
                                "$(cn[i][j])", " & ",
                                "$(round(rsp[i][j],3))", " & ",
                                "$(round(sp[i][j],3))", "\\\\"))
@@ -112,18 +114,18 @@ function savelatex(pcN,osN,sp,rsp)
         end
     end
 
-    savedata("../data/stab3x3.tsv",sp,rsp,cn);
+    savedata("../data/stab3x3.tsv",sp,rsp,cn,ged);
     clipboard(join(rv,"\n"))
     return join(rv,"\n")
 end
 
-function savedata(filename,sp,rsp,cn)
+function savedata(filename,sp,rsp,cn,ged)
     rv = (String)[]
-    push!(rv,"connectivity\tstability\tresampling\tcycle")
+    push!(rv,"connectivity\tstability\tresampling\tcycle\tdistance")
     for (i,plist) in enumerate(sp)
         for (j,p) in enumerate(plist)
             if p > 1e-3
-                  line = "$(i-1)\t$(sp[i][j])\t$(rsp[i][j])\t$(cn[i][j])"
+                  line = "$(i-1)\t$(sp[i][j])\t$(rsp[i][j])\t$(cn[i][j])\t$(ged[i][j])"
                   push!(rv,line)
             end
         end
@@ -225,6 +227,47 @@ function countorbitsize(M::Array{Int64,2})
     end
 
     return count
+end
+
+function genperms(M1::Array{Int64,2})
+    N = size(M1)[1]
+
+    # generate matrix representations of permutations
+    permmats = [matrix(Permutation(i)) for i in permutations([1:N])]
+    allbinmats = genbinmats(N)
+    permM1 = (Array{Int64,2})[]
+
+    for (i,mat) in enumerate(allbinmats)
[…]

### 2014-09-20 code/rds3x3.jl
add labels to figs

@@ -400,19 +400,27 @@ function combinegnuplot()
     horiz = 2
     vert = 3
     rv = (String)[]
-    headstring = "set terminal svg size $(600*horiz),$(400*vert) dynamic enhanced fname 'HelveticaNeue'  fsize 16\n
-set output '../fig/combinedfigs.svg'\n
-set multiplot layout $(vert),$(horiz)\n
-"
+    headstring = "set terminal svg size $(600*horiz),$(400*vert) dynamic enhanced fname 'HelveticaNeue'  fsize 16
+set output '../fig/combinedfigs.svg'
+set multiplot layout $(vert),$(horiz)
+set tmargin 3
+set bmargin 3
+set lmargin 10
+set rmargin 5\n"
     push!(rv,headstring)
 
-    for file in filelist
+    for (i,file) in enumerate(filelist)
         # f = open(joinpath("../data",file),"r")
+        push!(rv,"\n############# plot $(i) ##############\n\n")
         lines = readlines(`grep -v -P 'output|terminal' ../data/$(file)`)
         # close(f)
-        for l in lines
+        for (j,l) in enumerate(lines)
+            if j==length(lines)-1
+                push!(rv,"set label 1 \"$(string(char(64+i)))\" font \"HelveticaNeue, 23\" at graph -0.25, graph 1.20\n")
+            end
             push!(rv,l)
         end
+
     end
 
     tailstring = "unset multiplot\n"
@@ -421,6 +429,10 @@ set multiplot layout $(vert),$(horiz)\n
     f = open("../data/combinedfigs.p","w")
     write(f,join(rv,""))
     close(f)
+
+    cd("../data")
+    spawn(`./plot.sh combinedfigs`)
+    cd("../code")
 
     return rv
 end

### 2014-09-22 paper.tex
compute graph edit distance and combine figures
add section bounds on robustness

@@
 
+\todo{Generalize to open systems and say how we combine them.}
+
 In order to accommodate the possibility of uncertainty in our modelling,
@@
 
+Corresponding to this decomposition we can construct a graph $\hier (G)$.
+Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
+There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ iff there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
+Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
+
+One can also perform this construction in the oppsite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
+a strongly connected graph $C_n$.
+To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
+The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
+
 Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
@@
 \section{Stability and robustness analysis of multi-component systems}
-Suppose that $M$ is an $n \times n$ matrix with real coefficients.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample the $k^{th}$ element of this matrix, the resulting matrix will correspond to a stable matrix which is given by a special case of \ref{eq:robustness}
+Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$.
+We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.
+We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix which is given by a special case of \ref{eq:robustness}
 \begin{align}\label{eq:condprobgen}
- R (stab) =
-  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{-1}^{1} dy \, \mathbf{1}_{stab}(x_1 \cdots x_k \cdots x_d) \mathbf{1}_{stab}(x_1 \cdots y \cdots x_d)}
+ R (stab, k_1, \ldots k_m) =
+  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\, 
+    \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) 
+    \mathbf{1}_{stab}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)}
   {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
@@
 
-As previously stated, we call the matrix $M$ stable if all its eigenvalues have negative real part.
+\todo{Generalize to resampling a random link and to resampling multiple edges.}
+
+In order to evaluate these integrals, one needs to describe the stable region.  
+The theory of how to do so in terms of conditions on the coefficients of the 
+characteristic polynomials was developed by various mathematicians during the 
+second half of the nineteenth century and is summarized in chapter XV of 
+\cite{Gantmacher1959}.  We now summarize some of these results which are 
+relevant to our calculations.
+
 Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$. 
-Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
+Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$, 
+writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
 Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
@@
 This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
-A simple example of this is given by the graph in \ref{fig:scc}C which has two strongly connected components.
+Some simple examples of this are given by the graphs in \ref{fig:scc}.
 
@@
 
-An important consequence of this is that stability is invariant to perturbations to the edges that go between different strongly connected components.
-Hence, if we resample a matrix entry corresponding to such an edge, the conditional probability of the matrix remaining stable upon resampling is 1.
-If our graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$ where $C_i$ has $v_i$ vertices and connectivity $c_1$ and there are $\ell_{ij}$ links between components $i$ and $j$, then the total robustness to resampling a random entry is
+Using these observations, we may express the robustness of a graph in terms of the robustnesses of its strongly connected components.
+If the components $C_i$ of our graph $G$ has $v_i$ vertices connectivity $\connectivity_i$, and robustness $R_i$ and there are $\ell_{ij}$ links between components $i$ and $j$, then the total robustness to resampling a random entry is
 \begin{align}
-R(G) = \frac{\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i R_i}
-            {\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i}, \label{eq:sccrobustness}
+R(G) = \frac{\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i R(C_i)}
+            {\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i}. \label{eq:sccrobustness}
 \end{align}
-where $R_i$ refers to the independent robustness of component $i$.
+
+Next, we ask what is the maximally robust graph that has a given set of connected components. 
 \begin{align}
@@
 
+\section{Bounds on robustness}
+
+Suppose that $S$ is a region in $[0,1]^2$.  Then, the robustness of $S$ under resampling is given as
+\begin{align*}
+R &= \frac{\int_0^1 dx \int_0^1 dy \int_0^1 dx'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x',y) +
+           \int_0^1 dx \int_0^1 dy \int_0^1 dx\,dy\,dy'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x,y')}
+          {2 \int_0^1 dx \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)} \\
+  &= \frac{\int_0^1 dy\, (\int_0^1 dx\, \mathbf{1}_{S} (x,y))^2 +
+           \int_0^1 dx\, (\int_0^1 dy\, \mathbf{1}_{S} (x,y))^2}
+          {2 \int_0^1 dx\, \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)}
+\end{align*}
+We will show that $R \le \int_{H^2} dx\,dy\, \mathbf{1}_{S} (x,y)$ with equality occurring when $S$ is rearrangement equivalent to either the rectangle of height 1 or a rectangle of width 1.
+
+First, we start with the special case when $S$ is a rectangle with one vertex at the origin and sides parallel to the coordinate axes.
+If $a$ is the height of our rectangle and $b$ is its width, we have
+
+Next, consider the case where $R$ consists of a union of squares of a rectangular grid obtained by subdividing the $x$ and $y$ axes of our bounding square into $2N$ portions of equal length.
+Given two integers $i$ and $j$
+
+Suppose we restrict to rectangular regions in an $n$-dimensional hypercube and resample $m$ variables.
+Then we have $R = s_m (a_1, \ldots a_n)$ where $s_m$ is the $m$-th elementary symmetric polynomial.
+As before, we have the constraint that $\prod_i a_i = s_n (a_1, \ldots a_n)$ is constant.
+Using Lagrange multipliers, we obtain the equation
+\[
+\frac{\partial s_m (a_1, \ldots a_n)}{a_k} = \lambda \frac{\partial s_n (a_1, \ldots a_n)}{a_k} .
+\]
+Since the elementary symmetic polynomial is linear in each of its variables, this reduces to
+\[
+s_{m-1} (a_1, \ldots {\hat a}_k \ldots a_n) = \lambda \partial s_{n-1} (a_1, \ldots {\hat a}_k \ldots a_n) .
+\]
+Dividing through, this becomes
+\[
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \lambda
+\]
+or, eliminating $\lambda$, we have
+\[
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots \frac{1}{a_n}\right) = 
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
+\]
+for all $j,k$.  Suppose that $j < k$.  Then, we may expaand both sides,
+\begin{align*}
+&\frac{1}{a_k} s_{m-2} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \\ \qquad
+&\frac{1}{a_j} s_{m-2} s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
+\end{align*}
+and conclude that $a_j = a_k$.
+Hence, the only critical point lies at the point $a_1 = a_2 = \cdots = a_n$.
+By the second derivative test, that point is a minimum, so the maximum must lie on the boundary.
+
 \section{Discussion}
+
+\section{Acknowlegements}  The authors would like to thank Jay Sulzberger and Ximo Pechuan for useful suggestions and discussions.
 
@@
 
+\input{tex/figuressupp}
+
 \end{document}

### 2014-09-22 paper.tex
combine figures characterizing stability and robustness

@@
 Using this approach the probability of \emph{a priori} stability and of robustness for all two component systems is given in Table \ref{tab:structstabmat}.
-Similarly, the analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:stab3x3}.
+Similarly, the analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.
 
@@
 For two component systems, system stability is constant across the entire range of connectivities.
-For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
-The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:stab3x3}.
-If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \ref{fig:stab3x3}.
+For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:combined}A.
+The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:combined}B.
+If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \ref{fig:combined}B.
 % \section{Cycle number is inversely correlated with robustness}
-\ref{fig:cycle3x3} shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+\ref{fig:combined}C shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
 There appears to be a weak negative correlation between robustness and the number of simple cycles.
@@
 % \section{Cycle number and connectivity classify robust systems}
-The combination of connectivity and cycle number as shown in \ref{fig:connectcycle3D3x3} provides a better classification of the dependence of robustness upon network topology.
+The combination of connectivity and cycle number as shown in \ref{fig:combined}E provides a better classification of the dependence of robustness upon network topology.
 Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity.

### 2014-09-22 tex/figures.tex
combine figures characterizing stability and robustness

@@
 \centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/apstab3x3.pdf}
-\caption{{\bf Plot of stability versus connectivity for three component systems.} }
-\label{fig:apstab3x3}
+\noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
+\caption{{\bf Characterization of stability and robustness} (A) (B) (C) (D) (E) (F) }
+\label{fig:combined}
 \end{figure}
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
+% Figure translations
+% apstab3x3 4 -> A
+% stab3x3 6 -> B
+% cycle3x3 7 -> C
+% New -> D
+% connectcycle3D3x3 5 -> E
+% New -> F
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\columnwidth]{fig/apstab3x3.pdf}
+% \caption{{\bf Plot of stability versus connectivity for three component systems.} }
+% \label{fig:apstab3x3}
+% \end{figure}
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
+% \caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
+% \label{fig:connectcycle3D3x3}
+% \end{figure}
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
+% \caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
+% \label{fig:stab3x3}
+% \end{figure}
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
+% \caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
+% \label{fig:cycle3x3}
+% \end{figure}

### 2014-09-22 paper.tex
formatting

@@
 A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc}.
+The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
+
+The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
+Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
+Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
+Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
+This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
 
@@
 We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.
-We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix which is given by a special case of \ref{eq:robustness}
-\begin{align}\label{eq:condprobgen}
- R (stab, k_1, \ldots k_m) =
+We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix.
+This is a special case of \ref{eq:robustness}:
+\begin{equation}\label{eq:condprobgen}
+ R (\mathrm{stab}, k_1, \ldots k_m) =
   \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
-    \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity)
-    \mathbf{1}_{stab}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)}
+    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
   {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
-\end{align}
+\end{equation}
+where we made the abbreviations
+\begin{align*}
+\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
+\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
+\end{align*}
+i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.
 
-\todo{Generalize to resampling a random link and to resampling multiple edges.}
+While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.
+The result of this operation corresponds to averaging our quantity over subsets of links:
+\begin{equation}
+R (\hbox{avg stab}, m) =
+\frac{1}{\binom{\connectivity}{m}}
+\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
+R (\mathrm{stab}, k_1, \ldots k_m)
+\end{equation}
 
@@
 \section{Component decomposition and robustness}
-
-The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of matrices whose connectivity is given by that digraph.
-Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
-Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
-Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
-This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
-Some simple examples of this are given by the graphs in \ref{fig:scc}.
 

### 2014-09-22 tex/figures.tex
fix stability and robustness figure caption

@@
 \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness} (A) (B) (C) (D) (E) (F) }
+\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus graph edit distance, (E) robustness versus number of cycles and connectivity, and (F) robustness versus graph edit distance and connectivity for three component systems.}
 \label{fig:combined}

### 2014-09-22 tex/abstract.tex
edit abstract and intro
update abstract

@@
 %!TEX root = ../paper.tex
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution of structural stability over system connectivity and cycle number, parameters that have significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.
-We show that structural stability is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most stable systems are also the most hierarchical.
-This results in the classification of the dynaical robustness of biological networks based upon a purely topological property.
+The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components.
+When the details of these interactions are unknown for systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints that can be considered to apply to all such systems.
+Here we analyze the stability and robustness of a large class of dynamical systems.
+In particular, we precisely determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.
+We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.
+We also demonstrate that permutation of strongly connected components is a fundamental symmetry of system robustness.
+This results in the classification of the dynamical robustness of biological networks based upon a purely topological property.

### 2014-09-22 paper.tex
update title
edit introduction

@@
 {\Large
-\textbf{Ensemble Stability of Biological Networks}
+\textbf{Hierarchical Network Structure Promotes Dynamical Robustness}
 }
@@
 
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
+The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
+Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
 In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
@@
 We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two or three interacting components.
-We then proceed to determine the probability of \emph{structural stability} or \emph{robustness} which is the distribution of stability to perturbation in the system structure for systems which are determined to be stable \cite{Smale1967}.
+We then proceed to determine the probability of \emph{robustness} or \emph{structural stability} which is the distribution of stability to perturbation in the system structure for systems which are determined to be stable \cite{Smale1967}.
 This quantity is of interest at all scales of the biological hierarchy, and has been previously addressed in the study of the robustness of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.

### 2014-09-22 paper.tex
move bounds on robustness section to end
reorganize and add todos

@@
 % \section{Todo}
-% \listoftodos
+\listoftodos
+\tableofcontents
 
@@
 
-\section{Stability analysis of biological networks}\label{sec:stabanalbn}
+\section{Dynamical systems on biological networks}
 In the construction of a class of potential models for biological systems at any level of the biological hierarchy from metabolic to ecosystem-level networks, it is common to first attempt to define a collection of observable phenomena of interest and determine a domain (such as binary numbers, integers, or real numbers) in which each observable can be quantified.
@@
 having connectivity $\connectivity$ equal to the number of edges of the graph, in this case $\connectivity = 3$.
-For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
+For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
+ \begin{displaymath}
+   \adj(G_X)_{ij} = \left\{
+     \begin{array}{ll}
+       1, & F_i \hbox{ depends on } x_j\\
+       0, & F_i \hbox{ does not depend on } x_j
+     \end{array}.
+   \right.
+\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
 \begin{align*}
@@
 
-This description can be formalized as a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components.
-A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
-
-The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
-Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
-Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
-Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
-This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
-
-Corresponding to this decomposition we can construct a graph $\hier (G)$.
-Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
-There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ iff there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
-
-One can also perform this construction in the oppsite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
-a strongly connected graph $C_n$.
-To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
-The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
+\section{Stability analysis of biological networks}
+Linearization of the $F_i$ about any fixed point $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
 \begin{equation}\label{eq:lineardynsys}
@@
 
-\section{Stability and robustness analysis of multi-component systems}
+\todo{Finish writing these criteria with reference to Gantmacher}
+Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
+Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$,
+writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
+Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
+$M$ be stable is equivalent to asking that the roots of $\chi$ all have
+negative real part.  Hence, let us call a polynomial stable if all its
+roots have negative real part.  We will formulate conditions for this to
+happen in terms of the coefficients of $\chi$ \cite{Gantmacher1959}.
+
+\begin{lemma}
+If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
+\end{lemma}
+\begin{lemma}
+If a point lies on the boundary of the set of stable polynomials, then
+either $s_n = 0$ and $\Delta = 0$.
+\end{lemma}
+\begin{lemma}
+If $n < 6$, then a polynomial $p$ is stable if and only
+if $\Delta > 0$ and $s_1 > 0$ for all $i$.
+\end{lemma}
+
+\section{Network hierarchy and strongly connected components}
+
+A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components.
+A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
+The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
+\todo{add hierarchy diagrams to \ref{fig:scc}}
+
+The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
+Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
+Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
+Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
+This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
+
+Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$.
+Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
+There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
+Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
+
+One can also perform this construction in the oppsite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
+a strongly connected graph $C_n$.
+To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
+The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
+\todo{explain how robustness in \ref{eq:robustness} is symmetric under permutations of the SCCs in $G$ or equivalently of the nodes of $\hier{G}$.}
+
+\todo{add panel to \ref{fig:scc} showing that many graphs get mapped to the same SCC DAG}
+
+\section{Stability and robustness analysis of particular system ensembles}
 Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$.
@@
 relevant to our calculations.
-
-Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
-Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$,
-writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
-Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
-$M$ be stable is equivalent to asking that the roots of $\chi$ all have
-negative real part.  Hence, let us call a polynomial stable if all its
-roots have negative real part.  We will formulate conditions for this to
-happen in terms of the coefficients of $\chi$ \cite{Gantmacher1959}.
-
-\begin{lemma}
-If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
-\end{lemma}
-\begin{lemma}
-If a point lies on the boundary of the set of stable polynomials, then
-either $s_n = 0$ and $\Delta = 0$.
-\end{lemma}
-\begin{lemma}
-If $n < 6$, then a polynomial $p$ is stable if and only
-if $\Delta > 0$ and $s_1 > 0$ for all $i$.
-\end{lemma}
 
@@
 
-\section{Dependency of robustness upon connectivity and cycle number}
 It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
@@
 
-\section{Component decomposition and robustness}
+\section{Component decomposition, hierarchy, and the symmetries of robustness}
 
@@
 \end{align}
+\todo{explain how robustness in \ref{eq:sccrobustness} is symmetric under transformations on the the SCCs}
 
-\section{Strongly connected components, robustness, and hierarchy}
+% \section{Strongly connected components, robustness, and hierarchy}
 The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
@@
 % \ref{fig:scc} shows systems with varying levels of hierarchy and modularity.
-
-\section{Bounds on robustness}
-
-Suppose that $S$ is a region in $[0,1]^2$.  Then, the robustness of $S$ under resampling is given as
-\begin{align*}
-R &= \frac{\int_0^1 dx \int_0^1 dy \int_0^1 dx'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x',y) +
-           \int_0^1 dx \int_0^1 dy \int_0^1 dx\,dy\,dy'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x,y')}
-          {2 \int_0^1 dx \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)} \\
-  &= \frac{\int_0^1 dy\, (\int_0^1 dx\, \mathbf{1}_{S} (x,y))^2 +
-           \int_0^1 dx\, (\int_0^1 dy\, \mathbf{1}_{S} (x,y))^2}
-          {2 \int_0^1 dx\, \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)}
-\end{align*}
-We will show that $R \le \int_{H^2} dx\,dy\, \mathbf{1}_{S} (x,y)$ with equality occurring when $S$ is rearrangement equivalent to either the rectangle of height 1 or a rectangle of width 1.
-
-First, we start with the special case when $S$ is a rectangle with one vertex at the origin and sides parallel to the coordinate axes.
-If $a$ is the height of our rectangle and $b$ is its width, we have
-
-Next, consider the case where $R$ consists of a union of squares of a rectangular grid obtained by subdividing the $x$ and $y$ axes of our bounding square into $2N$ portions of equal length.
-Given two integers $i$ and $j$
-
-Suppose we restrict to rectangular regions in an $n$-dimensional hypercube and resample $m$ variables.
-Then we have $R = s_m (a_1, \ldots a_n)$ where $s_m$ is the $m$-th elementary symmetric polynomial.
-As before, we have the constraint that $\prod_i a_i = s_n (a_1, \ldots a_n)$ is constant.
-Using Lagrange multipliers, we obtain the equation
-\[
-\frac{\partial s_m (a_1, \ldots a_n)}{a_k} = \lambda \frac{\partial s_n (a_1, \ldots a_n)}{a_k} .
-\]
-Since the elementary symmetic polynomial is linear in each of its variables, this reduces to
-\[
-s_{m-1} (a_1, \ldots {\hat a}_k \ldots a_n) = \lambda \partial s_{n-1} (a_1, \ldots {\hat a}_k \ldots a_n) .
-\]
-Dividing through, this becomes
-\[
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \lambda
-\]
-or, eliminating $\lambda$, we have
-\[
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots \frac{1}{a_n}\right) =
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
-\]
-for all $j,k$.  Suppose that $j < k$.  Then, we may expaand both sides,
-\begin{align*}
-&\frac{1}{a_k} s_{m-2} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \\ \qquad
-&\frac{1}{a_j} s_{m-2} s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
-\end{align*}
-and conclude that $a_j = a_k$.
-Hence, the only critical point lies at the point $a_1 = a_2 = \cdots = a_n$.
-By the second derivative test, that point is a minimum, so the maximum must lie on the boundary.
 
@@
 \section{Stability criteria}
-\todo{Since there is no point in publishing proofs of results contemporary
-with Darwin's Origin of Species, this section will be replaced with a
-reference to Gantmacher.}
+% \todo{Since there is no point in publishing proofs of results contemporary with Darwin's Origin of Species, this section will be replaced with a reference to Gantmacher.}
 Suppose that $M$ is an $n \times n$ matrix with real coefficients.
-As described in \ref{sec:stabanalbn}, we call
+As previously described, we call
 the matrix $M$ stable if all its eigenvalues have negative real part.
@@
 \section{Paths and eigenvalues}
-\todo{I am thinking that we might want to jettison this section from
-the published version and leave it as private notes for future work.
-As a substitute for the published version, I made a section in front
-of it which takes a more direct route to proving what is needed for
-the purposes of showing that the characteristic polynomial factors and
-making our arguments about total orders being the most stable.  We
-should discuss this point tomorrow.}
+% \todo{I am thinking that we might want to jettison this section from
+% the published version and leave it as private notes for future work.
+% As a substitute for the published version, I made a section in front
+% of it which takes a more direct route to proving what is needed for
+% the purposes of showing that the characteristic polynomial factors and
+% making our arguments about total orders being the most stable.  We
+% should discuss this point tomorrow.}
 
@@
 
-\todo{Start talking about Abelianizing the category and the algebra, say
-more after decomposing into components.}
+% \todo{Start talking about Abelianizing the category and the algebra, say
+% more after decomposing into components.}
 
-\todo{Characterize strongly connected components in terms of
-fibration over poset.}
+% \todo{Characterize strongly connected components in terms of
+% fibration over poset.}
 
@@
 
+\section{Bounds on robustness}
+
+Suppose that $S$ is a region in $[0,1]^2$.  Then, the robustness of $S$ under resampling is given as
+\begin{align*}
+R &= \frac{\int_0^1 dx \int_0^1 dy \int_0^1 dx'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x',y) +
+           \int_0^1 dx \int_0^1 dy \int_0^1 dx\,dy\,dy'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x,y')}
+          {2 \int_0^1 dx \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)} \\
+  &= \frac{\int_0^1 dy\, (\int_0^1 dx\, \mathbf{1}_{S} (x,y))^2 +
+           \int_0^1 dx\, (\int_0^1 dy\, \mathbf{1}_{S} (x,y))^2}
+          {2 \int_0^1 dx\, \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)}
+\end{align*}
+We will show that $R \le \int_{H^2} dx\,dy\, \mathbf{1}_{S} (x,y)$ with equality occurring when $S$ is rearrangement equivalent to either the rectangle of height 1 or a rectangle of width 1.
+
+First, we start with the special case when $S$ is a rectangle with one vertex at the origin and sides parallel to the coordinate axes.
+If $a$ is the height of our rectangle and $b$ is its width, we have
+
+Next, consider the case where $R$ consists of a union of squares of a rectangular grid obtained by subdividing the $x$ and $y$ axes of our bounding square into $2N$ portions of equal length.
+Given two integers $i$ and $j$
+
+Suppose we restrict to rectangular regions in an $n$-dimensional hypercube and resample $m$ variables.
+Then we have $R = s_m (a_1, \ldots a_n)$ where $s_m$ is the $m$-th elementary symmetric polynomial.
+As before, we have the constraint that $\prod_i a_i = s_n (a_1, \ldots a_n)$ is constant.
+Using Lagrange multipliers, we obtain the equation
+\[
+\frac{\partial s_m (a_1, \ldots a_n)}{a_k} = \lambda \frac{\partial s_n (a_1, \ldots a_n)}{a_k} .
+\]
+Since the elementary symmetic polynomial is linear in each of its variables, this reduces to
+\[
+s_{m-1} (a_1, \ldots {\hat a}_k \ldots a_n) = \lambda \partial s_{n-1} (a_1, \ldots {\hat a}_k \ldots a_n) .
+\]
+Dividing through, this becomes
+\[
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \lambda
+\]
+or, eliminating $\lambda$, we have
+\[
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots \frac{1}{a_n}\right) =
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
+\]
+for all $j,k$.  Suppose that $j < k$.  Then, we may expaand both sides,
+\begin{align*}
+&\frac{1}{a_k} s_{m-2} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \\ \qquad
+&\frac{1}{a_j} s_{m-2} s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
+s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
+\end{align*}
+and conclude that $a_j = a_k$.
+Hence, the only critical point lies at the point $a_1 = a_2 = \cdots = a_n$.
+By the second derivative test, that point is a minimum, so the maximum must lie on the boundary.
+
 \input{tex/figuressupp}

### 2014-09-22 paper.tex
link notation of special cases of computing robustness back to the more general formula

@@
 
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
+% Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
 This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
+% $$
+% P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
+% $$
+% where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
 In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
+Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$.
+We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.
+We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix.
+This is a special case of \ref{eq:robustness}:
+\begin{equation}\label{eq:condprobgen}
+ R (\mathrm{stab}, k_1, \ldots k_m) =
+  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
+    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
+  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
+\end{equation}
+where we made the abbreviations
+\begin{align*}
+\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
+\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
+\end{align*}
+i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.
+
+While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.
+The result of this operation corresponds to averaging our quantity over subsets of links:
+\begin{equation}
+R (\hbox{avg stab}, m) =
+\frac{1}{\binom{\connectivity}{m}}
+\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
+R (\mathrm{stab}, k_1, \ldots k_m)
+\end{equation}
+
+In order to evaluate these integrals, one needs to describe the stable region.
+The theory of how to do so in terms of conditions on the coefficients of the
+characteristic polynomials was developed by various mathematicians during the
+second half of the nineteenth century and is summarized in chapter XV of
+\cite{Gantmacher1959}.  We now summarize some of these results which are
+relevant to our calculations.
 
@@
 \section{Stability and robustness analysis of particular system ensembles}
-Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.
-We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix.
-This is a special case of \ref{eq:robustness}:
-\begin{equation}\label{eq:condprobgen}
- R (\mathrm{stab}, k_1, \ldots k_m) =
-  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
-    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
-  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
-\end{equation}
-where we made the abbreviations
-\begin{align*}
-\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
-\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
-\end{align*}
-i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.
-
-While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.
-The result of this operation corresponds to averaging our quantity over subsets of links:
-\begin{equation}
-R (\hbox{avg stab}, m) =
-\frac{1}{\binom{\connectivity}{m}}
-\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
-R (\mathrm{stab}, k_1, \ldots k_m)
-\end{equation}
-
-In order to evaluate these integrals, one needs to describe the stable region.
-The theory of how to do so in terms of conditions on the coefficients of the
-characteristic polynomials was developed by various mathematicians during the
-second half of the nineteenth century and is summarized in chapter XV of
-\cite{Gantmacher1959}.  We now summarize some of these results which are
-relevant to our calculations.
-
+For systems having two components, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}.
+For those having three components, we can estimate these same quantities using Monte Carlo simulations.
+Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis.
 For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
@@
 where $a + c < 0$ and $ac > bd$.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$.
-Suppose that we resample $b$.
-We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is a special case of \ref{eq:condprobgen}
+For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},k)$ where $k=1 \ldots 4$.
+By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $R(\mathrm{stab},1)$ and $R(\mathrm{stab},2)$.
+Suppose that we resample $b$ to compute $R(\mathrm{stab},2)$.
 % \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
+% \begin{align}\label{eq:condprob}
+% P\left(\begin{pmatrix}
+% a & b' \\
+% d & c
+% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right)
+% & = \frac{P\left(\begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
+% a & b' \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
+% a & b \\
+% d & c
+% \end{pmatrix} \textrm{stable } \right)}.
+% \end{align}
+The denominator of \ref{eq:condprobgen} in this case is given by
 \begin{align*}
@@
 Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
+Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprobgen}
 \begin{align*}

### 2014-09-22 paper.tex
remove extra occurrence of Table

@@
 
-Using this approach the probability of \emph{a priori} stability and of robustness for all two component systems is given in Table \ref{tab:structstabmat}.
+Using this approach the probability of \emph{a priori} stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.
 Similarly, the analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.

### 2014-09-22 paper.tex
add todos regarding the definition of graph edit distance and hierarchy

@@
 
+\todo{add definition of graph edit distance to the most hierarchical network and define this as our measure of hierarchy}
+
 \section{Stability and robustness analysis of particular system ensembles}
@@
 This network is the most hierarchical of all three component systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \ref{fig:scc}A has a block triangular structure.
+
+\todo{add description of \ref{fig:scc}D and \ref{fig:scc}F referring to the monotonic correlation between the upper bound of robustness and degree of hierarchy as defined by graph edit distance in a preceeding section}
 

### 2014-09-23 paper.tex
organization
add definition of graph edit distance and remove unnecessary todo notes

@@
 The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
-\todo{add hierarchy diagrams to \ref{fig:scc}}
 
@@
 
-One can also perform this construction in the oppsite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
+One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
 a strongly connected graph $C_n$.
@@
 The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
-\todo{explain how robustness in \ref{eq:robustness} is symmetric under permutations of the SCCs in $G$ or equivalently of the nodes of $\hier{G}$.}
 
-\todo{add panel to \ref{fig:scc} showing that many graphs get mapped to the same SCC DAG}
+We note from the definition that the mapping from $G$ to $\hier(G)$ is many-to-one as exemplified in \ref{fig:scc}B and C.
+\todo{explain how robustness in \ref{eq:robustness} is symmetric under permutations of the SCCs in $G$ or equivalently of the nodes of $\hier(G)$.}
 
-\todo{add definition of graph edit distance to the most hierarchical network and define this as our measure of hierarchy}
+\todo{add reference for graph edit distance}
+The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
+The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \ref{fig:scc}A.
+This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
+The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second.
+This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical.
+In this work we take this distance to be the definition of hierarchy.
 
@@
 
+The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
+
 Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomials factors as the product of the charactericstic polynomials of its diagonal blocks.
@@
 
+Inspection of \ref{eq:sccrobustness} demonstrates that robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
+
+\todo{write down analogs of \ref{eq:sccrobustness} and \ref{eq:sccmaxrobustness} for resampling a fixed proportion or an arbitrary number of components}
+
 % \section{Strongly connected components, robustness, and hierarchy}
-The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
-Inspection of \ref{eq:sccrobustness} demonstrates that robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
+
 % \section{Hierarchical systems are the most robust}

### 2014-09-23 paper.tex
add example of robustness symmetry

@@
 
-\todo{Generalize to open systems and say how we combine them.}
+In order to discuss how different parts of the system influence each other, we shall generalize our definition to include open systems.
+We do this by having some of our variables be control variables whose values are given as specified functions of time rather than determined by the dynamics.
+Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
+For details and formalization in terms of operads see \cite{Vagner2014}.
+Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
+However, for most purposes, such a decomposition would be uninteresting and what we instead want is a decomposition into relatively self-contained modules.
+In the section on connected components, we shall propose a way of accomplishing this based upon the topology of the system graph.
 
@@
 \section{Stability analysis of biological networks}
+The class of dynamical quantities on which we shall focus in this investigation involve stability of equilibria.
+Suppose that $\vec x$ is a point on our phase space which depends upon the parameters.
 Linearization of the $F_i$ about any fixed point $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
@@
 The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
+By combining our decomposition of the graph with our analysis of matrices, we will deduce a general result which relates robustness to hierarchy.
 
@@
 Using these observations, we may express the robustness of a graph in terms of the robustnesses of its strongly connected components.
-If the components $C_i$ of our graph $G$ has $v_i$ vertices connectivity $\connectivity_i$, and robustness $R_i$ and there are $\ell_{ij}$ links between components $i$ and $j$, then the total robustness to resampling a random entry is
+If the components $C_i$ of our graph $G$ has $v_i$ vertices connectivity $\connectivity_i$ and there are $\ell_{ij}$ links between components $i$ and $j$, then the total robustness to resampling a random entry is
+\begin{align}\label{eq:sccrobustness}
+R(G) = \frac{\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i R(C_i)}{\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i}.
+\end{align}
+For instance, if our graph is the one in \ref{fig:scc}B or C, then we have two connected components, one with two nodes, and one with one node.
+From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and and robustness $0.62$.
+The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
+Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is
+\[
+R(G) = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
+\]
+which agrees with the value computed in \ref{tab:structstabmat}.
+
+Next, we derive an upper bound on the robustness of any graph constructed from a given set of connected components.
+Since $R(C_i) \le 1$ for all $i$, it follws that increasing $\ell_{ij}$ will increase $R(G)$.
+Given two connected components $C_i$ and $C_j$ with $v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$ links going from $C_i$ to $C_j$.
+Hence, $\ell_{ij} \le v_i v_j$, so, we conclude that
 \begin{align}
-R(G) = \frac{\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i R(C_i)}
-            {\sum_{i=1}^n \sum_{j=1}^n \ell_{ij} + \sum_{i=1}^n c_i}. \label{eq:sccrobustness}
+R(G) \le \frac{\sum_{i,j \in \hier(G)} v_i v_j + \sum_i \connectivity_i R(C_i)}
+{\sum_{i,j \in \hier(G)} v_i v_j + \sum_i \connectivity_i}, \label{eq:sccmaxrobustness}
 \end{align}
 
-Next, we ask what is the maximally robust graph that has a given set of connected components.
+Since every acyclic digraph can be embedded into a totally ordered set, we may assume without loss of generality that our components have been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i < j$.
+Hence, after substituting into our expression and simplifying using the identity
+$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} \right)^2-\frac{1}{2} \sum_{i=1}^{n} v_i^2,$$
+we arrive at the following upper bound on robustness in terms of the connected components:
 \begin{align}
-R_{\mathrm{max}} = \frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i R_i}{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i c_i}, \label{eq:sccmaxrobustness}
+R_{\mathrm{max}}(C_1, \ldots C_n) =
+\frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i \connectivity_i R(C_i)}
+{\frac{1}{2} ((\sum_i v_i)^2 - \sum_i v_i^2) + \sum_i \connectivity_i}, \label{eq:sccmaxrobustness}
 \end{align}
-\todo{explain how robustness in \ref{eq:sccrobustness} is symmetric under transformations on the the SCCs}
+Furthermore, this bound is attained.
+Suppose that $G_{\mathrm{tot}}$ is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.
+Then, by the construction described in the previous section on network hierachy, we have a graph $G_{\mathrm{max}}$ such that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.
+By our formula, $R(G_{\mathrm{max}}) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
 
-Inspection of \ref{eq:sccrobustness} demonstrates that robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
+The preceding argument demonstrates the manner in which robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
+And, the graph that has the latter property is that associated to the total ordering, $G_{\mathrm{tot}}$, which aslo possess the highest degree of hierarchy.
+The fact that robustness is equivalent for any configuration of strongly connected components with an equivalent number of links between them demonstrates that it is invariant to permutations of strongly connected components.
+This result does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled because the arguments presented here are purely topological in nature.
+The result that dynamical robustness is correlated with system hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles primarily studied here.
 
@@
 \section{Discussion}
+\todo{write Discussion}
 

### 2014-09-23 paper.tex
added reference for graph edit distance

@@
 
-\todo{add reference for graph edit distance}
 The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
@@
 This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second.
+The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
 This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical.
@@
 \frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i \connectivity_i R(C_i)}
-{\frac{1}{2} ((\sum_i v_i)^2 - \sum_i v_i^2) + \sum_i \connectivity_i}, \label{eq:sccmaxrobustness}
+{\frac{1}{2} ((\sum_i v_i)^2 - \sum_i v_i^2) + \sum_i \connectivity_i}, \label{eq:exsccmaxrobustness}
 \end{align}

### 2014-09-23 paper.tex
add description of hierarchy figures

@@
 
-Using this approach the probability of \emph{a priori} stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.
+Using this approach the probability of stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.
 Similarly, the analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.
@@
 For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:combined}A.
+
 The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:combined}B.
@@
 
-\todo{add description of \ref{fig:scc}D and \ref{fig:scc}F referring to the monotonic correlation between the upper bound of robustness and degree of hierarchy as defined by graph edit distance in a preceeding section}
+This observation suggested that graph edit distance from \ref{fig:scc}A, hierarchy, might provide a better characterization of system robustness.
+\ref{fig:scc}D shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
+There is a monotonic correlation between the upper bound of robustness and hierarchy.
+\ref{fig:scc}F shows system robustness as a function of both hierarchy and connectivity.
+The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:combined}E.
 

### 2014-09-23 paper.tex
add todo
clarify linear stability analysis

@@
 
-In order to discuss how different parts of the system influence each other, we shall generalize our definition to include open systems.
-We do this by having some of our variables be control variables whose values are given as specified functions of time rather than determined by the dynamics.
-Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
-For details and formalization in terms of operads see \cite{Vagner2014}.
-Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
-However, for most purposes, such a decomposition would be uninteresting and what we instead want is a decomposition into relatively self-contained modules.
-In the section on connected components, we shall propose a way of accomplishing this based upon the topology of the system graph.
+In order to discuss how different parts of the system influence each
+other, we shall generalize our definition to include open systems.  We
+do this by having some of our variables be control variables whose
+values are given as specified functions of time rather than determined
+by the dynamics.  Given several such open systems, we may combine them
+to produce a larger system by setting the control variables of a
+subsystem equal to the dynamical variables of another system.  For
+details and a discussion of the operad structure of composition of
+open systems see \cite{Vagner2014}.  Taking this notion to its logical
+extreme, one can dissect a system into a collection of one open system
+for each dynamical variable.  However, for most purposes, such a
+decomposition would be uninteresting and what we instead want is a
+decomposition into relatively self-contained modules.  In the section
+on connected components, we shall propose a way of accomplishing this
+based upon the topology of the system graph.
 
-In order to accommodate the possibility of uncertainty in our modelling,
-we will generalize our notion of a system on a network to that of a
-system with random parameters.  Again, this will involve three
-steps. First, we provide a measure space $S, \mathcal{B}(S)$ which represents the
-values over which our parameters can vary. Next, instead of a single
-vector field, we consider a family of vector fields parameterized by
-this space. That is to say, we introduce a mapping $V \colon S \to \Gamma T(\prod_i M_m)$.
-(Here $\Gamma$ denotes the space of sections of the tangent bundle.)
-As before, for each point $p \in S$, the vector field $V(p)$ must be consistent with
+In order to accommodate the possibility of uncertainty in our
+modelling, we will generalize our notion of a system on a network to
+that of a system with random parameters.  Again, this will involve
+three steps. First, we provide a measure space $S, \mathcal{B}(S)$
+which represents the values over which our parameters can vary. Next,
+instead of a single vector field, we consider a family of vector
+fields parameterized by this space. That is to say, we introduce a
+mapping $V \colon S \to \Gamma T(\prod_i M_m)$.  (Here $\Gamma$
+denotes the space of sections of the tangent bundle.)  As before, for
+each point $p \in S$, the vector field $V(p)$ must be consistent with
 the system graph. Finally, we select a probability measure $\mu \colon
@@
 
-Another question one can ask is that of robustness, which measures the
-probability that, if some property holds for a set of parameters, it will
-continue to hold if we make a random perturbation about those parameter values.
-To define robustness, we will introduce, in addition to the probability
-distribution $\mu$ described above, a family of probability distributions
-$\mu' \colon S \times \mathcal{B}(S) \to [0,1]$.  Then, given a binary random
-variable $q$, we define its robustness as the following conditional proability:
+A quality which is of interest to us is robustness or structural
+stability \cite{Smale1967}, which requires the determination of
+whether or not a given dynamical system that is determined to be
+stable remains stable under a perturbation to one or more of its
+defining parameters.  We will quantify this as the probability that,
+if some property holds for a set of parameters, it will continue to
+hold if we make a random perturbation about those parameter values.
+To do so, we will introduce, in addition to the probability
+distribution $\mu$ described above, a family of probability
+distributions $\mu' \colon S \times \mathcal{B}(S) \to [0,1]$ such
+that $\mu'(x,y)$ encodes the conditional probability that a system
+with a parameter value $x$ will have its parameter values changed to
+$y$ under a perturbation.  For our biological models, we will be
+interested in large perturbations rather than the small or
+infinitessimal perturbations usually considered in the theory of
+dynamical systems.  Then, given a binary random variable $q$,
+we define its robustness as the following conditional proability:
 \begin{align}\label{eq:robustness}
@@
 \section{Stability analysis of biological networks}
-The class of dynamical quantities on which we shall focus in this investigation involve stability of equilibria.
-Suppose that $\vec x$ is a point on our phase space which depends upon the parameters.
-Linearization of the $F_i$ about any fixed point $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
+The class of dynamical quantities on which we shall focus in this
+investigation involve stability of equilibria.  Suppose that $\vec x$
+is a point on our phase space which depends upon the parameters.  Then
+we may take $q$ to be a binary random variable which describes whether
+or not $\vec x$ is a fixed point of the system (i.e. $q$ is true if and only if
+$F_i(\vec{x}^0)=0$ for all $i$) and perform a probabilistic analysis
+of the sort discussed above.
+
+Furthermore, if it turns out that $\vec x$ is indeed a fixed point, we
+may proceed to ask whether it is a dynamically stable fixed point.
+Intuitively, dynamic stability means that, if one chooses the initial
+conditions sufficiently close to the fixeed point, the solution will
+stay close to the fixed point.  Physically, this is important because,
+if a fixed point ${\vec x}_0$ is unstable, we have zero proability of
+observing the solution ${\vec x}(t) = {\vec x}_0$.
+
+To determine stability, we linearize the equations of motion about the
+fixed point $\vec{x}^0$:
 \begin{equation}\label{eq:lineardynsys}
@@
 
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
+Even though the determination of stability only requires examining
+linearized equations of motion, it is worth noting that knowing the
+stability of a fixed point can yield information about the non-linear
+dynamics of a system.  For instance, in two dimensions,
+Poincare-Bendixson theory implies that every limit cycle must encircle
+at least one fixed point.  Furthermore, in the case where the cycle
+contains a single fixed point (such as in the Volterra-Lotke model of
+conflicting populations), the stability of the fixed point will be
+determined by whether the limit cycle is attractive or repulsive.
 
-% Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
+%% To determine whether a randomly selected dynamical system evaluated at
+%% a random critical point, given by a matrix such as $A$, is stable, it
+%% is sufficient to check the above condition. Integrating over the
+%% region of the parameter space defining $A$ enables the determination
+%% of the probability of stability to perturbations in state $\vec{x}$
+%% over an ensemble of dynamical systems.
+
+% Another form of stability, referred to as structural
+% stability \cite{Smale1967}, requires the determination of whether or
+% not a given dynamical system that is determined to be stable remains
+% stable under a perturbation to one of its defining parameters. This
+% can be formalized as the conditional probability distribution
 % $$
@@
 % $$
-% where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-Suppose that $M$ is an $n \times n$ matrix with real coefficients and connectivity $\connectivity$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin, and having volume $r^\connectivity$.
-We want to know the probability that, if we resample the elements $k_1, k_2, \ldots k_m$ of this matrix, the resulting matrix will correspond to a stable matrix.
-This is a special case of \ref{eq:robustness}:
+% where $A'$ represents a matrix derived from some perturbation
+% applied to the matrix $A$ given above. In the simple case where
+% $A \in \mathbb{R}^{n \times n}$ the perturbation which involves
+% sampling uniformly over the $a_{ij}$ defining $A$ the value of a
+% particular $a_{ij}$ from a given probability distribution.
+
+Suppose that our Jacobian $A$ is an $n \times n$ matrix with real
+coefficients and connectivity $\connectivity$ and denote the
+proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.  Then we may
+compute the robustness of this proposition using \ref{eq:robustness}
+once we have specified the proability densities $\mu$ and $\mu'$.  In
+general, these will depend upon the parameters of the non-linear
+system.  For the puprpose of the current investigation, we shall make
+the ansatz that these are such that the distribution of entries of $A$
+is (at least approximately) the uniform distribution
+$\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube,
+$H^\connectivity$, of edge length $r=2$, centered about the origin.
+To model the perturbations, we shall resample the elements $k_1,
+k_2, \ldots k_m$ of this matrix from the same uniform distribution.
+Under these assumptions, the expression for robustness becomes
+of \ref{eq:robustness}:
 \begin{equation}\label{eq:condprobgen}
@@
     \mathbf{1}_{stab} \mathbf{1}_{stab'}}
-  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)}
+  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)},
 \end{equation}
@@
 
-We note from the definition that the mapping from $G$ to $\hier(G)$ is many-to-one as exemplified in \ref{fig:scc}B and C.
-\todo{explain how robustness in \ref{eq:robustness} is symmetric under permutations of the SCCs in $G$ or equivalently of the nodes of $\hier(G)$.}
+This map $\hier$ is many-to-one and so there is a large class of
+operations which leaves $\hier(G)$ invariant for a given graph $G$.
+For instance, we may interchange the positions of the strongly
+connected components relative to each other.  Leaving the components
+fixed, we may move links between nodes in a component or betwwen
+components, or even add or delete links.  Some examples of these
+operations are illustrated in \ref{fig:scc}B and C.  As we shall see
+later, these operations leave invariant key dynamical quantities, so
+we may regard them as a symmetry groupoid of our system.
 
@@
 
-Using these observations, we may express the robustness of a graph in terms of the robustnesses of its strongly connected components.
-If the components $C_i$ of our graph $G$ has $v_i$ vertices connectivity $\connectivity_i$ and there are $\ell_{ij}$ links between components $i$ and $j$, then the total robustness to resampling a random entry is
+Using these observations, we may express the robustness of a graph in
+terms of the robustnesses of its strongly connected components.  If
+the components $C_i$ of our graph $G$ has $v_i$ vertices connectivity
+$\connectivity_i$ and there are $\ell_{ij}$ links between components
+$i$ and $j$, then there are a total of
+$\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i$
+links in $G$.  If we resample a randomly chosen link in a component
+$C_i$, the probability that the system will remain stable after
+resamling is $R(\hbox{avg stab} C_i, 1)$.  If we resample a link
+between components $C_i$ and $C_j$, the system will remain stable with
+probability 1.  Hence, the probability that the system will remain
+stable upon resampling a random link is the weighted average of these
+probabilities:
 \begin{align}\label{eq:sccrobustness}
-R(G) = \frac{\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i R(C_i)}{\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i}.
+R(\hbox{avg stab} G, 1) =
+\frac{\sum_{i,j \in \hier(G)} \ell_{ij} +
+      \sum_{i=1}^n \connectivity_i R(C_i)}
+     {\sum_{i,j \in \hier(G)} \ell_{ij} +
+      \sum_{i=1}^n \connectivity_i}.
 \end{align}
@@
 By our formula, $R(G_{\mathrm{max}}) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+
+This argument also works when we resample more than one entry,
+although the notation becomes more complicated.  Suppose that we
+resample $m$ nodes.  Then, given a partition $m = \sum_{i=0}^n m_i$,
+there are ${m \choose m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$
+links from $C_i$ and $m_0$ links between strongly connected
+components.  Hence, our weighted average becomes
+\begin{equation}
+R(\hbox{avg stab} C_i, m) =
+\frac{sum_{m_0, m_1, \ldots m_n} {m \choose m_0, m_1, \ldots. m_n}
+      m_0 + \sum_{i=1}^n m_i C_i}
+     {}
+\end{equation}
 
@@
 \section{Tables}
+\todo{Fill out notation table}
 \input{tex/tablenotation}

### 2014-09-23 paper.tex
add error estimate for Monte Carlo simulation

@@
 Poincare-Bendixson theory implies that every limit cycle must encircle
-at least one fixed point.  Furthermore, in the case where the cycle
-contains a single fixed point (such as in the Volterra-Lotke model of
+at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
 conflicting populations), the stability of the fixed point will be
@@
 
-\todo{Say a few words about Monte Carlo integration and error bars.}
+If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
+We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}.
+The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
+$$
+\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
+$$
+where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
+$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
+which corresponds in this case to the maximum likelihood estimate
+$$
+\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
+$$
+This estimate is characterized by the variance of the posterior Beta distribution
+$$
+\mathrm{var}(\theta | \mathcal{D}) = \frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
+$$
+Since for the chosen prior $a=b=1 \ll N$ this simplifies to
+$$
+\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
+$$
+yielding the error estimate given by the associated standard deviation.
+In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
 

### 2014-09-23 paper.tex
add formula for robustness under resampling multiple interaction weights

@@
 Hence, after substituting into our expression and simplifying using the identity
-$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} \right)^2-\frac{1}{2} \sum_{i=1}^{n} v_i^2,$$
+$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n} v_i^2,$$
 we arrive at the following upper bound on robustness in terms of the connected components:
@@
 This argument also works when we resample more than one entry,
-although the notation becomes more complicated.  Suppose that we
-resample $m$ nodes.  Then, given a partition $m = \sum_{i=0}^n m_i$,
+although the notation becomes more complicated.  Suppose that, as
+before, we have a graph $G$ with strongly connected components $C_i$,
+but that now we
+resample $m$ nodes.  Then, given a partition $m = \sum_{i=0}^n m_i$
+such that $m_0 \le \sum_{i=1}^{n-1}\sum_{j=i}^n \ell_{ij}$ and $m_i 
+\le \connectivity_i$ when $1 \le i \le n$,
 there are ${m \choose m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$
 links from $C_i$ and $m_0$ links between strongly connected
-components.  Hence, our weighted average becomes
+components.  Hence, if we denote the set of partitions which satisfy
+our inequalities as $M$, our weighted average becomes
 \begin{equation}
-R(\hbox{avg stab} C_i, m) =
-\frac{sum_{m_0, m_1, \ldots m_n} {m \choose m_0, m_1, \ldots. m_n}
-      m_0 + \sum_{i=1}^n m_i C_i}
-     {}
+R(\hbox{avg stab} (G), m) =
+\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M} {m \choose m_0, m_1, \ldots. m_n}
+      \left(m_0 + \sum_{i=1}^n m_i R(\hbox{avg stab} (C_i), m_i)\right)}
+     {m \sum\limits_{(m_0, m_1, \ldots m_n) \in M} {m \choose m_0, m_1, \ldots. m_n}}
 \end{equation}

### 2014-09-24 paper.tex
update description of resampling multiple interaction weights

@@
 links in $G$.  If we resample a randomly chosen link in a component
-$C_i$, the probability that the system will remain stable after
-resamling is $R(\hbox{avg stab} C_i, 1)$.  If we resample a link
+$C_i$, then the probability that the system will remain stable after
+resampling is $R(\hbox{avg stab} (C_i), 1)$.  If we resample a link
 between components $C_i$ and $C_j$, the system will remain stable with
@@
 \begin{align}\label{eq:sccrobustness}
-R(\hbox{avg stab} G, 1) =
-\frac{\sum_{i,j \in \hier(G)} \ell_{ij} +
-      \sum_{i=1}^n \connectivity_i R(C_i)}
-     {\sum_{i,j \in \hier(G)} \ell_{ij} +
-      \sum_{i=1}^n \connectivity_i}.
+R(\hbox{avg stab} (G), 1) =
+\frac{\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
+      \sum\limits_{i=1}^n \connectivity_i R(C_i)}
+     {\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
+      \sum\limits_{i=1}^n \connectivity_i}.
 \end{align}
@@
 \]
-which agrees with the value computed in \ref{tab:structstabmat}.
+which agrees with the value computed in \ref{tab:structstabmat} up to
+sampling error.
 
-Next, we derive an upper bound on the robustness of any graph constructed from a given set of connected components.
-Since $R(C_i) \le 1$ for all $i$, it follws that increasing $\ell_{ij}$ will increase $R(G)$.
-Given two connected components $C_i$ and $C_j$ with $v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$ links going from $C_i$ to $C_j$.
-Hence, $\ell_{ij} \le v_i v_j$, so, we conclude that
+Next, we derive an upper bound on the robustness of any graph
+constructed from a given set of connected components.  Since
+$R(C_i) \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
+increase $R(G)$.  Given two connected components $C_i$ and $C_j$ with
+$v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$
+links going from $C_i$ to $C_j$.  Hence, $\ell_{ij} \le v_i v_j$, so,
+we conclude that
 \begin{align}
-R(G) \le \frac{\sum_{i,j \in \hier(G)} v_i v_j + \sum_i \connectivity_i R(C_i)}
-{\sum_{i,j \in \hier(G)} v_i v_j + \sum_i \connectivity_i}, \label{eq:sccmaxrobustness}
+R(G) \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
+               \sum\limits_{i=1}^n \connectivity_i R(C_i)}
+              {\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
+               \sum\limits_{i=1}^n \connectivity_i},
 \end{align}
 
-Since every acyclic digraph can be embedded into a totally ordered set, we may assume without loss of generality that our components have been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i < j$.
-Hence, after substituting into our expression and simplifying using the identity
-$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n} v_i^2,$$
-we arrive at the following upper bound on robustness in terms of the connected components:
-\begin{align}
+Since every acyclic digraph can be embedded into a totally ordered
+set, we may assume without loss of generality that our components have
+been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i <
+j$.  Hence, after substituting into our expression and simplifying
+using the identity
+$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
+v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n}
+v_i^2,$$
+we arrive at the following upper bound on robustness in terms
+of the connected components:
+\begin{align} \label{eq:sccmaxrobustness}
 R_{\mathrm{max}}(C_1, \ldots C_n) =
-\frac{\frac{1}{2} ((\sum_i n_i)^2 - \sum_i n_i^2) + \sum_i \connectivity_i R(C_i)}
-{\frac{1}{2} ((\sum_i v_i)^2 - \sum_i v_i^2) + \sum_i \connectivity_i}, \label{eq:exsccmaxrobustness}
+\frac{\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
+                    \sum_{i=1}^n \connectivity_i R(\hbox{avg stab}(C_i),1)}
+     {\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
+                    \sum_{i=1}^n \connectivity_i}.
 \end{align}
-Furthermore, this bound is attained.
-Suppose that $G_{\mathrm{tot}}$ is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.
-Then, by the construction described in the previous section on network hierachy, we have a graph $G_{\mathrm{max}}$ such that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.
-By our formula, $R(G_{\mathrm{max}}) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
+is the graph on $n$ nodes with a link from node $i$ to node $j$
+whenever $i < j$.  Then, by the construction described in the previous
+section on network hierachy, we have a graph $G_{\mathrm{max}}$ such
+that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and
+$\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula,
+$R(\hbox{avg stab} (G_{\mathrm{max}}), 1) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
 
 This argument also works when we resample more than one entry,
-although the notation becomes more complicated.  Suppose that, as
-before, we have a graph $G$ with strongly connected components $C_i$,
-but that now we
-resample $m$ nodes.  Then, given a partition $m = \sum_{i=0}^n m_i$
-such that $m_0 \le \sum_{i=1}^{n-1}\sum_{j=i}^n \ell_{ij}$ and $m_i 
-\le \connectivity_i$ when $1 \le i \le n$,
-there are ${m \choose m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$
-links from $C_i$ and $m_0$ links between strongly connected
-components.  Hence, if we denote the set of partitions which satisfy
-our inequalities as $M$, our weighted average becomes
+although the notation becomes more complicated.  Suppose that we
+resample $m$ nodes.  Define
+\begin{equation*}
+M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
+m = \sum_{i=0}^n m_i \quad\&\quad
+m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
+\forall i \in \{1, \ldots, n\} \, m_i < \connectivity_i \right\}.
+\end{equation*}
+Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
+m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
+$m_0$ links between strongly connected components.  Hence, our
+weighted average becomes
 \begin{equation}
 R(\hbox{avg stab} (G), m) =
-\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M} {m \choose m_0, m_1, \ldots. m_n}
-      \left(m_0 + \sum_{i=1}^n m_i R(\hbox{avg stab} (C_i), m_i)\right)}
-     {m \sum\limits_{(m_0, m_1, \ldots m_n) \in M} {m \choose m_0, m_1, \ldots. m_n}}
+\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
+      {m \choose m_0, m_1, \ldots. m_n}
+      \left(m_0 + \sum\limits_{i=1}^n m_i R(\hbox{avg stab} (C_i), m_i) \right)}
+     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
+      {m \choose m_0, m_1, \ldots. m_n} m}
 \end{equation}
 
-The preceding argument demonstrates the manner in which robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
-And, the graph that has the latter property is that associated to the total ordering, $G_{\mathrm{tot}}$, which aslo possess the highest degree of hierarchy.
-The fact that robustness is equivalent for any configuration of strongly connected components with an equivalent number of links between them demonstrates that it is invariant to permutations of strongly connected components.
-This result does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled because the arguments presented here are purely topological in nature.
-The result that dynamical robustness is correlated with system hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles primarily studied here.
+As before, since $R(\hbox{avg stab} (C_i), m_i) \le 1$, we may
+increase $R(\hbox{avg stab} (G), m)$ by increasing the maximum
+possible value of $m_0$ while keeping the strongly connected
+components the same.  Again, if we fix $\hier(G)$, the maximum
+possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
+whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
+v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
+G_{tot}$.  Hence, we conclude that $R(\hbox{avg stab} (G), m) \le
+R(\hbox{avg stab} (G_{\mathrm{max}}), m)$.
 
-\todo{write down analogs of \ref{eq:sccrobustness} and \ref{eq:sccmaxrobustness} for resampling a fixed proportion or an arbitrary number of components}
+The preceding argument demonstrates the manner in which robustness is
+maximized by maximizing the number of edges \emph{between} the strongly
+connected components of the graph underlying system interdependencies.
+And, the graph that has the latter property is that associated to the
+ordering, $G_{\mathrm{tot}}$, which aslo possess the highest degree of
+hierarchy. The fact that robustness is equivalent for any configuration
+of strongly connected components with an equivalent number of links
+between them demonstrates that it is invariant to permutations of
+strongly connected components. This result does not depend at all upon
+any particular details such as the probability distribution from which
+the component interaction strengths are sampled because the arguments
+presented here are purely topological in nature. The result that
+dynamical robustness is correlated with system hierarchy therefore
+applies to an even broader class of dynamical systems than the particular
+random ensembles primarily studied here.
 

### 2014-09-24 paper.tex
update bibliography style

@@
 % \section{Todo}
-\listoftodos
-\tableofcontents
+% \listoftodos
+% \tableofcontents
 
@@
 
-\bibliographystyle{amsplain}
+\bibliographystyle{pnas}
 \bibliography{bib/books,bib/papers}
@@
 
+\beginsupplement
+\setcounter{secnumdepth}{4}
+\section{Supplementary Material}
+
 \section{Tables}
-\todo{Fill out notation table}
-\input{tex/tablenotation}
+% \todo{Fill out notation table}
+% \input{tex/tablenotation}
 \input{tex/table2x2stab}
@@
 
-\section{Stability criteria}
-% \todo{Since there is no point in publishing proofs of results contemporary with Darwin's Origin of Species, this section will be replaced with a reference to Gantmacher.}
-Suppose that $M$ is an $n \times n$ matrix with real coefficients.
-As previously described, we call
-the matrix $M$ stable if all its eigenvalues have negative real part.
-Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
-Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$,
-writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
-Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
-$M$ be stable is equivalent to asking that the roots of $\chi$ all have
-negative real part.  Hence, let us call a polynomial stable if all its
-roots have negative real part.  We will formulate conditions for this to
-happen in terms of the coefficients of $\chi$.
-
-\begin{lemma}
-If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
-\end{lemma}
-
-\begin{proof}
-Assume that $M$ is stable and that it has $2m$ complex roots which come in
-complex conjugate pairs and $n-m$ real roots.  Let $r_1 \ldots, r_{2m}$ be
-the complex roots with $r_{2k+1}$ being the complex conjugate of $r_{2k}$
-and let $r_{2m+1}, \ldots,  r_{n}$ be the real roots. Then we can write the
-factorization as follows:
-\[
-\chi(M)(z) = \prod_{k=1}^{m} (z^2 - (r_{2k} + r_{2k+1}) z + r_{2k} r_{2k+1})
-             \prod_{k=2m+1}^{n} (z - r_k)
-\]
-Since $r_{2k}$ and $r_{2k+1}$ are complex conjugates, $r_{2k} + r_{2k+1} =
-\Re r_{2k} = \Re r_{2k+1}$ and $r_{k} r_{2k+1} > 0$.   Since all the
-roots have negative real part, all the coefficients of the terms in
-the products above are positive so, when we multiply out the product,
-we will have all positive terms.  Hence, we conclude that a necessary
-condition for $M$ to be stable is that $s_k > 0$ for all $k$.
-\end{proof}
-
-To state the next criterion, we will require the quantity $\Delta$, which
-is defined as $\prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j)$.  Since
-this expression is invariant under permuting the roots, it may be expressed
-as a function of the coefficients $s_i$; for small values of $n$, the
-resulting expression looks as follows:
-\begin{align*}
-n=2 \qquad &\Delta = s_1 \\
-n=3 \qquad &\Delta = s_3 - s_1 s_2 \\
-n=4 \qquad &\Delta = s_1 s_2 s_3 - s_3^2 - s_1^2 s_4 \\
-n=5 \qquad &\Delta = s_1 s_2 s_3 s_4 - s_3^2 s_4 - s_1^2 s_4^2 -
-                     s_1 s_2^2 s_5 + s_2 s_3 s_5 + 2 s_1 s_4 s_5 - s_5^2
-\end{align*}
-
-\begin{lemma}
-If a point lies on the boundary of the set of stable polynomials, then
-either $s_n = 0$ and $\Delta = 0$.
-\end{lemma}
-
-\begin{proof}
-Let $S_n$ denote the symmetric group on $n$ objects and let $f \colon
-\mathbb{C}^n/S_n \to \mathbb{C}^n$ be the map that sends a set of roots
-to the n-tuplet of coefficients of a polynomial.  This map is explicitly
-given by the elemetary symmetric polynomials.  By the fundamental theorem
-of algebra, it is invertible and, by ???, its inverse is continuous.
-
-Suppose that $p$ is a stable polynomial.  By definition, this means that
-$\Re r > 0$ for every root $r$ of $p$.  Since the set $\{r \mid \Re r > 0\}$
-is an open subset of $\mathbb{C}^n/S_n$, the continuity of $f^{-1}$ implies
-that the set of stable polynomials is open.  Likewise, the set of unstable
-polynomials none of whose roots have real part equal to zero is open.
-
-Hence, if a point lies on the boundary of the stable region, it must
-correspond to a polynomial with at least one root whose real part is zero.
-There are two ways this could happen.  If the root is real, then having
-zero real part is the same as the root being zero, hence $s_n = 0$.  If
-the root is complex, then having zero real part means that the root and
-its complex conjugate are purely imaginary and sum to zero, hence $\Delta = 0$.
-\end{proof}
-
-\begin{lemma}
-If $n < 6$, then a polynomial $p$ is stable if and only
-if $\Delta > 0$ and $s_1 > 0$ for all $i$.
-\end{lemma}
-
-\begin{proof}
-Suppose that, to the contrary, there exists a polynomial which satisfies
-the inequalities $\Delta > 0$ and $s_1 > 0$ but which is not stable.
-Consider a straight line connecting this polynomial to a stable polynomial,
-say $(x - 1)^n$.
-
-Suppose that $p$ is a polynomial for which $\Delta = 0$ and $s_i > 0$ but
-which is not on the boundary of the stable region.  In order to have
-$\Delta = 0$, we must either have a pair of real roots which sum to zero
-or a pair of complex conjugate roots with zero real part.  In the former
-case, we would have $x^2 - r^2$ as a factor of our polynomial and in the
-latter, we would have $x^2 + r^2$ as a factor.
-
-Suppose that $x^2 - r^2$ is a factor.  Then we have
-\[
-p = (x^2 - r^2) \left(x^{n-2} + \sum_{k=1}^{n-3} c_k x^{n-k-2} \right) =
-x^n + \sum_{k=3}^{n-3} (c_k - r^2 c_{k-2}) x^{n-k} +
-\]
-
-Suppose that $x^2 + r^2$ is a factor. Then we have one of the following
-when $2 < n < 6$
-\begin{align*}
-(x^2 + r^2)(x + a_1) &=
-x^3 + a_1 x^3 + r^2 x^2 + a_1 r^2 \\
-(x^2 + r^2)(x^2 + a_1 x + a_2) &=
-x^4 + a_1 x^3 + (a_2 + r^2) x^2 + a_1 r^2 x + a_2 r^2 \\
-(x^2 + r^2)(x^3 + a_1 x^2 + a_2 x + a_3) &=
-x^5 + a_1 x^4 + (a_2 + r^2) x^3 + (a_3 + a_1 r^2) x^2 + a_2 r^2 x + a_3 r^2
-\end{align*}
-Inspecting these, we see that, if the coefficients of $p$ are positive,
-then $a_i > 0$ for all $i$.
-
-$\ldots$
-\end{proof}
-
-For $n = 6$, this argument breaks down because, for instance, we have
-the polynomial $x^6 + x^5 +\frac{1}{2} x^4 + 2x^3 + \frac{1}{2} x^2 +
-x + 1$  which factors as  $(x^2 + 1)(x^4 + x^3 - \frac{1}{2} x^2 + x + 1)$.
-This polynomial has positive cofficients and $\Delta = 0$ but does not
-lie on the boundary of the stable region because the second factor has
-a negative coefficient, hence has a root with positive real part.  Thus,
-for $n = 6$ and higher, we need to supplement our conditions with
-additional inequalities (which may be taken to be linear) to exclude
-such possibilitiies.  However, we will not pursue this topic further
-in this work because we will only be making computations with matrices
-of size less than $6$.
-
-\section{Paths and eigenvalues}
-% \todo{I am thinking that we might want to jettison this section from
-% the published version and leave it as private notes for future work.
-% As a substitute for the published version, I made a section in front
-% of it which takes a more direct route to proving what is needed for
-% the purposes of showing that the characteristic polynomial factors and
-% making our arguments about total orders being the most stable.  We
-% should discuss this point tomorrow.}
-
-Next, we will want to relate prpoperties of our matrix and its eigenvalues
-to  properties of the associated graph $G_A$.  To do this, we will first
-examine the category of paths on $G_A$, and derive a formula expressing
-the chararacteristic polynomial as a sum over cycles.
-
-Given a digraph $G$, let $\Path (a,b)$ denote the set of paths from the
-vertex $a$ to the vertex $b$ and let $\Path(G) = \bigcup_{a,b \in G} \Path (a,b)$.
-Given a path $p_1$ from $a$ to $b$ and a path $p_2$ from $b$ to $c$, we may
-concatenate them to obtain a path $p_1 \circ p_2$ from $a$ to $c$.  Since
-this operation is associative and has identity elements,
-$(\Vertex(G), \Path(G), \circ)$ forms a category.  Define
-$\mathcal{A}(G)$ to be the set of all linear combinations of elements of
-$\Path(G)$ with coefficients in $\mathbb{R}$.  Then $\mathcal{A}(G)$ forms
-an associative algebra under the multiplication induced by $\circ$.
-
-Next, we introduce two basic functors from $\Path (G)$.  One of these is
-$\length \colon \Path(G) \to \mathbb{N}$ (regarding the natural numbers
-as a category with a single object) which sends each path to its length.
-The other functor is the weight product functor.  Given an assignement
-$w \colon \Path(G) \to \mathbb(R)$ of weights to the edges of our digraph,
-define $\wprod \colon \Path(G) \to \mathbb{R}$ to be the function which sends
-each path to the product of the weights associated to the edges which
-that path traverses.  Given the way that graphs are associated to matrices,
-weights on edges correspond to matrix entries and, given a path
-$v_1, v_2, v_3 \ldots$, we have $\wprod(v_1, v_2, v_3 \ldots) =
-m_{v_1 v_2} m_{v_2 v_3} \cdots$.  Going in the opposite direction, we
-note that any functor from $\Path(G)$ to $\mathbb{R}$ may be regarded as
-a weighted product for a suitable choice of weights.  We extend $\wprod$ to
-$\mathcal{A}(G)$ by linearity; the result is an algebra homomorphism.
-
-% \todo{Start talking about Abelianizing the category and the algebra, say
-% more after decomposing into components.}
-
-% \todo{Characterize strongly connected components in terms of
-% fibration over poset.}
-
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored
-according to the strongly connected components of the graph $G$.
-\[
-\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}
-\]
-where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the
-matrix $M$ with $z$ a complex variable.  This formula is readily verified when
-$M$ is diagonal and has the nice property of being additive over decompositions
-of the matrix. Making the change of variables $x=\frac{1}{z}$, we have
-\[
-\tr (I - xM)^n =
-\sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} \wprod(p) x^{\length(p)}
-\]
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.  ; $F$ is the functor from this category
-to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-As an illustration of this formula, we will use it to compute the characteristic
-polynomial of the matrix associated to the labelled graph in figure ??.  We begin
-by noting that the closed paths of this graph can be described by the following
-regular expressions:
-\begin{align*}
-\Path (1,1) &: (e + bd + bca)^* \\
-\Path (2,2) &: (ae^*(be^*d)^*bc)^* \\
-\Path (3,3) &: ((ca + d)e^*b)^* \\
-\end{align*}
-Since these regular expressions are non-redundant, we can obtain generating
-functions for the paths by considering $a,b,c,d,e$ as non-commutative
-variables and replacing the Kleene star $x^*$ by $(1 - x)^{-1}$:
-\begin{align*}
-\Path (1,1) &: (1 - e - bd - bca)^{-1} \\
-\Path (2,2) &: (1 - a(1-e)^{-1}(1 - b(1-e)^{-1}d)^{-1}bc)^{-1} \\
-\Path (3,3) &: (1 - (ca + d)(1 - e)^{-1}b)^{-1} \\
-\end{align*}
-To obtain the characteristic polynomial, we make the substitutions
-$a \mapsto m_{12}/z, b \mapsto m_{31}/z, c \mapsto m_{23}/z, d \mapsto m_{13}/z,
-e \mapsto m_{11}/z$ and simplify, using the fact that the matrix entries
-commute:
-\begin{align*}
-\Path (1,1) &: \frac{z^3}
-{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} } \\
-\Path (2,2) &: \frac{z(z^2 - m_{11} z - m_{13} m_{31})}
-{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} }  \\
-\Path (3,3) &: \frac{z^2(z^2 - m_{11})}
-{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} } \\
-\end{align*}
-Adding these three quantitites together, we obtain
-\[
-\frac{z(3 z^2 - 2 m_{11} z - m_{13} m_{31})}
-{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} },
-\]
-which we recognize as $\chi'(z)/\chi(z)$, where
-$\chi(z) = z^3−m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31}$.
-finally, we note that this polynomial is indeed the characteristic
-polynomial of the matrix
-\[
- \begin{pmatrix}
-   m_{11} & m_{12} & m_{13} \\
-   0      & 0      & m_{23} \\
-   m_{31} & 0      & 0 \\
- \end{pmatrix}
-\]
-corresponding to our digraph.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Bounds on robustness}
-
-Suppose that $S$ is a region in $[0,1]^2$.  Then, the robustness of $S$ under resampling is given as
-\begin{align*}
-R &= \frac{\int_0^1 dx \int_0^1 dy \int_0^1 dx'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x',y) +
-           \int_0^1 dx \int_0^1 dy \int_0^1 dx\,dy\,dy'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x,y')}
-          {2 \int_0^1 dx \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)} \\
-  &= \frac{\int_0^1 dy\, (\int_0^1 dx\, \mathbf{1}_{S} (x,y))^2 +
-           \int_0^1 dx\, (\int_0^1 dy\, \mathbf{1}_{S} (x,y))^2}
-          {2 \int_0^1 dx\, \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)}
-\end{align*}
-We will show that $R \le \int_{H^2} dx\,dy\, \mathbf{1}_{S} (x,y)$ with equality occurring when $S$ is rearrangement equivalent to either the rectangle of height 1 or a rectangle of width 1.
-
-First, we start with the special case when $S$ is a rectangle with one vertex at the origin and sides parallel to the coordinate axes.
-If $a$ is the height of our rectangle and $b$ is its width, we have
-
-Next, consider the case where $R$ consists of a union of squares of a rectangular grid obtained by subdividing the $x$ and $y$ axes of our bounding square into $2N$ portions of equal length.
-Given two integers $i$ and $j$
-
-Suppose we restrict to rectangular regions in an $n$-dimensional hypercube and resample $m$ variables.
-Then we have $R = s_m (a_1, \ldots a_n)$ where $s_m$ is the $m$-th elementary symmetric polynomial.
-As before, we have the constraint that $\prod_i a_i = s_n (a_1, \ldots a_n)$ is constant.
-Using Lagrange multipliers, we obtain the equation
-\[
-\frac{\partial s_m (a_1, \ldots a_n)}{a_k} = \lambda \frac{\partial s_n (a_1, \ldots a_n)}{a_k} .
-\]
-Since the elementary symmetic polynomial is linear in each of its variables, this reduces to
-\[
-s_{m-1} (a_1, \ldots {\hat a}_k \ldots a_n) = \lambda \partial s_{n-1} (a_1, \ldots {\hat a}_k \ldots a_n) .
-\]
-Dividing through, this becomes
-\[
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \lambda
-\]
-or, eliminating $\lambda$, we have
-\[
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots \frac{1}{a_n}\right) =
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
-\]
-for all $j,k$.  Suppose that $j < k$.  Then, we may expaand both sides,
-\begin{align*}
-&\frac{1}{a_k} s_{m-2} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \\ \qquad
-&\frac{1}{a_j} s_{m-2} s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
-s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
-\end{align*}
-and conclude that $a_j = a_k$.
-Hence, the only critical point lies at the point $a_1 = a_2 = \cdots = a_n$.
-By the second derivative test, that point is a minimum, so the maximum must lie on the boundary.
+\input{tex/extra}
 

### 2014-09-24 tex/extra.tex
add supplement section and reformat figures

\section{Stability criteria}
% \todo{Since there is no point in publishing proofs of results contemporary with Darwin's Origin of Species, this section will be replaced with a reference to Gantmacher.}
Suppose that $M$ is an $n \times n$ matrix with real coefficients.
As previously described, we call
the matrix $M$ stable if all its eigenvalues have negative real part.
Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$,
writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
$M$ be stable is equivalent to asking that the roots of $\chi$ all have
negative real part.  Hence, let us call a polynomial stable if all its
roots have negative real part.  We will formulate conditions for this to
happen in terms of the coefficients of $\chi$.

\begin{lemma}
If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
\end{lemma}

\begin{proof}
Assume that $M$ is stable and that it has $2m$ complex roots which come in
complex conjugate pairs and $n-m$ real roots.  Let $r_1 \ldots, r_{2m}$ be
the complex roots with $r_{2k+1}$ being the complex conjugate of $r_{2k}$
and let $r_{2m+1}, \ldots,  r_{n}$ be the real roots. Then we can write the
factorization as follows:
\[
\chi(M)(z) = \prod_{k=1}^{m} (z^2 - (r_{2k} + r_{2k+1}) z + r_{2k} r_{2k+1})
             \prod_{k=2m+1}^{n} (z - r_k)
\]
Since $r_{2k}$ and $r_{2k+1}$ are complex conjugates, $r_{2k} + r_{2k+1} =
\Re r_{2k} = \Re r_{2k+1}$ and $r_{k} r_{2k+1} > 0$.   Since all the
roots have negative real part, all the coefficients of the terms in
the products above are positive so, when we multiply out the product,
we will have all positive terms.  Hence, we conclude that a necessary
condition for $M$ to be stable is that $s_k > 0$ for all $k$.
\end{proof}

To state the next criterion, we will require the quantity $\Delta$, which
is defined as $\prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j)$.  Since
this expression is invariant under permuting the roots, it may be expressed
as a function of the coefficients $s_i$; for small values of $n$, the
resulting expression looks as follows:
\begin{align*}
n=2 \qquad &\Delta = s_1 \\
n=3 \qquad &\Delta = s_3 - s_1 s_2 \\
n=4 \qquad &\Delta = s_1 s_2 s_3 - s_3^2 - s_1^2 s_4 \\
n=5 \qquad &\Delta = s_1 s_2 s_3 s_4 - s_3^2 s_4 - s_1^2 s_4^2 -
                     s_1 s_2^2 s_5 + s_2 s_3 s_5 + 2 s_1 s_4 s_5 - s_5^2
\end{align*}

\begin{lemma}
If a point lies on the boundary of the set of stable polynomials, then
either $s_n = 0$ and $\Delta = 0$.
\end{lemma}

\begin{proof}
Let $S_n$ denote the symmetric group on $n$ objects and let $f \colon
\mathbb{C}^n/S_n \to \mathbb{C}^n$ be the map that sends a set of roots
to the n-tuplet of coefficients of a polynomial.  This map is explicitly
given by the elemetary symmetric polynomials.  By the fundamental theorem
of algebra, it is invertible and, by ???, its inverse is continuous.

Suppose that $p$ is a stable polynomial.  By definition, this means that
$\Re r > 0$ for every root $r$ of $p$.  Since the set $\{r \mid \Re r > 0\}$
is an open subset of $\mathbb{C}^n/S_n$, the continuity of $f^{-1}$ implies
that the set of stable polynomials is open.  Likewise, the set of unstable
polynomials none of whose roots have real part equal to zero is open.

Hence, if a point lies on the boundary of the stable region, it must
correspond to a polynomial with at least one root whose real part is zero.
There are two ways this could happen.  If the root is real, then having
zero real part is the same as the root being zero, hence $s_n = 0$.  If
the root is complex, then having zero real part means that the root and
its complex conjugate are purely imaginary and sum to zero, hence $\Delta = 0$.
\end{proof}

\begin{lemma}
If $n < 6$, then a polynomial $p$ is stable if and only
if $\Delta > 0$ and $s_1 > 0$ for all $i$.
\end{lemma}

\begin{proof}
Suppose that, to the contrary, there exists a polynomial which satisfies
the inequalities $\Delta > 0$ and $s_1 > 0$ but which is not stable.
Consider a straight line connecting this polynomial to a stable polynomial,
say $(x - 1)^n$.

Suppose that $p$ is a polynomial for which $\Delta = 0$ and $s_i > 0$ but
which is not on the boundary of the stable region.  In order to have
$\Delta = 0$, we must either have a pair of real roots which sum to zero
or a pair of complex conjugate roots with zero real part.  In the former
case, we would have $x^2 - r^2$ as a factor of our polynomial and in the
latter, we would have $x^2 + r^2$ as a factor.

Suppose that $x^2 - r^2$ is a factor.  Then we have
\[
p = (x^2 - r^2) \left(x^{n-2} + \sum_{k=1}^{n-3} c_k x^{n-k-2} \right) =
x^n + \sum_{k=3}^{n-3} (c_k - r^2 c_{k-2}) x^{n-k} +
\]

Suppose that $x^2 + r^2$ is a factor. Then we have one of the following
when $2 < n < 6$
\begin{align*}
(x^2 + r^2)(x + a_1) &=
x^3 + a_1 x^3 + r^2 x^2 + a_1 r^2 \\
(x^2 + r^2)(x^2 + a_1 x + a_2) &=
x^4 + a_1 x^3 + (a_2 + r^2) x^2 + a_1 r^2 x + a_2 r^2 \\
(x^2 + r^2)(x^3 + a_1 x^2 + a_2 x + a_3) &=
x^5 + a_1 x^4 + (a_2 + r^2) x^3 + (a_3 + a_1 r^2) x^2 + a_2 r^2 x + a_3 r^2
\end{align*}
Inspecting these, we see that, if the coefficients of $p$ are positive,
then $a_i > 0$ for all $i$.

$\ldots$
\end{proof}

For $n = 6$, this argument breaks down because, for instance, we have
the polynomial $x^6 + x^5 +\frac{1}{2} x^4 + 2x^3 + \frac{1}{2} x^2 +
x + 1$  which factors as  $(x^2 + 1)(x^4 + x^3 - \frac{1}{2} x^2 + x + 1)$.
This polynomial has positive cofficients and $\Delta = 0$ but does not
lie on the boundary of the stable region because the second factor has
a negative coefficient, hence has a root with positive real part.  Thus,
for $n = 6$ and higher, we need to supplement our conditions with
additional inequalities (which may be taken to be linear) to exclude
such possibilitiies.  However, we will not pursue this topic further
in this work because we will only be making computations with matrices
of size less than $6$.

\section{Paths and eigenvalues}
% \todo{I am thinking that we might want to jettison this section from
% the published version and leave it as private notes for future work.
% As a substitute for the published version, I made a section in front
% of it which takes a more direct route to proving what is needed for
% the purposes of showing that the characteristic polynomial factors and
% making our arguments about total orders being the most stable.  We
% should discuss this point tomorrow.}

Next, we will want to relate prpoperties of our matrix and its eigenvalues
to  properties of the associated graph $G_A$.  To do this, we will first
examine the category of paths on $G_A$, and derive a formula expressing
the chararacteristic polynomial as a sum over cycles.

Given a digraph $G$, let $\Path (a,b)$ denote the set of paths from the
vertex $a$ to the vertex $b$ and let $\Path(G) = \bigcup_{a,b \in G} \Path (a,b)$.
Given a path $p_1$ from $a$ to $b$ and a path $p_2$ from $b$ to $c$, we may
concatenate them to obtain a path $p_1 \circ p_2$ from $a$ to $c$.  Since
this operation is associative and has identity elements,
$(\Vertex(G), \Path(G), \circ)$ forms a category.  Define
$\mathcal{A}(G)$ to be the set of all linear combinations of elements of
$\Path(G)$ with coefficients in $\mathbb{R}$.  Then $\mathcal{A}(G)$ forms
an associative algebra under the multiplication induced by $\circ$.

Next, we introduce two basic functors from $\Path (G)$.  One of these is
$\length \colon \Path(G) \to \mathbb{N}$ (regarding the natural numbers
as a category with a single object) which sends each path to its length.
The other functor is the weight product functor.  Given an assignement
$w \colon \Path(G) \to \mathbb(R)$ of weights to the edges of our digraph,
define $\wprod \colon \Path(G) \to \mathbb{R}$ to be the function which sends
each path to the product of the weights associated to the edges which
that path traverses.  Given the way that graphs are associated to matrices,
weights on edges correspond to matrix entries and, given a path
$v_1, v_2, v_3 \ldots$, we have $\wprod(v_1, v_2, v_3 \ldots) =
m_{v_1 v_2} m_{v_2 v_3} \cdots$.  Going in the opposite direction, we
note that any functor from $\Path(G)$ to $\mathbb{R}$ may be regarded as
a weighted product for a suitable choice of weights.  We extend $\wprod$ to
$\mathcal{A}(G)$ by linearity; the result is an algebra homomorphism.

% \todo{Start talking about Abelianizing the category and the algebra, say
% more after decomposing into components.}

% \todo{Characterize strongly connected components in terms of
% fibration over poset.}

The characteristic polynomial $\chi(A)$ associated to $A$ can be factored
according to the strongly connected components of the graph $G$.
\[
\frac{\chi(M)'(z)}{\chi(M)(z)} = \tr (zI - M)^{-1}
\]
where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the
matrix $M$ with $z$ a complex variable.  This formula is readily verified when
$M$ is diagonal and has the nice property of being additive over decompositions
of the matrix. Making the change of variables $x=\frac{1}{z}$, we have
\[
\tr (I - xM)^n =
\sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} \wprod(p) x^{\length(p)}
\]
where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
paths of length $n$ from $a$ to $b$.  ; $F$ is the functor from this category
to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.

As an illustration of this formula, we will use it to compute the characteristic
polynomial of the matrix associated to the labelled graph in figure ??.  We begin
by noting that the closed paths of this graph can be described by the following
regular expressions:
\begin{align*}
\Path (1,1) &: (e + bd + bca)^* \\
\Path (2,2) &: (ae^*(be^*d)^*bc)^* \\
\Path (3,3) &: ((ca + d)e^*b)^* \\
\end{align*}
Since these regular expressions are non-redundant, we can obtain generating
functions for the paths by considering $a,b,c,d,e$ as non-commutative
variables and replacing the Kleene star $x^*$ by $(1 - x)^{-1}$:
\begin{align*}
\Path (1,1) &: (1 - e - bd - bca)^{-1} \\
\Path (2,2) &: (1 - a(1-e)^{-1}(1 - b(1-e)^{-1}d)^{-1}bc)^{-1} \\
\Path (3,3) &: (1 - (ca + d)(1 - e)^{-1}b)^{-1} \\
\end{align*}
To obtain the characteristic polynomial, we make the substitutions
$a \mapsto m_{12}/z, b \mapsto m_{31}/z, c \mapsto m_{23}/z, d \mapsto m_{13}/z,
e \mapsto m_{11}/z$ and simplify, using the fact that the matrix entries
commute:
\begin{align*}
\Path (1,1) &: \frac{z^3}
{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} } \\
\Path (2,2) &: \frac{z(z^2 - m_{11} z - m_{13} m_{31})}
{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} }  \\
\Path (3,3) &: \frac{z^2(z^2 - m_{11})}
{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} } \\
\end{align*}
Adding these three quantitites together, we obtain
\[
\frac{z(3 z^2 - 2 m_{11} z - m_{13} m_{31})}
{z^3 - m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31} },
\]
which we recognize as $\chi'(z)/\chi(z)$, where
$\chi(z) = z^3−m_{11} z^2 - m_{13} m_{31} z - m_{12} m_{23} m_{31}$.
finally, we note that this polynomial is indeed the characteristic
polynomial of the matrix
\[
 \begin{pmatrix}
   m_{11} & m_{12} & m_{13} \\
   0      & 0      & m_{23} \\
   m_{31} & 0      & 0 \\
 \end{pmatrix}
\]
corresponding to our digraph.

We now make use of the additivity.  Let $\hier$ be the functor which
sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
for the fibers of the points in the hierarchy which corresponds to a
factorization of the characteristic polynomial.

\section{Bounds on robustness}

Suppose that $S$ is a region in $[0,1]^2$.  Then, the robustness of $S$ under resampling is given as
\begin{align*}
R &= \frac{\int_0^1 dx \int_0^1 dy \int_0^1 dx'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x',y) +
           \int_0^1 dx \int_0^1 dy \int_0^1 dx\,dy\,dy'\, \mathbf{1}_{S} (x,y) \mathbf{1}_{S} (x,y')}
          {2 \int_0^1 dx \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)} \\
  &= \frac{\int_0^1 dy\, (\int_0^1 dx\, \mathbf{1}_{S} (x,y))^2 +
           \int_0^1 dx\, (\int_0^1 dy\, \mathbf{1}_{S} (x,y))^2}
          {2 \int_0^1 dx\, \int_0^1 dx\,dy\, \mathbf{1}_{S} (x,y)}
\end{align*}
We will show that $R \le \int_{H^2} dx\,dy\, \mathbf{1}_{S} (x,y)$ with equality occurring when $S$ is rearrangement equivalent to either the rectangle of height 1 or a rectangle of width 1.

First, we start with the special case when $S$ is a rectangle with one vertex at the origin and sides parallel to the coordinate axes.  If $a$ is the height of our rectangle and $b$ is its width, we have

Next, consider the case where $R$ consists of a union of squares of a rectangular grid obtained by subdividing the $x$ and $y$ axes of our bounding square into $2N$ portions of equal length.  Given two integers $i$ and $j$

Suppose we restrict to rectangular regions in an $n$-dimensional hypercube and resample $m$ variables.  Then we have $R = s_m (a_1, \ldots a_n)$ where $s_m$ is the $m$-th elementary symmetric polynomial.  As before, we have the constraint that $\prod_i a_i = s_n (a_1, \ldots a_n)$ is constant.  Using Lagrange multipliers, we obtain the equation
\[
\frac{\partial s_m (a_1, \ldots a_n)}{a_k} = \lambda \frac{\partial s_n (a_1, \ldots a_n)}{a_k} .
\]
Since the elementary symmetic polynomial is linear in each of its variables, this reduces to
\[
s_{m-1} (a_1, \ldots {\hat a}_k \ldots a_n) = \lambda \partial s_{n-1} (a_1, \ldots {\hat a}_k \ldots a_n) .
\]
Dividing through, this becomes
\[
s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \lambda
\]
or, eliminating $\lambda$, we have
\[
s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots \frac{1}{a_n}\right) =
s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
\]
for all $j,k$.  Suppose that $j < k$.  Then, we may expaand both sides,
\begin{align*}
&\frac{1}{a_k} s_{m-2} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) = \\ \qquad
&\frac{1}{a_j} s_{m-2} s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right) +
s_{m-1} \left(\frac{1}{a_1}, \ldots {\hat a}_j \ldots {\hat a}_k \ldots \frac{1}{a_n}\right)
\end{align*}
and conclude that $a_j = a_k$.  Hence, the only critical point lies at the point $a_1 = a_2 = \cdots = a_n$.  By the second derivative test, that point is a minimum, so the maximum must lie on the boundary.

### 2014-09-24 code/rds3x3.jl

@@ -397,10 +397,10 @@ function combinegnuplot()
                 "dist3x3.p",
                 "connectcycle3D3x3.p",
                 "connectdist3D3x3.p",]
-    horiz = 2
-    vert = 3
+    horiz = 3
+    vert = 2
     rv = (String)[]
-    headstring = "set terminal svg size $(600*horiz),$(400*vert) dynamic enhanced fname 'HelveticaNeue'  fsize 16
+    headstring = "set terminal svg size $(600*horiz),$(400*vert) dynamic enhanced fname 'HelveticaNeue'  fsize 20
 set output '../fig/combinedfigs.svg'
 set multiplot layout $(vert),$(horiz)
 set tmargin 3
@@ -416,7 +416,7 @@ set rmargin 5\n"
         # close(f)
         for (j,l) in enumerate(lines)
             if j==length(lines)-1
-                push!(rv,"set label 1 \"$(string(char(64+i)))\" font \"HelveticaNeue, 23\" at graph -0.25, graph 1.20\n")
+                push!(rv,"set label 1 \"$(string(char(64+i)))\" font \"HelveticaNeue, 25\" at graph -0.25, graph 1.20\n")
             end
             push!(rv,l)
         end

### 2014-09-24 data/avgdata.jl
average overlapping points on 3D plots

@@ -1,0 +1,57 @@
+function avgconnectcycle(infilename)
+  rv = (String)[]
+  push!(rv,"connectivity\tstability\tresampling\tcycle\tdistance")
+
+  dat, head = readdlm(infilename,header=true)
+
+  for i in unique(dat[:,1]), j in unique(dat[:,4])
+
+    ind_1 = find(x->x==i,dat[:,1])
+    ind_2 = find(x->x==j,dat[:,4])
+    inds = intersect(ind_1,ind_2)
+    if length(inds)>0
+        avgstab = round(mean(dat[inds,2:3],1),4)
+        line = "$(i)\t$(avgstab[1])\t$(avgstab[2])\t$(j)\t$(dat[inds[1],5])"
+        # line = "$(i)\t$(avgstab[1])\t$(avgstab[2])\t$(dat[inds[1],4])\t$(j)"
+        push!(rv,line)
+    end
+
+  end
+
+  rv = join(rv,"\n")
+  f = open("avgconnectcycle.tsv","w")
+  write(f,rv)
+  close(f)
+  return rv
+
+end
+
+function avgconnectdist(infilename)
+  rv = (String)[]
+  push!(rv,"connectivity\tstability\tresampling\tcycle\tdistance")
+
+  dat, head = readdlm(infilename,header=true)
+
+  for i in unique(dat[:,1]), j in unique(dat[:,5])
+
+    ind_1 = find(x->x==i,dat[:,1])
+    ind_2 = find(x->x==j,dat[:,5])
+    inds = intersect(ind_1,ind_2)
+
+    if length(inds)>0
+        avgstab = round(mean(dat[inds,2:3],1),4)
+
+        # line = "$(i)\t$(avgstab[1])\t$(avgstab[2])\t$(j)\t$(dat[inds[1],5])"
+        line = "$(i)\t$(avgstab[1])\t$(avgstab[2])\t$(dat[inds[1],4])\t$(j)"
+        push!(rv,line)
+    end
+
+  end
+
+  rv = join(rv,"\n")
+  f = open("avgconnectdist.tsv","w")
+  write(f,rv)
+  close(f)
+  return rv
+
+end

### 2014-09-24 paper.tex
add supplement section and reformat figures
start discussion

@@
 \todo{write Discussion}
+The relationship between structure and function is fundamental to biology.
+Structure-function relationships exist at every level of the biological hierarchy.
+In all such networks, the ability of systems to persist over long periods of time, is dependent upon their dynamical stability.
+Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more stable and thus more likely to persist in the evolutionary process.
+This relationship transcends levels of the biological hierarchy in the sense that it is formulated mathematically and at a level of abstraction that allows its conclusions to apply independently to each of these levels.
+
+The fact that such a relationship exists has the potential to unify our understanding of biological networks.
+Our development suggests two predictions having empirical impact capable of testing this theory.
+One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
+In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
+The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
 
@@
 
-\newpage
+\pagebreak
 
@@
 
-\newpage
+\pagebreak
 \FloatBarrier
@@
 
-\input{tex/extra}
+% \input{tex/extra}
 
-\input{tex/figuressupp}
+% \input{tex/figuressupp}
 

### 2014-09-25 paper.tex
remove excess information

@@
 In order to evaluate these integrals, one needs to describe the stable region.
-The theory of how to do so in terms of conditions on the coefficients of the
-characteristic polynomials was developed by various mathematicians during the
-second half of the nineteenth century and is summarized in chapter XV of
-\cite{Gantmacher1959}.  We now summarize some of these results which are
-relevant to our calculations.
-
-\todo{Finish writing these criteria with reference to Gantmacher}
-Define the characteristic polynomial as $\chi(M)(z) = \det(zI - M)$.
-Denote the coefficients of $\chi(M)$ as $s_j$ and its roots as $r_j$,
-writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n (x-r_k)$.
-Then, since the roots of $\chi$ are the eigenvalues of $M$, asking that
-$M$ be stable is equivalent to asking that the roots of $\chi$ all have
-negative real part.  Hence, let us call a polynomial stable if all its
-roots have negative real part.  We will formulate conditions for this to
-happen in terms of the coefficients of $\chi$ \cite{Gantmacher1959}.
-
-\begin{lemma}
-If $M$ is stable, then $s_i > 0$ for all $i$ between $1$ and $n$.
-\end{lemma}
-\begin{lemma}
-If a point lies on the boundary of the set of stable polynomials, then
-either $s_n = 0$ and $\Delta = 0$.
-\end{lemma}
-\begin{lemma}
-If $n < 6$, then a polynomial $p$ is stable if and only
-if $\Delta > 0$ and $s_1 > 0$ for all $i$.
-\end{lemma}
+This can be done in terms of conditions on the coefficients of the
+characteristic polynomials associated to the systems described by matrices $A$ above \cite{Gantmacher1959}.
 

### 2014-09-25 paper.tex
update section on dynamical systems

@@
 
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V \in \Gamma T(\prod_i M_i)$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
+For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
 For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
@@
 
-In order to discuss how different parts of the system influence each
-other, we shall generalize our definition to include open systems.  We
-do this by having some of our variables be control variables whose
-values are given as specified functions of time rather than determined
-by the dynamics.  Given several such open systems, we may combine them
-to produce a larger system by setting the control variables of a
-subsystem equal to the dynamical variables of another system.  For
-details and a discussion of the operad structure of composition of
-open systems see \cite{Vagner2014}.  Taking this notion to its logical
-extreme, one can dissect a system into a collection of one open system
-for each dynamical variable.  However, for most purposes, such a
-decomposition would be uninteresting and what we instead want is a
-decomposition into relatively self-contained modules.  In the section
-on connected components, we shall propose a way of accomplishing this
-based upon the topology of the system graph.
-
 In order to accommodate the possibility of uncertainty in our
-modelling, we will generalize our notion of a system on a network to
+modeling, we will generalize our notion of a system on a network to
 that of a system with random parameters.  Again, this will involve
-three steps. First, we provide a measure space $S, \mathcal{B}(S)$
+three steps. First, we provide a measure space $S$
 which represents the values over which our parameters can vary. Next,
 instead of a single vector field, we consider a family of vector
-fields parameterized by this space. That is to say, we introduce a
-mapping $V \colon S \to \Gamma T(\prod_i M_m)$.  (Here $\Gamma$
-denotes the space of sections of the tangent bundle.)  As before, for
+fields parameterized by this space.  As before, for
 each point $p \in S$, the vector field $V(p)$ must be consistent with
-the system graph. Finally, we select a probability measure $\mu \colon
-\mathcal{B}(S) \to [0,1]$ which represents our understanding of which values of the parameters
+the system graph. Finally, we select a probability measure, $\mu$, which represents our understanding of which values of the parameters
 are likely. In accord with Bayesian statistics, we might revise this
@@
 
-A quality which is of interest to us is robustness or structural
-stability \cite{Smale1967}, which requires the determination of
-whether or not a given dynamical system that is determined to be
-stable remains stable under a perturbation to one or more of its
-defining parameters.  We will quantify this as the probability that,
-if some property holds for a set of parameters, it will continue to
-hold if we make a random perturbation about those parameter values.
-To do so, we will introduce, in addition to the probability
-distribution $\mu$ described above, a family of probability
-distributions $\mu' \colon S \times \mathcal{B}(S) \to [0,1]$ such
-that $\mu'(x,y)$ encodes the conditional probability that a system
-with a parameter value $x$ will have its parameter values changed to
-$y$ under a perturbation.  For our biological models, we will be
-interested in large perturbations rather than the small or
-infinitessimal perturbations usually considered in the theory of
-dynamical systems.  Then, given a binary random variable $q$,
-we define its robustness as the following conditional proability:
+A quality which is of interest to us is robustness or structural stability \cite{Smale1967}, which requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters.
+We will quantify this as the probability that, if some property holds for a set of parameters, it will continue to hold if we make a random perturbation about those parameter values.
+To do so, we will introduce, in addition to the probability distribution $\mu$ described above, a family of probability distributions $\mu'$ such that $\mu'(x,y)$ encodes the conditional probability that a system with a parameter value $x$ will have its parameter values changed to $y$ under a perturbation.
+For our biological models, we will be interested in large perturbations rather than the small or infinitesimal perturbations usually considered in the theory of dynamical systems.
+Then, given a binary random variable $q$, we define its robustness as the following conditional probability:
 \begin{align}\label{eq:robustness}
@@
 
-A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components.
+Open systems are distinguished by having some of the variables be control variables whose values are given as specified functions of time rather than determined by the dynamics.
+Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
+The composition and decomposition of open systems has algebraic structure equivalent to certain a kind of operad \cite{Vagner2014}.
+Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
+However, for most purposes, such a decomposition would be uninteresting and what we instead want is a decomposition into relatively self-contained modules.
+One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.
+
 A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.

### 2014-09-25 paper.tex
edits

@@
 stay close to the fixed point.  Physically, this is important because,
-if a fixed point ${\vec x}_0$ is unstable, we have zero proability of
+if a fixed point ${\vec x}_0$ is unstable, we have zero probability of
 observing the solution ${\vec x}(t) = {\vec x}_0$.
@@
 for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
+This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.
 
@@
 at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
-conflicting populations), the stability of the fixed point will be
-determined by whether the limit cycle is attractive or repulsive.
+conflicting populations), the stability of the fixed point will determine whether the limit cycle is attractive or repulsive.
 
@@
 
-Suppose that our Jacobian $A$ is an $n \times n$ matrix with real
-coefficients and connectivity $\connectivity$ and denote the
-proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.  Then we may
-compute the robustness of this proposition using \ref{eq:robustness}
-once we have specified the proability densities $\mu$ and $\mu'$.  In
-general, these will depend upon the parameters of the non-linear
-system.  For the puprpose of the current investigation, we shall make
-the ansatz that these are such that the distribution of entries of $A$
-is (at least approximately) the uniform distribution
-$\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube,
-$H^\connectivity$, of edge length $r=2$, centered about the origin.
-To model the perturbations, we shall resample the elements $k_1,
-k_2, \ldots k_m$ of this matrix from the same uniform distribution.
-Under these assumptions, the expression for robustness becomes
-of \ref{eq:robustness}:
+Suppose that our Jacobian $A$ is an $n \times n$ matrix with real coefficients, connectivity $\connectivity$, and denote the proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.
+Then we may compute the robustness of this proposition using \ref{eq:robustness} once we have specified the probability densities $\mu$ and $\mu'$.
+In general, these will depend upon the parameters of the non-linear system.
+For the puprpose of the current investigation, we shall make the ansatz that these are such that the distribution of entries of $A$ is (at least approximately) the uniform distribution $\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin.
+To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution.
+Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
 \begin{equation}\label{eq:condprobgen}
@@
 The result of this operation corresponds to averaging our quantity over subsets of links:
-\begin{equation}
+\begin{equation}\label{eq:robavgoverlinks}
 R (\hbox{avg stab}, m) =
@@
 \end{equation}
-
-In order to evaluate these integrals, one needs to describe the stable region.
-This can be done in terms of conditions on the coefficients of the
-characteristic polynomials associated to the systems described by matrices $A$ above \cite{Gantmacher1959}.
+We are fundamentally interested in the relationship between \ref{eq:condprobgen} and \ref{eq:robavgoverlinks} and topological properties of systems that derive from their associated graphs.
+In order to investigate this further, we must characterize properties of system graphs in order to evaluate correlations of those properties with dynamical robustness.
 
@@
 
-Open systems are distinguished by having some of the variables be control variables whose values are given as specified functions of time rather than determined by the dynamics.
+Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems.
+One such kind of system composition and decomposition is given by the notion of open systems, which are distinguished by having some of the variables specified as control variables whose values are given as autonomous functions of time rather than determined by the dynamics via intrinsic interactions \cite{Vagner2014}.
 Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
-The composition and decomposition of open systems has algebraic structure equivalent to certain a kind of operad \cite{Vagner2014}.
 Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
-However, for most purposes, such a decomposition would be uninteresting and what we instead want is a decomposition into relatively self-contained modules.
+However, this decomposition is trivial since it is equivalent to the underlying system graph and what we instead want is an intermediate decomposition into relatively self-contained modules.
 One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.
@@
 The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
-
 The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
@@
 
+\todo{add figure showing all symmetries under $\hier$}
 This map $\hier$ is many-to-one and so there is a large class of
@@
 connected components relative to each other.  Leaving the components
-fixed, we may move links between nodes in a component or betwwen
+fixed, we may move links between nodes in a component or between
 components, or even add or delete links.  Some examples of these
 operations are illustrated in \ref{fig:scc}B and C.  As we shall see
-later, these operations leave invariant key dynamical quantities, so
+later, some of these operations leave invariant key dynamical quantities, so
 we may regard them as a symmetry groupoid of our system.
@@
 Using this approach the probability of stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.
-Similarly, the analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.
 
+The analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.
+This process is associated with some error relative to the exact integration described above.
 If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
@@
 This observation suggested that graph edit distance from \ref{fig:scc}A, hierarchy, might provide a better characterization of system robustness.
-\ref{fig:scc}D shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
+\ref{fig:combined}D shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
 There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\ref{fig:scc}F shows system robustness as a function of both hierarchy and connectivity.
+\ref{fig:combined}F shows system robustness as a function of both hierarchy and connectivity.
 The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:combined}E.
@@
 The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
-By combining our decomposition of the graph with our analysis of matrices, we will deduce a general result which relates robustness to hierarchy.
+By combining our decomposition of the graph with our analysis of the robustness of random ensembles of dynamical systems, we will deduce a general result which relates robustness to hierarchy.
 
@@
 terms of the robustnesses of its strongly connected components.  If
-the components $C_i$ of our graph $G$ has $v_i$ vertices connectivity
-$\connectivity_i$ and there are $\ell_{ij}$ links between components
+the components $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
+$\connectivity_i$, and there are $\ell_{ij}$ links between components
 $i$ and $j$, then there are a total of
@@
 And, the graph that has the latter property is that associated to the
-ordering, $G_{\mathrm{tot}}$, which aslo possess the highest degree of
+ordering, $G_{\mathrm{tot}}$, which also possess the highest degree of
 hierarchy. The fact that robustness is equivalent for any configuration

### 2014-09-25 paper.tex
draft discussion

@@
 
-Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems.
+Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems \ref{fig:examplesystemmodules}.
 One such kind of system composition and decomposition is given by the notion of open systems, which are distinguished by having some of the variables specified as control variables whose values are given as autonomous functions of time rather than determined by the dynamics via intrinsic interactions \cite{Vagner2014}.
@@
 \section{Discussion}
-\todo{write Discussion}
 The relationship between structure and function is fundamental to biology.
@@
 In all such networks, the ability of systems to persist over long periods of time, is dependent upon their dynamical stability.
-Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more stable and thus more likely to persist in the evolutionary process.
+Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist in the evolutionary process.
 This relationship transcends levels of the biological hierarchy in the sense that it is formulated mathematically and at a level of abstraction that allows its conclusions to apply independently to each of these levels.
@@
 The fact that such a relationship exists has the potential to unify our understanding of biological networks.
-Our development suggests two predictions having empirical impact capable of testing this theory.
+Our development suggests two predictions that could be used to test the theory described here.
 One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
@@
 The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
+In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.
+
+Future work may be able to deepen our understanding of the relationship between dynamical robustness and the strongly connected components of the underlying network topology.
+One approach is to attempt to compute robustness for systems of larger size.
+Using existing methods, this will require large-scale Monte Carlo simulations.
+The conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
+In addition to experimental investigation, a deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
 
@@
 
-\section{Tables}
+% \section{Tables}
 % \todo{Fill out notation table}
@@
 \input{tex/table2x2stab}
+\pagebreak
 \input{tex/tablenographs}
-
-% \newpage
+\pagebreak
+\input{tex/figuressupp}
 
 % \input{tex/tablewithgraphs}
-
 % \input{tex/extra}
-
-% \input{tex/figuressupp}
 

### 2014-09-25 paper.tex
update discussion

@@
 The fact that such a relationship exists has the potential to unify our understanding of biological networks.
+In order to pursue this, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the strongly connected components of the underlying network topology.
+One approach is to attempt to compute robustness for systems of larger size.
+Using existing methods, this would require large-scale Monte Carlo simulations.
+However, it may be possible to prove a bound on the robustness of individual strongly connected components, which could provide more definite knowledge and help to circumvent such expensive simulations.
+Nevertheless, the conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
+A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
+
 Our development suggests two predictions that could be used to test the theory described here.
@@
 The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
+Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
 In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.
-
-Future work may be able to deepen our understanding of the relationship between dynamical robustness and the strongly connected components of the underlying network topology.
-One approach is to attempt to compute robustness for systems of larger size.
-Using existing methods, this will require large-scale Monte Carlo simulations.
-The conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
-In addition to experimental investigation, a deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
 

### 2014-09-25 tex/figures.tex
add scc's to scc figure
add supplement section and reformat figures
move stable region figure to supplement
separate 3D from 2D plots

--- 2014-09-22 tex/figures.tex
@@
 \centering
-\noindent\includegraphics[width=0.6\columnwidth]{fig/examplesystemmodules.pdf}
+\noindent\includegraphics[width=0.4\columnwidth]{fig/examplesystemmodules.pdf}
 \caption{{\bf Example of the combination of open system modules to construct closed systems.} (A) Example of combining two open modules to construct a closed system of two components (B) Analogous example for combining three open modules to construct a closed system with three components}
@@
 
+\pagebreak
+
 \begin{figure}[!ht]
 \centering
-\noindent\includegraphics[width=0.7\columnwidth]{fig/scc.pdf}
+\noindent\includegraphics[width=0.4\columnwidth]{fig/scc2.pdf}
 \caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/region2and3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
-\label{fig:region2and3}
-\end{figure}
+\pagebreak
 
@@
 \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness} (A) (B) (C) (D) (E) (F) }
+\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus graph edit distance%, (E) robustness versus number of cycles and connectivity, and (F) robustness versus graph edit distance and connectivity
+ for three component systems.
+}
 \label{fig:combined}
@@
 
-% \begin{figure}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-% \caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-% \label{fig:connectcycle3D3x3}
-% \end{figure}
+\pagebreak
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.7\columnwidth]{fig/connectcycle3D3x3.pdf}
+\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
+\label{fig:connectcycle3D3x3}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.7\columnwidth]{fig/connectdist3D3x3.pdf}
+\caption{{\bf Plot of robustness versus graph edit distance and connectivity for three component systems.} }
+\label{fig:connectdist3D3x3}
+\end{figure}
 

### 2014-09-25 code/rds3x3.jl

@@ -394,10 +394,10 @@ function combinegnuplot()
     filelist = ["apstab3x3.p",
                 "stab3x3.p",
                 "cycle3x3.p",
-                "dist3x3.p",
-                "connectcycle3D3x3.p",
-                "connectdist3D3x3.p",]
-    horiz = 3
+                "dist3x3.p",]
+                # "connectcycle3D3x3.p",
+                # "connectdist3D3x3.p",]
+    horiz = 2
     vert = 2
     rv = (String)[]
     headstring = "set terminal svg size $(600*horiz),$(400*vert) dynamic enhanced fname 'HelveticaNeue'  fsize 20

### 2014-09-26 paper.tex
update references to separated figures

@@
 % \section{Cycle number and connectivity classify robust systems}
-The combination of connectivity and cycle number as shown in \ref{fig:combined}E provides a better classification of the dependence of robustness upon network topology.
+The combination of connectivity and cycle number as shown in \ref{fig:connectcycle3D3x3} provides a better classification of the dependence of robustness upon network topology.
 Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity.
@@
 There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\ref{fig:combined}F shows system robustness as a function of both hierarchy and connectivity.
-The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:combined}E.
+\ref{fig:connectdist3D3x3} shows system robustness as a function of both hierarchy and connectivity.
+The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:connectcycle3D3x3}.
 

### 2014-09-26 paper.tex
add figure showing symmetries under Hier

@@
 
-A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
+A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
 The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
@@
 
-\todo{add figure showing all symmetries under $\hier$}
 This map $\hier$ is many-to-one and so there is a large class of
@@
 For instance, we may interchange the positions of the strongly
-connected components relative to each other.  Leaving the components
-fixed, we may move links between nodes in a component or between
-components, or even add or delete links.  Some examples of these
-operations are illustrated in \ref{fig:scc}B and C.  As we shall see
-later, some of these operations leave invariant key dynamical quantities, so
-we may regard them as a symmetry groupoid of our system.
+connected components relative to each other \ref{fig:hiertransformations}A.
+Leaving the components fixed, we may move links between nodes in a component \ref{fig:hiertransformations}B or between components, or even add or delete links \ref{fig:hiertransformations}C.
+As we shall see later, some of these operations leave invariant key dynamical quantities, so we may regard them as a symmetry groupoid of our system.
 

### 2014-09-26 tex/figures.tex
add figure showing symmetries under Hier

@@
 \label{fig:examplesystemmodules}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.7\columnwidth]{fig/hiertransformations.pdf}
+\caption{{\bf Symmetries of the $\hier$ transformation between graphs and SCCs.} The transformation A represents an interchange of SCCs, B moving a link between nodes in a component and C adding a link.
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only A and B are symmetries of robustness.}
+\label{fig:hiertransformations}
 \end{figure}
@@
 \noindent\includegraphics[width=0.7\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
+\caption{{\bf Stability to perturbations versus number of cycles and connectivity for three component systems.} }
 \label{fig:connectcycle3D3x3}
@@
 \noindent\includegraphics[width=0.7\columnwidth]{fig/connectdist3D3x3.pdf}
-\caption{{\bf Plot of robustness versus graph edit distance and connectivity for three component systems.} }
+\caption{{\bf Robustness versus graph edit distance and connectivity for three component systems.} }
 \label{fig:connectdist3D3x3}

### 2014-09-26 paper.tex
add figure showing symmetries of robustness

@@
 
+As can be seen from \ref{eq:sccrobustness}, all that is required to fix robustenss is a set of SCC sizes and the number of links between them.
+This demonstrates that robustness is symmetric under the transformations shown in \ref{fig:hiertransformations}A and B.
+Since transformations of the type \ref{fig:hiertransformations}C change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}.
+Nevertheless, a considerable degree of symmetry remains.
+For example, \ref{fig:robustnesssymmetries} shows the symmetry groupoid of robustness for the case in which we fix connected component sizes $\{2,1,1\}$ with a total of $3$ links between them.
+
 Next, we derive an upper bound on the robustness of any graph

### 2014-09-26 tex/figures.tex
add figure showing symmetries of robustness

@@
 \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus graph edit distance%, (E) robustness versus number of cycles and connectivity, and (F) robustness versus graph edit distance and connectivity
+\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus hierarchy%, (E) robustness versus number of cycles and connectivity, and (F) robustness versus graph edit distance and connectivity
  for three component systems.
@@
 \noindent\includegraphics[width=0.7\columnwidth]{fig/connectdist3D3x3.pdf}
-\caption{{\bf Robustness versus graph edit distance and connectivity for three component systems.} }
+\caption{{\bf Robustness versus hierarchy and connectivity for three component systems.} }
 \label{fig:connectdist3D3x3}
+\end{figure}
+
+\pagebreak
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.9\columnwidth]{fig/robustnesssymmetries.pdf}
+\caption{{\bf Example symmetries of robustness.} In this example, the connected component sizes are fixed at $\{2,1,1\}$ with a total of $3$ links between them.
+Red arrows correspond to transformations like \ref{fig:hiertransformations}A where SCCs are swapped whereas purple arrows correspond to transformations like \ref{fig:hiertransformations}B where links are moved between nodes within a SCC.}
+\label{fig:robustnesssymmetries}
 \end{figure}

### 2014-09-26 tex/figuressupp.tex
move stable region figure to supplement
add figure showing symmetries of robustness

@@
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
+%!TEX root = ../paper.tex
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.7\columnwidth]{fig/region2and3.pdf}
+% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
+% \label{fig:region2and3}
+% \end{figure}
+
+% \pagebreak
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
+% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
 Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
+% \label{fig:region2x2}
+% \end{figure}
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
+% \pagebreak
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
+% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
 This is the plane $s_3=1$.}
-\label{fig:region3x3}
-\end{figure}
+% \label{fig:region3x3}
+% \end{figure}

### 2014-09-26 tex/figures.tex
add max/min robustness graphs to 3D plot

@@
 \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus hierarchy%, (E) robustness versus number of cycles and connectivity, and (F) robustness versus graph edit distance and connectivity
- for three component systems.
+\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus hierarchy for three component systems.
 }
@@
 \end{figure}
-
-% Figure translations
-% apstab3x3 4 -> A
-% stab3x3 6 -> B
-% cycle3x3 7 -> C
-% New -> D
-% connectcycle3D3x3 5 -> E
-% New -> F
-
-% \begin{figure}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.8\columnwidth]{fig/apstab3x3.pdf}
-% \caption{{\bf Plot of stability versus connectivity for three component systems.} }
-% \label{fig:apstab3x3}
-% \end{figure}
 
@@
 \centering
-\noindent\includegraphics[width=0.7\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Stability to perturbations versus number of cycles and connectivity for three component systems.} }
+\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
+\caption{{\bf Robustness versus number of cycles and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given number of cycles and connectivity.}
 \label{fig:connectcycle3D3x3}
@@
 \centering
-\noindent\includegraphics[width=0.7\columnwidth]{fig/connectdist3D3x3.pdf}
-\caption{{\bf Robustness versus hierarchy and connectivity for three component systems.} }
+\noindent\includegraphics[width=0.8\columnwidth]{fig/connectdist3D3x3.pdf}
+\caption{{\bf Robustness versus hierarchy and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given hierarchy and connectivity.}
 \label{fig:connectdist3D3x3}
@@
 \caption{{\bf Example symmetries of robustness.} In this example, the connected component sizes are fixed at $\{2,1,1\}$ with a total of $3$ links between them.
-Red arrows correspond to transformations like \ref{fig:hiertransformations}A where SCCs are swapped whereas purple arrows correspond to transformations like \ref{fig:hiertransformations}B where links are moved between nodes within a SCC.}
+Red arrows correspond to transformations like \ref{fig:hiertransformations}A where SCCs are swapped whereas purple arrows correspond to transformations like \ref{fig:hiertransformations}B where links are moved between nodes within a SCC.
+(A) shows all underlying graphs while (B) shows the number of nodes in each SCC and the number of links between the SCCs.}
 \label{fig:robustnesssymmetries}
 \end{figure}
-
-% \begin{figure}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-% \caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-% \label{fig:stab3x3}
-% \end{figure}
-
-% \begin{figure}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-% \caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-% \label{fig:cycle3x3}
-% \end{figure}

### 2014-09-26 paper.tex
define indicator function

@@
 For our biological models, we will be interested in large perturbations rather than the small or infinitesimal perturbations usually considered in the theory of dynamical systems.
-Then, given a binary random variable $q$, we define its robustness as the following conditional probability:
+Then, given a binary random variable $q$, we define its robustness as the following conditional probability where $\mathbf{1}_q(x)$ is the standard indicator function equal to $1$ when $x$ satisfies $q$ and $0$ otherwise:
 \begin{align}\label{eq:robustness}

### 2014-10-05 tex/abstract-cited.tex
add cited abstract

--- 2014-09-22 tex/abstract.tex
@@
 %!TEX root = ../paper.tex
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components.
-When the details of these interactions are unknown for systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints that can be considered to apply to all such systems.
-Here we analyze the stability and robustness of a large class of dynamical systems.
+The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+When the details of these interactions are unknown for systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints that can be considered to apply to all such systems \cite{Gardner1970,May1972,Cohen1984,May1972a,Radius2014}.
+Here we analyze the stability and robustness of a large class of dynamical systems \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
 In particular, we precisely determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.

### 2014-10-06 tex/abstract-cited.tex
move potential

@@
-%!TEX root = ../paper.tex
+%!TEX root = ../paper-abstract.tex
 The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
@@
 Here we analyze the stability and robustness of a large class of dynamical systems \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
-In particular, we precisely determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.
+In particular, we determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics.
+We do so precisely for systems with two and three interacting components.
 We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.

### 2014-10-06 paper.tex
update main text

@@
 \begin{align}\label{eq:robustness}
-  R (q) =
+  R (q,\mu') =
   \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
@@
 or not $\vec x$ is a fixed point of the system (i.e. $q$ is true if and only if
-$F_i(\vec{x}^0)=0$ for all $i$) and perform a probabilistic analysis
+$F_i(\vec{x})=0$ for all $i$) and perform a probabilistic analysis
 of the sort discussed above.
@@
 stay close to the fixed point.  Physically, this is important because,
-if a fixed point ${\vec x}_0$ is unstable, we have zero probability of
-observing the solution ${\vec x}(t) = {\vec x}_0$.
+if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
+observing the solution ${\vec x}(t) = {\vec x}^0$.
 
@@
 For the puprpose of the current investigation, we shall make the ansatz that these are such that the distribution of entries of $A$ is (at least approximately) the uniform distribution $\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin.
-To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution.
+To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution yielding
+$$
+\mu_{k_1,\ldots,k_m}(x,y) = \prod_{y \notin \{x_1, \ldots, x_m\} } \delta(x_j-y_j) \prod_{y \notin \{x_1,\cdots,x_m\}} \mathbf{1}_{[-1,1]} (y_j) \mathbf{1}_{[-1,1]} (x_j)
+$$
 Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
 \begin{equation}\label{eq:condprobgen}
- R (\mathrm{stab}, k_1, \ldots k_m) =
+ R (\mathrm{stab}, \mu'_{k_1, \ldots, k_m}) =
   \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
@@
 \begin{equation}\label{eq:robavgoverlinks}
-R (\hbox{avg stab}, m) =
+\langle R (\hbox{stab}, \mu'_{k_1, \ldots, k_m}) \rangle_m =
 \frac{1}{\binom{\connectivity}{m}}
 \sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
-R (\mathrm{stab}, k_1, \ldots k_m)
+R (\hbox{stab}, \mu'_{k_1, \ldots, k_m})
 \end{equation}
@@
 
+Our development suggests two predictions that could be used to test the theory described here.
+One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
+In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
+The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
+Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
+In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.
+
 The fact that such a relationship exists has the potential to unify our understanding of biological networks.
@@
 A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
-
-Our development suggests two predictions that could be used to test the theory described here.
-One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
-In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
-The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
-Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
-In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.
 

### 2014-10-07 tex/abstract-cited.tex
reorder authors and update cited abstract

@@
 When the details of these interactions are unknown for systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints that can be considered to apply to all such systems \cite{Gardner1970,May1972,Cohen1984,May1972a,Radius2014}.
-Here we analyze the stability and robustness of a large class of dynamical systems \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
+Here we analyze the stability and robustness of a large class of dynamical systems \cite{WADDINGTON1942a,Wagner1997,Rutherford1998,VanNimwegen1999,Siegal2002,Bergman2003,Ciliberti2007b,Ciliberti2007,Draghi2010,Wagner2013}.
 In particular, we determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics.

### 2014-10-07 paper.tex
reorder authors and update cited abstract
update notation with angle bracket averages

@@
 \\[.5cm]
+Cameron Smith$^{1}$,
+Raymond S. Puzio$^{1}$,
 Aviv Bergman$^{1,2,3,4}$,
-Raymond S. Puzio$^{1}$,
-Cameron Smith$^{1}$,
 \\[.5cm]
@@
 \begin{equation}\label{eq:robavgoverlinks}
-\langle R (\hbox{stab}, \mu'_{k_1, \ldots, k_m}) \rangle_m =
+\langle R (\hbox{stab}, \mu') \rangle_m =
 \frac{1}{\binom{\connectivity}{m}}
@@
 $$
-\begin{pmatrix}
+\begin{bmatrix}
 a & b \\
 d & c
-\end{pmatrix}
+\end{bmatrix}
 $$
 where $a + c < 0$ and $ac > bd$.
-For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},k)$ where $k=1 \ldots 4$.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $R(\mathrm{stab},1)$ and $R(\mathrm{stab},2)$.
-Suppose that we resample $b$ to compute $R(\mathrm{stab},2)$.
+For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},\mu'_k)$ where $k=1 \ldots 4$.
+By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $ R(\mathrm{stab},\mu'_1)$ and $R(\mathrm{stab},\mu'_2)$.
+Suppose that we resample $b$ to compute $R(\mathrm{stab},\mu'_2)$.
 % \begin{strip}
@@
 \begin{align*}
-P\left(\begin{pmatrix}
+P\left(\textrm{stab } \left( \begin{bmatrix}
 a & b \\
 d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
+\end{bmatrix} \right) \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
 \end{align*}
@@
 \begin{align*}
-P\left(\begin{pmatrix}
+P\left(\textrm{stab} \left( \begin{bmatrix}
 a & b \\
 d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
 a & b' \\
 d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
+\end{bmatrix} \right) \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
 \end{align*}
@@
 \begin{align*}
-P\left(\begin{pmatrix}
+P\left(\textrm{stab} \left( \begin{bmatrix}
 a & b \\
 d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
+\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
 a' & b \\
 d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
+\end{bmatrix} \right) \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
 \end{align*}
@@
 $C_i$, then the probability that the system will remain stable after
-resampling is $R(\hbox{avg stab} (C_i), 1)$.  If we resample a link
+resampling is $\langle R(\hbox{stab} (C_i), \mu') \rangle_1$.  If we resample a link
 between components $C_i$ and $C_j$, the system will remain stable with
@@
 \begin{align}\label{eq:sccrobustness}
-R(\hbox{avg stab} (G), 1) =
+\langle R(\hbox{stab} (G), \mu') \rangle_1 =
 \frac{\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i R(C_i)}
+      \sum\limits_{i=1}^n \connectivity_i \langle R(\textrm{stab}(C_i),\mu') \rangle_1}
      {\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
@@
 \[
-R(G) = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
+\langle R(\textrm{stab}(G),\mu') \rangle_1 = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
 \]
@@
 constructed from a given set of connected components.  Since
-$R(C_i) \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
-increase $R(G)$.  Given two connected components $C_i$ and $C_j$ with
+$\langle R(\textrm{stab}(C_i), \mu') \rangle_1 \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
+increase $\langle R(\textrm{stab}(G),\mu') \rangle_1$.  Given two connected components $C_i$ and $C_j$ with
 $v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$
@@
 \begin{align}
-R(G) \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i R(C_i)}
+\langle R(\textrm{stab}(G),\mu') \rangle_1 \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
+               \sum\limits_{i=1}^n \connectivity_i \langle R(C_i,\mu') \rangle_1}
               {\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
@@
 \frac{\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i R(\hbox{avg stab}(C_i),1)}
+                    \sum_{i=1}^n \connectivity_i \langle R(\hbox{stab}(C_i),\mu') \rangle_1}
      {\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
@@
 $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula,
-$R(\hbox{avg stab} (G_{\mathrm{max}}), 1) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+$\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
 
@@
 \begin{equation}
-R(\hbox{avg stab} (G), m) =
+\langle R(\hbox{stab} (G), \mu') \rangle_m =
 \frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
       {m \choose m_0, m_1, \ldots. m_n}
-      \left(m_0 + \sum\limits_{i=1}^n m_i R(\hbox{avg stab} (C_i), m_i) \right)}
+      \left(m_0 + \sum\limits_{i=1}^n m_i \langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \right)}
      {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
@@
 
-As before, since $R(\hbox{avg stab} (C_i), m_i) \le 1$, we may
-increase $R(\hbox{avg stab} (G), m)$ by increasing the maximum
+As before, since $\langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \le 1$, we may
+increase $\langle R(\hbox{stab} (G), \mu') \rangle_m$ by increasing the maximum
 possible value of $m_0$ while keeping the strongly connected
@@
 v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
-G_{tot}$.  Hence, we conclude that $R(\hbox{avg stab} (G), m) \le
-R(\hbox{avg stab} (G_{\mathrm{max}}), m)$.
+G_{tot}$.  Hence, we conclude that $\langle R(\hbox{stab} (G), \mu') \rangle_m \le
+\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_m$.
 

### 2014-10-10 tex/abstract-cited.tex
update cited abstract

--- 2014-10-06 tex/abstract-cited.tex
@@
 %!TEX root = ../paper-abstract.tex
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-When the details of these interactions are unknown for systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints that can be considered to apply to all such systems \cite{Gardner1970,May1972,Cohen1984,May1972a,Radius2014}.
-Here we analyze the stability and robustness of a large class of dynamical systems \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
-In particular, we determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics.
-We do so precisely for systems with two and three interacting components.
-We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.
-We also demonstrate that permutation of strongly connected components is a fundamental symmetry of system robustness.
-This results in the classification of the dynamical robustness of biological networks based upon a purely topological property.
+The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components~\cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+When the details of these interactions are unknown for biological systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by evolutionary constraints that may apply to all such systems~\cite{Gardner1970,May1972,Cohen1984,May1972a,Radius2014}.
+One such constraint is that of dynamical robustness, the probability of a stable system to remain stable under perturbation~\cite{WADDINGTON1942a,Wagner1997,Rutherford1998,VanNimwegen1999,Siegal2002,Bergman2003,Ciliberti2007b,Ciliberti2007,Draghi2010,Wagner2013}.
+Despite previous investigation, the relationship between dynamical robustness, an important functional characteristic of many biological systems, and network structure is poorly understood.
+Here we analyze the stability and robustness of a large class of dynamical systems and demonstrate that the most hierarchical network structures are the most robust.
+In particular, we determine the probability distribution of robustness over system connectivity and show that robustness is maximized by maximizing the number of links between strongly connected components of the graph representing the underlying system connectivity.
+We also demonstrate that permutation of strongly connected components is a fundamental symmetry of system robustness, which applies to networks of any number of components.
+The classification of dynamical robustness based upon a purely topological property provides a fundamental organizing principle that can be used in the context of experimental validation to select among models that break or preserve network hierarchy.
+This result contributes to an explanation of the observation of hierarchical modularity across levels of the biological hierarchy~\cite{Zhao2006,Ravasz2002,Bhardwaj2010,Colm,Corominas-Murtra2013}.

### 2014-10-13 paper.tex
move methods to supplement; factor out intro, results and discussion; combine figures

--- 2014-09-26 paper.tex
@@
 {\begin{quote} \bf
-\input{tex/abstract}
+\input{tex/abstract-cited}
 \end{quote}}
@@
 \section{Introduction}
+\input{tex/introduction}
 
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to structural perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two or three interacting components.
-We then proceed to determine the probability of \emph{robustness} or \emph{structural stability} which is the distribution of stability to perturbation in the system structure for systems which are determined to be stable \cite{Smale1967}.
-This quantity is of interest at all scales of the biological hierarchy, and has been previously addressed in the study of the robustness of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
-We find that stability to structural perturbations of this class of dynamical systems is correlated with connectivity, number of cycles, and the number of links between strongly connected components of the underlying interdependency graph.
-The latter derives from the fact that the permutation of strongly connected components is a symmetry of system robustness.
-This work clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Dynamical systems on biological networks}
-In the construction of a class of potential models for biological systems at any level of the biological hierarchy from metabolic to ecosystem-level networks, it is common to first attempt to define a collection of observable phenomena of interest and determine a domain (such as binary numbers, integers, or real numbers) in which each observable can be quantified.
-Next, it is necessary to establish the interdependencies among system components.
-The specific manner in which the components depend upon one another must be clarified, which is often done by determining a particular system of mathematical functions that represents a hypothesis about how the quantified observables evolve in time.
-To the degree to which there is uncertainty about the interactions among system components, parameters are introduced to broaden the class of models under consideration.
-Finally, whatever model class remains may be compared to empirical observations to determine how capable the model is of representing the phenomena of interest.
-
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
-For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
-If the system under consideration has the graph given in \ref{fig:examplesystemmodules}A
-% \begin{center}
-% \input{tex/examplehamdyngraph}
-% \end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-having connectivity $\connectivity$ equal to the number of edges of the graph, in this case $\connectivity = 3$.
-For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}.
-   \right.
-\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-
-In the more general case of a system with $n$ components we would have an $n$-dimensional vector of observables
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-
-In order to accommodate the possibility of uncertainty in our
-modeling, we will generalize our notion of a system on a network to
-that of a system with random parameters.  Again, this will involve
-three steps. First, we provide a measure space $S$
-which represents the values over which our parameters can vary. Next,
-instead of a single vector field, we consider a family of vector
-fields parameterized by this space.  As before, for
-each point $p \in S$, the vector field $V(p)$ must be consistent with
-the system graph. Finally, we select a probability measure, $\mu$, which represents our understanding of which values of the parameters
-are likely. In accord with Bayesian statistics, we might revise this
-distribution as data comes in or use it to estimate parameters and error
-bars from experimental results.
-
-Having done this, we are now in a position to do a probabilistic analysis
-of our system and its properties.  Given some quantity $q$ characterizing
-our system, this quantity becomes a random variable which may be discrete or
-continuous depending on the quantity under consideration.  For example, if
-the quantity is the time it takes for a particle to travel between two points,
-we have a real-valued random variable; if the quantity is the number of fixed
-points, we have an integer-valued random variable; and if the quantity is
-whether or not the system posesses a limit cycle, we have a binary random
-variable.
-
-A quality which is of interest to us is robustness or structural stability \cite{Smale1967}, which requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters.
-We will quantify this as the probability that, if some property holds for a set of parameters, it will continue to hold if we make a random perturbation about those parameter values.
-To do so, we will introduce, in addition to the probability distribution $\mu$ described above, a family of probability distributions $\mu'$ such that $\mu'(x,y)$ encodes the conditional probability that a system with a parameter value $x$ will have its parameter values changed to $y$ under a perturbation.
-For our biological models, we will be interested in large perturbations rather than the small or infinitesimal perturbations usually considered in the theory of dynamical systems.
-Then, given a binary random variable $q$, we define its robustness as the following conditional probability where $\mathbf{1}_q(x)$ is the standard indicator function equal to $1$ when $x$ satisfies $q$ and $0$ otherwise:
-\begin{align}\label{eq:robustness}
-  R (q) =
-  \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
-  {\int_S d\mu(x) \mathbf{1}_q(x)}
-\end{align}
-
-\section{Stability analysis of biological networks}
-The class of dynamical quantities on which we shall focus in this
-investigation involve stability of equilibria.  Suppose that $\vec x$
-is a point on our phase space which depends upon the parameters.  Then
-we may take $q$ to be a binary random variable which describes whether
-or not $\vec x$ is a fixed point of the system (i.e. $q$ is true if and only if
-$F_i(\vec{x}^0)=0$ for all $i$) and perform a probabilistic analysis
-of the sort discussed above.
-
-Furthermore, if it turns out that $\vec x$ is indeed a fixed point, we
-may proceed to ask whether it is a dynamically stable fixed point.
-Intuitively, dynamic stability means that, if one chooses the initial
-conditions sufficiently close to the fixeed point, the solution will
-stay close to the fixed point.  Physically, this is important because,
-if a fixed point ${\vec x}_0$ is unstable, we have zero probability of
-observing the solution ${\vec x}(t) = {\vec x}_0$.
-
-To determine stability, we linearize the equations of motion about the
-fixed point $\vec{x}^0$:
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.
-
-Even though the determination of stability only requires examining
-linearized equations of motion, it is worth noting that knowing the
-stability of a fixed point can yield information about the non-linear
-dynamics of a system.  For instance, in two dimensions,
-Poincare-Bendixson theory implies that every limit cycle must encircle
-at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
-conflicting populations), the stability of the fixed point will determine whether the limit cycle is attractive or repulsive.
-
-%% To determine whether a randomly selected dynamical system evaluated at
-%% a random critical point, given by a matrix such as $A$, is stable, it
-%% is sufficient to check the above condition. Integrating over the
-%% region of the parameter space defining $A$ enables the determination
-%% of the probability of stability to perturbations in state $\vec{x}$
-%% over an ensemble of dynamical systems.
-
-% Another form of stability, referred to as structural
-% stability \cite{Smale1967}, requires the determination of whether or
-% not a given dynamical system that is determined to be stable remains
-% stable under a perturbation to one of its defining parameters. This
-% can be formalized as the conditional probability distribution
-% $$
-% P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-% $$
-% where $A'$ represents a matrix derived from some perturbation
-% applied to the matrix $A$ given above. In the simple case where
-% $A \in \mathbb{R}^{n \times n}$ the perturbation which involves
-% sampling uniformly over the $a_{ij}$ defining $A$ the value of a
-% particular $a_{ij}$ from a given probability distribution.
-
-Suppose that our Jacobian $A$ is an $n \times n$ matrix with real coefficients, connectivity $\connectivity$, and denote the proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.
-Then we may compute the robustness of this proposition using \ref{eq:robustness} once we have specified the probability densities $\mu$ and $\mu'$.
-In general, these will depend upon the parameters of the non-linear system.
-For the puprpose of the current investigation, we shall make the ansatz that these are such that the distribution of entries of $A$ is (at least approximately) the uniform distribution $\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin.
-To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution.
-Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
-\begin{equation}\label{eq:condprobgen}
- R (\mathrm{stab}, k_1, \ldots k_m) =
-  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
-    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
-  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)},
-\end{equation}
-where we made the abbreviations
-\begin{align*}
-\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
-\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
-\end{align*}
-i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.
-
-While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.
-The result of this operation corresponds to averaging our quantity over subsets of links:
-\begin{equation}\label{eq:robavgoverlinks}
-R (\hbox{avg stab}, m) =
-\frac{1}{\binom{\connectivity}{m}}
-\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
-R (\mathrm{stab}, k_1, \ldots k_m)
-\end{equation}
-We are fundamentally interested in the relationship between \ref{eq:condprobgen} and \ref{eq:robavgoverlinks} and topological properties of systems that derive from their associated graphs.
-In order to investigate this further, we must characterize properties of system graphs in order to evaluate correlations of those properties with dynamical robustness.
-
-\section{Network hierarchy and strongly connected components}
-
-Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems \ref{fig:examplesystemmodules}.
-One such kind of system composition and decomposition is given by the notion of open systems, which are distinguished by having some of the variables specified as control variables whose values are given as autonomous functions of time rather than determined by the dynamics via intrinsic interactions \cite{Vagner2014}.
-Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
-Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
-However, this decomposition is trivial since it is equivalent to the underlying system graph and what we instead want is an intermediate decomposition into relatively self-contained modules.
-One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.
-
-A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
-The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
-Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
-Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
-Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
-This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
-
-Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$.
-Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
-There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
-
-One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
-a strongly connected graph $C_n$.
-To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
-The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
-
-This map $\hier$ is many-to-one and so there is a large class of
-operations which leaves $\hier(G)$ invariant for a given graph $G$.
-For instance, we may interchange the positions of the strongly
-connected components relative to each other \ref{fig:hiertransformations}A.
-Leaving the components fixed, we may move links between nodes in a component \ref{fig:hiertransformations}B or between components, or even add or delete links \ref{fig:hiertransformations}C.
-As we shall see later, some of these operations leave invariant key dynamical quantities, so we may regard them as a symmetry groupoid of our system.
-
-The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
-The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \ref{fig:scc}A.
-This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
-This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical.
-In this work we take this distance to be the definition of hierarchy.
-
-\section{Stability and robustness analysis of particular system ensembles}
-For systems having two components, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}.
-For those having three components, we can estimate these same quantities using Monte Carlo simulations.
-Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis.
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant. Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},k)$ where $k=1 \ldots 4$.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $R(\mathrm{stab},1)$ and $R(\mathrm{stab},2)$.
-Suppose that we resample $b$ to compute $R(\mathrm{stab},2)$.
-% \begin{strip}
-% \begin{align}\label{eq:condprob}
-% P\left(\begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)
-% & = \frac{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}.
-% \end{align}
-The denominator of \ref{eq:condprobgen} in this case is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprobgen}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-Using this approach the probability of stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.
-
-The analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.
-This process is associated with some error relative to the exact integration described above.
-If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
-We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}.
-The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
-$$
-\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
-$$
-where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
-$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
-which corresponds in this case to the maximum likelihood estimate
-$$
-\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
-$$
-This estimate is characterized by the variance of the posterior Beta distribution
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
-$$
-Since for the chosen prior $a=b=1 \ll N$ this simplifies to
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
-$$
-yielding the error estimate given by the associated standard deviation.
-In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
-
-It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
-For small system sizes such as the two and three component systems, studied here the situation is not so clear cut.
-For two component systems, system stability is constant across the entire range of connectivities.
-For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:combined}A.
-
-The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:combined}B.
-If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \ref{fig:combined}B.
-% \section{Cycle number is inversely correlated with robustness}
-\ref{fig:combined}C shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-There appears to be a weak negative correlation between robustness and the number of simple cycles.
-
-% \section{Cycle number and connectivity classify robust systems}
-The combination of connectivity and cycle number as shown in \ref{fig:connectcycle3D3x3} provides a better classification of the dependence of robustness upon network topology.
-Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity.
-The network with the highest robustness for three component systems is that of \ref{fig:scc}A.
-This network is the most hierarchical of all three component systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \ref{fig:scc}A has a block triangular structure.
-
-This observation suggested that graph edit distance from \ref{fig:scc}A, hierarchy, might provide a better characterization of system robustness.
-\ref{fig:combined}D shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
-There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\ref{fig:connectdist3D3x3} shows system robustness as a function of both hierarchy and connectivity.
-The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:connectcycle3D3x3}.
-
-\section{Component decomposition, hierarchy, and the symmetries of robustness}
-
-The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
-By combining our decomposition of the graph with our analysis of the robustness of random ensembles of dynamical systems, we will deduce a general result which relates robustness to hierarchy.
-
-Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomials factors as the product of the charactericstic polynomials of its diagonal blocks.
-Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.
-Note that this condition does not depend upon the entries off the diagonal (which correspond to links between strongly connected components) and does not depend upon what order the components appear.
-
-Using these observations, we may express the robustness of a graph in
-terms of the robustnesses of its strongly connected components.  If
-the components $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
-$\connectivity_i$, and there are $\ell_{ij}$ links between components
-$i$ and $j$, then there are a total of
-$\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i$
-links in $G$.  If we resample a randomly chosen link in a component
-$C_i$, then the probability that the system will remain stable after
-resampling is $R(\hbox{avg stab} (C_i), 1)$.  If we resample a link
-between components $C_i$ and $C_j$, the system will remain stable with
-probability 1.  Hence, the probability that the system will remain
-stable upon resampling a random link is the weighted average of these
-probabilities:
-\begin{align}\label{eq:sccrobustness}
-R(\hbox{avg stab} (G), 1) =
-\frac{\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i R(C_i)}
-     {\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i}.
-\end{align}
-For instance, if our graph is the one in \ref{fig:scc}B or C, then we have two connected components, one with two nodes, and one with one node.
-From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and and robustness $0.62$.
-The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
-Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is
-\[
-R(G) = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
-\]
-which agrees with the value computed in \ref{tab:structstabmat} up to
-sampling error.
-
-As can be seen from \ref{eq:sccrobustness}, all that is required to fix robustenss is a set of SCC sizes and the number of links between them.
-This demonstrates that robustness is symmetric under the transformations shown in \ref{fig:hiertransformations}A and B.
-Since transformations of the type \ref{fig:hiertransformations}C change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}.
-Nevertheless, a considerable degree of symmetry remains.
-For example, \ref{fig:robustnesssymmetries} shows the symmetry groupoid of robustness for the case in which we fix connected component sizes $\{2,1,1\}$ with a total of $3$ links between them.
-
-Next, we derive an upper bound on the robustness of any graph
-constructed from a given set of connected components.  Since
-$R(C_i) \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
-increase $R(G)$.  Given two connected components $C_i$ and $C_j$ with
-$v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$
-links going from $C_i$ to $C_j$.  Hence, $\ell_{ij} \le v_i v_j$, so,
-we conclude that
-\begin{align}
-R(G) \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i R(C_i)}
-              {\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i},
-\end{align}
-
-Since every acyclic digraph can be embedded into a totally ordered
-set, we may assume without loss of generality that our components have
-been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i <
-j$.  Hence, after substituting into our expression and simplifying
-using the identity
-$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
-v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n}
-v_i^2,$$
-we arrive at the following upper bound on robustness in terms
-of the connected components:
-\begin{align} \label{eq:sccmaxrobustness}
-R_{\mathrm{max}}(C_1, \ldots C_n) =
-\frac{\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i R(\hbox{avg stab}(C_i),1)}
-     {\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i}.
-\end{align}
-Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
-is the graph on $n$ nodes with a link from node $i$ to node $j$
-whenever $i < j$.  Then, by the construction described in the previous
-section on network hierachy, we have a graph $G_{\mathrm{max}}$ such
-that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and
-$\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula,
-$R(\hbox{avg stab} (G_{\mathrm{max}}), 1) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
-
-This argument also works when we resample more than one entry,
-although the notation becomes more complicated.  Suppose that we
-resample $m$ nodes.  Define
-\begin{equation*}
-M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
-m = \sum_{i=0}^n m_i \quad\&\quad
-m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
-\forall i \in \{1, \ldots, n\} \, m_i < \connectivity_i \right\}.
-\end{equation*}
-Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
-m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
-$m_0$ links between strongly connected components.  Hence, our
-weighted average becomes
-\begin{equation}
-R(\hbox{avg stab} (G), m) =
-\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
-      {m \choose m_0, m_1, \ldots. m_n}
-      \left(m_0 + \sum\limits_{i=1}^n m_i R(\hbox{avg stab} (C_i), m_i) \right)}
-     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
-      {m \choose m_0, m_1, \ldots. m_n} m}
-\end{equation}
-
-As before, since $R(\hbox{avg stab} (C_i), m_i) \le 1$, we may
-increase $R(\hbox{avg stab} (G), m)$ by increasing the maximum
-possible value of $m_0$ while keeping the strongly connected
-components the same.  Again, if we fix $\hier(G)$, the maximum
-possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
-whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
-v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
-G_{tot}$.  Hence, we conclude that $R(\hbox{avg stab} (G), m) \le
-R(\hbox{avg stab} (G_{\mathrm{max}}), m)$.
-
-The preceding argument demonstrates the manner in which robustness is
-maximized by maximizing the number of edges \emph{between} the strongly
-connected components of the graph underlying system interdependencies.
-And, the graph that has the latter property is that associated to the
-ordering, $G_{\mathrm{tot}}$, which also possess the highest degree of
-hierarchy. The fact that robustness is equivalent for any configuration
-of strongly connected components with an equivalent number of links
-between them demonstrates that it is invariant to permutations of
-strongly connected components. This result does not depend at all upon
-any particular details such as the probability distribution from which
-the component interaction strengths are sampled because the arguments
-presented here are purely topological in nature. The result that
-dynamical robustness is correlated with system hierarchy therefore
-applies to an even broader class of dynamical systems than the particular
-random ensembles primarily studied here.
-
-% \section{Strongly connected components, robustness, and hierarchy}
-
-% \section{Hierarchical systems are the most robust}
-% \ref{fig:scc} shows systems with varying levels of hierarchy and modularity.
+\section{Results}
+\input{tex/results}
 
 \section{Discussion}
-The relationship between structure and function is fundamental to biology.
-Structure-function relationships exist at every level of the biological hierarchy.
-In all such networks, the ability of systems to persist over long periods of time, is dependent upon their dynamical stability.
-Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist in the evolutionary process.
-This relationship transcends levels of the biological hierarchy in the sense that it is formulated mathematically and at a level of abstraction that allows its conclusions to apply independently to each of these levels.
-
-The fact that such a relationship exists has the potential to unify our understanding of biological networks.
-In order to pursue this, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the strongly connected components of the underlying network topology.
-One approach is to attempt to compute robustness for systems of larger size.
-Using existing methods, this would require large-scale Monte Carlo simulations.
-However, it may be possible to prove a bound on the robustness of individual strongly connected components, which could provide more definite knowledge and help to circumvent such expensive simulations.
-Nevertheless, the conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
-A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
-
-Our development suggests two predictions that could be used to test the theory described here.
-One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
-In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
-The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
-Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
-In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.
+\input{tex/discussion}
 
@@
 
+\input{tex/supplement}
+\pagebreak
+
+\input{tex/figuressupp}
+\pagebreak
+
 % \section{Tables}
@@
 \input{tex/tablenographs}
-\pagebreak
-\input{tex/figuressupp}
 

### 2014-10-13 tex/discussion.tex
move methods to supplement; factor out intro, results and discussion; combine figures

%!TEX root = ../paper.tex
The relationship between structure and function is fundamental to biology. Structure-function relationships exist at every level of the biological hierarchy. In all such networks, the ability of systems to persist over long periods of time, is dependent upon their dynamical stability. Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist in the evolutionary process. This relationship transcends levels of the biological hierarchy in the sense that it is formulated mathematically and at a level of abstraction that allows its conclusions to apply independently to each of these levels.

Our development suggests two predictions that could be used to test the theory described here. One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies. In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}. The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness. Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.  In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.

The fact that such a relationship exists has the potential to unify our understanding of biological networks. In order to pursue this, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the strongly connected components of the underlying network topology. One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However, it may be possible to prove a bound on the robustness of individual strongly connected components, which could provide more definite knowledge and help to circumvent such expensive simulations. Nevertheless, the conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space. A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.

### 2014-10-13 tex/figures.tex
combine figures 1-3
move methods to supplement; factor out intro, results and discussion; combine figures

@@
 %!TEX root = ../paper.tex
+
 \begin{figure}[!ht]
 \centering
-\noindent\includegraphics[width=0.4\columnwidth]{fig/examplesystemmodules.pdf}
-\caption{{\bf Example of the combination of open system modules to construct closed systems.} (A) Example of combining two open modules to construct a closed system of two components (B) Analogous example for combining three open modules to construct a closed system with three components}
-\label{fig:examplesystemmodules}
+\noindent\includegraphics[width=0.9\columnwidth]{fig/biomodelexamples.pdf}
+\caption{{\bf Dynamical models in systems biology.} (A) (B) (C)}
+\label{fig:biomodelexamples}
 \end{figure}
@@
 \centering
-\noindent\includegraphics[width=0.7\columnwidth]{fig/hiertransformations.pdf}
-\caption{{\bf Symmetries of the $\hier$ transformation between graphs and SCCs.} The transformation A represents an interchange of SCCs, B moving a link between nodes in a component and C adding a link.
-All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only A and B are symmetries of robustness.}
-\label{fig:hiertransformations}
+\noindent\includegraphics[width=0.9\columnwidth]{fig/modsccsym.pdf}
+\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
+(B) strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
+The most hierarchical system in the top panel has the highest possible number of connected components, three, whereas the system in the bottom panel containing a single cycle and therefore posessing no hierarchy contains only one connected component.
+The two panels in the middle represent examples of hierarchical modular systems that posess both modularity and hierarchy.
+(C) Symmetries of the $\hier$ transformation between graphs and SCCs.
+The transformation a represents an interchange of SCCs, b moving a link between nodes in a component and c adding a link.
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only a and b are symmetries of robustness.}
+\label{fig:modsccsym}
 \end{figure}
@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.4\columnwidth]{fig/scc2.pdf}
-\caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.4\columnwidth]{fig/examplesystemmodules.pdf}
+% \caption{{\bf Example of the combination of open system modules to construct closed systems.} (A) Example of combining two open modules to construct a closed system of two components (B) Analogous example for combining three open modules to construct a closed system with three components}
+% \label{fig:examplesystemmodules}
+% \end{figure}
+
+% \pagebreak
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.7\columnwidth]{fig/hiertransformations.pdf}
+% \caption{{\bf Symmetries of the $\hier$ transformation between graphs and SCCs.} The transformation A represents an interchange of SCCs, B moving a link between nodes in a component and C adding a link.
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only A and B are symmetries of robustness.}
+% \label{fig:hiertransformations}
+% \end{figure}
+
+% \pagebreak
+
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.4\columnwidth]{fig/scc2.pdf}
+% \caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
 Note that the most hierarchical system in (A) has the highest possible number of connected components, three, whereas the system containing a single cycle and therefore posessing no hierarchy contains only one connected component.
 Systems (B) and (C) represent examples of hierarchical modular systems that posess both modularity and hierarchy.}
-\label{fig:scc}
-\end{figure}
+% \label{fig:scc}
+% \end{figure}
 
-\pagebreak
+% \pagebreak
 
@@
 \caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus hierarchy for three component systems.
-}
+}1,500 words long
 \label{fig:combined}
@@
 \end{figure}
-
-\pagebreak
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.9\columnwidth]{fig/robustnesssymmetries.pdf}
-\caption{{\bf Example symmetries of robustness.} In this example, the connected component sizes are fixed at $\{2,1,1\}$ with a total of $3$ links between them.
-Red arrows correspond to transformations like \ref{fig:hiertransformations}A where SCCs are swapped whereas purple arrows correspond to transformations like \ref{fig:hiertransformations}B where links are moved between nodes within a SCC.
-(A) shows all underlying graphs while (B) shows the number of nodes in each SCC and the number of links between the SCCs.}
-\label{fig:robustnesssymmetries}
-\end{figure}

### 2014-10-13 tex/figuressupp.tex
move methods to supplement; factor out intro, results and discussion; combine figures

@@
 %!TEX root = ../paper.tex
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.9\columnwidth]{fig/robustnesssymmetries.pdf}
+\caption{{\bf Example symmetries of robustness.} In this example, the connected component sizes are fixed at $\{2,1,1\}$ with a total of $3$ links between them.
+Red arrows correspond to transformations like \reffighiertransformations a where SCCs are swapped whereas purple arrows correspond to transformations like \reffighiertransformations b where links are moved between nodes within a SCC.
+(A) shows all underlying graphs while (B) shows the number of nodes in each SCC and the number of links between the SCCs.}
+\label{fig:robustnesssymmetries}
+\end{figure}
+
 % \begin{figure}[!ht]

### 2014-10-13 tex/introduction.tex
move methods to supplement; factor out intro, results and discussion; combine figures

--- 2014-09-12 journal.tex
@@
-%------------------------------------------------------------------------------
-% Beginning of journal.tex
-%------------------------------------------------------------------------------
-%
-% AMS-LaTeX version 2 sample file for journals, based on amsart.cls.
-%
-%        ***     DO NOT USE THIS FILE AS A STARTER.      ***
-%        ***  USE THE JOURNAL-SPECIFIC *.TEMPLATE FILE.  ***
-%
-% Replace amsart by the documentclass for the target journal, e.g., tran-l.
-%
-\documentclass{amsart}
-
-%     If your article includes graphics, uncomment this command.
-\usepackage{graphicx}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\theoremstyle{definition}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-\newtheorem{xca}[theorem]{Exercise}
-
-\theoremstyle{remark}
-\newtheorem{remark}[theorem]{Remark}
-
-\numberwithin{equation}{section}
-
-%    Absolute value notation
-\newcommand{\abs}[1]{\lvert#1\rvert}
-
-%    Blank box placeholder for figures (to avoid requiring any
-%    particular graphics capabilities for printing this document).
-\newcommand{\blankbox}[2]{%
-  \parbox{\columnwidth}{\centering
-%    Set fboxsep to 0 so that the actual size of the box will match the
-%    given measurements more closely.
-    \setlength{\fboxsep}{0pt}%
-    \fbox{\raisebox{0pt}[#2]{\hspace{#1}}}%
-  }%
-}
-
-% user added
-\usepackage{longtable}
-\usepackage{dot2texi}
-\usepackage{tikz}
-\usetikzlibrary{automata,shapes,arrows}
-
-\def\tr{\mathrm{tr}}
-\def\Path{\mathrm{Path}}
-\def\hier{\mathrm{Hier}}
-\def\Vertex{\mathrm{Vertex}}
-\def\adj{\mathrm{adj}}
-
-\begin{document}
-
-\title{Ensemble Stability of Biological Networks}
-
-%    Information for first author
-\author{Aviv Bergman}
-%    Address of record for the research reported here
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{aviv@einstein.yu.edu}
-%    \thanks will become a 1st page footnote.
-% \thanks{The first author was supported in part by NSF Grant \#000000.}
-
-%    Information for second author
-\author{Raymond S. Puzio}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{rsp@novres.org}
-% \thanks{Support information for the second author.}
-
-%    Information for third author
-\author{Cameron Smith}
-\address{Department of Systems and Computational Biology, Albert Einstein College of Medicine, Bronx, New York 10461}
-%    Current address
-\curraddr{}
-\email{cameron.smith@med.einstein.yu.edu}
-% \thanks{Support information for the second author.}
-
-%    General info
-% \subjclass[2000]{Primary 54C40, 14E20; Secondary 46E25, 20C20}
-
-% \date{January 1, 2001 and, in revised form, June 22, 2001.}
-
-% \dedicatory{This paper is dedicated to our advisors.}
-
-% \keywords{Differential geometry, algebraic geometry}
-
-\begin{abstract}
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having many parameters and thus many degrees of freedom.
-When the details of a system are unknown, one effective approach is to study the dynamical properties of a collection of models determined by constraints applying to all such systems.
-Here we analyze the stability of a large class of dynamical systems to perturbations in the underlying structure of the system: a property referred to as \emph{structural stability}.
-In particular, we precisely determine the probability distribution over system connectivity, a parameter which has significant implications from the study of gene-regulatory networks to ecosystem dynamics, of structural stability for systems with two and three interacting components.
-We show for these cases that structural stability has a non-monotonic relationship with system connectivity.
-We also demonstrate that networks with a hierarchical structure are more likely to be stable to perturbations than those with a more entangled heterarchical structure.
-These results support future work attempting to characterize the scaling relationship between structural stability and system connectivity.
-The latter investigation is necessary to evaluate several conjectured but ultimately untested hypotheses about biological networks.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to ``fit the data'' \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014}.
+%!TEX root = ../paper.tex
+The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
+Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
 In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
@@
 However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
-
-Here we build on work aimed at investigating the stability of randomly connected dynamical systems to state-based perturbations.
-We precisely compute the probability of stability as a distribution over system connectivity for all systems containing two interacting components.
-We then proceed to determine the probability of structural stability over the same domain.
-We find that while stability to perturbations in the state of this class of dynamical systems is independent of system connectivity, perturbations to the magnitude of connections is correlated with connectivity in a non-linear manner.
-This work reestablishes and clarifies a thread of research that is extremely important for the continuing development of systems biology.
-
-\section{Stability analysis of biological networks}
-
-Consider an n-dimensional vector
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-We can associate a directed graph $G_X$ that describes the manner in which each of the variables depends upon one another given by an adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}
-   \right.
-\end{displaymath}
-For example, given the system $X = \{\{x_1,x_2\},\{F_1,F_2\}\}$
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-we have the graph
-\begin{center}
-\begin{tikzpicture}[
-every state/.style={draw=red!50,very thick,fill=red!20}]
-\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
-digraph G {
-d2ttikzedgelabels = true;
-node [style="state"];
-edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-x_1 -> x_2;
-x_2 -> x_1;
-x_2 -> x_2 [topath="loop above"];
-}
-\end{dot2tex}
-\end{tikzpicture}
-\end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-
-Linearization of the $F_i$ about any equilibrium solution $\vec{x}^0$ such that $F_i(\vec{x}^0)=0$ leads to the system
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to equation \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-
-To determine whether a randomly selected dynamical system evaluated at a random critical point, given by a matrix such as $A$, is stable, it is sufficient to check the above condition.
-Integrating over the region of the parameter space defining $A$ enables the determination of the probability of stability to perturbations in state $\vec{x}$ over an ensemble of dynamical systems.
-
-Another form of stability, referred to as structural stability \cite{Smale1967}, requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one of its defining parameters.
-This can be formalized as the conditional probability distribution
-$$
-P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-$$
-where $A'$ represents a matrix derived from some perturbation applied to the matrix $A$ given above.
-In the simple case where $A \in \mathbb{R}^{n \times n}$ the perturbation which involves sampling uniformly over the $a_{ij}$ defining $A$ the value of a particular $a_{ij}$ from a given probability distribution.
-
-\section{System graph and decomposition into strongly connected components}
-We will show how the stability of the dynamical system associated to Jacobian $A$ is characterized by properties of the associated graph $G_A$.
-The characteristic polynomial $\chi(A)$ associated to $A$ can be factored according to the strongly connected components of the graph $G$.
-$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$ where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of the matrix $M$ with $z$ a complex variable.
-This formula is readily verified when $M$ is diagonal and has the nice property of being additive over decompositions of the matrix.
-Making the change of variables $x=\frac{1}{z}$, we have $$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p \in \Path_n (G)(a,a)} F(p)$$
-where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
-paths of length $n$ from $a$ to $b$.
-$(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.
-
-We now make use of the additivity.  Let $\hier$ be the functor which
-sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
-in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
-for the fibers of the points in the hierarchy which corresponds to a
-factorization of the characteristic polynomial.
-
-\section{Stability criteria}
-We call a matrix $M$ stable if all its eigenvalues have negative real
-part.  Define the polynomial $p(x) = \det(xI + M)$.  Then, since the
-roots of $p$ are minus the eigenvalues of $M$, asking that $M$ be
-stable is equivalent to asking that the roots of $p$ all have real
-part.  We will formulate conditions for this to happen in terms of the coefficients of $p$.
-
-Let us start by fixing some notation.  Suppose that $M$ is of size $n
-\times n$.  Denote the coefficients of $p$ as $s_j$ and its roots as
-$r_j$, writing $p(x) = x^n + \sum_{k=1}^n s_k x^{n-k} = \prod_{k=1}^n
-(x-r_k)$.
-
-Introduce $V$ as the parameter space of all polynomials of degree $n$
-with real coefficients.  In other words, $V$ is a real vector space of
-dimension $n$ whose coordinates we will identify with the coefficients
-$s_k$ of our polynomial.  We now partition this space into regions
-corresponding to specific types of polynomials.  Let $R_{abcd}$ be the
-subset of $V$ such that $x \in R_{abcd}$ if and only if the polynomial
-corresponding to $x$ has $a$ positive real roots, $b$ negative real
-roots, $c$ complex roots with positive real part and $d$ complex roots
-with negative real part.  Once we characterize these regions using
-invariants, we will have solved the original problem because a matrix
-is stable if the point of $V$ corresponding to it lies in a region of
-the form $R_{a0c0}$.
-
-Next, we introduce some products:
-\begin{align}
-D &= \prod_{k=1}^n r_n \cr
-\delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 - r_j)^2 \cr
-\Delta &= \prod_{j=1}^{n-1} \prod_{k=j+1}^n (r_1 + r_j) \cr
-\end{align}
-
-Since these quantities are invariant under permutation of the roots
-$r_j$, they may be expressed as functions of the coefficients $s_j$.
-The explicit expressions may be obtained by such techniques as long
-division and determinants. These quantities are of interest because they describe the boundaries between the regions introduced above.
-
-The simplest case is $n=2$.  There we have the regions $R_{2000},
-R_{1100}, R_{0200}, R_{0020}, R_{0002}$ and our invariants are $D =
-s_2, \delta = s_1^2 - 4s_2, \Delta = s_1$.  The relation between these
-is summarized by the conditions
-\begin{align}
-R_{2000} \qquad &D > 0, \delta > 0, \Delta > 0, \cr
-R_{1100} \qquad &D < 0, \delta > 0, \cr
-R_{0200} \qquad &D > 0, \delta > 0, \Delta < 0, \cr
-R_{0020} \qquad &\delta < 0, \Delta > 0, \cr
-R_{0002} \qquad &\delta < 0, \Delta < 0. \cr
-\end{align}
-These regions used for the classification of stability of two-component systems are show in Fig.
-\ref{fig:region2x2}.
-The stable regions are given by $R_{2000}$ and $R_{0020}$.
-All of the other regions represent unstable matrices.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
-Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
-\label{fig:region2x2}
-\end{figure}
-
-Moving to $n=3$, our invariants now look as follows: $D = s_3, \delta
-= 18 s_1 s_2 s_2 + s_1^2 s_2^2 - 4 s_2^3 - 4 s_1^3 s_3 - 27 s_3,
-\Delta = s_1 s_2 - s_3$.  However, now we can no longer simply use their
-signs to distinguish regions.  For instance, consider the regions
-$R_{3000}$ and $R_{1200}$.  These both have $D > 0$ and $\delta > 0$
-but to tell them apart, we need to note that the region $\delta > 0$
-has two connected components and use the auxiliary condition that $s_1 > 0$ to
-specify which of the components contains our region. The regions analogous to Fig. \ref{fig:region2x2} for three component systems where $D=1$ is given in Fig. \ref{fig:region3x3}.
-%(Looking at the sign of $\Delta$ will not help here because, while $\Delta > 0$ in the region $R_{3000}$, it is also positive for some elements of $R_{1200}$ as well.)
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-\caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
-This is the plane $s_3=1$.}
-\label{fig:region3x3}
-\end{figure}
-
-\section{Two component systems}
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant.
-
-Suppose we have a stable matrix
-$$
-\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-We will consider the case in which each of the parameters defining the matrix is sampled from the uniform distribution $\mathcal{U}(-1,1)$ so that the parameter space corresponds to the $d$-dimensional hypercube, $H^d$, of edge length $r=2$, centered about the origin, and having volume $r^d$.
-We want to know the probability that, if we resample an element of this matrix, the result will be stable.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent
-to resampling $d$.
-
-Suppose that we resample $b$.  We want to compute the probability that a random stable matrix will remain stable upon resampling $b$, which is
-% \begin{strip}
-\begin{align}\label{eq:condprob}
-P\left(\begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)
-& = \frac{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right)}.
-\end{align}
-The denominator of \ref{eq:condprob} is given by
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{ac>bd \atop a+c<0} \atop H^4} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprob}
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a & b' \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix} \textrm{stable and } \begin{pmatrix}
-a' & b \\
-d & c
-\end{pmatrix} \textrm{stable } \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-\begin{table}[h]
-\begin{tabular}{ c || c | c | c }
-\hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
-\hline
-  $\begin{pmatrix}
-a & b \\
-d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
-  $\begin{pmatrix}
-a & b \\
-d & 0
-\end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
-  $\begin{pmatrix}
-a & 0 \\
-d & c
-\end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
-$\begin{pmatrix}
-a & 0 \\
-0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
-\end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all two component systems is given in Table \ref{tab:structstabmat}.
-
-\section{Three component systems}
-
-The probability of \emph{a priori} stability and of stability to perturbations for all three component systems is given in Table \ref{tab:structstabmat3}.
-Fig.
-\ref{fig:connectcycle3D3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity and number of cycles.
-Fig.
-\ref{fig:stab3x3} shows the probability of stability to perturbations for all three component systems as a function of connectivity.
-Fig.
-\ref{fig:cycle3x3} shows the probability of stability to perturbations for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles and connectivity for three component systems.} }
-\label{fig:connectcycle3D3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/stab3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus connectivity for three component systems.} }
-\label{fig:stab3x3}
-\end{figure}
-
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf Plot of stability to perturbations versus number of cycles for three component systems.} }
-\label{fig:cycle3x3}
-\end{figure}
-
-\newpage
-
-\input{tablenographs}
-
-\newpage
-
-\input{tablewithgraphs}
-
-\bibliographystyle{amsplain}
-\bibliography{bib/books,bib/papers}
-
-\section{Appendix A}\label{sec:appA}
-Suppose that we have a family
-of polynomials whose roots are of one type but turn into another type.
-There are three ways this could happen:
-\begin{itemize}
-\item{1.} A real root could go from being positive to being negative
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a zero root.  For that intermediate value,
-  we have $D = 0$.
-\item{2.} A pair of real roots could turn into a pair of complex roots
-  or vice versa.  For this to happen, we must pass through an
-  intermediate value with a repeated root.  For that intermediate
-  value, we have $\delta = 0$.
-\item{3.} A conjugate pair of complex roots could go from having
-  positive real part to having negative real part or vice versa.  For
-  this to happen, we must pass through an intermediate value in which
-  the real part is zero.  For that intermediate value, the two roots
-  are imaginary so, bing conjugates, they sum to zero, hence we have
-  $\Delta = 0$.
-\end{itemize}
-\noindent  Thus, we conclude that the boundaries of our regions
-$R_{abcd}$ consist of portions of the surfaces $D=0, \delta=0,
-\Delta=0$ and we will characterize them by studying these surfaces.
-
-\end{document}
-
-%------------------------------------------------------------------------------
-% End of journal.tex
-%------------------------------------------------------------------------------

### 2014-10-13 tex/introduction.tex
refactor introduction

@@
 The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-Recent work has demonstrated that as a result of sloppiness in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
+All of these models utilize essentially equivalent mathematical structure \ref{fig:biomodelexamples} \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
+
+Recent work has demonstrated that as a result of \emph{sloppiness} in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
 In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
@@
 Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is so general that to restrict its applicability to ecosystem dynamics is certainly not necessary and perhaps even inefficient with respect to the goal of understanding how biological networks operate across the various relevant levels of organization.
-For example, the study of chemical reaction, population, and gene-regulatory networks all utilize essentially equivalent mathematical structure as that discussed by May in the context of ecosystem dynamics \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
-This incredibly generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed \cite{Cohen1984,May1972a,Radius2014}.
+The class of dynamical systems studied by May is not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the biological networks represented in \ref{fig:biomodelexamples}.
+In particular, May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014}.
+This relationship implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
+
+% The generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
+However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed.

### 2014-10-14 tex/supplement.tex
add supplement file

%!TEX root = ../paper.tex
\subsection{Dynamical systems on biological networks}
In the construction of a class of potential models for biological systems at any level of the biological hierarchy from metabolic to ecosystem-level networks, it is common to first attempt to define a collection of observable phenomena of interest and determine a domain (such as binary numbers, integers, or real numbers) in which each observable can be quantified. Next, it is necessary to establish the interdependencies among system components. The specific manner in which the components depend upon one another must be clarified, which is often done by determining a particular system of mathematical functions that represents a hypothesis about how the quantified observables evolve in time. To the degree to which there is uncertainty about the interactions among system components, parameters are introduced to broaden the class of models under consideration. Finally, whatever model class remains may be compared to empirical observations to determine how capable the model is of representing the phenomena of interest.

For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}. For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$. If the system under consideration has the graph given in \reffigexamplesystemmodules
% \begin{center}
% \input{tex/examplehamdyngraph}
% \end{center}
with adjacency matrix
$$
\adj(G_X) = \begin{pmatrix}
0 & 1 \\
1 & 1
\end{pmatrix}
$$
having connectivity $\connectivity$ equal to the number of edges of the graph, in this case $\connectivity = 3$. For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
 \begin{displaymath}
   \adj(G_X)_{ij} = \left\{
     \begin{array}{ll}
       1, & F_i \hbox{ depends on } x_j\\
       0, & F_i \hbox{ does not depend on } x_j
     \end{array}.
   \right.
\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
\begin{align*}
\frac{dx_1}{dt} & = F_1(x_2)\\
\frac{dx_2}{dt} & = F_2(x_1,x_2)
\end{align*}

In the more general case of a system with $n$ components we would have an $n$-dimensional vector of observables
$$
x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
$$
whose components are solutions to the arbitrary first order system
$$
\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
$$
where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.

In order to accommodate the possibility of uncertainty in our
modeling, we will generalize our notion of a system on a network to
that of a system with random parameters.  Again, this will involve
three steps. First, we provide a measure space $S$
which represents the values over which our parameters can vary. Next,
instead of a single vector field, we consider a family of vector
fields parameterized by this space.  As before, for
each point $p \in S$, the vector field $V(p)$ must be consistent with
the system graph. Finally, we select a probability measure, $\mu$, which represents our understanding of which values of the parameters
are likely. In accord with Bayesian statistics, we might revise this
distribution as data comes in or use it to estimate parameters and error
bars from experimental results.

Having done this, we are now in a position to do a probabilistic analysis
of our system and its properties.  Given some quantity $q$ characterizing
our system, this quantity becomes a random variable which may be discrete or
continuous depending on the quantity under consideration.  For example, if
the quantity is the time it takes for a particle to travel between two points,
we have a real-valued random variable; if the quantity is the number of fixed
points, we have an integer-valued random variable; and if the quantity is
whether or not the system posesses a limit cycle, we have a binary random
variable.

A quality which is of interest to us is robustness or structural stability \cite{Smale1967}, which requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters.  We will quantify this as the probability that, if some property holds for a set of parameters, it will continue to hold if we make a random perturbation about those parameter values. To do so, we will introduce, in addition to the probability distribution $\mu$ described above, a family of probability distributions $\mu'$ such that $\mu'(x,y)$ encodes the conditional probability that a system with a parameter value $x$ will have its parameter values changed to $y$ under a perturbation.  For our biological models, we will be interested in large perturbations rather than the small or infinitesimal perturbations usually considered in the theory of dynamical systems.  Then, given a binary random variable $q$, we define its robustness as the following conditional probability where $\mathbf{1}_q(x)$ is the standard indicator function equal to $1$ when $x$ satisfies $q$ and $0$ otherwise:
\begin{align}\label{eq:robustness}
  R (q,\mu') =
  \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
  {\int_S d\mu(x) \mathbf{1}_q(x)}
\end{align}

\subsection{Stability analysis of biological networks}
The class of dynamical quantities on which we shall focus in this
investigation involve stability of equilibria.  Suppose that $\vec x$
is a point on our phase space which depends upon the parameters.  Then
we may take $q$ to be a binary random variable which describes whether
or not $\vec x$ is a fixed point of the system (i.e. $q$ is true if and only if
$F_i(\vec{x})=0$ for all $i$) and perform a probabilistic analysis
of the sort discussed above.

Furthermore, if it turns out that $\vec x$ is indeed a fixed point, we
may proceed to ask whether it is a dynamically stable fixed point.
Intuitively, dynamic stability means that, if one chooses the initial
conditions sufficiently close to the fixeed point, the solution will
stay close to the fixed point.  Physically, this is important because,
if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
observing the solution ${\vec x}(t) = {\vec x}^0$.

To determine stability, we linearize the equations of motion about the
fixed point $\vec{x}^0$:
\begin{equation}\label{eq:lineardynsys}
\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
\end{equation}
where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
$$
a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
$$
To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
 \begin{displaymath}
   \adj(G_A)_{ij} = \left\{
     \begin{array}{lr}
       1, & a_{ij} \neq 0\\
       0, & a_{ij} = 0
     \end{array}
   \right.
\end{displaymath}
The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.

The spectral abscissa of the matrix $A$ is defined as
$$
\eta(A) = \max_i \{\Re(\lambda_i)\}
$$
where $\lambda_i$ are the eigenvalues of $A$. The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$. This is because the general solution to \ref{eq:lineardynsys} is
$$
y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
$$
for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$. This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.

Even though the determination of stability only requires examining
linearized equations of motion, it is worth noting that knowing the
stability of a fixed point can yield information about the non-linear
dynamics of a system.  For instance, in two dimensions,
Poincare-Bendixson theory implies that every limit cycle must encircle
at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
conflicting populations), the stability of the fixed point will determine whether the limit cycle is attractive or repulsive.

%% To determine whether a randomly selected dynamical system evaluated at
%% a random critical point, given by a matrix such as $A$, is stable, it
%% is sufficient to check the above condition. Integrating over the
%% region of the parameter space defining $A$ enables the determination
%% of the probability of stability to perturbations in state $\vec{x}$
%% over an ensemble of dynamical systems.

% Another form of stability, referred to as structural
% stability \cite{Smale1967}, requires the determination of whether or
% not a given dynamical system that is determined to be stable remains
% stable under a perturbation to one of its defining parameters. This
% can be formalized as the conditional probability distribution
% $$
% P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
% $$
% where $A'$ represents a matrix derived from some perturbation
% applied to the matrix $A$ given above. In the simple case where
% $A \in \mathbb{R}^{n \times n}$ the perturbation which involves
% sampling uniformly over the $a_{ij}$ defining $A$ the value of a
% particular $a_{ij}$ from a given probability distribution.

Suppose that our Jacobian $A$ is an $n \times n$ matrix with real coefficients, connectivity $\connectivity$, and denote the proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.  Then we may compute the robustness of this proposition using \ref{eq:robustness} once we have specified the probability densities $\mu$ and $\mu'$.  In general, these will depend upon the parameters of the non-linear system.  For the purpose of the current investigation, we shall make the ansatz that these are such that the distribution of entries of $A$ is (at least approximately) the uniform distribution $\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin. To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution yielding
$$
\mu_{k_1,\ldots,k_m}(x,y) = \prod_{y \notin \{x_1, \ldots, x_m\} } \delta(x_j-y_j) \prod_{y \notin \{x_1,\cdots,x_m\}} \mathbf{1}_{[-1,1]} (y_j) \mathbf{1}_{[-1,1]} (x_j)
$$
Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
\begin{equation}\label{eq:condprobgen}
 R (\mathrm{stab}, \mu'_{k_1, \ldots, k_m}) =
  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)},
\end{equation}
where we made the abbreviations
\begin{align*}
\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
\end{align*}
i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.

While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.  The result of this operation corresponds to averaging our quantity over subsets of links:
\begin{equation}\label{eq:robavgoverlinks}
\langle R (\hbox{stab}, \mu') \rangle_m =
\frac{1}{\binom{\connectivity}{m}}
\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
R (\hbox{stab}, \mu'_{k_1, \ldots, k_m})
\end{equation}
We are fundamentally interested in the relationship between \ref{eq:condprobgen} and \ref{eq:robavgoverlinks} and topological properties of systems that derive from their associated graphs. In order to investigate this further, we must characterize properties of system graphs in order to evaluate correlations of those properties with dynamical robustness.

\subsection{Network hierarchy and strongly connected components}

Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems \reffigexamplesystemmodules. One such kind of system composition and decomposition is given by the notion of open systems, which are distinguished by having some of the variables specified as control variables whose values are given as autonomous functions of time rather than determined by the dynamics via intrinsic interactions \cite{Vagner2014}.  Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.  Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.  However, this decomposition is trivial since it is equivalent to the underlying system graph and what we instead want is an intermediate decomposition into relatively self-contained modules.  One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.

A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}. The strongly connected components of some examples of three component systems are outlined in \reffigscc along with their adjacency matrices and hierarchy diagrams (to be defined below).
The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.  Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.  Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.  Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.  This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.

Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$.  Each node of $\hier (G)$ corresponds to a strongly connected component of $G$. There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.  Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.

One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
a strongly connected graph $C_n$.  To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.  The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.

This map $\hier$ is many-to-one and so there is a large class of
operations which leaves $\hier(G)$ invariant for a given graph $G$.
For instance, we may interchange the positions of the strongly
connected components relative to each other \reffighiertransformations a.  Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations b or between components, or even add or delete links \reffighiertransformations c.  As we shall see later, some of these operations leave invariant key dynamical quantities, so we may regard them as a symmetry groupoid of our system.

The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents. The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc (top panel). This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components. The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}. This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical. In this work we take this distance to be the definition of hierarchy.

\subsection{Stability and robustness analysis of particular system ensembles}
For systems having two components, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}. For those having three components, we can estimate these same quantities using Monte Carlo simulations. Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis. For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
0$ where $T$ and $D$ denote the trace and the determinant. Suppose we have a stable matrix
$$
\begin{bmatrix}
a & b \\
d & c
\end{bmatrix}
$$
where $a + c < 0$ and $ac > bd$.  For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},\mu'_k)$ where $k=1 \ldots 4$. By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $ R(\mathrm{stab},\mu'_1)$ and $R(\mathrm{stab},\mu'_2)$. Suppose that we resample $b$ to compute $R(\mathrm{stab},\mu'_2)$.
% \begin{strip}
% \begin{align}\label{eq:condprob}
% P\left(\begin{pmatrix}
% a & b' \\
% d & c
% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable } \right)
% & = \frac{P\left(\begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
% a & b' \\
% d & c
% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
% a & b \\
% d & c
% \end{pmatrix} \textrm{stable } \right)}.
% \end{align}
The denominator of \ref{eq:condprobgen} in this case is given by
\begin{align*}
P\left(\textrm{stab } \left( \begin{bmatrix}
a & b \\
d & c
\end{bmatrix} \right) \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
\end{align*}
Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant. Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprobgen}
\begin{align*}
P\left(\textrm{stab} \left( \begin{bmatrix}
a & b \\
d & c
\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
a & b' \\
d & c
\end{bmatrix} \right) \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
\end{align*}
The analogous equation for resampling $a$ is
\begin{align*}
P\left(\textrm{stab} \left( \begin{bmatrix}
a & b \\
d & c
\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
a' & b \\
d & c
\end{bmatrix} \right) \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
\end{align*}

Using this approach the probability of stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.

The analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B. This process is associated with some error relative to the exact integration described above. If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution. We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}. The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
$$
\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
$$
where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
which corresponds in this case to the maximum likelihood estimate
$$
\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
$$
This estimate is characterized by the variance of the posterior Beta distribution
$$
\mathrm{var}(\theta | \mathcal{D}) = \frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
$$
Since for the chosen prior $a=b=1 \ll N$ this simplifies to
$$
\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
$$
yielding the error estimate given by the associated standard deviation. In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.

It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}. For small system sizes such as the two and three component systems, studied here the situation is not so clear cut. For two component systems, system stability is constant across the entire range of connectivities. For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:combined}A.

The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:combined}B. If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \ref{fig:combined}B.
% \subsection{Cycle number is inversely correlated with robustness}
\ref{fig:combined}C shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}. There appears to be a weak negative correlation between robustness and the number of simple cycles.

% \subsection{Cycle number and connectivity classify robust systems}
The combination of connectivity and cycle number as shown in \ref{fig:connectcycle3D3x3} provides a better classification of the dependence of robustness upon network topology. Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity. The network with the highest robustness for three component systems is that of \reffigscc (top panel). This network is the most hierarchical of all three component systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \reffigscc (top panel) has a block triangular structure.

This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of system robustness. \ref{fig:combined}D shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0. There is a monotonic correlation between the upper bound of robustness and hierarchy. \ref{fig:connectdist3D3x3} shows system robustness as a function of both hierarchy and connectivity. The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:connectcycle3D3x3}.

\subsection{Component decomposition, hierarchy, and the symmetries of robustness}

The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.  By combining our decomposition of the graph with our analysis of the robustness of random ensembles of dynamical systems, we will deduce a general result which relates robustness to hierarchy.

Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomials factors as the product of the charactericstic polynomials of its diagonal blocks.  Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.  Note that this condition does not depend upon the entries off the diagonal (which correspond to links between strongly connected components) and does not depend upon what order the components appear.

Using these observations, we may express the robustness of a graph in
terms of the robustnesses of its strongly connected components.  If
the components $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
$\connectivity_i$, and there are $\ell_{ij}$ links between components
$i$ and $j$, then there are a total of
$\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i$
links in $G$.  If we resample a randomly chosen link in a component
$C_i$, then the probability that the system will remain stable after
resampling is $\langle R(\hbox{stab} (C_i), \mu') \rangle_1$.  If we resample a link
between components $C_i$ and $C_j$, the system will remain stable with
probability 1.  Hence, the probability that the system will remain
stable upon resampling a random link is the weighted average of these
probabilities:
\begin{align}\label{eq:sccrobustness}
\langle R(\hbox{stab} (G), \mu') \rangle_1 =
\frac{\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
      \sum\limits_{i=1}^n \connectivity_i \langle R(\textrm{stab}(C_i),\mu') \rangle_1}
     {\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
      \sum\limits_{i=1}^n \connectivity_i}.
\end{align}
For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.  From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and and robustness $0.62$.  The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.  Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is
\[
\langle R(\textrm{stab}(G),\mu') \rangle_1 = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
\]
which agrees with the value computed in \ref{tab:structstabmat} up to
sampling error.

As can be seen from \ref{eq:sccrobustness}, all that is required to fix robustenss is a set of SCC sizes and the number of links between them. This demonstrates that robustness is symmetric under the transformations shown in \reffighiertransformations a and b. Since transformations of the type \reffighiertransformations c change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}. Nevertheless, a considerable degree of symmetry remains. For example, \ref{fig:robustnesssymmetries} shows the symmetry groupoid of robustness for the case in which we fix connected component sizes $\{2,1,1\}$ with a total of $3$ links between them.

Next, we derive an upper bound on the robustness of any graph
constructed from a given set of connected components.  Since
$\langle R(\textrm{stab}(C_i), \mu') \rangle_1 \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
increase $\langle R(\textrm{stab}(G),\mu') \rangle_1$.  Given two connected components $C_i$ and $C_j$ with
$v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$
links going from $C_i$ to $C_j$.  Hence, $\ell_{ij} \le v_i v_j$, so,
we conclude that
\begin{align}
\langle R(\textrm{stab}(G),\mu') \rangle_1 \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
               \sum\limits_{i=1}^n \connectivity_i \langle R(C_i,\mu') \rangle_1}
              {\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
               \sum\limits_{i=1}^n \connectivity_i},
\end{align}

Since every acyclic digraph can be embedded into a totally ordered
set, we may assume without loss of generality that our components have
been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i <
j$.  Hence, after substituting into our expression and simplifying
using the identity
$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n}
v_i^2,$$
we arrive at the following upper bound on robustness in terms
of the connected components:
\begin{align} \label{eq:sccmaxrobustness}
R_{\mathrm{max}}(C_1, \ldots C_n) =
\frac{\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
                    \sum_{i=1}^n \connectivity_i \langle R(\hbox{stab}(C_i),\mu') \rangle_1}
     {\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
                    \sum_{i=1}^n \connectivity_i}.
\end{align}
Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
is the graph on $n$ nodes with a link from node $i$ to node $j$
whenever $i < j$.  Then, by the construction described in the previous
section on network hierachy, we have a graph $G_{\mathrm{max}}$ such
that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and
$\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula,
$\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.

This argument also works when we resample more than one entry,
although the notation becomes more complicated.  Suppose that we
resample $m$ nodes.  Define
\begin{equation*}
M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
m = \sum_{i=0}^n m_i \quad\&\quad
m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
\forall i \in \{1, \ldots, n\} \, m_i < \connectivity_i \right\}.
\end{equation*}
Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
$m_0$ links between strongly connected components.  Hence, our
weighted average becomes
\begin{equation}
\langle R(\hbox{stab} (G), \mu') \rangle_m =
\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
      {m \choose m_0, m_1, \ldots. m_n}
      \left(m_0 + \sum\limits_{i=1}^n m_i \langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \right)}
     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
      {m \choose m_0, m_1, \ldots. m_n} m}
\end{equation}

As before, since $\langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \le 1$, we may
increase $\langle R(\hbox{stab} (G), \mu') \rangle_m$ by increasing the maximum
possible value of $m_0$ while keeping the strongly connected
components the same.  Again, if we fix $\hier(G)$, the maximum
possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
G_{tot}$.  Hence, we conclude that $\langle R(\hbox{stab} (G), \mu') \rangle_m \le
\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_m$.

The preceding argument demonstrates the manner in which robustness is
maximized by maximizing the number of edges \emph{between} the strongly
connected components of the graph underlying system interdependencies.
And, the graph that has the latter property is that associated to the
ordering, $G_{\mathrm{tot}}$, which also possess the highest degree of
hierarchy. The fact that robustness is equivalent for any configuration
of strongly connected components with an equivalent number of links
between them demonstrates that it is invariant to permutations of
strongly connected components. This result does not depend at all upon
any particular details such as the probability distribution from which
the component interaction strengths are sampled because the arguments
presented here are purely topological in nature. The result that
dynamical robustness is correlated with system hierarchy therefore
applies to an even broader class of dynamical systems than the particular
random ensembles primarily studied here.

### 2014-10-15 tex/figures.tex
recombine figures

@@
 \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Stability versus connectivity, (B) robustness versus connectivity, (C) robustness versus number of cycles, (D) robustness versus hierarchy for three component systems.
-}1,500 words long
+\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Robustness versus connectivity, (B) robustness versus hierarchy, (C) Robustness versus number of cycles and connectivity for three component systems.
+Each point represents the average robustness of all graphs having a given number of cycles and connectivity.
+(D) Robustness versus hierarchy and connectivity for three component systems.
+Each point represents the average robustness of all graphs having a given hierarchy and connectivity.
+}
 \label{fig:combined}
@@
 
-\pagebreak
+% \pagebreak
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-\caption{{\bf Robustness versus number of cycles and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given number of cycles and connectivity.}
-\label{fig:connectcycle3D3x3}
-\end{figure}
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
+% \caption{{\bf Robustness versus number of cycles and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given number of cycles and connectivity.}
+% \label{fig:connectcycle3D3x3}
+% \end{figure}
 
-\pagebreak
+% \pagebreak
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\columnwidth]{fig/connectdist3D3x3.pdf}
-\caption{{\bf Robustness versus hierarchy and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given hierarchy and connectivity.}
-\label{fig:connectdist3D3x3}
-\end{figure}
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\columnwidth]{fig/connectdist3D3x3.pdf}
+% \caption{{\bf Robustness versus hierarchy and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given hierarchy and connectivity.}
+% \label{fig:connectdist3D3x3}
+% \end{figure}

### 2014-10-15 tex/supplement.tex
recombine figures

@@
 
-The analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \ref{fig:combined}B.
+The analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \reffigrobustconnect.
 This process is associated with some error relative to the exact integration described above.
@@
 For two component systems, system stability is constant across the entire range of connectivities.
-For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:combined}A.
+For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
 
-The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \ref{fig:combined}B.
-If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \ref{fig:combined}B.
+The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \reffigrobustconnect.
+If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \reffigrobustconnect.
 % \subsection{Cycle number is inversely correlated with robustness}
-\ref{fig:combined}C shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+\ref{fig:cycle3x3} shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
 There appears to be a weak negative correlation between robustness and the number of simple cycles.
@@
 % \subsection{Cycle number and connectivity classify robust systems}
-The combination of connectivity and cycle number as shown in \ref{fig:connectcycle3D3x3} provides a better classification of the dependence of robustness upon network topology.
+The combination of connectivity and cycle number as shown in \reffigconnectcycle3D3x3 provides a better classification of the dependence of robustness upon network topology.
 Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity.
@@
 This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of system robustness.
-\ref{fig:combined}D shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
+\reffigrobusthierarchy shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
 There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\ref{fig:connectdist3D3x3} shows system robustness as a function of both hierarchy and connectivity.
-The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \ref{fig:connectcycle3D3x3}.
+\reffigconnectdist3D3x3 shows system robustness as a function of both hierarchy and connectivity.
+The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \reffigconnectcycle3D3x3.
 

### 2014-10-15 code/rds3x3.jl

@@ -391,12 +391,12 @@ end
 
 function combinegnuplot()
 
-    filelist = ["apstab3x3.p",
+    filelist = [#"apstab3x3.p",
                 "stab3x3.p",
-                "cycle3x3.p",
-                "dist3x3.p",]
-                # "connectcycle3D3x3.p",
-                # "connectdist3D3x3.p",]
+                #"cycle3x3.p",
+                "dist3x3.p",
+                "connectcycle3D3x3.p",
+                "connectdist3D3x3.p",]
     horiz = 2
     vert = 2
     rv = (String)[]

### 2014-10-15 tex/figures.tex
update results

@@
 \noindent\includegraphics[width=0.9\columnwidth]{fig/biomodelexamples.pdf}
-\caption{{\bf Dynamical models in systems biology.} (A) (B) (C)}
+\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) a gene regulatory network (GRN) and an ecological regulatory network (ERN) in terms of the graphical methods specific to each field.
+The second row represents the interaction graph, which provides a unified representation for networks across these fields.
+The third row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
+The fourth row shows the general form of a system of differential equations that can be used to model any network architecture within the three domains of application exemplified here.}
 \label{fig:biomodelexamples}

### 2014-10-15 tex/abstract-cited.tex
revise

@@
 In particular, we determine the probability distribution of robustness over system connectivity and show that robustness is maximized by maximizing the number of links between strongly connected components of the graph representing the underlying system connectivity.
-We also demonstrate that permutation of strongly connected components is a fundamental symmetry of system robustness, which applies to networks of any number of components.
+We also demonstrate that permutation of strongly connected components is a fundamental symmetry of dynamical robustness, which applies to networks of any number of components.
 The classification of dynamical robustness based upon a purely topological property provides a fundamental organizing principle that can be used in the context of experimental validation to select among models that break or preserve network hierarchy.

### 2014-10-15 tex/discussion.tex
revise

@@
 %!TEX root = ../paper.tex
-The relationship between structure and function is fundamental to biology.
-Structure-function relationships exist at every level of the biological hierarchy.
-In all such networks, the ability of systems to persist over long periods of time, is dependent upon their dynamical stability.
+The relationship between structure and function is fundamental to networks at every level of the biological hierarchy.
+Equally fundamental is the ability of systems to persist over long periods of time, which is dependent upon their dynamical robustness.
 Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist in the evolutionary process.
-This relationship transcends levels of the biological hierarchy in the sense that it is formulated mathematically and at a level of abstraction that allows its conclusions to apply independently to each of these levels.
+Moreover, this result is formulated at a level of abstraction that allows its conclusions to apply independently to each of these levels.
 
 Our development suggests two predictions that could be used to test the theory described here.
-One is that, if this theory is correct, any ensemble of systems where robustness has been selected for over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
+One is that, if this theory is correct, any ensemble of systems where robustness has been positively selected over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
 In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
@@
 The fact that such a relationship exists has the potential to unify our understanding of biological networks.
-In order to pursue this, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the strongly connected components of the underlying network topology.
+In order to pursue this, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the underlying network topology.
 One approach is to attempt to compute robustness for systems of larger size.

### 2014-10-15 tex/introduction.tex
draft results
revise

@@
 %!TEX root = ../paper.tex
-The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to consider a particular example, derive a system of differential equations to model that example, try to fit the model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-All of these models utilize essentially equivalent mathematical structure \ref{fig:biomodelexamples} \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
+All of these models utilize essentially equivalent mathematical structures \ref{fig:biomodelexamples} \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
 Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
 
-Recent work has demonstrated that as a result of \emph{sloppiness} in the dependence of qualitative dynamic phenomena on the geometry of parameter space that this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
-In the face of uncertainty about the structure of such biological networks, to model the components under consideration as randomly interlinked becomes a reasonable approximation.
-This approach enables one to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
+Recent work has demonstrated that as a result of inhomogeneity in the geometry of the sensitivity of the dynamics over the parameter space for such models, this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
+In addition, there is often uncertainty about the very structure of such biological networks.
+In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
+This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems that, in comparison to one another, appear to have components that are randomly interlinked.
 
 Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the biological networks represented in \ref{fig:biomodelexamples}.
-In particular, May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014}.
-This relationship implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
+The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the biological networks represented in \ref{fig:biomodelexamples}.
+May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014}.
+This implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
 

### 2014-10-15 tex/supplement.tex
revise

@@
 
-This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of system robustness.
-\reffigrobusthierarchy shows system robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
+This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of dynamical robustness.
+\reffigrobusthierarchy shows dynamical robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
 There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\reffigconnectdist3D3x3 shows system robustness as a function of both hierarchy and connectivity.
+\reffigconnectdist3D3x3 shows dynamical robustness as a function of both hierarchy and connectivity.
 The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \reffigconnectcycle3D3x3.

### 2014-10-16 tex/figures.tex
split bibliographies and update figures

--- 2014-10-15 tex/figures.tex
@@
 \noindent\includegraphics[width=0.9\columnwidth]{fig/biomodelexamples.pdf}
-\caption{{\bf Dynamical models in systems biology.} (A) (B) (C)}
+\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
+The third row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here \cite{Shinar2010,Karlebach2008,Shinar2010}.
+The fourth row shows the general form of a system of differential equations that can be used to model any network architecture of any size within each of the three domains of application exemplified here.}
 \label{fig:biomodelexamples}

### 2014-10-16 tex/figures.tex
update fig 1 legend

@@
 \caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
-The third row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here \cite{Shinar2010,Karlebach2008,Shinar2010}.
-The fourth row shows the general form of a system of differential equations that can be used to model any network architecture of any size within each of the three domains of application exemplified here.}
+The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here \cite{Shinar2010,Karlebach2008,Shinar2010}.
+The third row shows the general form of a system of differential equations that can be used to model any network architecture of any size within each of the three domains of application exemplified here.}
 \label{fig:biomodelexamples}

### 2014-10-16 tex/discussion.tex
update discussion

@@
 %!TEX root = ../paper.tex
+% Our development suggests two predictions that could be used to test the theory described here. One is that,
+Our theory predicts that any ensemble of systems where robustness has been positively selected over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
+At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been demonstrated to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
+In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
+%The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
+Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
+In the long term, this prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that differentially select for robustness \cite{Leroi1994}.
+
+%The fact that such a relationship exists has the potential to unify our understanding of biological networks.
+In order to further this work from a theoretical perspective, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the underlying network topology.
+%One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However,
+It may be possible to prove a bound on the robustness of individual strongly connected components, which could provide more definite knowledge and circumvent the need for expensive simulations.
+The conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
+A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
+
 The relationship between structure and function is fundamental to networks at every level of the biological hierarchy.
@@
 Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist in the evolutionary process.
-Moreover, this result is formulated at a level of abstraction that allows its conclusions to apply independently to each of these levels.
-
-Our development suggests two predictions that could be used to test the theory described here.
-One is that, if this theory is correct, any ensemble of systems where robustness has been positively selected over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
-In the short term, this prediction may be checked by looking at both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
-The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
-Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
-In the long term, each of these predictions may be evaluated using experimental evolution by tracking the evolution of gene regulatory network topology in the context of environmental perturbations \cite{Leroi1994}.
-
-The fact that such a relationship exists has the potential to unify our understanding of biological networks.
-In order to pursue this, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the underlying network topology.
-One approach is to attempt to compute robustness for systems of larger size.
-Using existing methods, this would require large-scale Monte Carlo simulations.
-However, it may be possible to prove a bound on the robustness of individual strongly connected components, which could provide more definite knowledge and help to circumvent such expensive simulations.
-Nevertheless, the conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
-A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
+%Moreover, this result is formulated at a level of abstraction that allows its conclusions to apply independently to each of these levels.

### 2014-10-17 tex/discussion.tex
use SCC acroynym

@@
 %One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However,
-It may be possible to prove a bound on the robustness of individual strongly connected components, which could provide more definite knowledge and circumvent the need for expensive simulations.
-The conservation of robustness with respect to nontrivial symmetries including the interchange of strongly connected components and permutation within strongly connected components suggests the existence of an evolutionary neutral space.
+It may be possible to prove a bound on the robustness of individual SCCs, which could provide more definite knowledge and circumvent the need for expensive simulations.
+The conservation of robustness with respect to nontrivial symmetries including the interchange of SCCs and permutation within SCCs suggests the existence of an evolutionary neutral space.
 A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.

### 2014-10-17 tex/supplement.tex
update modsccsym fig with explicit robustness computations
comment sentences from supplement moved to main text

@@
 For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.
-From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and and robustness $0.62$.
+From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.
 The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
@@
 between them demonstrates that it is invariant to permutations of
-strongly connected components. This result does not depend at all upon
-any particular details such as the probability distribution from which
-the component interaction strengths are sampled because the arguments
-presented here are purely topological in nature. The result that
-dynamical robustness is correlated with system hierarchy therefore
-applies to an even broader class of dynamical systems than the particular
-random ensembles primarily studied here.
+strongly connected components.
+% This result does not depend at all uponany particular details such as the probability distribution from which the component interaction strengths are sampled because the arguments presented here are purely topological in nature.
+The result that dynamical robustness is correlated with system hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles primarily studied here.

### 2014-10-17 tex/discussion.tex
update main text

@@
 % Our development suggests two predictions that could be used to test the theory described here. One is that,
-Our theory predicts that any ensemble of systems where robustness has been positively selected over a sufficiently long period of time, should exhibit a bias toward more hierarchical network topologies.
-At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been demonstrated to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
+Our theory predicts that any ensemble of systems where robustness has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
+For example, at the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
 In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
@@
 The conservation of robustness with respect to nontrivial symmetries including the interchange of SCCs and permutation within SCCs suggests the existence of an evolutionary neutral space.
-A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may help to characterize a fundamental evolutionary constraint \cite{Golubitsky2006}.
+A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may thus help to characterize this potential evolutionary constraint \cite{Golubitsky2006}.
 

### 2014-10-17 tex/figures.tex
update main text

@@
 \caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
-The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here \cite{Shinar2010,Karlebach2008,Shinar2010}.
-The third row shows the general form of a system of differential equations that can be used to model any network architecture of any size within each of the three domains of application exemplified here.}
+The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
+The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
 \label{fig:biomodelexamples}
@@
 \caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
-(B) strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
-The most hierarchical system in the top panel has the highest possible number of connected components, three, whereas the system in the bottom panel containing a single cycle and therefore posessing no hierarchy contains only one connected component.
-The two panels in the middle represent examples of hierarchical modular systems that posess both modularity and hierarchy.
+(B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
+The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
+We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
+The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e.
+SCCs with more than one variable) and hierarchy.
 (C) Symmetries of the $\hier$ transformation between graphs and SCCs.
-The transformation a represents an interchange of SCCs, b moving a link between nodes in a component and c adding a link.
-All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only a and b are symmetries of robustness.}
+The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link.
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
 \label{fig:modsccsym}
@@
 % \noindent\includegraphics[width=0.4\columnwidth]{fig/scc2.pdf}
-% \caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three component systems.
+% \caption{{\bf Example of strongly connected components.} (A) - (D) show strongly connected components highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
 Note that the most hierarchical system in (A) has the highest possible number of connected components, three, whereas the system containing a single cycle and therefore posessing no hierarchy contains only one connected component.
@@
 \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness according to properties of system structure} (A) Robustness versus connectivity, (B) robustness versus hierarchy, (C) Robustness versus number of cycles and connectivity for three component systems.
-Each point represents the average robustness of all graphs having a given number of cycles and connectivity.
-(D) Robustness versus hierarchy and connectivity for three component systems.
-Each point represents the average robustness of all graphs having a given hierarchy and connectivity.
+\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.
+The red colored points represent averages for each connectivity.
+The lowest and highest robustness network architectures are labelled.
+Other network architectures are shown in \ref{tab:structstabmat3}.
+(B) Robustness versus hierarchy.
+(C) Number of cycles and (D) hierarchy vs connectivity and robustness.
+The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
 }
@@
 % \noindent\includegraphics[width=0.8\columnwidth]{fig/connectcycle3D3x3.pdf}
-% \caption{{\bf Robustness versus number of cycles and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given number of cycles and connectivity.}
+% \caption{{\bf Robustness versus number of cycles and connectivity for three variable systems.} Each point represents the average robustness of all graphs having a given number of cycles and connectivity.}
 % \label{fig:connectcycle3D3x3}
@@
 % \noindent\includegraphics[width=0.8\columnwidth]{fig/connectdist3D3x3.pdf}
-% \caption{{\bf Robustness versus hierarchy and connectivity for three component systems.} Each point represents the average robustness of all graphs having a given hierarchy and connectivity.}
+% \caption{{\bf Robustness versus hierarchy and connectivity for three variable systems.} Each point represents the average robustness of all graphs having a given hierarchy and connectivity.}
 % \label{fig:connectdist3D3x3}

### 2014-10-17 tex/introduction.tex
update main text

@@
 The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the biological networks represented in \ref{fig:biomodelexamples}.
-May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014}.
-This implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
+May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
 

### 2014-10-17 tex/supplement.tex
update main text

@@
 A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three component systems are outlined in \reffigscc along with their adjacency matrices and hierarchy diagrams (to be defined below).
+The strongly connected components of some examples of three variable systems are outlined in \reffigscc along with their adjacency matrices and hierarchy diagrams (to be defined below).
 The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
@@
 Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis.
-For two-component systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
+For two-variable systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
 0$ where $T$ and $D$ denote the trace and the determinant. Suppose we have a stable matrix
@@
 
-Using this approach the probability of stability and of robustness for all two component systems is given in \ref{tab:structstabmat}.
+Using this approach the probability of stability and of robustness for all two variable systems is given in \ref{tab:structstabmat}.
 
-The analogous results for all three component systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \reffigrobustconnect.
+The analogous results for all three variable systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \reffigrobustconnect.
 This process is associated with some error relative to the exact integration described above.
@@
 It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
-For small system sizes such as the two and three component systems, studied here the situation is not so clear cut.
-For two component systems, system stability is constant across the entire range of connectivities.
-For three component systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
+For small system sizes such as the two and three variable systems, studied here the situation is not so clear cut.
+For two variable systems, system stability is constant across the entire range of connectivities.
+For three variable systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
 
-The relationship between connectivity and robustness for two component systems is shown in \ref{tab:structstabmat} and likewise for three component systems in \ref{tab:structstabmat3} and \reffigrobustconnect.
+The relationship between connectivity and robustness for two variable systems is shown in \ref{tab:structstabmat} and likewise for three variable systems in \ref{tab:structstabmat3} and \reffigrobustconnect.
 If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \reffigrobustconnect.
 % \subsection{Cycle number is inversely correlated with robustness}
-\ref{fig:cycle3x3} shows the robustness for all three component systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
+\ref{fig:cycle3x3} shows the robustness for all three variable systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
 There appears to be a weak negative correlation between robustness and the number of simple cycles.
@@
 The combination of connectivity and cycle number as shown in \reffigconnectcycle3D3x3 provides a better classification of the dependence of robustness upon network topology.
-Here the robustness of three component systems with a given number of cycles, increases monotonically with connectivity.
-The network with the highest robustness for three component systems is that of \reffigscc (top panel).
-This network is the most hierarchical of all three component systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \reffigscc (top panel) has a block triangular structure.
+Here the robustness of three variable systems with a given number of cycles, increases monotonically with connectivity.
+The network with the highest robustness for three variable systems is that of \reffigscc (top panel).
+This network is the most hierarchical of all three variable systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \reffigscc (top panel) has a block triangular structure.
 
@@
 
-The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
+The fact that the most hierarchical of networks is the most robust in the case of three variable systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
 By combining our decomposition of the graph with our analysis of the robustness of random ensembles of dynamical systems, we will deduce a general result which relates robustness to hierarchy.

### 2014-10-17 tex/table2x2stab.tex
add orbit size to 3x3 table
update main text

@@
 \begin{table}[h]
+\begin{center}
 \begin{tabular}{ c || c | c | c }
 \hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & probability of stability\\
+matrix & connectivity & \specialcell{probability of stability\\to perturbation} & \specialcell{probability\\of stability}\\
 \hline
@@
 \end{tabular}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two component systems derived analytically}.
+\end{center}
+\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two variable systems derived analytically}.
 All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}

### 2014-10-17 tex/tablewithgraphs.tex
reorganize
update main text

--- 2014-09-12 tablewithgraphs.tex
@@
+%!TEX root = ../paper.tex
 \begin{longtable}{ c || c | c | c | c }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
-2 -> 2 [topath="loop above"];;
-3 -> 3 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
+2 -> 2 [topath="loop above"];
+3 -> 3 [topath="loop above"];
 }
@@
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
-2 -> 2 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
+2 -> 2 [topath="loop above"];
 3 -> 2;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
@@
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 2 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
-2 -> 2 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
+2 -> 2 [topath="loop above"];
 3 -> 2;
 2 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 2 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 2 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 2 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
@@
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 1 -> 3;
 2 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 1 -> 2;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 2 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 2 -> 1;
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 2 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 1 -> 2;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 2 -> 1;
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 2 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 2 -> 1;
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
 1 -> 3;
-3 -> 3 [topath="loop above"];;
+3 -> 3 [topath="loop above"];
 }
@@
 1 -> 2;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 2 -> 1;
 3 -> 1;
-2 -> 2 [topath="loop above"];;
+2 -> 2 [topath="loop above"];
 3 -> 2;
@@
 edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
-1 -> 1 [topath="loop above"];;
+1 -> 1 [topath="loop above"];
 2 -> 1;
@@
 \end{tikzpicture} & 9 & 5 & 0.638 & 0.106\\
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3graph}
+\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three variable systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3graph}
 \end{longtable}

### 2014-10-17 tex/supplement.tex
update supplement

@@
 For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
-If the system under consideration has the graph given in \reffigexamplesystemmodules
+If the system under consideration has the graph given in \reffigexamplesystemmodules$\,$
 % \begin{center}
@@
    \right.
-\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
+\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is a function of $x_2$ alone and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
 \begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
+\frac{dx_1}{dt} & = F_1(x_2),\\
+\frac{dx_2}{dt} & = F_2(x_1,x_2).
 \end{align*}
@@
 each point $p \in S$, the vector field $V(p)$ must be consistent with
-the system graph. Finally, we select a probability measure, $\mu$, which represents our understanding of which values of the parameters
+the system graph. Finally, we select a probability measure $\mu$ which represents our understanding of which values of the parameters
 are likely. In accord with Bayesian statistics, we might revise this
@@
 \begin{align*}
-\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
-\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
+\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity), \\
+\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d),
 \end{align*}
@@
 A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three variable systems are outlined in \reffigscc along with their adjacency matrices and hierarchy diagrams (to be defined below).
+The strongly connected components of some examples of three variable systems are outlined in \reffigscc$\,$ along with their adjacency matrices and hierarchy diagrams (to be defined below).
 The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
@@
 
-Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$.
+Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$ or the condensed graph \cite{Corominas-Murtra2013}.
 Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
@@
 For instance, we may interchange the positions of the strongly
-connected components relative to each other \reffighiertransformations a.
-Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations b or between components, or even add or delete links \reffighiertransformations c.
+connected components relative to each other \reffighiertransformations$a$.
+Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations$b$ or between components, or even add or delete links \reffighiertransformations$c$.
 As we shall see later, some of these operations leave invariant key dynamical quantities, so we may regard them as a symmetry groupoid of our system.
@@
 This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
+The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
 This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical.
-In this work we take this distance to be the definition of hierarchy.
+In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.
 
 \subsection{Stability and robustness analysis of particular system ensembles}
-For systems having two components, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}.
-For those having three components, we can estimate these same quantities using Monte Carlo simulations.
+For systems having two variables, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}.
+For those having three variables, we can estimate these same quantities using Monte Carlo simulations.
 Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis.
@@
 \end{align*}
-
 Using this approach the probability of stability and of robustness for all two variable systems is given in \ref{tab:structstabmat}.
@@
 This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of dynamical robustness.
-\reffigrobusthierarchy shows dynamical robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
+\reffigrobusthierarchy$\,$ shows dynamical robustness as a function of hierarchy.
 There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\reffigconnectdist3D3x3 shows dynamical robustness as a function of both hierarchy and connectivity.
+\reffigconnectdist3D3x3$\,$ shows dynamical robustness as a function of both hierarchy and connectivity.
 The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \reffigconnectcycle3D3x3.
@@
 
-Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomials factors as the product of the charactericstic polynomials of its diagonal blocks.
+Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomial factors as the product of the charactericstic polynomials of its diagonal blocks.
 Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.
@@
 terms of the robustnesses of its strongly connected components.  If
-the components $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
+each component $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
 $\connectivity_i$, and there are $\ell_{ij}$ links between components
@@
 As can be seen from \ref{eq:sccrobustness}, all that is required to fix robustenss is a set of SCC sizes and the number of links between them.
-This demonstrates that robustness is symmetric under the transformations shown in \reffighiertransformations a and b.
-Since transformations of the type \reffighiertransformations c change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}.
+This demonstrates that robustness is symmetric under the transformations shown in \reffighiertransformations $a$ and $b$.
+Since transformations of the type \reffighiertransformations $c$ change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}.
 Nevertheless, a considerable degree of symmetry remains.
@@
 increase $\langle R(\textrm{stab}(G),\mu') \rangle_1$.  Given two connected components $C_i$ and $C_j$ with
-$v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$
+$v_i$ and $v_j$ nodes respectively, we have a maximum of $v_i v_j$
 links going from $C_i$ to $C_j$.  Hence, $\ell_{ij} \le v_i v_j$, so,
@@
 m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
-\forall i \in \{1, \ldots, n\} \, m_i < \connectivity_i \right\}.
+(\forall i \in \{1, \ldots, n\}) \; m_i < \connectivity_i \right\}.
 \end{equation*}

### 2014-10-20 tex/abstract-cited.tex
update abstract

@@
 One such constraint is that of dynamical robustness, the probability of a stable system to remain stable under perturbation~\cite{WADDINGTON1942a,Wagner1997,Rutherford1998,VanNimwegen1999,Siegal2002,Bergman2003,Ciliberti2007b,Ciliberti2007,Draghi2010,Wagner2013}.
-Despite previous investigation, the relationship between dynamical robustness, an important functional characteristic of many biological systems, and network structure is poorly understood.
+Despite previous investigations, the relationship between dynamical robustness---an important functional characteristic of many biological systems---and network structure is poorly understood.
 Here we analyze the stability and robustness of a large class of dynamical systems and demonstrate that the most hierarchical network structures are the most robust.

### 2014-10-20 tex/abstract-cited.tex
update abstract

@@
 The classification of dynamical robustness based upon a purely topological property provides a fundamental organizing principle that can be used in the context of experimental validation to select among models that break or preserve network hierarchy.
-This result contributes to an explanation of the observation of hierarchical modularity across levels of the biological hierarchy~\cite{Zhao2006,Ravasz2002,Bhardwaj2010,Colm,Corominas-Murtra2013}.
+This result contributes to an explanation for the observation of hierarchical modularity in biological networks at all scales~\cite{Zhao2006,Ravasz2002,Bhardwaj2010,Colm,Corominas-Murtra2013}.

### 2014-11-25 paper.tex
update acknowledgements
refactor introduction
recombine figures
revise
split bibliographies and update figures
update abstract
modify tables and imported packages to work with revtex

@@
 
-\title{Ensemble Stability of Biological Networks}
-
 \pagenumbering{arabic}
 
-% Title must be 150 characters or less
-\begin{center}
-{\Large
-\textbf{Hierarchical Network Structure Promotes Dynamical Robustness}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\[.5cm]
-Cameron Smith$^{1}$,
+% \twocolumn[
+% \begin{@twocolumnfalse}
+% \title{Ensemble Stability of Biological Networks}
+
+\title{\bf Hierarchical Network Structure Promotes Dynamical Robustness}
+
+\author{Cameron Smith$^{1}$,
 Raymond S. Puzio$^{1}$,
-Aviv Bergman$^{1,2,3,4}$,
-\\[.5cm]
-$^1$Department of Systems and Computational Biology,\\
+Aviv Bergman$^{1,2,3,4,*}$}
+
+\affiliation{$^1$Department of Systems and Computational Biology,\\
 $^2$Dominick P. Purpura Department of Neuroscience,\\
@@
 $^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
-\\[.5cm]
-% $\ast$To whom correspondence should be addressed; E-mail: aviv@einstein.yu.edu.
-\end{center}
+\\
+$*$to whom correspondence should be addressed: aviv@einstein.yu.edu}
 
-{\begin{quote} \bf
-\input{tex/abstract-cited}
-\end{quote}}
+\date{\today}
+
+% Title must be 150 characters or less
+% \begin{center}
+% {\Large
+% \textbf{Hierarchical Network Structure Promotes Dynamical Robustness}
+% }
+% % Insert Author names, affiliations and corresponding author email.
+% \\[.5cm]
+% Cameron Smith$^{1}$,
+% Raymond S. Puzio$^{1}$,
+% Aviv Bergman$^{1,2,3,4,*}$,
+% \\[.5cm]
+% $^1$Department of Systems and Computational Biology,\\
+% $^2$Dominick P. Purpura Department of Neuroscience,\\
+% $^3$Department of Pathology, Albert Einstein College of Medicine,\\
+% 1301 Morris Park Ave, Bronx, NY 10461, USA\\
+% $^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
+% \\
+% $*$to whom correspondence should be addressed: aviv@einstein.yu.edu\\[.5cm]
+% % $\ast$To whom correspondence should be addressed; E-mail: aviv@einstein.yu.edu.
+% \end{center}
+% \end{@twocolumnfalse}
+% ]
+
+% \begin{multicols}{2}
+
+\begin{bibunit}[naturemag]
+
+% {\begin{quote} \bf
+% \input{tex/abstract-cited}
+% \end{quote}}
+% {\bf
+% \input{tex/abstract-cited}
+% }
+
+\begin{abstract}
+The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components~\cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+When the details of these interactions are unknown for biological systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by evolutionary constraints that may apply to all such systems~\cite{Gardner1970,May1972,Cohen1984,May1972a,Radius2014}.
+One such constraint is that of dynamical robustness, the probability of a stable system to remain stable under perturbation~\cite{WADDINGTON1942a,Wagner1997,Rutherford1998,VanNimwegen1999,Siegal2002,Bergman2003,Ciliberti2007b,Ciliberti2007,Draghi2010,Wagner2013}.
+Despite previous investigations, the relationship between dynamical robustness---an important functional characteristic of many biological systems---and network structure is poorly understood.
+Here we analyze the stability and robustness of a large class of dynamical systems and demonstrate that the most hierarchical network structures are the most robust.
+In particular, we determine the probability distribution of robustness over system connectivity and show that robustness is maximized by maximizing the number of links between strongly connected components of the graph representing the underlying system connectivity.
+We also demonstrate that permutation of strongly connected components is a fundamental symmetry of dynamical robustness, which applies to networks of any number of components.
+The classification of dynamical robustness based upon a purely topological property provides a fundamental organizing principle that can be used in the context of experimental validation to select among models that break or preserve network hierarchy.
+This result contributes to an explanation for the observation of hierarchical modularity in biological networks at all scales~\cite{Zhao2006,Ravasz2002,Bhardwaj2010,Colm,Corominas-Murtra2013}.
+\end{abstract}
+
+\maketitle
+
+\tableofcontents
 
@@
 
-\section{Introduction}
+% \section{Introduction}
 \input{tex/introduction}
 
-\section{Results}
+% \section{Results}
 \input{tex/results}
 
-\section{Discussion}
+% \section{Discussion}
 \input{tex/discussion}
 
-\section{Acknowlegements}  The authors would like to thank Jay Sulzberger and Ximo Pechuan for useful suggestions and discussions.
+\section{Acknowlegements}
+\input{tex/acknowledgements}
+% \pagebreak
+{\footnotesize
+% \bibliographystyle{naturemag}
+% \bibliography{bib/books,bib/papers}
+\putbib[bib/books,bib/papers]
+}
+
+% \end{multicols}
 
 \pagebreak
-
-\bibliographystyle{pnas}
-\bibliography{bib/books,bib/papers}
+\FloatBarrier
+\section{Figure Legends}
+\input{tex/figures}
+\end{bibunit}
 
@@
 
-\section{Figure Legends}
-\input{tex/figures}
-
-\pagebreak
-\FloatBarrier
+\begin{bibunit}[naturemag]
 
@@
 \input{tex/supplement}
+
+{\footnotesize
+% \bibliographystyle{naturemag}
+% \bibliography{bib/books,bib/papers}
+\putbib[bib/books,bib/papers]
+}
 \pagebreak
@@
 % \input{tex/extra}
-
+\end{bibunit}
 \end{document}

### 2014-11-25 tex/figures.tex
add LS lines and correlation coefficients to fig 3 A and B
move figures into main text and switch bibliography format

@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.9\columnwidth]{fig/biomodelexamples.pdf}
-\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.9\columnwidth]{fig/biomodelexamples.pdf}
+% \caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
 The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
 The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
-\label{fig:biomodelexamples}
-\end{figure}
+% \label{fig:biomodelexamples}
+% \end{figure}
 
-\pagebreak
+% \pagebreak
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=0.9\columnwidth]{fig/modsccsym.pdf}
-\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.9\columnwidth]{fig/modsccsym.pdf}
+% \caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
 (B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
@@
 All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
-\label{fig:modsccsym}
-\end{figure}
+% \label{fig:modsccsym}
+% \end{figure}
 
-\pagebreak
+% \pagebreak
 
@@
 
-\begin{figure}[!ht]
-\centering
-\noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.
-The red colored points represent averages for each connectivity.
+% \begin{figure}[!ht]
+% \centering
+% \noindent\includegraphics[width=1.0\columnwidth]{fig/combinedfigs.pdf}
+% \caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.
+The red line represents a best fit in the least-squares sense with Pearson product-moment correlation coefficient $r=0.29$.
 The lowest and highest robustness network architectures are labelled.
@@
 (B) Robustness versus hierarchy.
+Correlation coefficient $r=0.67$.
 (C) Number of cycles and (D) hierarchy vs connectivity and robustness.
 The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
-}
-\label{fig:combined}
-\end{figure}
+% }
+% \label{fig:combined}
+% \end{figure}
 

### 2014-11-25 tex/introduction.tex
move figures into main text and switch bibliography format

@@
 However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed.
+
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
+\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
+The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
+The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
+\label{fig:biomodelexamples}
+\end{figure*}

### 2014-11-25 tex/supplement.tex
move figures into main text and switch bibliography format

@@
 To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution yielding
+\begin{widetext}
 $$
@@
 $$
+\end{widetext}
 Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
+\begin{widetext}
 \begin{equation}\label{eq:condprobgen}
@@
 \end{equation}
+\end{widetext}
 where we made the abbreviations
@@
 The result of this operation corresponds to averaging our quantity over subsets of links:
+\begin{widetext}
 \begin{equation}\label{eq:robavgoverlinks}
@@
 \end{equation}
+\end{widetext}
 We are fundamentally interested in the relationship between \ref{eq:condprobgen} and \ref{eq:robavgoverlinks} and topological properties of systems that derive from their associated graphs.
@@
 One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.
+
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/modsccsym.pdf}
+\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
+(B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
+The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
+We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
+The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e.
+SCCs with more than one variable) and hierarchy.
+(C) Symmetries of the $\hier$ transformation between graphs and SCCs.
+The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link.
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
+\label{fig:modsccsym}
+\end{figure*}
 
@@
 Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprobgen}
+\begin{widetext}
 \begin{align*}
@@
 \end{align*}
+\end{widetext}
 The analogous equation for resampling $a$ is
+\begin{widetext}
 \begin{align*}
@@
 \end{align*}
+\end{widetext}
 Using this approach the probability of stability and of robustness for all two variable systems is given in \ref{tab:structstabmat}.
@@
 This estimate is characterized by the variance of the posterior Beta distribution
+\begin{widetext}
 $$
@@
 $$
+\end{widetext}
 Since for the chosen prior $a=b=1 \ll N$ this simplifies to
@@
 
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/combinedfigs.pdf}
+\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.
+The red line represents a best fit in the least-squares sense with Pearson product-moment correlation coefficient $r=0.29$.
+The lowest and highest robustness network architectures are labelled.
+Other network architectures are shown in \ref{tab:structstabmat3}.
+(B) Robustness versus hierarchy.
+Correlation coefficient $r=0.67$.
+(C) Number of cycles and (D) hierarchy vs connectivity and robustness.
+The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
+}
+\label{fig:combined}
+\end{figure*}
+
 \subsection{Component decomposition, hierarchy, and the symmetries of robustness}
@@
 probabilities:
+\begin{widetext}
 \begin{align}\label{eq:sccrobustness}
@@
 \end{align}
+\end{widetext}
 For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.
@@
 of the connected components:
+\begin{widetext}
 \begin{align} \label{eq:sccmaxrobustness}
@@
 \end{align}
+\end{widetext}
 Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
@@
 resample $m$ nodes.  Define
+\begin{widetext}
 \begin{equation*}
@@
 \end{equation*}
+\end{widetext}
 Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
@@
 weighted average becomes
+\begin{widetext}
 \begin{equation}
@@
 \end{equation}
+\end{widetext}
 

### 2014-11-26 paper.tex
move supplement to main text
move figures into main text and switch bibliography format
update text

--- 2014-10-13 paper.tex
@@
 
-\title{Ensemble Stability of Biological Networks}
-
 \pagenumbering{arabic}
 
-% Title must be 150 characters or less
-\begin{center}
-{\Large
-\textbf{Hierarchical Network Structure Promotes Dynamical Robustness}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\[.5cm]
-Cameron Smith$^{1}$,
+\title{\bf Hierarchical Network Structure Promotes Dynamical Robustness}
+
+\author{Cameron Smith$^{1}$,
 Raymond S. Puzio$^{1}$,
-Aviv Bergman$^{1,2,3,4}$,
-\\[.5cm]
-$^1$Department of Systems and Computational Biology,\\
+Aviv Bergman$^{1,2,3,4,*}$}
+
+\affiliation{$^1$Department of Systems and Computational Biology,\\
 $^2$Dominick P. Purpura Department of Neuroscience,\\
@@
 $^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
-\\[.5cm]
-% $\ast$To whom correspondence should be addressed; E-mail: aviv@einstein.yu.edu.
-\end{center}
+\\
+$*$to whom correspondence should be addressed: aviv@einstein.yu.edu}
 
-{\begin{quote} \bf
+\date{\today}
+
 \input{tex/abstract-cited}
-\end{quote}}
+
+\maketitle
 
@@
 
-\section{Acknowlegements}  The authors would like to thank Jay Sulzberger and Ximo Pechuan for useful suggestions and discussions.
+\section{Methods}
+{\footnotesize
+\input{tex/methods}
+}
 
-\pagebreak
+\section{Acknowlegements}
+\input{tex/acknowledgements}
 
-\bibliographystyle{pnas}
+\bibliographystyle{unsrtnat}
 \bibliography{bib/books,bib/papers}
 
-\pagebreak
-\FloatBarrier
-
-\section{Figure Legends}
-\input{tex/figures}
+\onecolumngrid
 
@@
 \setcounter{secnumdepth}{4}
-\section{Supplementary Material}
-
-\input{tex/supplement}
-\pagebreak
+\section{Supporting Information}
 
@@
 % \section{Tables}
-% \todo{Fill out notation table}
-% \input{tex/tablenotation}
 \input{tex/table2x2stab}
@@
 
-% \input{tex/tablewithgraphs}
-% \input{tex/extra}
-
 \end{document}
@@
 %------------------------------------------------------------------------------
-% End of journal.tex
+% End of paper.tex
 %------------------------------------------------------------------------------

### 2014-11-26 tex/abstract-cited.tex
update text

@@
-%!TEX root = ../paper-abstract.tex
+%!TEX root = ../paper.tex
+\begin{abstract}
 The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components~\cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
 When the details of these interactions are unknown for biological systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by evolutionary constraints that may apply to all such systems~\cite{Gardner1970,May1972,Cohen1984,May1972a,Radius2014}.
-One such constraint is that of dynamical robustness, the probability of a stable system to remain stable under perturbation~\cite{WADDINGTON1942a,Wagner1997,Rutherford1998,VanNimwegen1999,Siegal2002,Bergman2003,Ciliberti2007b,Ciliberti2007,Draghi2010,Wagner2013}.
+One such constraint is that of dynamical robustness, the probability of a stable system to remain stable under perturbations to the strength of interactions among its components~\cite{WADDINGTON1942a,Wagner1997,Rutherford1998,VanNimwegen1999,Siegal2002,Bergman2003,Ciliberti2007b,Ciliberti2007,Draghi2010,Wagner2013}.
 Despite previous investigations, the relationship between dynamical robustness---an important functional characteristic of many biological systems---and network structure is poorly understood.
@@
 In particular, we determine the probability distribution of robustness over system connectivity and show that robustness is maximized by maximizing the number of links between strongly connected components of the graph representing the underlying system connectivity.
-We also demonstrate that permutation of strongly connected components is a fundamental symmetry of dynamical robustness, which applies to networks of any number of components.
+We demonstrate that this can be understood in terms of the fact that permutation of strongly connected components is a fundamental symmetry of dynamical robustness, which applies to networks of any number of components and is independent of the distribution from which the strengths of interconnection among components are sampled.
 The classification of dynamical robustness based upon a purely topological property provides a fundamental organizing principle that can be used in the context of experimental validation to select among models that break or preserve network hierarchy.
 This result contributes to an explanation for the observation of hierarchical modularity in biological networks at all scales~\cite{Zhao2006,Ravasz2002,Bhardwaj2010,Colm,Corominas-Murtra2013}.
+\end{abstract}

### 2014-11-26 tex/introduction.tex
update text

@@
 
-Recent work has demonstrated that as a result of inhomogeneity in the geometry of the sensitivity of the dynamics over the parameter space for such models, this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
+Recent work has demonstrated that as a result of inhomogeneity in the sensitivity of the dynamics over the parameter space for such models, this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
 In addition, there is often uncertainty about the very structure of such biological networks.
@@
 The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the biological networks represented in \ref{fig:biomodelexamples}.
-May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
+May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
 
-% The generic applicability of gaining a better understanding of the class of models investigated by May demonstrates the unequivocal value of deeper investigation.
-However, work attempting to continue the development of the so-called May-Wigner stability theorem revealed that May's conjectured stability criterion was not as easy to demonstrate as was initially believed.
+Some conclusions derived from this approach may be sensitive to the probability distribution from which the strengths of interconnection are sampled, whereas others are independent of the specific form of this distribution.
+In the former case, discovering the constraints that contribute to determining the functional form of the distribution can be a matter for empirical investigation and results that are themselves functions of this distribution can be reinterpreted in light of ongoing updates from relevant empirical observations.
+On the other hand, results that are independent of the specific form of this distribution can be expected to hold in any case and a specific distribution may simply be chosen for the purposes of exemplifying the result.
 
@@
 \end{figure*}
+
+Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
+Robustness is of interest to biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
+Intuitively, robustness is the probability of stability to perturbation in the system structure or its parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
+We demonstrate that systems with the most hierarchical network topology exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.

### 2014-11-26 tex/methods.tex
update text

--- 2014-10-17 tex/supplement.tex
@@
 %!TEX root = ../paper.tex
-\subsection{Dynamical systems on biological networks}
-In the construction of a class of potential models for biological systems at any level of the biological hierarchy from metabolic to ecosystem-level networks, it is common to first attempt to define a collection of observable phenomena of interest and determine a domain (such as binary numbers, integers, or real numbers) in which each observable can be quantified.
-Next, it is necessary to establish the interdependencies among system components.
-The specific manner in which the components depend upon one another must be clarified, which is often done by determining a particular system of mathematical functions that represents a hypothesis about how the quantified observables evolve in time.
-To the degree to which there is uncertainty about the interactions among system components, parameters are introduced to broaden the class of models under consideration.
-Finally, whatever model class remains may be compared to empirical observations to determine how capable the model is of representing the phenomena of interest.
-
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
-For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
-If the system under consideration has the graph given in \reffigexamplesystemmodules
-% \begin{center}
-% \input{tex/examplehamdyngraph}
-% \end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-having connectivity $\connectivity$ equal to the number of edges of the graph, in this case $\connectivity = 3$.
-For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}.
-   \right.
-\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is constrained to be a function of $x_2$ and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2)\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2)
-\end{align*}
-
-In the more general case of a system with $n$ components we would have an $n$-dimensional vector of observables
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-
-In order to accommodate the possibility of uncertainty in our
-modeling, we will generalize our notion of a system on a network to
-that of a system with random parameters.  Again, this will involve
-three steps. First, we provide a measure space $S$
-which represents the values over which our parameters can vary. Next,
-instead of a single vector field, we consider a family of vector
-fields parameterized by this space.  As before, for
-each point $p \in S$, the vector field $V(p)$ must be consistent with
-the system graph. Finally, we select a probability measure, $\mu$, which represents our understanding of which values of the parameters
-are likely. In accord with Bayesian statistics, we might revise this
-distribution as data comes in or use it to estimate parameters and error
-bars from experimental results.
-
-Having done this, we are now in a position to do a probabilistic analysis
-of our system and its properties.  Given some quantity $q$ characterizing
-our system, this quantity becomes a random variable which may be discrete or
-continuous depending on the quantity under consideration.  For example, if
-the quantity is the time it takes for a particle to travel between two points,
-we have a real-valued random variable; if the quantity is the number of fixed
-points, we have an integer-valued random variable; and if the quantity is
-whether or not the system posesses a limit cycle, we have a binary random
-variable.
-
-A quality which is of interest to us is robustness or structural stability \cite{Smale1967}, which requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters.
-We will quantify this as the probability that, if some property holds for a set of parameters, it will continue to hold if we make a random perturbation about those parameter values.
-To do so, we will introduce, in addition to the probability distribution $\mu$ described above, a family of probability distributions $\mu'$ such that $\mu'(x,y)$ encodes the conditional probability that a system with a parameter value $x$ will have its parameter values changed to $y$ under a perturbation.
-For our biological models, we will be interested in large perturbations rather than the small or infinitesimal perturbations usually considered in the theory of dynamical systems.
-Then, given a binary random variable $q$, we define its robustness as the following conditional probability where $\mathbf{1}_q(x)$ is the standard indicator function equal to $1$ when $x$ satisfies $q$ and $0$ otherwise:
-\begin{align}\label{eq:robustness}
-  R (q,\mu') =
-  \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
-  {\int_S d\mu(x) \mathbf{1}_q(x)}
-\end{align}
-
-\subsection{Stability analysis of biological networks}
-The class of dynamical quantities on which we shall focus in this
-investigation involve stability of equilibria.  Suppose that $\vec x$
-is a point on our phase space which depends upon the parameters.  Then
-we may take $q$ to be a binary random variable which describes whether
-or not $\vec x$ is a fixed point of the system (i.e. $q$ is true if and only if
-$F_i(\vec{x})=0$ for all $i$) and perform a probabilistic analysis
-of the sort discussed above.
-
-Furthermore, if it turns out that $\vec x$ is indeed a fixed point, we
-may proceed to ask whether it is a dynamically stable fixed point.
-Intuitively, dynamic stability means that, if one chooses the initial
-conditions sufficiently close to the fixeed point, the solution will
-stay close to the fixed point.  Physically, this is important because,
-if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
-observing the solution ${\vec x}(t) = {\vec x}^0$.
-
-To determine stability, we linearize the equations of motion about the
-fixed point $\vec{x}^0$:
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.
-
-Even though the determination of stability only requires examining
-linearized equations of motion, it is worth noting that knowing the
-stability of a fixed point can yield information about the non-linear
-dynamics of a system.  For instance, in two dimensions,
-Poincare-Bendixson theory implies that every limit cycle must encircle
-at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
-conflicting populations), the stability of the fixed point will determine whether the limit cycle is attractive or repulsive.
-
-%% To determine whether a randomly selected dynamical system evaluated at
-%% a random critical point, given by a matrix such as $A$, is stable, it
-%% is sufficient to check the above condition. Integrating over the
-%% region of the parameter space defining $A$ enables the determination
-%% of the probability of stability to perturbations in state $\vec{x}$
-%% over an ensemble of dynamical systems.
-
-% Another form of stability, referred to as structural
-% stability \cite{Smale1967}, requires the determination of whether or
-% not a given dynamical system that is determined to be stable remains
-% stable under a perturbation to one of its defining parameters. This
-% can be formalized as the conditional probability distribution
-% $$
-% P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-% $$
-% where $A'$ represents a matrix derived from some perturbation
-% applied to the matrix $A$ given above. In the simple case where
-% $A \in \mathbb{R}^{n \times n}$ the perturbation which involves
-% sampling uniformly over the $a_{ij}$ defining $A$ the value of a
-% particular $a_{ij}$ from a given probability distribution.
-
-Suppose that our Jacobian $A$ is an $n \times n$ matrix with real coefficients, connectivity $\connectivity$, and denote the proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.
-Then we may compute the robustness of this proposition using \ref{eq:robustness} once we have specified the probability densities $\mu$ and $\mu'$.
-In general, these will depend upon the parameters of the non-linear system.
-For the purpose of the current investigation, we shall make the ansatz that these are such that the distribution of entries of $A$ is (at least approximately) the uniform distribution $\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin.
-To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution yielding
-$$
-\mu_{k_1,\ldots,k_m}(x,y) = \prod_{y \notin \{x_1, \ldots, x_m\} } \delta(x_j-y_j) \prod_{y \notin \{x_1,\cdots,x_m\}} \mathbf{1}_{[-1,1]} (y_j) \mathbf{1}_{[-1,1]} (x_j)
-$$
-Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
-\begin{equation}\label{eq:condprobgen}
- R (\mathrm{stab}, \mu'_{k_1, \ldots, k_m}) =
-  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
-    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
-  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)},
-\end{equation}
-where we made the abbreviations
-\begin{align*}
-\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity) \\
-\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d)
-\end{align*}
-i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.
-
-While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.
-The result of this operation corresponds to averaging our quantity over subsets of links:
-\begin{equation}\label{eq:robavgoverlinks}
-\langle R (\hbox{stab}, \mu') \rangle_m =
-\frac{1}{\binom{\connectivity}{m}}
-\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
-R (\hbox{stab}, \mu'_{k_1, \ldots, k_m})
-\end{equation}
-We are fundamentally interested in the relationship between \ref{eq:condprobgen} and \ref{eq:robavgoverlinks} and topological properties of systems that derive from their associated graphs.
-In order to investigate this further, we must characterize properties of system graphs in order to evaluate correlations of those properties with dynamical robustness.
-
-\subsection{Network hierarchy and strongly connected components}
-
-Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems \reffigexamplesystemmodules.
-One such kind of system composition and decomposition is given by the notion of open systems, which are distinguished by having some of the variables specified as control variables whose values are given as autonomous functions of time rather than determined by the dynamics via intrinsic interactions \cite{Vagner2014}.
-Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
-Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
-However, this decomposition is trivial since it is equivalent to the underlying system graph and what we instead want is an intermediate decomposition into relatively self-contained modules.
-One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.
-
-A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three variable systems are outlined in \reffigscc along with their adjacency matrices and hierarchy diagrams (to be defined below).
-The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
-Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
-Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
-Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
-This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
-
-Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$.
-Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
-There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
-
-One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
-a strongly connected graph $C_n$.
-To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
-The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
-
-This map $\hier$ is many-to-one and so there is a large class of
-operations which leaves $\hier(G)$ invariant for a given graph $G$.
-For instance, we may interchange the positions of the strongly
-connected components relative to each other \reffighiertransformations a.
-Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations b or between components, or even add or delete links \reffighiertransformations c.
-As we shall see later, some of these operations leave invariant key dynamical quantities, so we may regard them as a symmetry groupoid of our system.
-
-The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
-The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc (top panel).
-This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
-This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical.
-In this work we take this distance to be the definition of hierarchy.
-
-\subsection{Stability and robustness analysis of particular system ensembles}
-For systems having two components, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}.
-For those having three components, we can estimate these same quantities using Monte Carlo simulations.
-Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis.
-For two-variable systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant. Suppose we have a stable matrix
-$$
-\begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},\mu'_k)$ where $k=1 \ldots 4$.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $ R(\mathrm{stab},\mu'_1)$ and $R(\mathrm{stab},\mu'_2)$.
-Suppose that we resample $b$ to compute $R(\mathrm{stab},\mu'_2)$.
-% \begin{strip}
-% \begin{align}\label{eq:condprob}
-% P\left(\begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)
-% & = \frac{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}.
-% \end{align}
-The denominator of \ref{eq:condprobgen} in this case is given by
-\begin{align*}
-P\left(\textrm{stab } \left( \begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix} \right) \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprobgen}
-\begin{align*}
-P\left(\textrm{stab} \left( \begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
-a & b' \\
-d & c
-\end{bmatrix} \right) \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-The analogous equation for resampling $a$ is
-\begin{align*}
-P\left(\textrm{stab} \left( \begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
-a' & b \\
-d & c
-\end{bmatrix} \right) \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-
-Using this approach the probability of stability and of robustness for all two variable systems is given in \ref{tab:structstabmat}.
-
-The analogous results for all three variable systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \reffigrobustconnect.
-This process is associated with some error relative to the exact integration described above.
+\subsection{Monte carlo integration}
 If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
@@
 yielding the error estimate given by the associated standard deviation.
-In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
-
-It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
-For small system sizes such as the two and three variable systems, studied here the situation is not so clear cut.
-For two variable systems, system stability is constant across the entire range of connectivities.
-For three variable systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
-
-The relationship between connectivity and robustness for two variable systems is shown in \ref{tab:structstabmat} and likewise for three variable systems in \ref{tab:structstabmat3} and \reffigrobustconnect.
-If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \reffigrobustconnect.
-% \subsection{Cycle number is inversely correlated with robustness}
-\ref{fig:cycle3x3} shows the robustness for all three variable systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-There appears to be a weak negative correlation between robustness and the number of simple cycles.
-
-% \subsection{Cycle number and connectivity classify robust systems}
-The combination of connectivity and cycle number as shown in \reffigconnectcycle3D3x3 provides a better classification of the dependence of robustness upon network topology.
-Here the robustness of three variable systems with a given number of cycles, increases monotonically with connectivity.
-The network with the highest robustness for three variable systems is that of \reffigscc (top panel).
-This network is the most hierarchical of all three variable systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \reffigscc (top panel) has a block triangular structure.
-
-This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of dynamical robustness.
-\reffigrobusthierarchy shows dynamical robustness as a function of hierarchy, where the most hierarchical system has a hierarchy of 0.
-There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\reffigconnectdist3D3x3 shows dynamical robustness as a function of both hierarchy and connectivity.
-The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \reffigconnectcycle3D3x3.
-
-\subsection{Component decomposition, hierarchy, and the symmetries of robustness}
-
-The fact that the most hierarchical of networks is the most robust in the case of three variable systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
-By combining our decomposition of the graph with our analysis of the robustness of random ensembles of dynamical systems, we will deduce a general result which relates robustness to hierarchy.
-
-Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomials factors as the product of the charactericstic polynomials of its diagonal blocks.
-Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.
-Note that this condition does not depend upon the entries off the diagonal (which correspond to links between strongly connected components) and does not depend upon what order the components appear.
-
-Using these observations, we may express the robustness of a graph in
-terms of the robustnesses of its strongly connected components.  If
-the components $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
-$\connectivity_i$, and there are $\ell_{ij}$ links between components
-$i$ and $j$, then there are a total of
-$\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i$
-links in $G$.  If we resample a randomly chosen link in a component
-$C_i$, then the probability that the system will remain stable after
-resampling is $\langle R(\hbox{stab} (C_i), \mu') \rangle_1$.  If we resample a link
-between components $C_i$ and $C_j$, the system will remain stable with
-probability 1.  Hence, the probability that the system will remain
-stable upon resampling a random link is the weighted average of these
-probabilities:
-\begin{align}\label{eq:sccrobustness}
-\langle R(\hbox{stab} (G), \mu') \rangle_1 =
-\frac{\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i \langle R(\textrm{stab}(C_i),\mu') \rangle_1}
-     {\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i}.
-\end{align}
-For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.
-From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.
-The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
-Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is
-\[
-\langle R(\textrm{stab}(G),\mu') \rangle_1 = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
-\]
-which agrees with the value computed in \ref{tab:structstabmat} up to
-sampling error.
-
-As can be seen from \ref{eq:sccrobustness}, all that is required to fix robustenss is a set of SCC sizes and the number of links between them.
-This demonstrates that robustness is symmetric under the transformations shown in \reffighiertransformations a and b.
-Since transformations of the type \reffighiertransformations c change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}.
-Nevertheless, a considerable degree of symmetry remains.
-For example, \ref{fig:robustnesssymmetries} shows the symmetry groupoid of robustness for the case in which we fix connected component sizes $\{2,1,1\}$ with a total of $3$ links between them.
-
-Next, we derive an upper bound on the robustness of any graph
-constructed from a given set of connected components.  Since
-$\langle R(\textrm{stab}(C_i), \mu') \rangle_1 \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
-increase $\langle R(\textrm{stab}(G),\mu') \rangle_1$.  Given two connected components $C_i$ and $C_j$ with
-$v_i$ and $v_j$ nodes respectively, we have a maximum of $n_i n_j$
-links going from $C_i$ to $C_j$.  Hence, $\ell_{ij} \le v_i v_j$, so,
-we conclude that
-\begin{align}
-\langle R(\textrm{stab}(G),\mu') \rangle_1 \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i \langle R(C_i,\mu') \rangle_1}
-              {\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i},
-\end{align}
-
-Since every acyclic digraph can be embedded into a totally ordered
-set, we may assume without loss of generality that our components have
-been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i <
-j$.  Hence, after substituting into our expression and simplifying
-using the identity
-$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
-v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n}
-v_i^2,$$
-we arrive at the following upper bound on robustness in terms
-of the connected components:
-\begin{align} \label{eq:sccmaxrobustness}
-R_{\mathrm{max}}(C_1, \ldots C_n) =
-\frac{\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i \langle R(\hbox{stab}(C_i),\mu') \rangle_1}
-     {\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i}.
-\end{align}
-Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
-is the graph on $n$ nodes with a link from node $i$ to node $j$
-whenever $i < j$.  Then, by the construction described in the previous
-section on network hierachy, we have a graph $G_{\mathrm{max}}$ such
-that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and
-$\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula,
-$\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
-
-This argument also works when we resample more than one entry,
-although the notation becomes more complicated.  Suppose that we
-resample $m$ nodes.  Define
-\begin{equation*}
-M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
-m = \sum_{i=0}^n m_i \quad\&\quad
-m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
-\forall i \in \{1, \ldots, n\} \, m_i < \connectivity_i \right\}.
-\end{equation*}
-Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
-m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
-$m_0$ links between strongly connected components.  Hence, our
-weighted average becomes
-\begin{equation}
-\langle R(\hbox{stab} (G), \mu') \rangle_m =
-\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
-      {m \choose m_0, m_1, \ldots. m_n}
-      \left(m_0 + \sum\limits_{i=1}^n m_i \langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \right)}
-     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
-      {m \choose m_0, m_1, \ldots. m_n} m}
-\end{equation}
-
-As before, since $\langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \le 1$, we may
-increase $\langle R(\hbox{stab} (G), \mu') \rangle_m$ by increasing the maximum
-possible value of $m_0$ while keeping the strongly connected
-components the same.  Again, if we fix $\hier(G)$, the maximum
-possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
-whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
-v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
-G_{tot}$.  Hence, we conclude that $\langle R(\hbox{stab} (G), \mu') \rangle_m \le
-\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_m$.
-
-The preceding argument demonstrates the manner in which robustness is
-maximized by maximizing the number of edges \emph{between} the strongly
-connected components of the graph underlying system interdependencies.
-And, the graph that has the latter property is that associated to the
-ordering, $G_{\mathrm{tot}}$, which also possess the highest degree of
-hierarchy. The fact that robustness is equivalent for any configuration
-of strongly connected components with an equivalent number of links
-between them demonstrates that it is invariant to permutations of
-strongly connected components.
-% This result does not depend at all uponany particular details such as the probability distribution from which the component interaction strengths are sampled because the arguments presented here are purely topological in nature.
-The result that dynamical robustness is correlated with system hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles primarily studied here.

### 2014-11-26 tex/supplement.tex
update text

@@
 %!TEX root = ../paper.tex
+\subsection{Overview}
+The dynamical model given in terms of a system of differential equations for any network can be represented in terms of an interaction graph (\ref{fig:biomodelexamples} top row).
+These interaction graphs can be viewed as deriving from the combination of system modules that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).
+Symmetries are characterized by the ability to interchange these modules or their connectivity without changing some property of the system.
+
+The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see Supplementary Information) (\reffigscc).
+This map from the interaction graph of a biological network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
+These three symmetries represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated \reffigscc.
+Two of these three are also symmetries with respect to dynamical robustness.
+\ref{fig:robustnesssymmetries} shows an example of these symmetries applied to a specific interaction graph.
+
+We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a biological network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
+This expression is fully developed in \ref{eq:sccrobustness}, but can be schematized as in \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
+\begin{equation}\label{eq:robschematic}
+R_{tot} = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
+\end{equation}
+Examining this expression noting that $R_i$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R_{tot}$.
+This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
+This analytical result predicts that any biological network whose associated dynamical system has the interaction graph \reffigscc $\,$ top will be more robust than those associated to any of the other interaction graphs in \reffigscc.
+Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.
+The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
+
+To test this prediction, we computed the probability distribution of stability and dynamical robustness relative to biological network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
+For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).
+Accounting for the number of cycles in a network architecture reveals a strong correlation between robustness and connectivity that was hidden when networks with any number of cycles were considered together (\reffigconnectcycle3D3x3).
+While the most hierarchical network architecture will always lack cycles altogether, cycle number alone is clearly insufficient to account for robustness as the members of each class span nearly the entire range of possible robustness values.
+Consistent with our analysis of the symmetries of robustness, we found that the most hierarchical network architecture is the most robust (\reffigrobusthierarchy).
+Moreover, if we consider hierarchy partitioned by connectivity, we find that there is a monotonic increase in robustness following any line of increasing hierarchy in \reffigconnectdist3D3x3.
+
 \subsection{Dynamical systems on biological networks}
@@
 
-A quality which is of interest to us is robustness or structural stability \cite{Smale1967}, which requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters.
+A quality which is of interest to us is robustness whose evaluation requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters \footnote{We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the components as determined by the parameters and not only to perturbations of the state vector at a given point in time.
+For example, when the network of components and interactions corresponds to a gene-regulatory network, then mutation is one mechanism by which such perturbations may arise}.
 We will quantify this as the probability that, if some property holds for a set of parameters, it will continue to hold if we make a random perturbation about those parameter values.
@@
 \end{widetext}
-where we made the abbreviations
+where we use the abbreviations
 \begin{align*}
@@
 Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations$b$ or between components, or even add or delete links \reffighiertransformations$c$.
-As we shall see later, some of these operations leave invariant key dynamical quantities, so we may regard them as a symmetry groupoid of our system.
+As we shall see later, some of these operations leave invariant key dynamical quantities, in our case dynamical robustness, so we may regard them as a symmetry groupoid of our system with respect to the robustness property.
 
@@
 This process is associated with some error relative to the exact integration described above.
-If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
-We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}.
-The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
-$$
-\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
-$$
-where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
-$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
-which corresponds in this case to the maximum likelihood estimate
-$$
-\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
-$$
-This estimate is characterized by the variance of the posterior Beta distribution
-\begin{widetext}
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
-$$
-\end{widetext}
-Since for the chosen prior $a=b=1 \ll N$ this simplifies to
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
-$$
-yielding the error estimate given by the associated standard deviation.
-In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
+In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$ (see \emph{Methods}).
 

### 2014-11-26 tex/abstract-cited.tex
minor edits

@@
 Despite previous investigations, the relationship between dynamical robustness---an important functional characteristic of many biological systems---and network structure is poorly understood.
-Here we analyze the stability and robustness of a large class of dynamical systems and demonstrate that the most hierarchical network structures are the most robust.
+Here we analyze the stability and robustness of a large class of dynamical systems and demonstrate that the most hierarchical network structures, those equivalent to the total ordering, are the most robust.
 In particular, we determine the probability distribution of robustness over system connectivity and show that robustness is maximized by maximizing the number of links between strongly connected components of the graph representing the underlying system connectivity.

### 2014-11-26 tex/introduction.tex
minor edits

@@
 Robustness is of interest to biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
-Intuitively, robustness is the probability of stability to perturbation in the system structure or its parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
+Intuitively, robustness is the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
+The maximally hierarchical network is considered to be the graph associated to the total ordering, which is the maximally connected directed acyclic graph for a given system size.
+Here, hierarchy is then defined as the edit distance from this maximally hierarchical network.
 We demonstrate that systems with the most hierarchical network topology exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
+Moreover these results hold for networks of arbitrary size and independent of the probability distribution from which the strengths of interaction are sampled.

### 2014-11-27 tex/introduction.tex
fix reference alignment and footnoting

@@
 Intuitively, robustness is the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
-The maximally hierarchical network is considered to be the graph associated to the total ordering, which is the maximally connected directed acyclic graph for a given system size.
-Here, hierarchy is then defined as the edit distance from this maximally hierarchical network.
-We demonstrate that systems with the most hierarchical network topology exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
+The maximally hierarchical network is considered to be the graph associated to the total ordering, which is the complete directed acyclic graph for a given system size \cite{Cormen2009}.
+Here, hierarchy is then defined as the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
+We demonstrate that systems with the most hierarchical network topology according to this definition exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
 Moreover these results hold for networks of arbitrary size and independent of the probability distribution from which the strengths of interaction are sampled.

### 2014-11-28 tex/introduction.tex
minor changes

@@
 
-Recent work has demonstrated that as a result of inhomogeneity in the sensitivity of the dynamics over the parameter space for such models, this approach allows for a large variety of models to fit the data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
+Recent work has demonstrated that as a result of the existence of largely insensitive directions within the parameter space for such models, the approach outlined above often allows for a large variety of models to fit equivalent data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
 In addition, there is often uncertainty about the very structure of such biological networks.
@@
 Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
-Robustness is of interest to biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
+Robustness is of interest in biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
 Intuitively, robustness is the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.

### 2014-11-28 tex/discussion.tex
minor edits

@@
 % Our development suggests two predictions that could be used to test the theory described here. One is that,
-Our theory predicts that any ensemble of systems where robustness has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
+Our theory predicts that, in general, an ensemble of systems where robustness has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
 For example, at the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.

### 2014-11-28 tex/introduction.tex
minor edits

@@
 Recent work has demonstrated that as a result of the existence of largely insensitive directions within the parameter space for such models, the approach outlined above often allows for a large variety of models to fit equivalent data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
-In addition, there is often uncertainty about the very structure of such biological networks.
+In addition, there is often uncertainty about the very structure of such networks.
 In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
@@
 Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
-The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the biological networks represented in \ref{fig:biomodelexamples}.
+The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the networks represented in \ref{fig:biomodelexamples}.
 May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.

### 2014-12-01 tex/discussion.tex
minor edits

--- 2014-10-17 tex/discussion.tex
@@
 % Our development suggests two predictions that could be used to test the theory described here. One is that,
-Our theory predicts that any ensemble of systems where robustness has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
+Our theory predicts that, in general, an ensemble of systems where robustness has been the predominant object of selection and has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
 For example, at the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
@@
 Equally fundamental is the ability of systems to persist over long periods of time, which is dependent upon their dynamical robustness.
-Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist in the evolutionary process.
+Here we have demonstrated a structure-function relationship wherein biological networks that are more hierarchical are more robust and thus more likely to persist when this feature is the dominant object of selection in the evolutionary process.
 %Moreover, this result is formulated at a level of abstraction that allows its conclusions to apply independently to each of these levels.

### 2015-01-14 tex/abstract.tex
revise
shorten abstract

@@
 %!TEX root = ../paper.tex
-The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components.
+% The analysis of dynamical systems that attempts to model chemical reaction, gene-regulatory, population, and ecosystem networks all rely on models having interacting components.
 When the details of these interactions are unknown for systems of interest, one effective approach is to study the dynamical properties of an ensemble of models determined by constraints that can be considered to apply to all such systems.
@@
 We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.
-We also demonstrate that permutation of strongly connected components is a fundamental symmetry of system robustness.
+We also demonstrate that permutation of strongly connected components is a fundamental symmetry of dynamical robustness.
 This results in the classification of the dynamical robustness of biological networks based upon a purely topological property.
+
+\begin{abstract}
+Here we analyze the stability and robustness of a large class of dynamical systems.
+In particular, we precisely determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.
+We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.
+We also demonstrate that permutation of strongly connected components is a fundamental symmetry of dynamical robustness.
+This results in the classification of the dynamical robustness of biological networks based upon a purely topological property.
+\end{abstract}

### 2015-01-14 tex/figures.tex
redistribute text between main text and supplement

@@
 %!TEX root = ../paper.tex
+
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
+\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
+The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
+The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
+\label{fig:biomodelexamples}
+\end{figure*}
+
+\pagebreak
+
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/modsccsym.pdf}
+\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
+(B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
+The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
+We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
+The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e.
+SCCs with more than one variable) and hierarchy.
+(C) Symmetries of the $\hier$ transformation between graphs and SCCs.
+The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link.
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
+\label{fig:modsccsym}
+\end{figure*}
+
+\pagebreak
+
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/combinedfigs.pdf}
+\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.
+The red line represents a best fit in the least-squares sense with Pearson product-moment correlation coefficient $r=0.29$.
+The lowest and highest robustness network architectures are labelled.
+Other network architectures are shown in \ref{tab:structstabmat3}.
+(B) Robustness versus hierarchy.
+Correlation coefficient $r=0.67$.
+(C) Number of cycles and (D) hierarchy vs connectivity and robustness.
+The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
+}
+\label{fig:combined}
+\end{figure*}
 

### 2015-01-14 tex/introduction.tex
redistribute text between main text and supplement

@@
 
-\begin{figure*}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
-\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
+% \begin{figure*}[!ht]
+% \centering
+% \noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
+% \caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
 The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
 The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
-\label{fig:biomodelexamples}
-\end{figure*}
+% \label{fig:biomodelexamples}
+% \end{figure*}
 

### 2015-01-14 tex/overview.tex
redistribute text between main text and supplement

--- 2014-11-26 tex/supplement.tex
@@
 %!TEX root = ../paper.tex
-\subsection{Overview}
+
+% \subsection{Overview}
 The dynamical model given in terms of a system of differential equations for any network can be represented in terms of an interaction graph (\ref{fig:biomodelexamples} top row).
@@
 The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see Supplementary Information) (\reffigscc).
-This map from the interaction graph of a biological network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
+This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
 These three symmetries represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated \reffigscc.
@@
 
-We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a biological network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
+We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
 This expression is fully developed in \ref{eq:sccrobustness}, but can be schematized as in \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
@@
 \end{equation}
-Examining this expression noting that $R_i$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R_{tot}$.
-This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
-This analytical result predicts that any biological network whose associated dynamical system has the interaction graph \reffigscc $\,$ top will be more robust than those associated to any of the other interaction graphs in \reffigscc.
-Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.
-The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
-
-To test this prediction, we computed the probability distribution of stability and dynamical robustness relative to biological network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
-For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).
-Accounting for the number of cycles in a network architecture reveals a strong correlation between robustness and connectivity that was hidden when networks with any number of cycles were considered together (\reffigconnectcycle3D3x3).
-While the most hierarchical network architecture will always lack cycles altogether, cycle number alone is clearly insufficient to account for robustness as the members of each class span nearly the entire range of possible robustness values.
-Consistent with our analysis of the symmetries of robustness, we found that the most hierarchical network architecture is the most robust (\reffigrobusthierarchy).
-Moreover, if we consider hierarchy partitioned by connectivity, we find that there is a monotonic increase in robustness following any line of increasing hierarchy in \reffigconnectdist3D3x3.
-
-\subsection{Dynamical systems on biological networks}
-In the construction of a class of potential models for biological systems at any level of the biological hierarchy from metabolic to ecosystem-level networks, it is common to first attempt to define a collection of observable phenomena of interest and determine a domain (such as binary numbers, integers, or real numbers) in which each observable can be quantified.
-Next, it is necessary to establish the interdependencies among system components.
-The specific manner in which the components depend upon one another must be clarified, which is often done by determining a particular system of mathematical functions that represents a hypothesis about how the quantified observables evolve in time.
-To the degree to which there is uncertainty about the interactions among system components, parameters are introduced to broaden the class of models under consideration.
-Finally, whatever model class remains may be compared to empirical observations to determine how capable the model is of representing the phenomena of interest.
-
-For the case of continuous deterministic observables, the above process can be made more precise by associating a manifold $M_i$ to each observable, a directed graph, $G$, to the interdependencies, and a vector field $V$ over the space determined by taking the product of the manifolds associated to the collection of observables, satisfying these interdependencies for each observable  \cite{Deville}.
-For example, if we have two observables $\{x_1,x_2\}$ where the domains in which they are quantified are given by manifolds $\{M_1,M_2\}$ such that $x_1 \in M_1 \equiv \mathbb{R}^1$ and $x_2 \in M_2 \equiv \mathbb{R}^1$, a directed graph $G_X$ describing the dependencies between these observables and vector field $V$ with components $\{F_1,F_2\}$ defined on $M_1 \otimes M_2 \equiv \mathbb{R}^2$ satisfying the dependencies determined by $G_X$.
-If the system under consideration has the graph given in \reffigexamplesystemmodules$\,$
-% \begin{center}
-% \input{tex/examplehamdyngraph}
-% \end{center}
-with adjacency matrix
-$$
-\adj(G_X) = \begin{pmatrix}
-0 & 1 \\
-1 & 1
-\end{pmatrix}
-$$
-having connectivity $\connectivity$ equal to the number of edges of the graph, in this case $\connectivity = 3$.
-For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}.
-   \right.
-\end{displaymath} For the system $X = \{G_X, M_X, F_X\}$, where $M_X \equiv \{M_1,M_2\}$ and $F_X \equiv \{F_1,F_2\}$ such that $F_1$ is a function of $x_2$ alone and $F_2$ is a function of both $x_1$ and $x_2$ yielding the flow equations
-\begin{align*}
-\frac{dx_1}{dt} & = F_1(x_2),\\
-\frac{dx_2}{dt} & = F_2(x_1,x_2).
-\end{align*}
-
-In the more general case of a system with $n$ components we would have an $n$-dimensional vector of observables
-$$
-x(t) = (x_1(t), \ldots x_n(t)) = \vec{x}(t)
-$$
-whose components are solutions to the arbitrary first order system
-$$
-\frac{dx_i(t)}{dt} = F_i(\vec{x}(t)), \; (i=1,\ldots,n)
-$$
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables.
-
-In order to accommodate the possibility of uncertainty in our
-modeling, we will generalize our notion of a system on a network to
-that of a system with random parameters.  Again, this will involve
-three steps. First, we provide a measure space $S$
-which represents the values over which our parameters can vary. Next,
-instead of a single vector field, we consider a family of vector
-fields parameterized by this space.  As before, for
-each point $p \in S$, the vector field $V(p)$ must be consistent with
-the system graph. Finally, we select a probability measure $\mu$ which represents our understanding of which values of the parameters
-are likely. In accord with Bayesian statistics, we might revise this
-distribution as data comes in or use it to estimate parameters and error
-bars from experimental results.
-
-Having done this, we are now in a position to do a probabilistic analysis
-of our system and its properties.  Given some quantity $q$ characterizing
-our system, this quantity becomes a random variable which may be discrete or
-continuous depending on the quantity under consideration.  For example, if
-the quantity is the time it takes for a particle to travel between two points,
-we have a real-valued random variable; if the quantity is the number of fixed
-points, we have an integer-valued random variable; and if the quantity is
-whether or not the system posesses a limit cycle, we have a binary random
-variable.
-
-A quality which is of interest to us is robustness whose evaluation requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters \footnote{We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the components as determined by the parameters and not only to perturbations of the state vector at a given point in time.
-For example, when the network of components and interactions corresponds to a gene-regulatory network, then mutation is one mechanism by which such perturbations may arise}.
-We will quantify this as the probability that, if some property holds for a set of parameters, it will continue to hold if we make a random perturbation about those parameter values.
-To do so, we will introduce, in addition to the probability distribution $\mu$ described above, a family of probability distributions $\mu'$ such that $\mu'(x,y)$ encodes the conditional probability that a system with a parameter value $x$ will have its parameter values changed to $y$ under a perturbation.
-For our biological models, we will be interested in large perturbations rather than the small or infinitesimal perturbations usually considered in the theory of dynamical systems.
-Then, given a binary random variable $q$, we define its robustness as the following conditional probability where $\mathbf{1}_q(x)$ is the standard indicator function equal to $1$ when $x$ satisfies $q$ and $0$ otherwise:
-\begin{align}\label{eq:robustness}
-  R (q,\mu') =
-  \frac{\int_S d\mu(x) \int_S d\mu'(x,y) \mathbf{1}_q(x) \mathbf{1}_q(y)}
-  {\int_S d\mu(x) \mathbf{1}_q(x)}
-\end{align}
-
-\subsection{Stability analysis of biological networks}
-The class of dynamical quantities on which we shall focus in this
-investigation involve stability of equilibria.  Suppose that $\vec x$
-is a point on our phase space which depends upon the parameters.  Then
-we may take $q$ to be a binary random variable which describes whether
-or not $\vec x$ is a fixed point of the system (i.e. $q$ is true if and only if
-$F_i(\vec{x})=0$ for all $i$) and perform a probabilistic analysis
-of the sort discussed above.
-
-Furthermore, if it turns out that $\vec x$ is indeed a fixed point, we
-may proceed to ask whether it is a dynamically stable fixed point.
-Intuitively, dynamic stability means that, if one chooses the initial
-conditions sufficiently close to the fixeed point, the solution will
-stay close to the fixed point.  Physically, this is important because,
-if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
-observing the solution ${\vec x}(t) = {\vec x}^0$.
-
-To determine stability, we linearize the equations of motion about the
-fixed point $\vec{x}^0$:
-\begin{equation}\label{eq:lineardynsys}
-\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
-\end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
-$$
-a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
-$$
-To each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ we can associate a directed graph $G_A$ given by an adjacency matrix $\adj(G_A)$ where
- \begin{displaymath}
-   \adj(G_A)_{ij} = \left\{
-     \begin{array}{lr}
-       1, & a_{ij} \neq 0\\
-       0, & a_{ij} = 0
-     \end{array}
-   \right.
-\end{displaymath}
-The graphs $G_X$ and $G_A$ are equivalent because the condition $F_i$ independent of $x_j$ definitive of $\adj(G_X)$ corresponds precisely to $\frac{\partial F_i}{\partial x_j}=0$.
-
-The spectral abscissa of the matrix $A$ is defined as
-$$
-\eta(A) = \max_i \{\Re(\lambda_i)\}
-$$
-where $\lambda_i$ are the eigenvalues of $A$.
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
-This is because the general solution to \ref{eq:lineardynsys} is
-$$
-y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
-$$
-for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
-This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.
-
-Even though the determination of stability only requires examining
-linearized equations of motion, it is worth noting that knowing the
-stability of a fixed point can yield information about the non-linear
-dynamics of a system.  For instance, in two dimensions,
-Poincare-Bendixson theory implies that every limit cycle must encircle
-at least one fixed point \cite{Davis1962}.  Furthermore, in the case where the cycle contains a single fixed point (such as in the Lotka-Volterra model of
-conflicting populations), the stability of the fixed point will determine whether the limit cycle is attractive or repulsive.
-
-%% To determine whether a randomly selected dynamical system evaluated at
-%% a random critical point, given by a matrix such as $A$, is stable, it
-%% is sufficient to check the above condition. Integrating over the
-%% region of the parameter space defining $A$ enables the determination
-%% of the probability of stability to perturbations in state $\vec{x}$
-%% over an ensemble of dynamical systems.
-
-% Another form of stability, referred to as structural
-% stability \cite{Smale1967}, requires the determination of whether or
-% not a given dynamical system that is determined to be stable remains
-% stable under a perturbation to one of its defining parameters. This
-% can be formalized as the conditional probability distribution
-% $$
-% P(A' \, \textrm{stable}\, \big| \, A \, \textrm{stable})
-% $$
-% where $A'$ represents a matrix derived from some perturbation
-% applied to the matrix $A$ given above. In the simple case where
-% $A \in \mathbb{R}^{n \times n}$ the perturbation which involves
-% sampling uniformly over the $a_{ij}$ defining $A$ the value of a
-% particular $a_{ij}$ from a given probability distribution.
-
-Suppose that our Jacobian $A$ is an $n \times n$ matrix with real coefficients, connectivity $\connectivity$, and denote the proposition ``$A$ is stable'' as $\mathrm{stab}(A)$.
-Then we may compute the robustness of this proposition using \ref{eq:robustness} once we have specified the probability densities $\mu$ and $\mu'$.
-In general, these will depend upon the parameters of the non-linear system.
-For the purpose of the current investigation, we shall make the ansatz that these are such that the distribution of entries of $A$ is (at least approximately) the uniform distribution $\mathcal{U}(-1,1)$ on the $\connectivity$-dimensional hypercube, $H^\connectivity$, of edge length $r=2$, centered about the origin.
-To model the perturbations, we resample the elements $k_1, k_2, \ldots k_m$ of this matrix from the same uniform distribution yielding
-\begin{widetext}
-$$
-\mu_{k_1,\ldots,k_m}(x,y) = \prod_{y \notin \{x_1, \ldots, x_m\} } \delta(x_j-y_j) \prod_{y \notin \{x_1,\cdots,x_m\}} \mathbf{1}_{[-1,1]} (y_j) \mathbf{1}_{[-1,1]} (x_j)
-$$
-\end{widetext}
-Under these assumptions, the expression for robustness becomes a special case of \ref{eq:robustness}:
-\begin{widetext}
-\begin{equation}\label{eq:condprobgen}
- R (\mathrm{stab}, \mu'_{k_1, \ldots, k_m}) =
-  \frac{\int_{H^\connectivity} dx_1 \cdots dx_d \int_{H^m} {dx'}_{k_1} \cdots {dx'}_{k_m}\,
-    \mathbf{1}_{stab} \mathbf{1}_{stab'}}
-  {\int_{H^\connectivity} dx_1 \cdots dx_d  \, \mathbf{1}_{stab}(x_1 \cdots x_d)},
-\end{equation}
-\end{widetext}
-where we use the abbreviations
-\begin{align*}
-\mathbf{1}_{stab} &= \mathbf{1}_{stab}(x_1 \cdots x_{k_1} \cdots x_{k_m} \cdots x_\connectivity), \\
-\mathbf{1}_{stab'} &= \mathbf{1}_{stab'}(x_1 \cdots {x'}_{k_1} \cdots {x'}_{k_m}  \cdots x_d),
-\end{align*}
-i.e. $\mathbf{1}_{stab'}$ corresponds to replacing $x_{k_1}, \ldots x_{k_m}$ with their primed counterparts.
-
-While one can resample a fixed subset of the links, more typically we will be randomly selecting which links to resample with a uniform distribution on the links.
-The result of this operation corresponds to averaging our quantity over subsets of links:
-\begin{widetext}
-\begin{equation}\label{eq:robavgoverlinks}
-\langle R (\hbox{stab}, \mu') \rangle_m =
-\frac{1}{\binom{\connectivity}{m}}
-\sum_{\{k_1, k_2, \ldots k_m\} \subset \{1, 2,, \ldots \connectivity\}}
-R (\hbox{stab}, \mu'_{k_1, \ldots, k_m})
-\end{equation}
-\end{widetext}
-We are fundamentally interested in the relationship between \ref{eq:condprobgen} and \ref{eq:robavgoverlinks} and topological properties of systems that derive from their associated graphs.
-In order to investigate this further, we must characterize properties of system graphs in order to evaluate correlations of those properties with dynamical robustness.
-
-\subsection{Network hierarchy and strongly connected components}
-
-Dynamical networks of the kind described above can be viewed at the level of their interdependency graphs as being composed of more fundamental interacting subsystems \reffigexamplesystemmodules.
-One such kind of system composition and decomposition is given by the notion of open systems, which are distinguished by having some of the variables specified as control variables whose values are given as autonomous functions of time rather than determined by the dynamics via intrinsic interactions \cite{Vagner2014}.
-Given several such open systems, we may combine them to produce a larger system by setting the control variables of a subsystem equal to the dynamical variables of another system.
-Taking this notion to its logical extreme, one can dissect a system into a collection of one open system for each dynamical variable.
-However, this decomposition is trivial since it is equivalent to the underlying system graph and what we instead want is an intermediate decomposition into relatively self-contained modules.
-One method of accomplishing this based upon the topology of the system graph is the decomposition into strongly connected components.
-
-\begin{figure*}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\textwidth]{fig/modsccsym.pdf}
-\caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
-(B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
-The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
-We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
-The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e.
-SCCs with more than one variable) and hierarchy.
-(C) Symmetries of the $\hier$ transformation between graphs and SCCs.
-The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link.
-All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
-\label{fig:modsccsym}
-\end{figure*}
-
-A strongly connected component (SCC) of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three variable systems are outlined in \reffigscc$\,$ along with their adjacency matrices and hierarchy diagrams (to be defined below).
-The decomposition of a digraph into strongly connected components corresponds to a block triangular decomposition of its adjacency matrix.
-Say that the graph $G$ has strongly connected components $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
-Label the vertices in such a way that $v_1, \ldots, v_{n_1}$ belong to $C_1$, $v_{n_1 + 1}, \ldots, v_{n_2}$ belong to $C_2$, etc.
-Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
-This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
-
-Corresponding to this decomposition we can construct a directed acyclic graph $\hier (G)$ or the condensed graph \cite{Corominas-Murtra2013}.
-Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
-There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
-
-One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
-a strongly connected graph $C_n$.
-To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
-The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
-
-This map $\hier$ is many-to-one and so there is a large class of
-operations which leaves $\hier(G)$ invariant for a given graph $G$.
-For instance, we may interchange the positions of the strongly
-connected components relative to each other \reffighiertransformations$a$.
-Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations$b$ or between components, or even add or delete links \reffighiertransformations$c$.
-As we shall see later, some of these operations leave invariant key dynamical quantities, in our case dynamical robustness, so we may regard them as a symmetry groupoid of our system with respect to the robustness property.
-
-The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
-The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc (top panel).
-This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
-This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical.
-In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.
-
-\subsection{Stability and robustness analysis of particular system ensembles}
-For systems having two variables, we can analytically compute the probability of stability and robustness from \ref{eq:condprobgen}.
-For those having three variables, we can estimate these same quantities using Monte Carlo simulations.
-Systems of larger size can be analyzed using the symmetry properties of robustness extracted from this analysis.
-For two-variable systems having $2 \times 2$ Jacobian matrices, the aforementioned stability criteria result in the conditions $T < 0$ and $D >
-0$ where $T$ and $D$ denote the trace and the determinant. Suppose we have a stable matrix
-$$
-\begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix}
-$$
-where $a + c < 0$ and $ac > bd$.
-For the case in which $x_1=a,\,x_2=b,\,x_3=c,\,x_4=d$ we need to compute what corresponds to $R(\mathrm{stab},\mu'_k)$ where $k=1 \ldots 4$.
-By symmetry, there are two cases to consider; resampling $a$ is equivalent to resampling $c$ and resampling $b$ is equivalent to resampling $d$ so we only need to explicitly compute $ R(\mathrm{stab},\mu'_1)$ and $R(\mathrm{stab},\mu'_2)$.
-Suppose that we resample $b$ to compute $R(\mathrm{stab},\mu'_2)$.
-% \begin{strip}
-% \begin{align}\label{eq:condprob}
-% P\left(\begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)
-% & = \frac{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable and } \begin{pmatrix}
-% a & b' \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}{P\left(\begin{pmatrix}
-% a & b \\
-% d & c
-% \end{pmatrix} \textrm{stable } \right)}.
-% \end{align}
-The denominator of \ref{eq:condprobgen} in this case is given by
-\begin{align*}
-P\left(\textrm{stab } \left( \begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix} \right) \right) = \frac{\int_{\genfrac{}{}{0pt}{}{\genfrac{}{}{0pt}{}{ac>bd}{a+c<0}}{H^4}} da\,db\,dc\,dd\,1}{\int_{H^4} da\,db\,dc\,dd\,1}.
-\end{align*}
-Since the trace does not involve $b$, the $T<0$ condition will be satisfied automatically and we only need to examine the determinant.
-Thus, we have the inequalities $ac > b'd$ and $-1 < b' < 1$ in addition to the previous constraints leading to an expression for the numerator of \ref{eq:condprobgen}
-\begin{widetext}
-\begin{align*}
-P\left(\textrm{stab} \left( \begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
-a & b' \\
-d & c
-\end{bmatrix} \right) \right) = \frac{\int_{{{ac>b'd \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,db'\,1}{\int_{H^5} da\,db\,dc\,dd\,db'\,1}.
-\end{align*}
-\end{widetext}
-The analogous equation for resampling $a$ is
-\begin{widetext}
-\begin{align*}
-P\left(\textrm{stab} \left( \begin{bmatrix}
-a & b \\
-d & c
-\end{bmatrix} \right) \textrm{ and stab} \left( \begin{bmatrix}
-a' & b \\
-d & c
-\end{bmatrix} \right) \right) = \frac{\int_{{{{a'c>bd \atop a' + c < 0} \atop ac>bd} \atop a+c<0} \atop H^5} da\,db\,dc\,dd\,da'\,1}{\int_{H^5} da\,db\,dc\,dd\,da'\,1}.
-\end{align*}
-\end{widetext}
-Using this approach the probability of stability and of robustness for all two variable systems is given in \ref{tab:structstabmat}.
-
-The analogous results for all three variable systems are computed using Monte Carlo integration and shown in \ref{tab:structstabmat3} and \reffigrobustconnect.
-This process is associated with some error relative to the exact integration described above.
-In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$ (see \emph{Methods}).
-
-It has been stated previously on the basis of simulation that system stability decreases with connectivity as the system size goes to infinity \cite{May1972}.
-For small system sizes such as the two and three variable systems, studied here the situation is not so clear cut.
-For two variable systems, system stability is constant across the entire range of connectivities.
-For three variable systems, the trend shows a minor decrease from connectivity $4$ to $5$ followed by small fluctuations as shown in \ref{fig:apstab3x3}.
-
-The relationship between connectivity and robustness for two variable systems is shown in \ref{tab:structstabmat} and likewise for three variable systems in \ref{tab:structstabmat3} and \reffigrobustconnect.
-If we average over the different classes of matrices for a given connectivity we see there is a correlation between connectivity and robustness demonstrated by the red dots in \reffigrobustconnect.
-% \subsection{Cycle number is inversely correlated with robustness}
-\ref{fig:cycle3x3} shows the robustness for all three variable systems as a function of the number of simple cycles (elementary circuits) of length greater than one in the corresponding directed graph \cite{Johnson1975}.
-There appears to be a weak negative correlation between robustness and the number of simple cycles.
-
-% \subsection{Cycle number and connectivity classify robust systems}
-The combination of connectivity and cycle number as shown in \reffigconnectcycle3D3x3 provides a better classification of the dependence of robustness upon network topology.
-Here the robustness of three variable systems with a given number of cycles, increases monotonically with connectivity.
-The network with the highest robustness for three variable systems is that of \reffigscc (top panel).
-This network is the most hierarchical of all three variable systems in the sense that it represents a total ordering of the components of the network and its adjacency matrix also shown in \reffigscc (top panel) has a block triangular structure.
-
-This observation suggested that graph edit distance from \reffigscc (top panel), hierarchy, might provide a better characterization of dynamical robustness.
-\reffigrobusthierarchy$\,$ shows dynamical robustness as a function of hierarchy.
-There is a monotonic correlation between the upper bound of robustness and hierarchy.
-\reffigconnectdist3D3x3$\,$ shows dynamical robustness as a function of both hierarchy and connectivity.
-The monotonic correlation between hierarchy and robustness is refined by an underlying correlation between robustness and connectivity analogous to that of \reffigconnectcycle3D3x3.
-
-\begin{figure*}[!ht]
-\centering
-\noindent\includegraphics[width=0.8\textwidth]{fig/combinedfigs.pdf}
-\caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.
-The red line represents a best fit in the least-squares sense with Pearson product-moment correlation coefficient $r=0.29$.
-The lowest and highest robustness network architectures are labelled.
-Other network architectures are shown in \ref{tab:structstabmat3}.
-(B) Robustness versus hierarchy.
-Correlation coefficient $r=0.67$.
-(C) Number of cycles and (D) hierarchy vs connectivity and robustness.
-The color of each point represents the average robustness of all graphs having the parameters specified on the $x$ and $y$ axes.
-}
-\label{fig:combined}
-\end{figure*}
-
-\subsection{Component decomposition, hierarchy, and the symmetries of robustness}
-
-The fact that the most hierarchical of networks is the most robust in the case of three variable systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.
-By combining our decomposition of the graph with our analysis of the robustness of random ensembles of dynamical systems, we will deduce a general result which relates robustness to hierarchy.
-
-Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomial factors as the product of the charactericstic polynomials of its diagonal blocks.
-Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.
-Note that this condition does not depend upon the entries off the diagonal (which correspond to links between strongly connected components) and does not depend upon what order the components appear.
-
-Using these observations, we may express the robustness of a graph in
-terms of the robustnesses of its strongly connected components.  If
-each component $C_i$ of our graph $G$ has $v_i$ vertices, connectivity
-$\connectivity_i$, and there are $\ell_{ij}$ links between components
-$i$ and $j$, then there are a total of
-$\sum_{i,j \in \hier(G)} \ell_{ij} + \sum_{i=1}^n \connectivity_i$
-links in $G$.  If we resample a randomly chosen link in a component
-$C_i$, then the probability that the system will remain stable after
-resampling is $\langle R(\hbox{stab} (C_i), \mu') \rangle_1$.  If we resample a link
-between components $C_i$ and $C_j$, the system will remain stable with
-probability 1.  Hence, the probability that the system will remain
-stable upon resampling a random link is the weighted average of these
-probabilities:
-\begin{widetext}
-\begin{align}\label{eq:sccrobustness}
-\langle R(\hbox{stab} (G), \mu') \rangle_1 =
-\frac{\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i \langle R(\textrm{stab}(C_i),\mu') \rangle_1}
-     {\sum\limits_{(i,j) \in \hier(G)} \ell_{ij} +
-      \sum\limits_{i=1}^n \connectivity_i}.
-\end{align}
-\end{widetext}
 For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.
@@
 The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
-Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is
-\[
-\langle R(\textrm{stab}(G),\mu') \rangle_1 = \frac{2 + 0.5 + 4 \times 0.62}{2 + 1 + 4} = 0.714,
-\]
-which agrees with the value computed in \ref{tab:structstabmat} up to
-sampling error.
+Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.
 
-As can be seen from \ref{eq:sccrobustness}, all that is required to fix robustenss is a set of SCC sizes and the number of links between them.
-This demonstrates that robustness is symmetric under the transformations shown in \reffighiertransformations $a$ and $b$.
-Since transformations of the type \reffighiertransformations $c$ change the number of links between connected components, this symmetry of $\hier$ is broken by \ref{eq:sccrobustness}.
-Nevertheless, a considerable degree of symmetry remains.
-For example, \ref{fig:robustnesssymmetries} shows the symmetry groupoid of robustness for the case in which we fix connected component sizes $\{2,1,1\}$ with a total of $3$ links between them.
-
-Next, we derive an upper bound on the robustness of any graph
-constructed from a given set of connected components.  Since
-$\langle R(\textrm{stab}(C_i), \mu') \rangle_1 \le 1$ for all $i$, it follows that increasing $\ell_{ij}$ will
-increase $\langle R(\textrm{stab}(G),\mu') \rangle_1$.  Given two connected components $C_i$ and $C_j$ with
-$v_i$ and $v_j$ nodes respectively, we have a maximum of $v_i v_j$
-links going from $C_i$ to $C_j$.  Hence, $\ell_{ij} \le v_i v_j$, so,
-we conclude that
-\begin{align}
-\langle R(\textrm{stab}(G),\mu') \rangle_1 \le \frac{\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i \langle R(C_i,\mu') \rangle_1}
-              {\sum\limits_{(i,j) \in \hier(G)} v_i v_j +
-               \sum\limits_{i=1}^n \connectivity_i},
-\end{align}
-
+Examining this expression noting that $R_i$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R_{tot}$.
+Given two connected components $C_i$ and $C_j$ with $v_i$ and $v_j$ nodes respectively, we have a maximum of $v_i v_j$ links going from $C_i$ to $C_j$.
+Hence, $l \le \sum_{(ij) \in \hier(G)} v_i v_j$.
 Since every acyclic digraph can be embedded into a totally ordered
@@
 been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i <
-j$.  Hence, after substituting into our expression and simplifying
-using the identity
-$$\sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
+j$.  Hence, $l \le l_{max}$ where
+$$l_{max} = \sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
 v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n}
 v_i^2,$$
-we arrive at the following upper bound on robustness in terms
-of the connected components:
-\begin{widetext}
-\begin{align} \label{eq:sccmaxrobustness}
-R_{\mathrm{max}}(C_1, \ldots C_n) =
-\frac{\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i \langle R(\hbox{stab}(C_i),\mu') \rangle_1}
-     {\frac{1}{2} ((\sum_{i=1}^n v_i)^2 - \sum_{i=1}^n v_i^2) +
-                    \sum_{i=1}^n \connectivity_i}.
-\end{align}
-\end{widetext}
+Hence, we conclude that $R_{tot} \le R_{max}$ where $R_{max}$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$.
+
 Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
@@
 
-This argument also works when we resample more than one entry,
-although the notation becomes more complicated.  Suppose that we
-resample $m$ nodes.  Define
-\begin{widetext}
-\begin{equation*}
-M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
-m = \sum_{i=0}^n m_i \quad\&\quad
-m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
-(\forall i \in \{1, \ldots, n\}) \; m_i < \connectivity_i \right\}.
-\end{equation*}
-\end{widetext}
-Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
-m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
-$m_0$ links between strongly connected components.  Hence, our
-weighted average becomes
-\begin{widetext}
-\begin{equation}
-\langle R(\hbox{stab} (G), \mu') \rangle_m =
-\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
-      {m \choose m_0, m_1, \ldots. m_n}
-      \left(m_0 + \sum\limits_{i=1}^n m_i \langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \right)}
-     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
-      {m \choose m_0, m_1, \ldots. m_n} m}
-\end{equation}
-\end{widetext}
+This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
+This analytical result predicts that any network whose associated dynamical system has the interaction graph \reffigscc $\,$ top will be more robust than those associated to any of the other interaction graphs in \reffigscc.
+Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.
+The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
 
-As before, since $\langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \le 1$, we may
-increase $\langle R(\hbox{stab} (G), \mu') \rangle_m$ by increasing the maximum
-possible value of $m_0$ while keeping the strongly connected
-components the same.  Again, if we fix $\hier(G)$, the maximum
-possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
-whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
-v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
-G_{tot}$.  Hence, we conclude that $\langle R(\hbox{stab} (G), \mu') \rangle_m \le
-\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_m$.
-
-The preceding argument demonstrates the manner in which robustness is
-maximized by maximizing the number of edges \emph{between} the strongly
-connected components of the graph underlying system interdependencies.
-And, the graph that has the latter property is that associated to the
-ordering, $G_{\mathrm{tot}}$, which also possess the highest degree of
-hierarchy. The fact that robustness is equivalent for any configuration
-of strongly connected components with an equivalent number of links
-between them demonstrates that it is invariant to permutations of
-strongly connected components.
-% This result does not depend at all uponany particular details such as the probability distribution from which the component interaction strengths are sampled because the arguments presented here are purely topological in nature.
-The result that dynamical robustness is correlated with system hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles primarily studied here.
+To test this prediction, we computed the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
+For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).
+Accounting for the number of cycles in a network architecture reveals a strong correlation between robustness and connectivity that was hidden when networks with any number of cycles were considered together (\reffigconnectcycle3D3x3).
+While the most hierarchical network architecture will always lack cycles altogether, cycle number alone is clearly insufficient to account for robustness as the members of each class span nearly the entire range of possible robustness values.
+Consistent with our analysis of the symmetries of robustness, we found that the most hierarchical network architecture is the most robust (\reffigrobusthierarchy).
+Moreover, if we consider hierarchy partitioned by connectivity, we find that there is a monotonic increase in robustness following any line of increasing hierarchy in \reffigconnectdist3D3x3.

### 2015-01-15 tex/introduction.tex
reduce number of references

@@
 The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-All of these models utilize essentially equivalent mathematical structures \ref{fig:biomodelexamples} \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}.
+All of these models utilize essentially equivalent mathematical structures \ref{fig:biomodelexamples} \cite{RossCr2003,Palsson2011a,Sauro2012}.
 Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
+% \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}
 
-Recent work has demonstrated that as a result of the existence of largely insensitive directions within the parameter space for such models, the approach outlined above often allows for a large variety of models to fit equivalent data \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
+Recent work has demonstrated that as a result of the existence of largely insensitive directions within the parameter space for such models, the approach outlined above often allows for a large variety of models to fit equivalent data \cite{Machta2013,Hines2014,Prabakaran2014,Tonsing2014}.
 In addition, there is often uncertainty about the very structure of such networks.
@@
 This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems that, in comparison to one another, appear to have components that are randomly interlinked.
+% \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}
 
-Indeed, investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
+Investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
 The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the networks represented in \ref{fig:biomodelexamples}.
@@
 Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
-Robustness is of interest in biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}.
+Robustness is of interest in biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}.
 Intuitively, robustness is the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
@@
 Moreover these results hold for networks of arbitrary size and independent of the probability distribution from which the strengths of interaction are sampled.
+% \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}

### 2015-01-15 tex/methods.tex
minor edits
reduce number of references

@@
 This estimate is characterized by the variance of the posterior Beta distribution
-\begin{widetext}
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
-$$
-\end{widetext}
+% \begin{widetext}
+\begin{align*}
+&\mathrm{var}(\theta | \mathcal{D}) =\\
+&\frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
+\end{align*}
+% \end{widetext}
 Since for the chosen prior $a=b=1 \ll N$ this simplifies to
@@
 yielding the error estimate given by the associated standard deviation.
+In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.

### 2015-01-15 tex/overview.tex
reduce number of references

@@
 
-The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see Supplementary Information) (\reffigscc).
+The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{}) (\reffigscc).
 This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.

### 2015-01-15 tex/supplement.tex
reduce number of references

@@
 
-The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see Supplementary Information) (\reffigscc).
+The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{}) (\reffigscc).
 This map from the interaction graph of a biological network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.

### 2015-01-15 tex/introduction.tex
qualify introduction of robustness and hierarchy

@@
 Robustness is of interest in biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}.
-Intuitively, robustness is the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
+For the purpose of this investigation, we quantify robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
 The maximally hierarchical network is considered to be the graph associated to the total ordering, which is the complete directed acyclic graph for a given system size \cite{Cormen2009}.
-Here, hierarchy is then defined as the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
+Hence, we use a measure of hierarchy based upon the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
 We demonstrate that systems with the most hierarchical network topology according to this definition exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.

### 2015-01-15 tex/abstract.tex
minor changes to abstract

@@
 Here we analyze the stability and robustness of a large class of dynamical systems.
-In particular, we precisely determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics, for systems with two and three interacting components.
+We determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics.
 We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.

### 2015-01-16 tex/introduction.tex
add model example to main text

@@
 Robustness is of interest in biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}.
-For the purpose of this investigation, we quantify robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
+For the purpose of this investigation, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
 The maximally hierarchical network is considered to be the graph associated to the total ordering, which is the complete directed acyclic graph for a given system size \cite{Cormen2009}.

### 2015-01-16 tex/overview.tex
add model example to main text

@@
 %!TEX root = ../paper.tex
+
+In the general case of a dynamical system with $n$ components, where the components may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
+$
+whose components are solutions to the arbitrary first order system
+$$
+\frac{dx_i(t)}{dt} = F_i(\vec{x}(t), \vec{p}), \; (i=1,\ldots,n)
+$$
+where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p}$ is the vector of parameters of the $F_i$.
+These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks, or cooperativity exponents.
+For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$.
+
+Fixed points are the simplest class of solutions to the dynamical system characterizing its long-term behavior. If $\vec x$ is a fixed point (i.e. $F_i(\vec{x})=0$ for all $i$), we
+may proceed to ask whether it is dynamically stable.
+Intuitively, dynamic stability means that, if one chooses the initial
+conditions sufficiently close to the fixed point, the solution will
+stay nearby.  Physically, this is important because,
+if a fixed point ${\vec x}^0$ is unstable, we have zero probability of
+observing the solution ${\vec x}(t) = {\vec x}^0$ in the absence of coupling to another system.
+The Lotka-Volterra model has two fixed points: the trivial one of all zero species $n_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{N} b_{ij} n_j = 0$.
+
+To determine stability, we linearize the equations of motion about the
+fixed point $\vec{x}^0$:
+\begin{equation}\label{eq:lineardynsys}
+\frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
+\end{equation}
+where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
+$$
+a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
+$$
+The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix.
+The Lotka-Volterra model has
+ \begin{equation}\label{eq:lotkavolterrajacobian}
+   a_{ij} = \left\{
+     \begin{array}{lr}
+       r_i + b_{ii} n_i + \sum_{k=1}^{N} b_{ik} n_{k}, & i \neq j\\
+       b_{ii} n_i & i=j
+     \end{array}.
+   \right.
+\end{equation}
 
@@
 The dynamical model given in terms of a system of differential equations for any network can be represented in terms of an interaction graph (\ref{fig:biomodelexamples} top row).
+For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
+ \begin{displaymath}
+   \adj(G_X)_{ij} = \left\{
+     \begin{array}{ll}
+       1, & F_i \hbox{ depends on } x_j\\
+       0, & F_i \hbox{ does not depend on } x_j
+     \end{array}.
+   \right.
+\end{displaymath}
+For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be quite sparse.
+We can also associate a linearized interdependency graph $G_A$ given by an adjacency matrix $\adj(G_A)$ to each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ where
+ \begin{equation}\label{eq:lininterdepadj}
+   \adj(G_A)_{ij} = \left\{
+     \begin{array}{lr}
+       1, & a_{ij} \neq 0\\
+       0, & a_{ij} = 0
+     \end{array}.
+   \right.
+\end{equation}
+In general, the graph $G_A$ is a subgraph of $G_X$, however, for nearly all systems, $G_A$ is equivalent to $G_X$.
+We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
 These interaction graphs can be viewed as deriving from the combination of system modules that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).
-Symmetries are characterized by the ability to interchange these modules or their connectivity without changing some property of the system.
+Symmetries of the system are characterized by the ability to interchange these modules or their connectivity without changing some property of the system.
 
-The network architecture can be represented in terms of an adjacency matrix and further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{}) (\reffigscc).
+The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{}) (\reffigscc).
 This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
@@
 
-We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
-This expression is fully developed in \ref{eq:sccrobustness}, but can be schematized as in \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
+In the process of biological evolution, the parameter values $\vec{p}$ or the form of the functions $F_i$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due, for example, to mutations or environmental fluctuations.
+These changes may alter the capacity of a given system to achieve a stable steady state.
+In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$.
+This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of ${\vec{p}}\,'$.
+We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
+This probability is what we refer to as dynamical robustness.
+
+It is justified to consider resampling elements of $A$ to generate $A'$ as a proxy for resampling elements of $\vec{p}$ to produce $\vec{p}\,'$ if any matrix $A$ can be obtained for some $F_i$, $\vec{p}$ and $\vec{x}^0$.
+This holds for the $F_i$ defining the Lotka-Volterra model.
+This is due to the fact that for a specification of non-zero real numbers for the components of $\vec{n}^0$ and any real numbers for the components of $a_{ij}$, there is a choice of parameters $\vec{p}$ given by $b_{ij} = \frac{a_{ij}}{n_i^0}$ and $r_i = - \sum_{j=1}^N \frac{n_j^0}{n_i^0} a_{ij}$ that generates those particular $a_{ij}$ as the Jacobian matrix of the dynamical system.
+
+We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$ the entries of the Jacobian are sampled independently from a generic probability distribution.
+The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
 \begin{equation}\label{eq:robschematic}

### 2015-01-16 tex/overview.tex
minor edit

@@
 
-In the general case of a dynamical system with $n$ components, where the components may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
+In the general case of a dynamical system with $n$ components, where the components may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of state variables or observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
 $
@@
 \end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$-matrix $A$ has components
+where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$ Jacobian matrix $A$ has components
 $$

### 2015-01-16 tex/discussion.tex
minor edit

@@
 %One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However,
-It may be possible to prove a bound on the robustness of individual SCCs, which could provide more definite knowledge and circumvent the need for expensive simulations.
+It may be possible to prove bounds on the robustness of individual SCCs, which could provide more precise knowledge about the relative robustness of different network architectures.
 The conservation of robustness with respect to nontrivial symmetries including the interchange of SCCs and permutation within SCCs suggests the existence of an evolutionary neutral space.

### 2015-01-16 tex/discussion.tex
minor edits

--- 2014-12-01 tex/discussion.tex
@@
 In order to further this work from a theoretical perspective, it will be necessary to deepen our understanding of the relationship between dynamical robustness and the underlying network topology.
+Following May \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, this will involve improving the general understanding of the relationship between perturbations to a system's structure and the qualitative changes in the dynamical phenomena it can produce.
 %One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However,
-It may be possible to prove a bound on the robustness of individual SCCs, which could provide more definite knowledge and circumvent the need for expensive simulations.
+It may also be possible to prove bounds on the robustness of individual SCCs, which could provide more precise knowledge about the relative robustness of different network architectures.
 The conservation of robustness with respect to nontrivial symmetries including the interchange of SCCs and permutation within SCCs suggests the existence of an evolutionary neutral space.

### 2015-01-16 tex/introduction.tex
minor edits

@@
 The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-All of these models utilize essentially equivalent mathematical structures \ref{fig:biomodelexamples} \cite{RossCr2003,Palsson2011a,Sauro2012}.
+All of these models utilize essentially equivalent mathematical structures, \ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}.
 Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
@@
 Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
-Robustness is of interest in biological systems at all scales, and has been previously studied in the context of gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}.
+Robustness is of interest in biological systems at all scales, and has been previously studied, for example, in the context of gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}.
 For the purpose of this investigation, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.

### 2015-01-16 tex/overview.tex
minor edits

@@
 $$
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix.
+The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix (\refsupp{}).
 The Lotka-Volterra model has
@@
 The dynamical model given in terms of a system of differential equations for any network can be represented in terms of an interaction graph (\ref{fig:biomodelexamples} top row).
-For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where
- \begin{displaymath}
-   \adj(G_X)_{ij} = \left\{
-     \begin{array}{ll}
-       1, & F_i \hbox{ depends on } x_j\\
-       0, & F_i \hbox{ does not depend on } x_j
-     \end{array}.
-   \right.
-\end{displaymath}
+For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ for $F_i \hbox{ depends on } x_j$ and $0$ for $F_i \hbox{ does not depend on } x_j$.
+%  \begin{displaymath}
+%    \adj(G_X)_{ij} = \left\{
+%      \begin{array}{ll}
+%        1, & F_i \hbox{ depends on } x_j\\
+%        0, & F_i \hbox{ does not depend on } x_j
+%      \end{array}.
+%    \right.
+% \end{displaymath}
 For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be quite sparse.
@@
 
-The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be further abstracted by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{}) (\reffigscc).
+The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be further abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc).
 This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
@@
 In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$.
-This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of ${\vec{p}}\,'$.
+This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of ${\vec{p}}\,'$ (\refsupp{}).
 We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
 This probability is what we refer to as dynamical robustness.
-
-It is justified to consider resampling elements of $A$ to generate $A'$ as a proxy for resampling elements of $\vec{p}$ to produce $\vec{p}\,'$ if any matrix $A$ can be obtained for some $F_i$, $\vec{p}$ and $\vec{x}^0$.
-This holds for the $F_i$ defining the Lotka-Volterra model.
-This is due to the fact that for a specification of non-zero real numbers for the components of $\vec{n}^0$ and any real numbers for the components of $a_{ij}$, there is a choice of parameters $\vec{p}$ given by $b_{ij} = \frac{a_{ij}}{n_i^0}$ and $r_i = - \sum_{j=1}^N \frac{n_j^0}{n_i^0} a_{ij}$ that generates those particular $a_{ij}$ as the Jacobian matrix of the dynamical system.
 
@@
 v_i^2,$$
-Hence, we conclude that $R_{tot} \le R_{max}$ where $R_{max}$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$.
-
-Furthermore, this bound is attained.  Suppose that $G_{\mathrm{tot}}$
-is the graph on $n$ nodes with a link from node $i$ to node $j$
-whenever $i < j$.  Then, by the construction described in the previous
-section on network hierachy, we have a graph $G_{\mathrm{max}}$ such
-that the components of $G_{\mathrm{max}}$ are $C_1, \ldots C_n$ and
-$\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$.  By our formula,
-$\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+Hence, we conclude that $R_{tot} \le R_{max}$ where $R_{max}$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$ and the upper bound is attained (\refsupp{}).
 
@@
 
-To test this prediction, we computed the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
+To test the prediction of this analytical result, we computed the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
 For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).

### 2015-01-21 tex/methods.tex
derive Jacobian for crn

@@
 In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
+
+\subsection{Reaction Networks with Prescribed Connecctivity and Jacobians}\label{sec:reactionnetjacobian}
+
+To justify our consideration of arbitrary Jacobian matrices, we determine a simple ensemble of reaction networks for which arbitrary Jacobian matrices are realizable.
+This condition holds if one can solve for the parameter values of the system of equations corresponding to that ensemble in terms of the elements of an arbitrary Jacobian matrix.
+More precisely, we will show that, given an arbitrary directed graph $G$ where $G_{ii} = 1$ for all $i$, there exists a system of reactions having $G$ as its interaction graph and satisfying the following property: For any point $\vec{x}^0$ in the positive orthant and an arbitrary matrix $M$ whose interaction graph is $G$, there exists a choice of non-negative rates such that $\vec{x}^0$ is a fixed point of the network and the Jacobian equals $M$ at $\vec{x}^0$.
+
+We begin by noting that, since the form of the rate equations for reaction networks are invariant under rescaling the concentrations and rate constants, we can make the coordinates of the point $\vec{x}^0$ be $(1,1,\ldots,1)$.
+This will simplify the computation.
+
+Let $N$ be the number of nodes of $G$.
+Our reaction net will consist of $N$ species of reactants, $A_1, \ldots, A_N$, whose concentrations are $c_1, \ldots, c_N$.
+The reactions are defined as follows:
+% \begin{enumerate}
+% \item For every integer $1 \le i \le N$, we have the reactions $\emptyset \to A_i$, $A_i \to \emptyset$ and $2A_i \to 3A_i$.
+% \item For every pair of integers $1 \le i,j \le N$ such that $G_{ij} = 1$, we have the reactions $A_i + A_j \leftrightarrow A_j$.
+% \end{enumerate}
+\begin{equation}\label{eq:arbitraryjacobianreactionnetwork}
+\begin{aligned}
+\emptyset &\to A_i, & 1 \le i \le N,\\
+A_i &\to \emptyset, & 1 \le i \le N,\\
+2A_i &\to 3A_i, & 1 \le i \le N,\\
+A_i + A_j &\leftrightarrow A_j, & i \neq j,\, 1 \le i,j \le N,\, G_{ij} = 1.
+\end{aligned}
+\end{equation}
+
+The rate equations for such a system are:
+% \begin{widetext}
+\begin{equation}\label{eq:crnarbitraryjacobian}
+\begin{aligned}
+\frac{dc_i}{dt} = &F_i = k_{\emptyset \to A_i} - k_{A_i \to \emptyset} c_i + k_{2A_i \to 3A_i} c_i^2 \\
+&+ \sum_{\substack{1 \le j \le N \\ j \neq i \\ G_{ij} = 1}} k_{A_j \to A_i + A_j} c_j - k_{A_i + A_j \to A_j} c_i c_j
+\end{aligned}
+\end{equation}
+% \end{widetext}
+The Jacobian at $\vec{x}^0$ is given as
+% \begin{widetext}
+\begin{align*}
+\left.
+\frac{\partial F_i}{\partial c_i}\right|_{\vec{x}^0} &= - k_{A_i \to \emptyset} + 2 k_{2A_i \to 3A_i} - \sum_{\substack{1 \le j \le N \\ j \neq i \\ G_{ij} = 1}} k_{A_i + A_j \rightarrow A_j}, & \\
+\left. \frac{\partial F_i}{\partial c_j}\right|_{\vec{x}^0} &= k_{A_j \to A_i + A_j} - k_{A_i + A_j \to A_j}, &
+\end{align*}
+where $i \neq j$.
+% \end{widetext}
+By combining the equations $F_i(\vec{x}^0) = 0$ from \ref{eq:crnarbitraryjacobian} and $\frac{\partial F_i}{c_j}|_{\vec{x}^0} = M_{ij}$ we obtain the equivalent system of equations
+\begin{align}
+& k_{2A_i \to 3A_i} - k_{\emptyset \to A_i} = M_{ii} + \sum_{\substack{1 \le j \le N \\ j \neq i \\ G_{ij} = 1}} k_{A_j \to A_i + A_j} \label{eq:jacobianconstraint1}  \\
+& 2k_{2A_i \to 3A_i} - k_{A_i \to \emptyset} = M_{ii} + \sum_{\substack{1 \le j \le N \\ j \neq i \\ G_{ij} = 1}} k_{A_i + A_j \to A_j} \label{eq:jacobianconstraint2}\\
+& k_{A_j \to A_i + A_j} - k_{A_i + A_j \to A_j} = M_{ij} \label{eq:jacobianconstraint3}
+\end{align}
+We may solve these equations for the rate constants as follows.
+We begin by solving \ref{eq:jacobianconstraint3} by either choosing $k_{A_i + A_j \to A_j} \ge 0$ and setting $k_{A_j \to A_i + A_j} = M_{ij} + k_{A_i + A_j \to A_j}$ when $M_{ij} \ge 0$ or choosing $k_{A_j \to A_i + A_j} \ge 0$ and setting $k_{A_i + A_j \to A_j} = k_{A_j \to A_i + A_j} - M_{ij}$ when $M_{ij} < 0$.
+Pick
+% \begin{widetext}
+\begin{equation}
+\begin{aligned}
+k_{2A_i \to 3A_i} \ge \max \bigg(&0, M_{ii} + \sum_{\substack{1 \le j \le N \\ j \neq i \\ G_{ij} = 1}} k_{A_j \to A_i + A_j}, \\
+&M_{ii} + \sum_{\substack{1 \le j \le N \\ j \neq i \\ G_{ij} = 1}} k_{A_i + A_j \to A_j} \bigg).
+\end{aligned}
+\end{equation}
+% \end{widetext}
+Then we may solve \ref{eq:jacobianconstraint1} for $k_{\emptyset \to A_i}$ and \ref{eq:jacobianconstraint2} for $k_{A_i \to \emptyset}$ and obtain non-negative answers.
+This demonstrates that arbitrary Jacobian matrices can arise from reaction network ensembles that allow for the possibility of at least those reactions in \ref{eq:arbitraryjacobianreactionnetwork}.
+Note that \ref{eq:crnarbitraryjacobian}, \ref{eq:jacobianconstraint1}, \ref{eq:jacobianconstraint2}, and \ref{eq:jacobianconstraint3} are linear in the parameter values.
+Therefore, any probability distribution on the elements of the Jacobian can be obtained from a probability distribution on the parameter values.

### 2015-02-04 tex/discussion.tex
modify discussion and add full Taylor expansion

@@
 % Our development suggests two predictions that could be used to test the theory described here. One is that,
-Our theory predicts that, in general, an ensemble of systems where robustness has been the predominant object of selection and has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
-For example, at the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
+Our analysis predicts that, in general, an ensemble of systems where robustness has been the predominant object of selection and has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
 In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
+At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
 %The other prediction of the theory outlined here that may be able to be checked is the converse statement that an ensemble of systems demonstrating a positive or negative bias in their network topology toward hierarchy may have been under selection for, or respectively against, robustness.
@@
 Following May \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, this will involve improving the general understanding of the relationship between perturbations to a system's structure and the qualitative changes in the dynamical phenomena it can produce.
-%One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However,
-It may also be possible to prove bounds on the robustness of individual SCCs, which could provide more precise knowledge about the relative robustness of different network architectures.
 The conservation of robustness with respect to nontrivial symmetries including the interchange of SCCs and permutation within SCCs suggests the existence of an evolutionary neutral space.
 A deeper mathematical characterization of the full symmetry groupoid of dynamical robustness may thus help to characterize this potential evolutionary constraint \cite{Golubitsky2006}.
+%One approach is to attempt to compute robustness for systems of larger size. Using existing methods, this would require large-scale Monte Carlo simulations. However,
+%It may also be possible to prove bounds on the robustness of individual SCCs, which could provide more precise knowledge about the relative robustness of different network architectures.
+For some classes of systems, it may be possible to go beyond the linear approximation and corresponding local summary statistics of the phase space, such as dynamical robustness, to provide a more complete characterization of the relationship between network architecture and the global structure of the phase space of the corresponding ensemble of biological networks.
 

### 2015-02-04 tex/overview.tex
modify discussion and add full Taylor expansion

@@
 whose components are solutions to the arbitrary first order system
-$$
+\begin{equation}\label{eq:eom}
 \frac{dx_i(t)}{dt} = F_i(\vec{x}(t), \vec{p}), \; (i=1,\ldots,n)
-$$
+\end{equation}
 where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p}$ is the vector of parameters of the $F_i$.

### 2015-02-05 tex/introduction.tex
add robustness citations

@@
 In addition, there is often uncertainty about the very structure of such networks.
-In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems, which is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
+In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems.
+This is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
 This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems that, in comparison to one another, appear to have components that are randomly interlinked.
@@
 Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
-Robustness is of interest in biological systems at all scales, and has been previously studied, for example, in the context of gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}.
+Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical reaction networks \cite{Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}, and ecological networks \cite{Rohr2014}.
 For the purpose of this investigation, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.

### 2015-02-05 tex/introduction.tex
add figure reference to example of most hierarchical network

@@
 For the purpose of this investigation, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
-The maximally hierarchical network is considered to be the graph associated to the total ordering, which is the complete directed acyclic graph for a given system size \cite{Cormen2009}.
-Hence, we use a measure of hierarchy based upon the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
+The maximally hierarchical network is considered to be the graph associated to the reflexive total ordering \cite{Cormen2009}.
+An example of this for three system components is shown in \reffigscc{} top.
+We use a measure of hierarchy based upon the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
 We demonstrate that systems with the most hierarchical network topology according to this definition exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.

### 2015-02-05 tex/overview.tex
minor edits

@@
 
-To determine stability, we linearize the equations of motion about the
+To determine stability, we linearize the equations of motion (\refsupp{}) about the
 fixed point $\vec{x}^0$:
@@
 These interaction graphs can be viewed as deriving from the combination of system modules that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).
-Symmetries of the system are characterized by the ability to interchange these modules or their connectivity without changing some property of the system.
+Symmetries with respect to some property of the system are characterized by the ability to interchange these modules or their connectivity without changing that property.
 
@@
 In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$.
-This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of ${\vec{p}}\,'$ (\refsupp{}).
+This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of a different vector of parameters ${\vec{p}}\,'$ (\refsupp{}).
 We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.

### 2015-02-05 tex/overview.tex
add figure showing random Jacobians

@@
 
-We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$ the entries of the Jacobian are sampled independently from a generic probability distribution.
+We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
+The entries of the Jacobian are sampled independently from a generic probability distribution.
+Examples of vector fields that may result from this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}.
 The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)

### 2015-02-05 tex/overview.tex
minor edits

@@
 
-The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be further abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc).
+The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc).
 This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
@@
 \end{equation}
-For instance, if our graph is the one in \reffigscc (middle panels), then we have two connected components, one with two nodes, and one with one node.
+For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.
 From \ref{tab:structstabmat}, we know that the graph with two nodes has probability $0.25$ of being stable and robustness $0.62$.
 The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
-Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.
+Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc{}, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.
 
@@
 This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
-This analytical result predicts that any network whose associated dynamical system has the interaction graph \reffigscc $\,$ top will be more robust than those associated to any of the other interaction graphs in \reffigscc.
+This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the reflexive total ordering like that of \reffigscc{} top will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
 Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.

### 2015-02-05 tex/overview.tex
clarify explanation of symmetries

@@
 We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
-These interaction graphs can be viewed as deriving from the combination of system modules that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).
-Symmetries with respect to some property of the system are characterized by the ability to interchange these modules or their connectivity without changing that property.
+These interaction graphs can be viewed as deriving from the combination of system components that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).
 
 The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc).
-This map from the interaction graph of a network, referred to as $\hier$, has a collection of symmetries shown in \reffighiertransformations.
+This map from the interaction graph of a network, referred to as $\hier$, has a collection of intrinsic symmetries shown in \reffighiertransformations.
 These three symmetries represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated \reffigscc.
-Two of these three are also symmetries with respect to dynamical robustness.
-\ref{fig:robustnesssymmetries} shows an example of these symmetries applied to a specific interaction graph.
+Symmetries with respect to some property of the system are characterized by the ability to interchange these modules or their connectivity without changing that property.
+Two of these three intrinsic symmetries are also symmetries with respect to dynamical robustness.
+\ref{fig:robustnesssymmetries} shows an example of these latter symmetries applied to a specific interaction graph.
 

### 2015-02-05 tex/discussion.tex
minor edits

@@
 Our analysis predicts that, in general, an ensemble of systems where robustness has been the predominant object of selection and has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
-In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
+In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
 At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
@@
 Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
-In the long term, this prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that differentially select for robustness \cite{Leroi1994}.
+In the long term, this prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that differentially select for dynamical robustness \cite{Leroi1994}.
 

### 2015-02-05 tex/methods.tex
minor edits

@@
 %!TEX root = ../paper.tex
-\subsection{Monte carlo integration}
+\subsection{Monte carlo integration}\label{suppsec:montecarlo}
 If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
@@
 
-To justify our consideration of arbitrary Jacobian matrices, we determine a simple ensemble of reaction networks for which arbitrary Jacobian matrices are realizable.
+To justify our consideration of arbitrary Jacobian matrices in the case of reaction networks, we determine a simple ensemble for which arbitrary Jacobian matrices are realizable.
 This condition holds if one can solve for the parameter values of the system of equations corresponding to that ensemble in terms of the elements of an arbitrary Jacobian matrix.

### 2015-02-05 tex/overview.tex
relate interaction graph to Jacobian

@@
 % \subsection{Overview}
-The dynamical model given in terms of a system of differential equations for any network can be represented in terms of an interaction graph (\ref{fig:biomodelexamples} top row).
+The dynamical model given in terms of a system of differential equations for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
 For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ for $F_i \hbox{ depends on } x_j$ and $0$ for $F_i \hbox{ does not depend on } x_j$.
+These two conditions on global system interactions are expressed respectively in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ for all $\vec{x}$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$ for all $\vec{x}$.
 %  \begin{displaymath}
@@
 % \end{displaymath}
-For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be quite sparse.
-We can also associate a linearized interdependency graph $G_A$ given by an adjacency matrix $\adj(G_A)$ to each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ where
+For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be sparse.
+We can also associate a local interaction graph $G_A$ given by an adjacency matrix $\adj(G_A)$ to each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ where
  \begin{equation}\label{eq:lininterdepadj}
@@
 \end{equation}
-In general, the graph $G_A$ is a subgraph of $G_X$, however, for nearly all systems, $G_A$ is equivalent to $G_X$.
+In general, the graph $G_A$ is a subgraph of $G_X$, however, for nearly all systems, $G_A$ is equivalent to $G_X$ (\refsupp{}).
 We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.

### 2015-02-05 tex/overview.tex
minor edits

@@
 \end{equation}
-In general, the graph $G_A$ is a subgraph of $G_X$, however, for nearly all systems, $G_A$ is equivalent to $G_X$ (\refsupp{}).
+In general, the graph $G_A$ is a subgraph of $G_X$, however, $G_A$ is almost always equivalent to $G_X$ (\refsupp{}).
 We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
@@
 The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc).
-This map from the interaction graph of a network, referred to as $\hier$, has a collection of intrinsic symmetries shown in \reffighiertransformations.
-These three symmetries represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated \reffigscc.
+This map from the interaction graph of a network to its SCCs, referred to as $\hier$ in \ref{fig:modsccsym}, has a collection of intrinsic symmetries shown in \reffighiertransformations{}.
+These three symmetries, \reffigscc{}, represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated.
 Symmetries with respect to some property of the system are characterized by the ability to interchange these modules or their connectivity without changing that property.
@@
 The entries of the Jacobian are sampled independently from a generic probability distribution.
-Examples of vector fields that may result from this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}.
+Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}.
 The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)

### 2015-02-07 tex/abstract.tex
'Merge' abstract from 'science' branch

@@
 \begin{abstract}
-Here we analyze the stability and robustness of a large class of dynamical systems.
-We determine the probability distribution of robustness over system connectivity, which has significant implications from the study of metabolic to gene-regulatory to ecosystem network dynamics.
-We show that robustness is classified by the number of links between strongly connected components of the graph representing the underlying system connectivity leading to the conclusion that the most robust systems are also the most hierarchical.
-We also demonstrate that permutation of strongly connected components is a fundamental symmetry of dynamical robustness.
-This results in the classification of the dynamical robustness of biological networks based upon a purely topological property.
+The relationship between network topology and system dynamics has significant implications for unifying our understanding of the interplay among metabolic, gene-regulatory, and ecosystem network architecures.
+Here we analyze the stability and robustness of a large class of dynamics on such networks.
+We determine the probability distribution of robustness as a function of network topology and show that robustness is classified by the number of links between modules of the network.
+We also demonstrate that permutation of these modules is a fundamental symmetry of dynamical robustness.
+Analysis of these findings leads to the conclusion that the most robust systems have the most hierarchical structure.
+This relationship provides a means by which evolutionary selection for a purely dynamical phenomenon may shape network architectures across scales of the biological hierarchy.
 \end{abstract}

### 2015-02-07 tex/introduction.tex
'Merge' introduction from 'science' branch

@@
 
+% Some conclusions derived from this approach may be sensitive to the probability distribution from which the strengths of interconnection are sampled, whereas others are independent of the specific form of this distribution.
+In the former case, discovering the constraints that contribute to determining the functional form of the distribution can be a matter for empirical investigation and results that are themselves functions of this distribution can be reinterpreted in light of ongoing updates from relevant empirical observations.
+On the other hand, results that are independent of the specific form of this distribution can be expected to hold in any case and a specific distribution may simply be chosen for the purposes of exemplifying the result.
+
 Investigating generic properties of a large class of dynamical systems was the approach taken by May in models of ecosystem dynamics \cite{Gardner1970,May1972}.
@@
 May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
-
-Some conclusions derived from this approach may be sensitive to the probability distribution from which the strengths of interconnection are sampled, whereas others are independent of the specific form of this distribution.
-In the former case, discovering the constraints that contribute to determining the functional form of the distribution can be a matter for empirical investigation and results that are themselves functions of this distribution can be reinterpreted in light of ongoing updates from relevant empirical observations.
-On the other hand, results that are independent of the specific form of this distribution can be expected to hold in any case and a specific distribution may simply be chosen for the purposes of exemplifying the result.
 

### 2015-02-07 tex/figuressupp.tex
recombine figures
revise
update main text
minor edits
redistribute text between main text and supplement
'Merge' figuressupp from 'science' branch

@@
 %!TEX root = ../paper.tex
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.9\columnwidth]{fig/jacobianvectorfields.pdf}
+\caption{{\bf Vector fields resulting from the random sampling of two component systems.} System parameters of \ref{eq:arbitraryjacobianreactionnetwork} are rescaled to ensure the fixed point is located at $(1,1)$.
+The color of the dot located at the fixed point indicates whether it is stable (black) or unstable (gray).}
+\label{fig:jacobianvectorfields}
+\end{figure}
+
+\pagebreak
 
@@
 
+\pagebreak
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.6\columnwidth]{fig/apstab3x3.pdf}
+\caption{{\bf System stability as a function of connectivity.} The red points represent the average of system stability at each connectivity.}
+\label{fig:apstab3x3}
+\end{figure}
+
+\pagebreak
+\FloatBarrier
+
+\begin{figure}[!ht]
+\centering
+\noindent\includegraphics[width=0.6\columnwidth]{fig/cycle3x3.pdf}
+\caption{{\bf dynamical robustness as a function of number of cycles.} }
+\label{fig:cycle3x3}
+\end{figure}
+
 % \begin{figure}[!ht]
@@
 % \noindent\includegraphics[width=0.7\columnwidth]{fig/region2and3.pdf}
-% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
+% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three variable systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
 % \label{fig:region2and3}
@@
 % \noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
-% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
+% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two variable systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.
 Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
@@
 % \noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
-% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
+% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three variable systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$.
 This is the plane $s_3=1$.}

### 2015-02-07 tex/introduction.tex
uncomment inline figures

@@
 
-% \begin{figure*}[!ht]
-% \centering
-% \noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
-% \caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
+\begin{figure*}[!ht]
+\centering
+\noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
+\caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
 The second row represents a particular example of a system of differential equations that are used to model a biological network within each of the domains of application considered here.
 The third row shows the general form of a system of differential equations that can be used to model any network architecture within each domain.}
-% \label{fig:biomodelexamples}
-% \end{figure*}
+\label{fig:biomodelexamples}
+\end{figure*}
 

### 2015-02-09 tex/introduction.tex
add reference

--- 2015-02-07 tex/introduction.tex
@@
 Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
-Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical reaction networks \cite{Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Wagner2013}, and ecological networks \cite{Rohr2014}.
+Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical reaction networks \cite{Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}.
 For the purpose of this investigation, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.

### 2015-02-10 tex/introduction.tex
minor edit

@@
 In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems.
-This is necessary to understand in order to determine whether or not a particular dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
+This is necessary to understand in order to determine whether or not a given dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
 This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems that, in comparison to one another, appear to have components that are randomly interlinked.

### 2015-02-10 tex/overview.tex
add overview to paper.tex
minor edit

@@
 This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
-This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the reflexive total ordering like that of \reffigscc{} top will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
+This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the total ordering like that of \reffigscc{} top will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
 Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.

### 2015-02-10 tex/methods.tex
add definition of total order to supplement

@@
 %!TEX root = ../paper.tex
-\subsection{Monte carlo integration}\label{suppsec:montecarlo}
-If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
-We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}.
-The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
-$$
-\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
-$$
-where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
-$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
-which corresponds in this case to the maximum likelihood estimate
-$$
-\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
-$$
-This estimate is characterized by the variance of the posterior Beta distribution
-% \begin{widetext}
-\begin{align*}
-&\mathrm{var}(\theta | \mathcal{D}) =\\
-&\frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
-\end{align*}
-% \end{widetext}
-Since for the chosen prior $a=b=1 \ll N$ this simplifies to
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
-$$
-yielding the error estimate given by the associated standard deviation.
-In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
-
-\subsection{Reaction Networks with Prescribed Connecctivity and Jacobians}\label{sec:reactionnetjacobian}
+\section{Reaction Networks with Prescribed Connecctivity and Jacobians}\label{sec:reactionnetjacobian}
 
@@
 Therefore, any probability distribution on the elements of the Jacobian can be obtained from a probability distribution on the parameter values.
+
+\section{Monte carlo integration}\label{suppsec:montecarlo}
+If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
+We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}.
+The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
+$$
+\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
+$$
+where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
+$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
+which corresponds in this case to the maximum likelihood estimate
+$$
+\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
+$$
+This estimate is characterized by the variance of the posterior Beta distribution
+% \begin{widetext}
+\begin{align*}
+&\mathrm{var}(\theta | \mathcal{D}) =\\
+&\frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
+\end{align*}
+% \end{widetext}
+Since for the chosen prior $a=b=1 \ll N$ this simplifies to
+$$
+\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
+$$
+yielding the error estimate given by the associated standard deviation.
+In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
+
+\section{Hierarchy and Total ordering}\label{sec:totalordering}
+
+A directed graph $G=(V,E)$ is a set $V$ of nodes and a set $E$ of ordered pairs of nodes \cite{Cormen2009}.
+For example, if $V = \{1,2,3\}$ and $E = \{(1,1),(2,2),(3,3),(1,2),(1,3),(2,3)\}$ then $G=(V,E)$ is the graph depicted in \reffigscc{} top where the labels $1$, $2$, and $3$ have been respectively assigned to the nodes vertically from top to bottom.
+
+ We refer to the most hierarchical network architecture as the directed graph associated to a total ordering on the set of system components corresponding to the set of nodes, $V$, of the graph \cite{Cormen2009}.
+In general, a totally ordered set is a pair $(S,R)$ consisting of a set $S$ together with a total order relation $R$ on it.
+An example of a total ordering is the less than or equal to relation, $R \equiv \leq$, on the subset of natural numbers $S \equiv \{1,2,3\}$ given by $R \equiv \{1 \leq 1, 2 \leq 2, 3 \leq 3, 1 \leq 2, 1 \leq 3, 2 \leq 3\}$.
+The graph associated to this relation is equivalent to the graph shown in \reffigscc{} top and described algebraically in the preceding paragraph.
+More precisely, the conditions on $R$ for arbitrary elements $x,\,y,\,z \in S$ necessary for $(S,R)$ to be a totally ordered set are
+\begin{enumerate}
+\item If $x R y$ and $y R x$ then $x=y$ (antisymmetry)
+\item If $x R y$ and $y R z$ then $x R z$ (transitivity)
+\item $x R y$ or $y R x$ (totality)
+\end{enumerate}
+The totality condition implies $x R x$ (reflexivity) corresponding to the fact that the directed graph associated to the total ordering has, for each node, an edge whose source and target are the same node.

### 2015-05-19 tex/figures.tex
explain robustness definition in introduction

@@
 \label{fig:biomodelexamples}
+\end{figure*}
+
+\pagebreak
+
+\begin{figure*}[!htbp]
+\centering
+\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessconcept.pdf}
+\caption{{\bf Stability and robustness of biological networks.} Linearization around a steady-state of a model of a biological network allows for the assessment of the stability of that state.
+For two-dimensional systems, assessment of stability in terms of the trace and determinant of the Jacobian matrix associated to a particular steady-state geometrically partitions the plane into a stable region (S, blue) and an unstable region (U, red).
+The stochastic process induced on the stable and unstable regions that results from modifications to a given dynamical model over evolutionary timescales is represented by the state transition diagram (bottom middle).
+The probability of remaining in the stable region in the context of such modifications, given a previously existing stable network, is then provided by estimating $R$.}
+\label{fig:robustnessconcept}
 \end{figure*}

### 2015-05-19 tex/introduction.tex
minor edit
add definition of total order to supplement
explain robustness definition in introduction

@@
 Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
-Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical reaction networks \cite{Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}.
-For the purpose of this investigation, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the parameters for systems which have already been determined to be stable in the sense of linear stability analysis \cite{Davis1962}.
-The maximally hierarchical network is considered to be the graph associated to the reflexive total ordering \cite{Cormen2009}.
+Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical networks \cite{Alon1999,Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}.
+Over physiological timescales, the robustness of a particular network state may be evaluated by determining its linear stability \cite{Davis1962}.
+Network states that are linearly stable, are robust to perturbations in the states of their components.
+For example, in the case of gene-regulatory networks, a state vector containing protein molecule counts or concentrations of each gene that is stable with respect to the dynamics of the network will exponentially suppress any relatively small modifications to the state of a gene and return to the initial stable state.
+Here we are interested in evaluating robustness over evolutionary as opposed to physiological timescales.
+
+Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e.
+parameters, structure of the rate functions, state variables, etc.).
+It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states.
+However, what must remain invariant on such evolutionary timescales is the higher-level constraint that the system possess a stable attractor at all associated to the basin selected at the initiation of its dynamics.
+This is necessary in order for particular lower-level networks to remain \emph{observable} long enough to serve as the substrate out of which relatively higher-level ones are constructed \cite{Simon2002}.
+What is important over at least moderate evolutionary timescales is then the conditional probability $R$, given that a system possesses a stable state somewhere in its state space, that upon modifications to its structure, it continues to possess such a stable state, regardless of where that state is located within the state space \ref{fig:robustnessconcept}.
+For the purpose of this investigation then, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be stable in the sense of linear stability analysis.
+
+The maximally hierarchical network is considered to be the graph associated to the total ordering (\refsupp{} and \cite{Cormen2009}).
 An example of this for three system components is shown in \reffigscc{} top.

### 2015-05-20 tex/introduction.tex
update text

@@
 
-Here we determine the relationship between network hierarchy, a topological property, and the probability of \emph{robustness}, a dynamical property.
+Here we determine the relationship between network hierarchy, a topological property, and \emph{robustness}, a dynamical property.
 Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical networks \cite{Alon1999,Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}.
@@
 Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e.
-parameters, structure of the rate functions, state variables, etc.).
+parameters, structure of the rate functions, etc.).
 It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states.
-However, what must remain invariant on such evolutionary timescales is the higher-level constraint that the system possess a stable attractor at all associated to the basin selected at the initiation of its dynamics.
-This is necessary in order for particular lower-level networks to remain \emph{observable} long enough to serve as the substrate out of which relatively higher-level ones are constructed \cite{Simon2002}.
-What is important over at least moderate evolutionary timescales is then the conditional probability $R$, given that a system possesses a stable state somewhere in its state space, that upon modifications to its structure, it continues to possess such a stable state, regardless of where that state is located within the state space \ref{fig:robustnessconcept}.
-For the purpose of this investigation then, we quantify dynamical robustness as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be stable in the sense of linear stability analysis.
+However, what must remain invariant on such evolutionary timescales is the higher-level constraint that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process.
+This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon2002}.
+What is important over at least moderate evolutionary timescales is then the conditional probability, $R$, given a system is in a stable state, that upon modifications to its structure in the context of environmental fluctuations, it remains stable, regardless of where the stable state is located within the state space \ref{fig:robustnessconcept}.
+For the purpose of this investigation then, we quantify dynamical robustness in this way.
+% as the probability of stability to potentially large perturbations of the model specifying the system dynamics for systems which have already been determined to be linearly stable.
 
-The maximally hierarchical network is considered to be the graph associated to the total ordering (\refsupp{} and \cite{Cormen2009}).
+We demonstrate that systems exhibiting maximal robustness with respect to this definition have the most hierarchical network topology and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
+For the purpose of formulating this result, the maximally hierarchical network is considered to be the graph associated to the total ordering (\refsupp{} and \cite{Cormen2009}).
 An example of this for three system components is shown in \reffigscc{} top.
 We use a measure of hierarchy based upon the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
-We demonstrate that systems with the most hierarchical network topology according to this definition exhibit maximal robustness and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
-Moreover these results hold for networks of arbitrary size and independent of the probability distribution from which the strengths of interaction are sampled.
+Our results hold for networks of arbitrary size and are independent of the probability distribution from which the strengths of interaction are sampled.
 % \cite{WADDINGTON1942a,VanNimwegen1999,Siegal2002,Ciliberti2007b,Ciliberti2007,Wagner2013}

### 2015-05-20 tex/methods.tex
update text

@@
 % \end{widetext}
-By combining the equations $F_i(\vec{x}^0) = 0$ from \ref{eq:crnarbitraryjacobian} and $\frac{\partial F_i}{c_j}|_{\vec{x}^0} = M_{ij}$ we obtain the equivalent system of equations
+By combining the equations $F_i(\vec{x}^0) = 0$ from \ref{eq:crnarbitraryjacobian} and $\frac{\partial F_i}{\partial c_j}|_{\vec{x}^0} = M_{ij}$ we obtain the equivalent system of equations
 \begin{align}

### 2015-05-20 tex/overview.tex
update text

@@
 The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix (\refsupp{}).
-The Lotka-Volterra model has
+This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.
+As an example of a Jacobian matrix, the Lotka-Volterra model has
  \begin{equation}\label{eq:lotkavolterrajacobian}
@@
 These changes may alter the capacity of a given system to achieve a stable steady state.
-In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$.
+In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$ \ref{fig:robustnessconcept}.
 This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of a different vector of parameters ${\vec{p}}\,'$ (\refsupp{}).
@@
 
+This argument also works when we resample more than one entry,
+although the notation becomes more complicated.  Suppose that we
+resample $m$ nodes.  Define
+\begin{widetext}
+\begin{equation*}
+M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
+m = \sum_{i=0}^n m_i \quad\&\quad
+m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
+(\forall i \in \{1, \ldots, n\}) \; m_i < \connectivity_i \right\}.
+\end{equation*}
+\end{widetext}
+Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
+m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
+$m_0$ links between strongly connected components.  Hence, our
+weighted average becomes
+\begin{widetext}
+\begin{equation}
+\langle R(\hbox{stab} (G), \mu') \rangle_m =
+\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
+      {m \choose m_0, m_1, \ldots. m_n}
+      \left(m_0 + \sum\limits_{i=1}^n m_i \langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \right)}
+     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
+      {m \choose m_0, m_1, \ldots. m_n} m}
+\end{equation}
+\end{widetext}
+
+As before, since $\langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \le 1$, we may
+increase $\langle R(\hbox{stab} (G), \mu') \rangle_m$ by increasing the maximum
+possible value of $m_0$ while keeping the strongly connected
+components the same.  Again, if we fix $\hier(G)$, the maximum
+possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
+whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
+v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
+G_{tot}$.  Hence, we conclude that $\langle R(\hbox{stab} (G), \mu') \rangle_m \le
+\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_m$.
+
 This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.

### 2015-05-21 tex/overview.tex
describe Jacobian equivalence classes in main text

@@
 \end{equation}
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p}$ is the vector of parameters of the $F_i$.
+where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$.
 These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks, or cooperativity exponents.
 For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$.
+The set of all such dynamical systems, $\mathcal{D}$, is then
+\begin{equation}\label{eq:setdynsys}
+\mathcal{D} = \{ (F,\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
+\end{equation}
 
@@
 The Lotka-Volterra model has two fixed points: the trivial one of all zero species $n_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{N} b_{ij} n_j = 0$.
+The set of all such fixed points, $\mathcal{F}$, is then
+\begin{equation}\label{eq:setdynsys}
+\mathcal{F} = \{ (F,\vec{p},\vec{x}^0) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s, F(\vec{x}^0,\vec{p})=0 \}.
+\end{equation}
 
@@
 
+Jacobian matrices define an equivalence relation on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}',\vec{x}^{0'}) $ if and only if
+\begin{equation}\label{eq:jaceqrel}
+\left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} =
+\left. \frac{\partial F'(\vec{x},\vec{p}')}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^{0'}}.
+\end{equation}
+This relation then partitions the set of all fixed points into equivalence classes where the class $[A]$ of $\mathcal{F} / \sim$ associated to Jacobian matrix $A \in \mathbb{R}^{n \times n}$ is
+\begin{equation}\label{eq:jaceqs}
+[A] = \left\{ (F,\vec{p},\vec{x}^0) \; | \; \left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} = A \right\}.
+\end{equation}
+
 % \subsection{Overview}
 The dynamical model given in terms of a system of differential equations for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
-For a general system the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ for $F_i \hbox{ depends on } x_j$ and $0$ for $F_i \hbox{ does not depend on } x_j$.
-These two conditions on global system interactions are expressed respectively in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ for all $\vec{x}$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$ for all $\vec{x}$.
+For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$.
+These two conditions on global system interactions are expressed respectively for all $\vec{x}$ in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$.
 %  \begin{displaymath}
@@
 These interaction graphs can be viewed as deriving from the combination of system components that accept a given pattern of inputs and produce a given pattern of outputs (\reffigexamplesystemmodules).
+
+Each distinct directed graph $G$ selects a subset of $\mathcal{F} / \sim$
+\begin{equation}\label{eq:jaceqs}
+[G] = \left\{ [A] \; | \; G_A = G \right\}.
+\end{equation}
+The $G$-classes thereby partition the collection of fixed points, $\mathcal{F}$, over the space of dynamical systems, $\mathcal{D}$, according to the interactions among the variables of the dynamical system represented by the topology of $G$.
+This partition, $\mathcal{F} / G$, is a coarsening of $\mathcal{F} / \sim$.
 

### 2015-05-21 tex/figures.tex
add new robustnessconcept fig
update: incorporate references to new conceptual robustness figures

--- 2015-01-14 tex/figures.tex
@@
 \centering
-\noindent\includegraphics[width=0.8\textwidth]{fig/biomodelexamples.pdf}
+\noindent\includegraphics[width=1.0\textwidth]{fig/biomodelexamples.pdf}
 \caption{{\bf Dynamical models in systems biology.} The top row represents a chemical reaction network (CRN) \cite{Shinar2010}, a gene regulatory network (GRN) \cite{Karlebach2008}, and an ecological regulatory network (ERN) \cite{Rohr2014} in terms of the graphical methods specific to each field mapped into the interaction graph, which provides a unified representation for networks across these fields.
@@
 
+\begin{figure*}[!htbp]
+\centering
+\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessconceptvis.pdf}
+\caption{{\bf Stability and equivalence classes of fixed points.} Linearization around a steady-state of a model of a biological network allows for the assessment of the stability of that state.
+For two-dimensional systems, assessment of stability in terms of the trace and determinant of the Jacobian matrix associated to a particular steady-state geometrically partitions the plane into a stable region (S, blue) and an unstable region (U, red).
+For larger systems in higher dimensions this geometry is too complicated to visualize, but it is nevertheless well-defined and can be investigated algebraically and numerically.
+(top center) Jacobian matrices such as $A_1$ partition the space of fixed points into equivalence classes $[A_1]$, each of which is either stable or unstable.}
+\label{fig:robustnessconcept}
+\end{figure*}
+
+\pagebreak
+
+\begin{figure*}[!htbp]
+\centering
+\noindent\includegraphics[width=1.0\textwidth]{fig/robustnessprocess.pdf}
+\caption{{\bf Assessment of dynamical robustness.} (A) Any stochastic process induced on the stable and unstable regions of the space of fixed points can be represented as a state transition diagram.
+The probability of remaining in the stable region in the context of such modifications, given a previously existing stable state, is then provided by estimating $R$.
+(B) The stochastic process that results from modifications to a given dynamical model over evolutionary timescales induces a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ represented by the boxes in the top row.
+This in turn induces a process on the stable and unstable regions, which can be used to estimate robustness.
+(C) Each network architecture encoded by a directed graph, $\{G_1, \ldots, G_k \}$, that indicates the manner in which the variables of a biological network depend upon one another selects a particular subset of equivalence classes from $\mathcal{F}/{\sim}$ (see \ref{eq:jacgrapheqs}).
+Robustness can therefore be estimated for each network architecture independently.}
+\label{fig:robustnessprocess}
+\end{figure*}
+
+\pagebreak
+
 \begin{figure*}[!ht]
 \centering
-\noindent\includegraphics[width=0.8\textwidth]{fig/modsccsym.pdf}
+\noindent\includegraphics[width=1.0\textwidth]{fig/modsccsym.pdf}
 \caption{{\bf Open systems, strongly connected components and symmetries of robustness.} (A) Example of the combination of open system modules to construct closed systems.
@@
 \centering
-\noindent\includegraphics[width=0.8\textwidth]{fig/combinedfigs.pdf}
+\noindent\includegraphics[width=1.0\textwidth]{fig/combinedfigs.pdf}
 \caption{{\bf Characterization of stability and robustness according to properties of system structure for three variable systems} (A) Robustness versus connectivity.

### 2015-05-21 tex/introduction.tex
update: incorporate references to new conceptual robustness figures

@@
 The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-All of these models utilize essentially equivalent mathematical structures, \ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}.
-Developing unified mathematical descriptions of all of these is one of the paramount goals of systems biology.
+All of these models utilize essentially equivalent mathematical structures, \ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}, and developing unified mathematical descriptions of them is one of the paramount goals of systems biology.
 % \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}
@@
 For example, in the case of gene-regulatory networks, a state vector containing protein molecule counts or concentrations of each gene that is stable with respect to the dynamics of the network will exponentially suppress any relatively small modifications to the state of a gene and return to the initial stable state.
-Here we are interested in evaluating robustness over evolutionary as opposed to physiological timescales.
 
+Here, rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change.
 Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e.
@@
 It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states.
-However, what must remain invariant on such evolutionary timescales is the higher-level constraint that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process.
+However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process subject to environmental constraints.
 This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon2002}.

### 2015-05-21 tex/overview.tex
update: incorporate references to new conceptual robustness figures

@@
 \end{equation}
-where $F_i$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$.
-These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks, or cooperativity exponents.
+where $F=\{F_1,\ldots,F_i,\ldots,F_n\}$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$.
+These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks.
 For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$.
-The set of all such dynamical systems, $\mathcal{D}$, is then
+The set of all such dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $s$, is then
 \begin{equation}\label{eq:setdynsys}
@@
 The set of all such fixed points, $\mathcal{F}$, is then
-\begin{equation}\label{eq:setdynsys}
+\begin{equation}\label{eq:setfixedpoints}
 \mathcal{F} = \{ (F,\vec{p},\vec{x}^0) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s, F(\vec{x}^0,\vec{p})=0 \}.
@@
 
-Jacobian matrices define an equivalence relation on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}',\vec{x}^{0'}) $ if and only if
+Jacobian matrices define an equivalence relation, $\sim$, on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}',\vec{x}^{0'}) $ if and only if
 \begin{equation}\label{eq:jaceqrel}
@@
 \end{equation}
-This relation then partitions the set of all fixed points into equivalence classes where the class $[A]$ of $\mathcal{F} / \sim$ associated to Jacobian matrix $A \in \mathbb{R}^{n \times n}$ is
+This relation then partitions the set of all fixed points into equivalence classes $\mathcal{F}/{\sim}$, where the class $[A]$ associated to Jacobian matrix $A \in \mathbb{R}^{n \times n}$ is
 \begin{equation}\label{eq:jaceqs}
@@
 \end{equation}
+An example with two members of $\mathcal{F}$ from each of four different equivalence classes $[A_1]$, $[A_2]$, $[A_3]$, and $[A_4]$ of $\mathcal{F}/{\sim}$ is shown in \ref{fig:robustnessconcept}.
 
 % \subsection{Overview}
-The dynamical model given in terms of a system of differential equations for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
+The interactions among variables in a dynamical model for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
 For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$.
@@
 
-Each distinct directed graph $G$ selects a subset of $\mathcal{F} / \sim$
-\begin{equation}\label{eq:jaceqs}
+Each distinct directed graph $G$, of which there are $k=2^{n(n-1)}$ that could be associated to the interactions in a model defined on $n$ variables, selects a subset of $\mathcal{F} / {\sim}$ (see~\ref{fig:robustnessprocess}C)
+\begin{equation}\label{eq:jacgrapheqs}
 [G] = \left\{ [A] \; | \; G_A = G \right\}.
@@
 The $G$-classes thereby partition the collection of fixed points, $\mathcal{F}$, over the space of dynamical systems, $\mathcal{D}$, according to the interactions among the variables of the dynamical system represented by the topology of $G$.
-This partition, $\mathcal{F} / G$, is a coarsening of $\mathcal{F} / \sim$.
+This partition, $\mathcal{F} / G$, is a coarsening of $\mathcal{F} / {\sim}$.
+
+In the process of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
+These modifications induce a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ \ref{fig:robustnessprocess}B.
+These changes may likewise alter the capacity of a given system to achieve a stable steady state, thereby inducing a coarse-grained stochastic process on the stable regions of the space of fixed points \ref{fig:robustnessprocess}A and B.
+In order to evaluate this we consider a process whereby perturbations are applied to a given dynamical system and fixed point associated to a stable matrix $A$ leading to $A'$.
+In order to classify the properties of this process according to network architecture, we enforce the condition $\adj(G_A) = \adj(G_{A'})$, which results in analogous processes defined on each $G$-class of $\mathcal{F}/G$ \ref{fig:robustnessconcept}C.
+This corresponds to an ensemble of dynamical systems with equivalent connectivities, but where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{}).
+We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
+This conditional probability corresponds to the parameter $R$ of the two-state process depicted in \ref{fig:robustnessprocess}A, and it is what we refer to as dynamical robustness.
+Comparing the values of $R$ for each $G$-class in $\mathcal{F}/G$ places a partial ordering on network architectures, $G$, which allows for the determination of which network architectures would be expected to be enriched in a process selecting for higher $R$ values.
 
@@
 
-In the process of biological evolution, the parameter values $\vec{p}$ or the form of the functions $F_i$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due, for example, to mutations or environmental fluctuations.
-These changes may alter the capacity of a given system to achieve a stable steady state.
-In order to evaluate this we consider a stochastic process whereby for a given dynamical system associated to a stable matrix $A$, we consider a perturbation of the system leading to $A'$ where $\adj(G_A) = \adj(G_{A'})$ \ref{fig:robustnessconcept}.
-This corresponds to a dynamical system whose connectivity is preserved, but where the underlying model is defined in terms of a different vector of parameters ${\vec{p}}\,'$ (\refsupp{}).
-We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
-This probability is what we refer to as dynamical robustness.
-
 We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
 The entries of the Jacobian are sampled independently from a generic probability distribution.
-Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:jacobianvectorfields}.
+Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
 The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)

### 2015-05-22 tex/figures.tex
incorporate explicit definitions of relevant stochastic processes

@@
 \caption{{\bf Assessment of dynamical robustness.} (A) Any stochastic process induced on the stable and unstable regions of the space of fixed points can be represented as a state transition diagram.
-The probability of remaining in the stable region in the context of such modifications, given a previously existing stable state, is then provided by estimating $R$.
+The conditional probability of remaining in the stable region in the context of such modifications, given a previously existing stable state, is then provided by estimating $R$.
 (B) The stochastic process that results from modifications to a given dynamical model over evolutionary timescales induces a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ represented by the boxes in the top row.
 This in turn induces a process on the stable and unstable regions, which can be used to estimate robustness.
+We note that $R$ is distinct from the stationary probability of being within the stable region $p^{\infty}(S) = \frac{1-U}{2-R-U}$.
 (C) Each network architecture encoded by a directed graph, $\{G_1, \ldots, G_k \}$, that indicates the manner in which the variables of a biological network depend upon one another selects a particular subset of equivalence classes from $\mathcal{F}/{\sim}$ (see \ref{eq:jacgrapheqs}).
-Robustness can therefore be estimated for each network architecture independently.}
+Robustness can therefore be estimated independently for each network architecture.}
 \label{fig:robustnessprocess}

### 2015-05-22 tex/overview.tex
incorporate explicit definitions of relevant stochastic processes

@@
 For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$.
-The set of all such dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $s$, is then
+The set of all dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $s$, is then
 \begin{equation}\label{eq:setdynsys}
-\mathcal{D} = \{ (F,\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
+\mathcal{D} = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
 \end{equation}
@@
 \begin{equation}\label{eq:setfixedpoints}
-\mathcal{F} = \{ (F,\vec{p},\vec{x}^0) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s, F(\vec{x}^0,\vec{p})=0 \}.
+\mathcal{F} = \{ (d,\vec{x}^0) \; | \; d=(F(-,\vec{p}),\vec{p}) \in \mathcal{D}, \vec{x}^0 \in \mathbb{R}^n,F(\vec{x}^0,\vec{p})=0 \}.
 \end{equation}
@@
 The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix (\refsupp{}).
-This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$ \cite{Gantmacher1959}.
+This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$.
+In the 2-dimensional case, $\chi(A) = z^2 + Tr(A)z+Det(A)$ has solutions $z$ with negative real parts if $Tr(A)<0$ and $Det(A)>0$, which we make use of in examples.
+Generalized conditions for higher dimensions are available in \cite{Gantmacher1959}.
 As an example of a Jacobian matrix, the Lotka-Volterra model has
@@
 \end{equation}
+Evaluation of the stability criterion occurs on a space of two states $\mathcal{S} = \{ S, U \}$ where $S$ stands for stable and $U$ stands for unstable.
+The stability criterion defines an equivalence relation on the set of all Jacobian matrices $A \in \mathbb{R}^{n \times n}$ deriving from fixed points on $n$ variables that simply splits the set into two classes $S=\{ A \, | \, A \hbox{ is stable}  \}$ and $U~=~\{ A \, | \, A \hbox{ is unstable} \}$.
 
-Jacobian matrices define an equivalence relation, $\sim$, on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}',\vec{x}^{0'}) $ if and only if
+Jacobian matrices define an equivalence relation, $\sim$, on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}\,',\vec{x}^{0'}) $ if and only if
 \begin{equation}\label{eq:jaceqrel}
 \left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} =
-\left. \frac{\partial F'(\vec{x},\vec{p}')}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^{0'}}.
+\left. \frac{\partial F'(\vec{x},\vec{p}\,')}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^{0'}}.
 \end{equation}
@@
 
-In the process of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
-These modifications induce a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ \ref{fig:robustnessprocess}B.
-These changes may likewise alter the capacity of a given system to achieve a stable steady state, thereby inducing a coarse-grained stochastic process on the stable regions of the space of fixed points \ref{fig:robustnessprocess}A and B.
-In order to evaluate this we consider a process whereby perturbations are applied to a given dynamical system and fixed point associated to a stable matrix $A$ leading to $A'$.
-In order to classify the properties of this process according to network architecture, we enforce the condition $\adj(G_A) = \adj(G_{A'})$, which results in analogous processes defined on each $G$-class of $\mathcal{F}/G$ \ref{fig:robustnessconcept}C.
+In the course of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
+The stochastic process by which these modifications occur induces one on the set of fixed points that results in the assignment of a probability $P(f^T)$ for each history of length $T$, $f^T = \{ f_1,f_2,\ldots,f_T \} \in \mathcal{F}^T$.
+% $$
+% P(f',t+1 | f,t) = \mathcal{T}_{\mathcal{F}}(f',f)
+% $$
+% for all $f,f' \in \mathcal{F}$.
+These dynamics induce a stochastic process on the equivalence classes of fixed points $\mathcal{F}/{\sim}$ indexed by Jacobian matrices given by
+$$
+P([A]^T) = \sum_{ \{ f^T \mid f_i^T \in [A]_i^T \} } P(f^T)
+$$
+for each history of length $T$, $[A]^T = \{ [A_1], [A_2], \ldots, [A_T]\} \in (\mathcal{F}/{\sim})^T$
+% $$
+% P([A'],t+1 | [A],t) = \mathcal{T}_{\mathcal{F}/{\sim}}([A'],[A])
+% $$
+% for all $A,A' \in \mathcal{F}/{\sim}$
+as visualized in \ref{fig:robustnessprocess}B.
+These changes can alter the capacity of a given system to achieve a stable steady state, thereby inducing an even more coarse-grained stochastic process on the stable regions of the space of fixed points given by
+$$
+P(s^T) = \sum_{ \{ A^T \mid A_i^T \in s_i^T \} } P(A^T)
+$$
+for each history of length $T$, $s^T = \{ s_1, s_2, \ldots, s_T \} \in \mathcal{S}^T$
+as visualized in \ref{fig:robustnessprocess}A and B.
+In order to model this we consider a process whereby perturbations are applied to a given dynamical system and fixed point associated to a stable matrix $A$ leading to $A'$.
+In order to classify the properties of this process according to network architecture, we enforce the condition $\adj(G_A) = \adj(G_{A'})$, which results in an analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
+$$
+P([A]^T \mid G_A {=} G) = \sum_{ \{ f^T \mid f_i^T \in [A]_i^T \} } P(f^T)
+$$
+for each history of length $T$, $[A]^T = \{ [A_1], [A_2], \ldots, [A_T]\} \in [G]^T$
+as visualized in \ref{fig:robustnessprocess}C.
 This corresponds to an ensemble of dynamical systems with equivalent connectivities, but where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{}).
 We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
+This quantifies the intuitive statement that, over evolutionary timescales, it is not enough for a system to be stable, but rather that it have a reasonably high probability of remaining stable over continguous timeframes.
+In terms of a history $s^T$ over the states specifying the stability property, $\mathcal{S}$, this is given by
+$$
+r = \frac{\sum_{t=0}^{T-1}s_{t+1}^T s_{t}^T}{\sum_{t=0}^{T-1} s_{t}^T},
+$$
+where we make numerical substitutions for the values of $\mathcal{S}$ such that $S = 1$ and $U = 0$.
+Given that the process $P(s^T)$ has stationary probability $P^{\infty}(s)$ for $s \in \mathcal{S}$, the expectation of $r$ as $T \rightarrow \infty$ is approximated by
+$$
+R = E[r] \approx \frac{\sum_{s,s'} P^{\infty}(s)P^{\infty}(s' \mid s)}{P^{\infty}(S)}.
+$$
 This conditional probability corresponds to the parameter $R$ of the two-state process depicted in \ref{fig:robustnessprocess}A, and it is what we refer to as dynamical robustness.
@@
 Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
-The expression in \ref{eq:sccrobustness} holds for the case of simultaneously perturbing any number of elements of the Jacobian, but can be schematized for the case of perturbing a single element as \ref{eq:robschematic} (see \reffigscc$\,$ for examples demonstrating this expression)
+The result ultimately holds for the case of simultaneously perturbing any number of elements of the Jacobian.
+To anchor the intuition before stating the more general result, \ref{eq:robschematic} shows a schematized version for the case of perturbing a single element  (see \reffigscc$\,$ for examples demonstrating this expression)
 \begin{equation}\label{eq:robschematic}

### 2015-05-25 tex/overview.tex
minor changes

@@
 
-In the course of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
-The stochastic process by which these modifications occur induces one on the set of fixed points that results in the assignment of a probability $P(f^T)$ for each history of length $T$, $f^T = \{ f_1,f_2,\ldots,f_T \} \in \mathcal{F}^T$.
+In the course of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to the basins associated to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
+The stochastic process by which these modifications occur induces one on the set of fixed points that results in the assignment of a probability $P(f^T)$ to each history of length $T$, $f^T = ( f_1,f_2,\ldots,f_T ) \in \mathcal{F}^T$.
 % $$
@@
 $$
-for each history of length $T$, $[A]^T = \{ [A_1], [A_2], \ldots, [A_T]\} \in (\mathcal{F}/{\sim})^T$
+for each history of length $T$, $[A]^T = ( [A_1], [A_2], \ldots, [A_T] ) \in (\mathcal{F}/{\sim})^T$
 % $$
@@
 $$
-for each history of length $T$, $s^T = \{ s_1, s_2, \ldots, s_T \} \in \mathcal{S}^T$
+for each history of length $T$, $s^T = ( s_1, s_2, \ldots, s_T ) \in \mathcal{S}^T$
 as visualized in \ref{fig:robustnessprocess}A and B.
@@
 $$
-for each history of length $T$, $[A]^T = \{ [A_1], [A_2], \ldots, [A_T]\} \in [G]^T$
+for each history of length $T$, $[A]^T = ( [A_1], [A_2], \ldots, [A_T] ) \in [G]^T$
 as visualized in \ref{fig:robustnessprocess}C.

### 2015-05-25 tex/introduction.tex
minor changes
modify intro

@@
 The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
-All of these models utilize essentially equivalent mathematical structures, \ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}, and developing unified mathematical descriptions of them is one of the paramount goals of systems biology.
+Over evolutionary timescales, one expects to observe changes in the model of best fit.
+All of these models utilize essentially equivalent mathematical structures, instances of which are sampled in the evolutionary process (\ref{fig:biomodelexamples}, \cite{RossCr2003,Palsson2011a,Sauro2012}).
+Developing unified mathematical descriptions of each of these that can be embedded into models of the evolutionary process is one of the paramount goals of systems biology.
 % \cite{RossCr2003,Alon2006,Palsson2006,HamidBolouri2008,Palsson2011a,Voit2012,Sauro2012}
@@
 In addition, there is often uncertainty about the very structure of such networks.
+Since the evolutionary process results in modifications to the underlying model this fact may be used to characterize evolutionarily effective versus neutral spaces.
 In this context, it is crucial to gain insight into what dynamical phenomena are possible to observe within a given class of dynamical systems.
@@
 
-Here, rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change.
+Rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change.
 Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e.
-parameters, structure of the rate functions, etc.).
-It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states.
+parameters, structure of the rate functions, etc.), and these changes may occur at any level of the hierarchy including metabolic, gene-regulatory, population, and ecosystem.
+It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states and even to the geometry of their state spaces.
 However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process subject to environmental constraints.

### 2015-05-28 tex/overview.tex
define robustness in terms of limiting distributions on underlying stochastic processes

@@
 as visualized in \ref{fig:robustnessprocess}B.
-These changes can alter the capacity of a given system to achieve a stable steady state, thereby inducing an even more coarse-grained stochastic process on the stable regions of the space of fixed points given by
+These changes can alter the stability of a given system, thereby inducing an even more coarse-grained stochastic process on the stable regions of the space of fixed points given by
 $$
@@
 $$
-r = \frac{\sum_{t=0}^{T-1}s_{t+1}^T s_{t}^T}{\sum_{t=0}^{T-1} s_{t}^T},
+r = \frac{\sum_{t=0}^{T-1} \mathbf{1}_S (s_{t+1}^T) \mathbf{1}_S (s_{t}^T)}{\sum_{t=0}^{T-1} \mathbf{1}_S (s_{t}^T)},
 $$
-where we make numerical substitutions for the values of $\mathcal{S}$ such that $S = 1$ and $U = 0$.
-Given that the process $P(s^T)$ has stationary probability $P^{\infty}(s)$ for $s \in \mathcal{S}$, the expectation of $r$ as $T \rightarrow \infty$ is approximated by
+where we have introduced the indicator function $\mathbf{1}_S$ on $\mathcal{S}$ such that $\mathbf{1}_S(S) = 1$ and $\mathbf{1}_S(U) = 0$.
+If the process $P(s^T)$ has limits $\mu(s) = \lim_{t \rightarrow \infty} P(s_t^T=s)$ and $\tau(s' \mid s) = \lim_{t \rightarrow \infty} P(s_{t+1}^T = s' \mid s_t^T = s)$ for $s,s' \in \mathcal{S}$
+%stationary probability $P^{\infty}(s)$ for $s \in \mathcal{S}$
+, the expectation of $r$ as $T \rightarrow \infty$ is approximated by
 $$
-R = E[r] \approx \frac{\sum_{s,s'} P^{\infty}(s)P^{\infty}(s' \mid s)}{P^{\infty}(S)}.
+R(S,\mu,\tau) = E[r] \approx \frac{\sum_{s,s'} \mu(s) \tau(s' \mid s)}{\sum_{s=S} \mu(s)}.
 $$

### 2015-05-29 tex/overview.tex
begin robustness derivation in new notation

--- 2015-05-25 tex/overview.tex
@@
 
-To determine stability, we linearize the equations of motion (\refsupp{}) about the
-fixed point $\vec{x}^0$:
+% To determine stability, we linearize the equations of motion (\refsupp{}) about the
+% fixed point $\vec{x}^0$:
+% \begin{equation}\label{eq:lineardynsys}
+% \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
+% \end{equation}
+% where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$ Jacobian matrix $A$ has components
+% $$
+% a_{ij} = \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0}.
+% $$
+To determine stability, we use the Taylor series expansion of the equations of motion \ref{eq:eom} about the fixed point $\vec{x}^0$ where $\vec{y} = \vec{x} - \vec{x}^0$ by
+\begin{equation}\label{eq:taylorseries}
+\begin{aligned}
+\frac{dx_i(t)}{dt} \approx & F_i(\vec{x}^0)
++ \sum_{j=1}^{N} \left. \frac{\partial F_i}{\partial x_j} \right|_{\vec{x} = \vec{x}^0} y_j\\
+& + \frac{1}{2}\sum_{j,k=1}^{N} \left. \frac{\partial^2 F_i}{\partial x_j \partial x_k} \right|_{\vec{x} = \vec{x}^0} y_j y_k + \cdots
+\end{aligned}
+\end{equation}
+The zeroth order term vanishes since $F_i(\vec{x}^0)=0$ by definition and thus neglecting terms higher than first order from \ref{eq:taylorseries} results in
+% \begin{equation*}
+% \frac{d\vec{y}(t)}{dt} = A \vec{y}(t),
+% \end{equation*}
 \begin{equation}\label{eq:lineardynsys}
@@
 \end{equation}
-where $\vec{y} = \vec{x} - \vec{x}^0$ and the $n \times n$ Jacobian matrix $A$ has components
+where the $n \times n$-matrix $A$ has components
 $$
@@
 $$
-The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix (\refsupp{}).
+
+The system defined by $F_i$, $\vec{p}$, and $\vec{x}^0$ is dynamically stable if the eigenvalues of $A$ all have real parts less than zero and $A$ is then referred to as a stable matrix.
+The spectral abscissa of the matrix $A$ is defined as
+$$
+\eta(A) = \max_i \{\Re(\lambda_i)\}
+$$
+where $\lambda_i$ are the eigenvalues of $A$.
+The system defined by $F_i$ and $\vec{x}^0$ is dynamically stable if the spectral abscissa of $A$ is less than zero, equivalently, $\eta(A) < 0$.
+This is because the general solution to \ref{eq:lineardynsys} is
+$$
+y_i(t) = \sum_j b_{ij} e^{\lambda_j t}, \; (i=1,\ldots,n)
+$$
+for some matrix $B=(b_{ij})$ and thus all $\vec{y} = \vec{x} - \vec{x}^0$ decay to zero when all $\lambda_i < 0$.
+
 This criterion can be checked equivalently in terms of conditions on the coefficients of the characteristic polynomials $\chi(A)$ associated to the systems described by matrices $A$.
-In the 2-dimensional case, $\chi(A) = z^2 + Tr(A)z+Det(A)$ has solutions $z$ with negative real parts if $Tr(A)<0$ and $Det(A)>0$, which we make use of in examples.
+In the $2$-dimensional case, $\chi(A) = \lambda^2 + Tr(A)\lambda+Det(A)$ has solutions $\lambda$ with negative real parts if $Tr(A)<0$ and $Det(A)>0$, which we make use of in examples.
 Generalized conditions for higher dimensions are available in \cite{Gantmacher1959}.
@@
 \end{equation}
-Evaluation of the stability criterion occurs on a space of two states $\mathcal{S} = \{ S, U \}$ where $S$ stands for stable and $U$ stands for unstable.
+
+Evaluation of the stability criterion occurs on a space of two states inducing a mapping from matrices to binary values $\mathcal{S} \colon \mathbb{R}^{n \times n} \rightarrow \{ 1, 0 \}$ given by
+ \begin{equation}\label{eq:stabeval}
+   \mathcal{S}(A) = \left\{
+     \begin{array}{lr}
+       1, & \eta (A) < 0\\
+       0, & \eta (A) \geq 0
+     \end{array},
+   \right.
+\end{equation}
+where $1$ stands for $S$ or stable and $0$ stands for $U$ or unstable.
 The stability criterion defines an equivalence relation on the set of all Jacobian matrices $A \in \mathbb{R}^{n \times n}$ deriving from fixed points on $n$ variables that simply splits the set into two classes $S=\{ A \, | \, A \hbox{ is stable}  \}$ and $U~=~\{ A \, | \, A \hbox{ is unstable} \}$.
@@
 as visualized in \ref{fig:robustnessprocess}B.
-These changes can alter the capacity of a given system to achieve a stable steady state, thereby inducing an even more coarse-grained stochastic process on the stable regions of the space of fixed points given by
+These changes can alter the stability of a given system, thereby inducing an even more coarse-grained stochastic process on the stable regions of the space of fixed points given by
 $$
-P(s^T) = \sum_{ \{ A^T \mid A_i^T \in s_i^T \} } P(A^T)
+P(s^T) = \sum_{ \{ A^T \mid \mathcal{S}(A_i^T) = s_i^T \} } P(A^T)
 $$
-for each history of length $T$, $s^T = ( s_1, s_2, \ldots, s_T ) \in \mathcal{S}^T$
+for each history of length $T$, $s^T = ( s_1, s_2, \ldots, s_T ) \in \{0,1\}^T$
 as visualized in \ref{fig:robustnessprocess}A and B.
-In order to model this we consider a process whereby perturbations are applied to a given dynamical system and fixed point associated to a stable matrix $A$ leading to $A'$.
-In order to classify the properties of this process according to network architecture, we enforce the condition $\adj(G_A) = \adj(G_{A'})$, which results in an analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
+In order to model this we consider a process whereby perturbations applied to a given dynamical system and fixed point associated to a stable Jacobian matrix $A$ lead to another Jacobian matrix $A'$.
+This corresponds to an ensemble of fixed points of dynamical systems where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{}).
+We then ask what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
+This quantifies the intuitive statement that, over evolutionary timescales, it is not enough for a system to be stable, but rather that it have a reasonably high probability of remaining stable over continguous timeframes.
+In terms of a history $s^T$ over the states specifying the stability property, $\mathcal{S}$, this is given by
+$$
+r = \frac{\sum_{t=0}^{T-1} s_{t+1}^T s_{t}^T}{\sum_{t=0}^{T-1} s_{t}^T}.
+$$
+If the process $P(s^T)$ has limits $\mu(s) = \lim_{t \rightarrow \infty} P(s_t^T=s)$ and $\tau(s' \mid s) = \lim_{t \rightarrow \infty} P(s_{t+1}^T = s' \mid s_t^T = s)$ for $s,s' \in \{0,1\}$
+%stationary probability $P^{\infty}(s)$ for $s \in \mathcal{S}$
+, the expectation of $r$ as $T \rightarrow \infty$ is approximated by
+\begin{equation}\label{eq:robustnessfromstoch}
+R(S,\mu,\tau) = E[r] \approx \frac{\sum_{s,s'} \mu(s) \tau(s' \mid s)}{\sum_{s=1} \mu(s)}.
+\end{equation}
+This conditional probability corresponds to the parameter $R$ of the two-state process depicted in \ref{fig:robustnessprocess}A, and it is what we refer to as dynamical robustness.
+
+In terms of Jacobian matrices $A$ and $A'$, if the underlying process $P([A]^T)$ has analogous limits
+$$\mu(A) = \lim_{t \rightarrow \infty} P(A_t^T=A)$$
+and
+$$\tau(A' \mid A) = \lim_{t \rightarrow \infty} P(A_{t+1}^T = A' \mid A_t^T = A)$$
+for $A,A' \in \mathbb{R}^{n \times n}$, \ref{eq:robustnessfromstoch} becomes
+\begin{equation}
+R(S,\mu,\tau) = \frac{\int_{A,A' \in \mathbb{R}^{n \times n}} d\mu(A) d\tau(A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}{\int_{A \in \mathbb{R}^{n \times n}} d\mu(A) \mathcal{S}(A)}.
+\end{equation}
+
+In order to classify the properties of this process according to network architecture, we consider cases where the condition $\adj(G_A) = \adj(G_{A'})$ holds, which results in an analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
 $$
@@
 as visualized in \ref{fig:robustnessprocess}C.
-This corresponds to an ensemble of dynamical systems with equivalent connectivities, but where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{}).
-We then ask, for each distinct network topology on $n$ components, what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
-This quantifies the intuitive statement that, over evolutionary timescales, it is not enough for a system to be stable, but rather that it have a reasonably high probability of remaining stable over continguous timeframes.
-In terms of a history $s^T$ over the states specifying the stability property, $\mathcal{S}$, this is given by
-$$
-r = \frac{\sum_{t=0}^{T-1}s_{t+1}^T s_{t}^T}{\sum_{t=0}^{T-1} s_{t}^T},
-$$
-where we make numerical substitutions for the values of $\mathcal{S}$ such that $S = 1$ and $U = 0$.
-Given that the process $P(s^T)$ has stationary probability $P^{\infty}(s)$ for $s \in \mathcal{S}$, the expectation of $r$ as $T \rightarrow \infty$ is approximated by
-$$
-R = E[r] \approx \frac{\sum_{s,s'} P^{\infty}(s)P^{\infty}(s' \mid s)}{P^{\infty}(S)}.
-$$
-This conditional probability corresponds to the parameter $R$ of the two-state process depicted in \ref{fig:robustnessprocess}A, and it is what we refer to as dynamical robustness.
-Comparing the values of $R$ for each $G$-class in $\mathcal{F}/G$ places a partial ordering on network architectures, $G$, which allows for the determination of which network architectures would be expected to be enriched in a process selecting for higher $R$ values.
+This now corresponds to a collection of ensembles of dynamical systems each with equivalent connectivities corresponding to a graph $G$.
+For each graph, there is a potentially different value of robustness given by
+\begin{equation}\label{eq:robustnessbygraph}
+\begin{aligned}
+R(G,S,\mu,\tau)  & = \\
+& \frac{\int_{A,A' \in \mathbb{R}^{n \times n}, G_A = G} d\mu(A) d\tau(A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}{\int_{A \in \mathbb{R}^{n \times n}, G_A = G} d\mu(A) \mathcal{S}(A)}.
+\end{aligned}
+\end{equation}
+Comparing the values of $R$ for each $G$-class in $\mathcal{F}/G$ places a partial ordering on network architectures, $G$, which allows for the determination of which network architectures would be expected to be enriched relative to others in an evolutionary process where selection is imposed in a manner that results in a bias toward higher $R$ values.
 
@@
 
-We have derived an analytical expression for dynamical robustness, $R_{tot}$, of a network in terms of its interaction graph as a weighted average of the robustness of the SCCs, $R_i$, the corresponding number of links within each SCC, $d_i$, and the number of links between the SCCs, $l$.
-The entries of the Jacobian are sampled independently from a generic probability distribution.
+Now we derive an analytical expression for dynamical robustness, $R$, of a network in terms of its interaction graph, $G$, as a weighted average of the robustness, $R_{\alpha}$, of the SCCs, $C_{\alpha}$, the corresponding number of links within each SCC, $d_{\alpha}$, and the number of links between the SCCs, $l$ to give $d = l + \sum_{\alpha} d_{\alpha}$.
+The result ultimately holds for the case of simultaneously perturbing any number of elements of the Jacobian.
+To anchor the intuition before stating the more general result, we derive the expression for the case of perturbing a single element at each timestep of the process.
+Let the index $i$ range over the non-zero entries $a_i$ of $A$.
+The entries of the Jacobian are sampled independently from a generic probability distribution $\rho_i$ for each entry.
+Under the given assumptions then
+\begin{equation}
+\begin{aligned}
+\mu(A) &= \prod_i \rho_i(a_i),\\
+\tau(A' \mid A) &= \frac{1}{d} \sum_i \tau_i (A' \mid A),
+\end{aligned}
+\end{equation}
+where
+$$
+\tau_i(A',A) = \rho_i(a'_i)\rho_i(a_i) \prod_{j \neq i} \delta(a'_j - a_j).
+$$
+
+Since the determinant of a triangular matrix equals the product of the determinants of its diagonal blocks, it follows that the characteristic polynomial factors as the product of the charactericstic polynomials of its diagonal blocks.
+Hence, a block triangular matrix is stable if and only if its diagonal blocks are stable.
+Note that this condition does not depend upon the entries off the diagonal (which correspond to links between SCCs) and does not depend upon what order the components appear.
+This fact implies that the terms evaluating stability such as $\mathcal{S}(A)$ from \ref{eq:robustnessbygraph} decompose into products over the SCCs
+$$
+\mathcal{S}(A) = \prod_{C_\alpha \in \hier(G)} \mathcal{S}(C_{\alpha}).
+$$
+
 Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
-The result ultimately holds for the case of simultaneously perturbing any number of elements of the Jacobian.
-To anchor the intuition before stating the more general result, \ref{eq:robschematic} shows a schematized version for the case of perturbing a single element  (see \reffigscc$\,$ for examples demonstrating this expression)
+\ref{eq:robschematic} shows a schematized version for the case of perturbing a single element  (see \reffigscc$\,$ for examples demonstrating this expression)
 \begin{equation}\label{eq:robschematic}
-R_{tot} = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
+R = \frac{l+d_1 R_1 + d_2 R_2 + \cdots}{l+d_1 + d_2 + \cdots}.
 \end{equation}

### 2015-06-03 tex/overview.tex
update main text results sections

@@
 %!TEX root = ../paper.tex
+\section{Dynamical Systems on Biological Networks}
 
@@
 % $$
+\subsection{Stability analysis of biological networks}
 To determine stability, we use the Taylor series expansion of the equations of motion \ref{eq:eom} about the fixed point $\vec{x}^0$ where $\vec{y} = \vec{x} - \vec{x}^0$ by
@@
 
+\subsection{Equivalence classes of systems associated to Jacobian Matrices}
 Jacobian matrices define an equivalence relation, $\sim$, on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}\,',\vec{x}^{0'}) $ if and only if
@@
 % \subsection{Overview}
+\subsection{Interaction graphs encoding network architecture}
 The interactions among variables in a dynamical model for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
@@
 
+\section{Evolutionary processes sampling dynamical systems}
 In the course of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to the basins associated to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
-The stochastic process by which these modifications occur induces one on the set of fixed points that results in the assignment of a probability $P(f^T)$ to each history of length $T$, $f^T = ( f_1,f_2,\ldots,f_T ) \in \mathcal{F}^T$.
+The stochastic process by which these modifications occur induces one on the set of fixed points that results in the assignment of a probability $P(f^T)$ to each history of length $T$, $f^T = ( f_1,f_2,\ldots,f_T ) \in \mathcal{F}^T$ \cite{RobertM.Gray130}.
 % $$
@@
 $$
-If the process $P(s^T)$ has limits $\mu(s) = \lim_{t \rightarrow \infty} P(s_t^T=s)$ and $\tau(s' \mid s) = \lim_{t \rightarrow \infty} P(s_{t+1}^T = s' \mid s_t^T = s)$ for $s,s' \in \{0,1\}$
+If the process $P(s^T)$ has limits $\mu(s) = \lim_{t,T \rightarrow \infty} P(s_t^T=s)$ and $\tau(s' \mid s) = \lim_{t,T \rightarrow \infty} P(s_{t+1}^T = s' \mid s_t^T = s)$ for $s,s' \in \{0,1\}$
 %stationary probability $P^{\infty}(s)$ for $s \in \mathcal{S}$
@@
 In terms of Jacobian matrices $A$ and $A'$, if the underlying process $P([A]^T)$ has analogous limits
-$$\mu(A) = \lim_{t \rightarrow \infty} P(A_t^T=A)$$
+$$\mu(A) = \lim_{t,T \rightarrow \infty} P(A_t^T=A)$$
 and
-$$\tau(A' \mid A) = \lim_{t \rightarrow \infty} P(A_{t+1}^T = A' \mid A_t^T = A)$$
+$$\tau(A' \mid A) = \lim_{t,T \rightarrow \infty} P(A_{t+1}^T = A' \mid A_t^T = A)$$
 for $A,A' \in \mathbb{R}^{n \times n}$, \ref{eq:robustnessfromstoch} becomes
@@
 \begin{aligned}
-R(G,S,\mu,\tau)  & = \\
-& \frac{\int_{A,A' \in \mathbb{R}^{n \times n}, G_A = G} d\mu(A) d\tau(A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}{\int_{A \in \mathbb{R}^{n \times n}, G_A = G} d\mu(A) \mathcal{S}(A)}.
+R(G,S,\mu,\tau)  =
+ \frac{\int\limits_{A,A' \in \mathbb{R}^{n \times n} \atop G_A = G} d\mu(A) d\tau(A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}{\int\limits_{A \in \mathbb{R}^{n \times n} \atop G_A = G} d\mu(A) \mathcal{S}(A)}.
 \end{aligned}
@@
 
-The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs, see \refsupp{} and \reffigscc).
-This map from the interaction graph of a network to its SCCs, referred to as $\hier$ in \ref{fig:modsccsym}, has a collection of intrinsic symmetries shown in \reffighiertransformations{}.
+\section{Abstracting network architectures to their strongly connected components and the symmetries of robustness}
+The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs).
+A SCC of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
+The strongly connected components of some examples of three variable systems are outlined in \reffigscc{} along with their adjacency matrices.
+
+The map from the interaction graph of a network to its SCCs, referred to as $\hier$ in \ref{fig:modsccsym}, results in a decomposition of $G$ into its SCCs.
+Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
+There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
+Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
+
+One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $H$.  To each node $n$ of $H$ associate
+a strongly connected graph $C_n$.
+To each link $(i,j)$ of $H$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
+The result will be a graph $G$ such that $\hier(G) = H$ and furthermore, every graph $G$ such that $\hier(G) = H$ can be obtained in this manner.
+
+This map $\hier$ is many-to-one and so there is a large class of
+operations which leaves $\hier(G)$ invariant for a given graph $G$ \reffighiertransformations{}.
 These three symmetries, \reffigscc{}, represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated.
+For instance, we may interchange the positions of the strongly
+connected components relative to each other \reffighiertransformations$a$.
+Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations$b$ or between components, or even add or delete links \reffighiertransformations$c$.
+
 Symmetries with respect to some property of the system are characterized by the ability to interchange these modules or their connectivity without changing that property.
-Two of these three intrinsic symmetries are also symmetries with respect to dynamical robustness.
+Two of these three intrinsic symmetries of $\hier$ are also symmetries with respect to dynamical robustness.
 \ref{fig:robustnesssymmetries} shows an example of these latter symmetries applied to a specific interaction graph.
 
+\section{Derivation of the relationship between network architecture and robustness}
 Now we derive an analytical expression for dynamical robustness, $R$, of a network in terms of its interaction graph, $G$, as a weighted average of the robustness, $R_{\alpha}$, of the SCCs, $C_{\alpha}$, the corresponding number of links within each SCC, $d_{\alpha}$, and the number of links between the SCCs, $l$ to give $d = l + \sum_{\alpha} d_{\alpha}$.
@@
 To anchor the intuition before stating the more general result, we derive the expression for the case of perturbing a single element at each timestep of the process.
+\subsection{Independent modification process to individual network interactions}
 Let the index $i$ range over the non-zero entries $a_i$ of $A$.
@@
 Under the given assumptions then
-\begin{equation}
+\begin{equation}\label{eq:singleresamplemutau}
 \begin{aligned}
 \mu(A) &= \prod_i \rho_i(a_i),\\
-\tau(A' \mid A) &= \frac{1}{d} \sum_i \tau_i (A' \mid A),
+\tau^{(1)}(A' \mid A) &= \frac{1}{d} \sum_i \tau^{(1)}_i (A' \mid A),
 \end{aligned}
@@
 $$
-\tau_i(A',A) = \rho_i(a'_i)\rho_i(a_i) \prod_{j \neq i} \delta(a'_j - a_j).
+\tau^{(1)}_i(A' \mid A) = \rho_i(a'_i) \prod_{j \neq i} \delta(a'_j - a_j).
 $$
+
+The decomposition of a digraph into SCCs corresponds to a block triangular decomposition of its adjacency matrix.
+Say that the graph $G$ has SCCs $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
+Label the vertices in such a way that $V_1, \ldots, V_{n_1}$ belong to $C_1$, $V_{n_1 + 1}, \ldots, V_{n_2}$ belong to $C_2$, etc.
+Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
+This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.
 
@@
 This fact implies that the terms evaluating stability such as $\mathcal{S}(A)$ from \ref{eq:robustnessbygraph} decompose into products over the SCCs
-$$
-\mathcal{S}(A) = \prod_{C_\alpha \in \hier(G)} \mathcal{S}(C_{\alpha}).
-$$
+\begin{equation}\label{eq:stabfactor}
+\mathcal{S}(A) = \prod_{C_\alpha \in \hier(G)} \mathcal{S}(\pi_{C_{\alpha}}(A))
+\end{equation}
+where $\pi_{C_{\alpha}}(A)$ denotes the projection of the matrix $A$ onto the SCC $C_{\alpha}$.
 
-Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
-\ref{eq:robschematic} shows a schematized version for the case of perturbing a single element  (see \reffigscc$\,$ for examples demonstrating this expression)
+To relate the robustness of a graph to the robustness of its SCCs, we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessfromstoch}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
+Let $L$ denote the set of edges of $G$ that connect distinct strongly connected components.
+Then, if $i \in C_\alpha$ for some SCC $C_\alpha$, we have
+\begin{widetext}
+\begin{equation}\label{eq:robustncessforsccs}
+\begin{aligned}
+\frac{\int d\mu(A) d\tau^{(1)}_i (A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}
+       {\int d\mu(A) \mathcal{S}(A)}
+&= \frac{\begin{matrix}
+  \int \prod\limits_{k \in C_\alpha} da_k\,d{a'}_k\, \rho(a_k)
+    \rho_i({a'}_i) \prod\limits_{j \in C_\alpha \setminus i} \delta ({a'}_j - a_j)       
+    \mathcal{S}(\pi_{C_{\alpha}}(A)) \mathcal{S}(\pi_{C_{\alpha}}(A')) \times \\
+ \prod\limits_{C_\beta \in \hier(G) \setminus C_\alpha} \int 
+   \prod\limits_{j \in C_\beta} da_k\,d{a'}_k\, \rho({a'}_j) \delta ({a'}_j - a_j)
+      \mathcal{S}(\pi_{C_{\beta}}(A)) \mathcal{S}(\pi_{C_{\beta}}(A')) \times \\
+ \prod\limits_{l \in L} \int da_l\,d{a'}_l\,\rho({a'}_l) \delta({a'}_l - a_l) \end{matrix}}
+{\begin{matrix}\int \prod\limits_{k \in C_\alpha} da_k\,\rho(a_k) \mathcal{S}(\pi_{C_{\alpha}}(A)) \times \\
+ \prod_{C_\beta \in \hier(G) \setminus C_\alpha}
+   \int \prod\limits_{h \in C_\alpha} da_h,\rho(a_h) \mathcal{S}(\pi_{C_{\beta}}(A)) \times \\
+ \prod\limits_{l \in L} \int da_l\,\rho(a_l) \end{matrix}} \\
+&= \frac{\int \prod\limits_{k \in C_\alpha} da_k\,d{a'}_k\,\rho(a_k)
+    \rho_i({a'}_i) \prod\limits_{j \in C_\alpha \setminus i} \delta ({a'}_j - a_j)       
+    \mathcal{S}(\pi_{C_{\alpha}}(A)) \mathcal{S}(\pi_{C_{\alpha}}(A'))}
+{\int \prod\limits_{k \in C_\alpha} dk\, \rho(a_k) \mathcal{S}(\pi_{C_{\alpha}}(A))} \\
+&= R(C_{\alpha}, S, \mu_\alpha, \tau^{(1)}_\alpha)
+\end{aligned}
+\end{equation}
+where $\mu_\alpha$ and $\tau^{(1)}_\alpha$ refer to the analogues of \ref{eq:singleresamplemutau} for the subgraph $C_\alpha$ of $G$.
+\ref{eq:robustncessforsccs} shows that the terms in the sum over the elements of $A$, or equivalently the edges of $G$, that comprise a given SCC of $G$ reduce to the robustness of that SCC alone.
+Likewise, when $i \in L$, we have
+\begin{equation}
+\begin{aligned}
+\frac{\int d\mu(A) d\tau^{(1)}_i (A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}
+       {\int d\mu(A) \mathcal{S}(A)}
+&= \frac{\begin{matrix}
+  \prod\limits_{C_\alpha \in \hier(G)} \int 
+   \prod\limits_{j \in C_\beta} da_k\,d{a'}_k\, \rho({a'}_j) \delta ({a'}_j - a_j)
+      \mathcal{S}(\pi_{C_\alpha}(A)) \mathcal{S}(\pi_{C_\alpha}(A')) \times \\
+ \int da_i d{a'}_i \rho_i(a) \rho_a({a'}_i) \times
+ \prod\limits_{l \in L\setminus i} \int da_l\,d{a'}_l\,\rho({a'}_l) \delta({a'}_l - a_l) \end{matrix}}
+{\prod_{C_\alpha \in \hier(G)}
+   \int \prod\limits_{h \in C_\alpha} da_h,\rho(a_h) \mathcal{S}(\pi_{C_\alpha}(A)) \times
+ \int da_i \rho(a_i) \times
+ \prod\limits_{l \in L \setminus i} \int da_l\,\rho(a_l)} \\
+&= 1
+\end{aligned}
+\end{equation}
+\end{widetext}
+For each $C_{\alpha} \in \hier(G)$, there will be $d_\alpha$ values of $i$ such that $i \in C_{\alpha}$; likewise, there will be $l$ values of $i$ such that $i \in L$.
+Hence, when we perform the summation over $i$ to compute the robustness of $G$, we will obtain a weighted average:
 \begin{equation}\label{eq:robschematic}
@@
 \end{equation}
+Here $R_\alpha$ is shorthand for $R(C_\alpha,S,\mu_\alpha,\tau^{(1)}_\alpha)$ and $R$ is shorthand for $R(G,S,\mu,\tau^{(1)})$.
+Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
+\ref{eq:robschematic} shows a schematized version of \ref{eq:robustnessbygraph} for the case of perturbing a single element  (see \reffigscc$\,$ for examples demonstrating this expression).
 For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.
@@
 
-Examining this expression noting that $R_i$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R_{tot}$.
-Given two connected components $C_i$ and $C_j$ with $v_i$ and $v_j$ nodes respectively, we have a maximum of $v_i v_j$ links going from $C_i$ to $C_j$.
-Hence, $l \le \sum_{(ij) \in \hier(G)} v_i v_j$.
+Examining this expression noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
+Given two connected components $C_{\alpha}$ and $C_{\beta}$ with $v_{\alpha}$ and $v_{\beta}$ nodes respectively, we have a maximum of $v_{\alpha} v_{\beta}$ links going from $C_{\alpha}$ to $C_{\beta}$.
+Hence, $l \le \sum_{(\alpha,\beta) \in \hier(G)} v_{\alpha} v_{\beta}$.
 Since every acyclic digraph can be embedded into a totally ordered
 set, we may assume without loss of generality that our components have
-been ordered in a way such that, if $(i,j) \in \hier(G)$, then $i <
-j$.  Hence, $l \le l_{max}$ where
-$$l_{max} = \sum_{i=1}^{n-1}\sum_{j=i+1}^{n}v_i
-v_j~=~\frac{1}{2} \left( \sum_{i=1}^{n} v_i \right)^2-\frac{1}{2} \sum_{i=1}^{n}
-v_i^2,$$
-Hence, we conclude that $R_{tot} \le R_{max}$ where $R_{max}$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$ and the upper bound is attained (\refsupp{}).
-
+been ordered in a way such that, if $(\alpha,\beta) \in \hier(G)$, then $\alpha <
+\beta$.  Hence, $l \le l_{max}$ where
+$$l_{max} = \sum_{\alpha=1}^{n-1}\sum_{\beta=\alpha+1}^{n}v_{\alpha}
+v_{\beta}~=~\frac{1}{2} \left( \sum_{\alpha=1}^{n} v_{\alpha} \right)^2-\frac{1}{2} \sum_{\alpha=1}^{n}
+v_{\alpha}^2.$$
+Suppose we have a graph $G$ with SCCs $C_1,\ldots,C_n$ and that $G_{\mathrm{tot}}$
+is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.
+Then we have a graph $G_{\mathrm{max}}$ such that $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$, the components of $G_{\mathrm{max}}$ are also $C_1, \ldots, C_n$, but that includes all possible links between each pair of components.
+In this case $R(G) \le R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$
+\begin{widetext}
+\begin{align} \label{eq:sccmaxrobustness}
+R(G_{\mathrm{max}}) =
+\frac{\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
+                    \sum_{\alpha=1}^n \connectivity_{\alpha} 
+                    % \langle R(\hbox{stab}(C_i),\mu') \rangle_1
+                R(C_{\alpha},S,\mu_{\alpha},\tau_{\alpha})}
+     {\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
+                    \sum_{\alpha=1}^n \connectivity_{\alpha}}.
+\end{align}
+\end{widetext}
+% where $R(G_{\mathrm{max}},S,\mu,\tau) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+%This bound is attained by $G_{\mathrm{max}}$.  
+% $\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+\subsection{Independent modification process to arbitrary network interactions}
 This argument also works when we resample more than one entry,
 although the notation becomes more complicated.  Suppose that we
-resample $m$ nodes.  Define
+resample $m$ nodes.  Then the analogue of \ref{eq:singleresamplemutau} is
+\begin{equation}\label{eq:multipleresamplemutau}
+\begin{aligned}
+\tau^{(m)}(A' \mid A) &= \frac{1}{\binom{d}{m}} \sum_{i_1, \ldots, i_m}
+  \tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A) \\
+\tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A) &= 
+ \prod_{k=1}^m \rho_{i_k} (a'_{i_k})
+ \prod_{j \notin \{i_1, \ldots i_m\}} \delta(a'_j - a_j)
+\end{aligned}
+\end{equation}
+Define
 \begin{widetext}
@@
 \begin{widetext}
-\begin{equation}
-\langle R(\hbox{stab} (G), \mu') \rangle_m =
+\begin{equation}\label{eq:robustnessmultiple}
+R(G, S, \mu, \tau^{(m)}) =
 \frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
       {m \choose m_0, m_1, \ldots. m_n}
-      \left(m_0 + \sum\limits_{i=1}^n m_i \langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \right)}
+      \left(m_0 + \sum\limits_{\alpha=1}^n m_\alpha R(C_\alpha, S, \mu, \tau^{(m_\alpha)}) \right)}
      {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
@@
 
-As before, since $\langle R(\hbox{stab} (C_i), \mu') \rangle_{m_i} \le 1$, we may
-increase $\langle R(\hbox{stab} (G), \mu') \rangle_m$ by increasing the maximum
+As before, since $R(C_\alpha, S, \mu, \tau^{(m_i)}) \le 1$, we may
+increase $R(G, S, \mu, \tau^{(m)})$ by increasing the maximum
 possible value of $m_0$ while keeping the strongly connected
 components the same.  Again, if we fix $\hier(G)$, the maximum
-possible value of $m_0$ is $\sum_{(i,j) \in \hier(G)} v_i v_j$
-whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{i=1}^n
-v_i)^2 - \sum_{i=1}^n v_i^2)$, which is attained when $\hier(G_{max}) =
-G_{tot}$.  Hence, we conclude that $\langle R(\hbox{stab} (G), \mu') \rangle_m \le
-\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_m$.
+possible value of $m_0$ is $\sum_{(\alpha,\beta) \in \hier(G)} v_\alpha v_\beta$
+whereas, if we allow it to vary, the maximum is $\frac{1}{2} ((\sum_{\alpha=1}^n
+v_\alpha)^2 - \sum_{\alpha=1}^n v_\alpha^2)$, which is attained when $\hier(G_{max}) =
+G_{tot}$.  Hence, we conclude that $R(G, S, \mu, \tau^{(m)}) \le
+R(G_{\mathrm{max}}, S, \mu, \tau^{(m)})$.
 
@@
 
-To test the prediction of this analytical result, we computed the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
+To test the prediction of the analytical results in \ref{eq:robschematic} and \ref{eq:robustnessmultiple}, we computed the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
 For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).

### 2015-06-03 tex/discussion.tex
update discussion

@@
 % Our development suggests two predictions that could be used to test the theory described here. One is that,
+\section{Conclusion}
 Our analysis predicts that, in general, an ensemble of systems where robustness has been the predominant object of selection and has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.

### 2015-06-03 tex/methods.tex
minor changes
update supplement

@@
 %!TEX root = ../paper.tex
-\section{Reaction Networks with Prescribed Connecctivity and Jacobians}\label{sec:reactionnetjacobian}
+
+\section{Reaction Networks, Gene Regulatory Networks, and Ecological Networks with Prescribed Connectivity and Jacobians}\label{sec:reactionnetjacobian}
+
+The quality of interest in this paper is robustness, which is related to the concept of structural stability \cite{Smale1967}, whose evaluation requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters, its rate functions, or environmental constraints that restrict it to a subset of its basins of attraction.
+We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the components and not only to perturbations of the state vector at a given point in time.
+It is justified to consider resampling elements of $A$ to generate $A'$ as a proxy for resampling elements of $\vec{p}$ to produce $\vec{p}\,'$ if any matrix $A$ can be obtained for some $F_i$, $\vec{p}$ and $\vec{x}^0$.
+This holds for the $F_i$ defining the Lotka-Volterra model.
+This is due to the fact that for a specification of non-zero real numbers for the components of $\vec{n}^0$ and any real numbers for the components of $a_{ij}$, there is a choice of parameters $\vec{p}$ given by $b_{ij} = \frac{a_{ij}}{n_i^0}$ and $r_i = - \sum_{j=1}^N \frac{n_j^0}{n_i^0} a_{ij}$ that generates those particular $a_{ij}$ as the Jacobian matrix of the dynamical system.
+Checking this property of the domain of realizability of the Jacobian can be done for ensembles of systems other than the Lotka-Volterra ensemble.
+For arbitrary biochemical reaction and gene regulatory networks, this property is likely to hold so long as not too many types of transformations are constrained from possibility.
+For example, a simplified version of the general form of the gene regulatory network model presented in \ref{fig:biomodelexamples} center panel is given by the system
+\begin{equation}
+\frac{dg_i}{dt} = \sum_{j=1}^N k_{ij} g_j,
+\end{equation}
+with one parameter $k_{ij} \in \mathbb{R}$ for every pair $(g_i,g_j)$ of genes.
+The Jacobian of this system is $A_{ij} = k_{ij}$, and, therefore, sampling parameters of the model is precisely equivalent to sampling elements of the Jacobian.
+% For reaction networks, we demonstrate an ensemble from which arbitrary Jacobians can arise in \ref{sec:reactionnetjacobian}.
 
@@
 
-\section{Monte carlo integration}\label{suppsec:montecarlo}
-If we sample $N = N_{\mathrm{stab}} + N_{\mathrm{unstab}}$ matrices where each has some probability $\theta$ of being stable then $N_{\mathrm{stab}}$ has a binomial distribution.
-We can compute a sample estimate for $\theta$, $\hat{\theta} = \frac{N_{\mathrm{stab}}}{N}$ \cite{Murphy2012}.
-The posterior distribution in this case is known to be a Beta distribution as a result of Beta-Binomial conjugacy
-$$
-\mathrm{Beta}(\theta | \mathcal{D}) = \mathrm{Beta}(\theta | N_{\mathrm{stab}} + a, N_{\mathrm{unstab}} + b)
-$$
-where $a$ and $b$ are the hyperparameters of the Beta prior and we consider the uninformative uniform prior corresponding to $a=b=1$. We consider the maximum a posteriori estimate
-$$\hat{\theta}_{MAP} = \frac{a + N_{\mathrm{stab}} - 1}{a + b + N - 2}$$
-which corresponds in this case to the maximum likelihood estimate
-$$
-\hat{\theta}_{MLE} = \frac{N_{\mathrm{stab}}}{N}.
-$$
-This estimate is characterized by the variance of the posterior Beta distribution
-% \begin{widetext}
-\begin{align*}
-&\mathrm{var}(\theta | \mathcal{D}) =\\
-&\frac{(a+N_{\mathrm{stab}})(b+N_{\mathrm{unstab}})}{(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}})^2(a + N_{\mathrm{stab}} + b + N_{\mathrm{unstab}}+1)}
-\end{align*}
-% \end{widetext}
-Since for the chosen prior $a=b=1 \ll N$ this simplifies to
-$$
-\mathrm{var}(\theta | \mathcal{D}) = \frac{\hat{\theta}(1-\hat{\theta})}{N}
-$$
-yielding the error estimate given by the associated standard deviation.
-In all simulations we use $N~=~10000$ so that the maximum error for $\hat{\theta}~=~0.5$ is $\sigma~=~\sqrt{\mathrm{var}(\theta | \mathcal{D})} \approx 0.005$.
-
 \section{Hierarchy and Total ordering}\label{sec:totalordering}
@@
 The totality condition implies $x R x$ (reflexivity) corresponding to the fact that the directed graph associated to the total ordering has, for each node, an edge whose source and target are the same node.
+
+Corresponding to the SCC decomposition of $G$ we can construct a directed acyclic graph $\hier (G)$ or the condensed graph \cite{Corominas-Murtra2013}.
+Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
+There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
+Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
+
+% One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
+% a strongly connected graph $C_n$.
+To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
+The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
+
+% This map $\hier$ is many-to-one and so there is a large class of
+% operations which leaves $\hier(G)$ invariant for a given graph $G$.
+% For instance, we may interchange the positions of the strongly
+% connected components relative to each other \reffighiertransformations$a$.
+Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations$b$ or between components, or even add or delete links \reffighiertransformations$c$.
+Some of these operations leave dynamical robustness invariant, so we may regard them as a symmetry groupoid of the systems associated to a given graph $G$ with respect to the robustness property.
+
+The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
+The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc{} (top panel).
+This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
+The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
+This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being maximally hierarchical.
+In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.

### 2015-06-04 tex/overview.tex
update overview and figures

@@
 where $F=\{F_1,\ldots,F_i,\ldots,F_n\}$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$.
-These parameters typically represent reaction rates in chemical and gene-regulatory networks, interaction strengths in ecological networks.
+These parameters typically represent reaction rates or interaction strengths in chemical, gene-regulatory and ecological networks.
 For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$.
@@
 The entries of the Jacobian are sampled independently from a generic probability distribution $\rho_i$ for each entry.
-Under the given assumptions then
+Under this assumption then
 \begin{equation}\label{eq:singleresamplemutau}
@@
 \end{aligned}
-\end{equation}
+\end{equation}\label{eq:robustncessforsccconnect}
+\end{widetext}
 where $\mu_\alpha$ and $\tau^{(1)}_\alpha$ refer to the analogues of \ref{eq:singleresamplemutau} for the subgraph $C_\alpha$ of $G$.
@@
 Likewise, when $i \in L$, we have
+\begin{widetext}
 \begin{equation}
@@
 \end{widetext}
-For each $C_{\alpha} \in \hier(G)$, there will be $d_\alpha$ values of $i$ such that $i \in C_{\alpha}$; likewise, there will be $l$ values of $i$ such that $i \in L$.
-Hence, when we perform the summation over $i$ to compute the robustness of $G$, we will obtain a weighted average:
+For each $C_{\alpha} \in \hier(G)$, there will be $d_\alpha$ values of $i$ such that $i \in C_{\alpha}$ requiring instances of \ref{eq:robustncessforsccs}; likewise, there will be $l$ values of $i$ such that $i \in L$ requiring instances of \ref{eq:robustncessforsccconnect}.
+Hence, when we perform the summation over $i$ to compute the robustness of $\mathcal{F}/G$ for each $G$, we will obtain a weighted average:
 \begin{equation}\label{eq:robschematic}
@@
 Here $R_\alpha$ is shorthand for $R(C_\alpha,S,\mu_\alpha,\tau^{(1)}_\alpha)$ and $R$ is shorthand for $R(G,S,\mu,\tau^{(1)})$.
-Examples of vector fields that correspond to this sampling of Jacobian matrices are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
-\ref{eq:robschematic} shows a schematized version of \ref{eq:robustnessbygraph} for the case of perturbing a single element  (see \reffigscc$\,$ for examples demonstrating this expression).
+Examples of vector fields that correspond to the sampling of Jacobian matrices used in the computation of robustness for particular examples are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
+\ref{eq:robschematic} shows a schematized version of \ref{eq:robustnessbygraph} for the case of perturbing a single element at a time  (see \reffigscc$\,$ for examples demonstrating this expression).
 For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.
@@
 Then we have a graph $G_{\mathrm{max}}$ such that $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$, the components of $G_{\mathrm{max}}$ are also $C_1, \ldots, C_n$, but that includes all possible links between each pair of components.
-In this case $R(G) \le R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$
+In this case $R(G)~\le~R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$
 \begin{widetext}
@@
 % $\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
-\subsection{Independent modification process to arbitrary network interactions}
-This argument also works when we resample more than one entry,
+\subsection{Modifying multiple interactions simultaneously}
+This argument also works when relationships between multiple system components are perturbed simultaneously,
 although the notation becomes more complicated.  Suppose that we
-resample $m$ nodes.  Then the analogue of \ref{eq:singleresamplemutau} is
+resample $m$ interactions.  Then the analogue of \ref{eq:singleresamplemutau} is
 \begin{equation}\label{eq:multipleresamplemutau}
@@
 \tau^{(m)}(A' \mid A) &= \frac{1}{\binom{d}{m}} \sum_{i_1, \ldots, i_m}
-  \tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A) \\
+  \tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A), \\
 \tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A) &=
  \prod_{k=1}^m \rho_{i_k} (a'_{i_k})
- \prod_{j \notin \{i_1, \ldots i_m\}} \delta(a'_j - a_j)
+ \prod_{j \notin \{i_1, \ldots i_m\}} \delta(a'_j - a_j).
 \end{aligned}
 \end{equation}
-Define
+Now define
 \begin{widetext}
@@
 This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
-This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the total ordering like that of \reffigscc{} top will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
+This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the total ordering will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
+The graph associated to the total ordering is the most hierarchical network architecture for any given number of system components like that of \reffigscc{} top for three component systems where the \emph{highest} component in the hierarchy has directed links to all other nodes in the network, the second highest component has directed links to all other nodes in the network except the highest one, \emph{et cetera} (\refsupp{}).
 Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.
@@
 
-To test the prediction of the analytical results in \ref{eq:robschematic} and \ref{eq:robustnessmultiple}, we computed the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
+To test the prediction of the analytical results in \ref{eq:robschematic} and \ref{eq:robustnessmultiple}, we computed approximations to the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
 For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).

### 2015-06-04 tex/discussion.tex
update discussion

@@
 Our analysis predicts that, in general, an ensemble of systems where robustness has been the predominant object of selection and has been positively selected over a sufficiently long period of time should exhibit a bias toward more hierarchical network topologies.
-In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness \cite{Zhao2006,Bhardwaj2010,Colm}.
+Given the manner in which we define robustness in \ref{eq:robustnessfromstoch}, this is a very general constraint.
+In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness as we have defined it here \cite{Zhao2006,Bhardwaj2010,Colm}.
 At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
@@
 Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
-In the long term, this prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that differentially select for dynamical robustness \cite{Leroi1994}.
+In the long term, this prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that impose differential selection strengths for dynamical robustness \cite{Leroi1994}.
 

### 2015-07-13 tex/overview.tex
move equation label to correct location
clarify the definition of the set of dynamical systems
use abbreviation for SCCs

@@
 \begin{equation}\label{eq:setdynsys}
-\mathcal{D} = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
+\mathcal{D'} = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
 \end{equation}
+The space of all $F$ is too large to precisely define probabilities over, so we consider various subspaces $\mathcal{D} \subset \mathcal{D}'$.
+In some cases these may be given by parametrized models with a single fixed form of $F$ and then only $\vec{p}$ is allowed to vary.
 
@@
 A SCC of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
-The strongly connected components of some examples of three variable systems are outlined in \reffigscc{} along with their adjacency matrices.
+The SCCs of some examples of three variable systems are outlined in \reffigscc{} along with their adjacency matrices.
 
@@
 \end{aligned}
-\end{equation}\label{eq:robustncessforsccconnect}
+\end{equation}
 \end{widetext}
@@
 \begin{widetext}
-\begin{equation}
+\begin{equation}\label{eq:robustncessforsccconnect}
 \begin{aligned}

### 2015-07-14 tex/methods.tex
transform Lotka-Volterra variables from n's to x's

@@
 This holds for the $F_i$ defining the Lotka-Volterra model.
-This is due to the fact that for a specification of non-zero real numbers for the components of $\vec{n}^0$ and any real numbers for the components of $a_{ij}$, there is a choice of parameters $\vec{p}$ given by $b_{ij} = \frac{a_{ij}}{n_i^0}$ and $r_i = - \sum_{j=1}^N \frac{n_j^0}{n_i^0} a_{ij}$ that generates those particular $a_{ij}$ as the Jacobian matrix of the dynamical system.
+This is due to the fact that for a specification of non-zero real numbers for the components of $\vec{x}^0$ and any real numbers for the components of $a_{ij}$, there is a choice of parameters $\vec{p}$ given by $b_{ij} = \frac{a_{ij}}{x_i^0}$ and $r_i = - \sum_{j=1}^n \frac{x_j^0}{x_i^0} a_{ij}$ that generates those particular $a_{ij}$ as the Jacobian matrix of the dynamical system.
 Checking this property of the domain of realizability of the Jacobian can be done for ensembles of systems other than the Lotka-Volterra ensemble.

### 2015-07-14 tex/overview.tex
transform Lotka-Volterra variables from n's to x's

@@
 These parameters typically represent reaction rates or interaction strengths in chemical, gene-regulatory and ecological networks.
-For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (n_1, \ldots, n_N)$, $\vec{p}=(r_1,\ldots,r_N,b_{11},\ldots,b_{NN})$, and $F_i = r_i n_i + \sum_{j=1}^{N} b_{ij} n_i n_j$.
+For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (x_1, \ldots, x_n)$, $\vec{p}=(r_1,\ldots,r_n,b_{11},\ldots,b_{nn})$, and $F_i = r_i x_i + \sum_{j=1}^{n} b_{ij} x_i x_j$.
 The set of all dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $s$, is then
@@
 observing the solution ${\vec x}(t) = {\vec x}^0$ in the absence of coupling to another system.
-The Lotka-Volterra model has two fixed points: the trivial one of all zero species $n_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{N} b_{ij} n_j = 0$.
+The Lotka-Volterra model has two fixed points: the trivial one of all zero species $x_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{n} b_{ij} x_j = 0$.
 The set of all such fixed points, $\mathcal{F}$, is then
@@
      \begin{array}{lr}
-       r_i + b_{ii} n_i + \sum_{k=1}^{N} b_{ik} n_{k}, & i \neq j\\
-       b_{ii} n_i & i=j
+       r_i + b_{ii} x_i + \sum_{k=1}^{n} b_{ik} x_{k}, & i \neq j\\
+       b_{ii} x_i & i=j
      \end{array}.
@@
 
-In order to classify the properties of this process according to network architecture, we consider cases where the condition $\adj(G_A) = \adj(G_{A'})$ holds, which results in an analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
+We would like to determine whether selection for dynamical robustness would result in a preference for the maintenance of one kind of network architecture relative to another.
+In order to then classify the properties of this process according to network architecture, we consider cases where the condition $\adj(G_A) = \adj(G_{A'})$ holds, which results in an analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
 $$
@@
 
-\section{Abstracting network architectures to their strongly connected components and the symmetries of robustness}
-The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by mapping the interaction graph to the network of strongly connected components (SCCs).
+\section{Properties of network architecture}
+
+\subsection{Abstracting network architectures to their strongly connected components}
+We will ultimately find that it is possible to decompose the computation of dynamical robustness from \ref{eq:robustnessfromstoch} over a certain kind of modular structure intrinsic to any network architecture.
+Here we introduce the manner in which such a decomposition can be consistently performed for any architecture.
+The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by decomposing the interaction graph into its associated network of strongly connected components (SCCs).
 A SCC of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
@@
 The map from the interaction graph of a network to its SCCs, referred to as $\hier$ in \ref{fig:modsccsym}, results in a decomposition of $G$ into its SCCs.
-Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
+Each node of $\hier (G)$ corresponds to a SCC of $G$.
 There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
+Because of the maximality of SCCs, $\hier (G)$ is acyclic for any $G$.
 
-One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $H$.  To each node $n$ of $H$ associate
+One can also perform this construction in the opposite direction to produce one of the graphs consistent with a given specification of SCCs and their connectivity.
+Start with a directed acyclic graph $H$.
+To each node $n$ of $H$ associate
 a strongly connected graph $C_n$.
@@
 
+\subsection{Symmetries of SCCs and their relationship to robustness}
 This map $\hier$ is many-to-one and so there is a large class of
@@
 \section{Derivation of the relationship between network architecture and robustness}
-Now we derive an analytical expression for dynamical robustness, $R$, of a network in terms of its interaction graph, $G$, as a weighted average of the robustness, $R_{\alpha}$, of the SCCs, $C_{\alpha}$, the corresponding number of links within each SCC, $d_{\alpha}$, and the number of links between the SCCs, $l$ to give $d = l + \sum_{\alpha} d_{\alpha}$.
-The result ultimately holds for the case of simultaneously perturbing any number of elements of the Jacobian.
+Here we show that dynamical robustness is maximized by the most hierarchical of network architectures.
+In doing so, we restrict to a set of graphs which has the same SCCs but may have different links between the components, consider a stochastic process as a simple model for evolution that independently resamples individual elements of Jacobian matrices, and maximize robustness in that context.
+
+We demonstrate that the Jacobian matrix associated to a system at a fixed point is stable in the sense of \ref{eq:stabeval} if and only if each of the blocks corresponding to its SCCs are all independently stable.
+We derive an analytical expression for dynamical robustness, $R$, of a network in terms of its interaction graph, $G$, as a weighted average of the dynamical robustness values, $R_{\alpha}$, of the SCCs, $C_{\alpha}$, the corresponding number of links within each SCC, $d_{\alpha}$, and the number of links between the SCCs, $l$ to give a total connectivity $d = l + \sum_{\alpha} d_{\alpha}$.
+The network architecture having $l=l_{max}$ is the most hierarchical one.
+Comparing the robustness of a network architecture having $l = l_{max}$ to any other network architecture having any other value of $l$, we show that the most hierarchical network architecture maximizes dynamical robustness.
+The result ultimately holds for those processes involving simultaneous perturbation of any number of elements of the Jacobian.
 To anchor the intuition before stating the more general result, we derive the expression for the case of perturbing a single element at each timestep of the process.
-\subsection{Independent modification process to individual network interactions}
+\subsection{Decomposition of stability over SCCs}
 Let the index $i$ range over the non-zero entries $a_i$ of $A$.
@@
 
+\subsection{General computation of dynamical robustness}
+
 To relate the robustness of a graph to the robustness of its SCCs, we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessfromstoch}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
@@
    \int \prod\limits_{h \in C_\alpha} da_h,\rho(a_h) \mathcal{S}(\pi_{C_{\beta}}(A)) \times \\
- \prod\limits_{l \in L} \int da_l\,\rho(a_l) \end{matrix}} \\
+ \prod\limits_{l \in L} \int da_l\,\rho(a_l) \end{matrix}} \\[1em]
 &= \frac{\int \prod\limits_{k \in C_\alpha} da_k\,d{a'}_k\,\rho(a_k)
@@
 Here $R_\alpha$ is shorthand for $R(C_\alpha,S,\mu_\alpha,\tau^{(1)}_\alpha)$ and $R$ is shorthand for $R(G,S,\mu,\tau^{(1)})$.
+
+\subsection{Example computation of robustness}
 Examples of vector fields that correspond to the sampling of Jacobian matrices used in the computation of robustness for particular examples are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
@@
 
+\subsection{Most hierarchical network architecture maximizes robustness}
 Examining this expression noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
@@
 % $\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
-\subsection{Modifying multiple interactions simultaneously}
+\subsection{Simultaneous perturbation of multiple interaction strengths}
 This argument also works when relationships between multiple system components are perturbed simultaneously,
@@
 
-This implies that the interaction graphs for systems that are the most robust will maximize the number of links between SCCs as well as the overall number of SCCs with respect to a particular system size.
+This implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs.
 This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the total ordering will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
@@
 The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
+
+\subsection{Demonstration of the analytical relationship between hierarchy and robustness}
 

### 2015-07-14 tex/overview.tex
remind reader of reuse of n

@@
 The decomposition of a digraph into SCCs corresponds to a block triangular decomposition of its adjacency matrix.
-Say that the graph $G$ has SCCs $C_1, C_2, \ldots C_n$, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
+Say that the graph $G$ has SCCs $C_1, C_2, \ldots C_n$, where we now use $n$ with no index to specify the number of SCCs, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
 Label the vertices in such a way that $V_1, \ldots, V_{n_1}$ belong to $C_1$, $V_{n_1 + 1}, \ldots, V_{n_2}$ belong to $C_2$, etc.

### 2015-07-14 tex/overview.tex
use n prime rather than s to refer to the dimensionality of the parameter space

@@
 \end{equation}
-where $F=\{F_1,\ldots,F_i,\ldots,F_n\}$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^s$ is the vector of $s$ parameters of the $F_i$.
+where $F=\{F_1,\ldots,F_i,\ldots,F_n\}$ represent, potentially nonlinear, functions of the given vector of state variables and $\vec{p} \in \mathbb{R}^{n'}$ is the vector of $n'$ parameters of the $F_i$.
 These parameters typically represent reaction rates or interaction strengths in chemical, gene-regulatory and ecological networks.
 For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (x_1, \ldots, x_n)$, $\vec{p}=(r_1,\ldots,r_n,b_{11},\ldots,b_{nn})$, and $F_i = r_i x_i + \sum_{j=1}^{n} b_{ij} x_i x_j$.
-The set of all dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $s$, is then
+The set of all dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $n'$, is then
 \begin{equation}\label{eq:setdynsys}
-\mathcal{D'} = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^s \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
+\mathcal{D'} = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^{n'} \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
 \end{equation}
@@
 $$
-for each history of length $T$, $s^T = ( s_1, s_2, \ldots, s_T ) \in \{0,1\}^T$
-as visualized in \ref{fig:robustnessprocess}A and B.
+for each history of length $T$, $s^T = ( s_1, s_2, \ldots, s_T ) \in \{0,1\}^T$ where $s=1 \equiv S,\, s=0 \equiv U$ as visualized in \ref{fig:robustnessprocess}A and B.
 In order to model this we consider a process whereby perturbations applied to a given dynamical system and fixed point associated to a stable Jacobian matrix $A$ lead to another Jacobian matrix $A'$.
@@
 The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by decomposing the interaction graph into its associated network of strongly connected components (SCCs).
-A SCC of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}.
+A SCC of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other in some number (not necessarily single) number of steps \cite{Cormen2009}.
 The SCCs of some examples of three variable systems are outlined in \reffigscc{} along with their adjacency matrices.
@@
 
-To relate the robustness of a graph to the robustness of its SCCs, we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessfromstoch}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
+To relate the robustness of a graph to the robustness of its SCCs, we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessbygraph}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
 Let $L$ denote the set of edges of $G$ that connect distinct strongly connected components.

### 2015-07-16 tex/discussion.tex
add reference and discussion of Prill 2005

@@
 Given the manner in which we define robustness in \ref{eq:robustnessfromstoch}, this is a very general constraint.
-In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness as we have defined it here \cite{Zhao2006,Bhardwaj2010,Colm}.
+In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness as we have defined it here \cite{Zhao2006,Bhardwaj2010,Colm,Prill2005}.
+Indeed, Figure 2 from \cite{Prill2005} demonstrates an enrichment of various hierarchical network architectures consisting of multiple SCCs relative to those containing cycles with single SCCs for networks with three and four components in \emph{E.
+coli}, \emph{S.
+cerevisiae}, and \emph{D.
+melanogaster} transcription factor networks as well as the signal transduction knowledge environment network and the \emph{C.
+elegans} neural network.
+The network architecture that we predict to be enriched by selection for robustness for three variable networks ranks 4th out of 13 in overall abundance across these networks and within each network individually with the exception of the \emph{C.
+elegans} neural network, and the three more abundant network architectures are all proper subgraphs of this one.
+These results are essentially consistent with our prediction, though, of course, we would not claim that there could not be various alternative explanations for these observations.
 At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.

### 2015-07-16 tex/figures.tex
update figure legend
move definition of hierarchy in terms of graph edit distance from SI to main text

@@
 This in turn induces a process on the stable and unstable regions, which can be used to estimate robustness.
-We note that $R$ is distinct from the stationary probability of being within the stable region $p^{\infty}(S) = \frac{1-U}{2-R-U}$.
+We note that $R$ is distinct from the stationary probability of being within the stable region $P^{\infty}(S) = \frac{1-U}{2-R-U}$.
 (C) Each network architecture encoded by a directed graph, $\{G_1, \ldots, G_k \}$, that indicates the manner in which the variables of a biological network depend upon one another selects a particular subset of equivalence classes from $\mathcal{F}/{\sim}$ (see \ref{eq:jacgrapheqs}).
@@
 The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
-We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
+% We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
 The two panels in the middle represent examples of hierarchical modular systems that posess both modularity (i.e.

### 2015-07-16 tex/methods.tex
move definition of hierarchy in terms of graph edit distance from SI to main text

@@
 
-\section{Hierarchy and Total ordering}\label{sec:totalordering}
+\section{Graph associated to the total ordering}\label{sec:totalordering}
 
@@
 Some of these operations leave dynamical robustness invariant, so we may regard them as a symmetry groupoid of the systems associated to a given graph $G$ with respect to the robustness property.
-
-The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
-The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc{} (top panel).
-This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
-This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being maximally hierarchical.
-In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.

### 2015-07-16 tex/overview.tex
move definition of hierarchy in terms of graph edit distance from SI to main text

@@
 
+\subsection{Quantification of network hierarchy}
+
+The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
+The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc{} (top panel).
+This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
+The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
+This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being maximally hierarchical.
+In this work we take $max(ED) - ED$ to be the definition of hierarchy, where $max(ED)$ is the maximum edit distance for all graphs with a given number of nodes.
+
 \subsection{Symmetries of SCCs and their relationship to robustness}

### 2015-07-16 tex/overview.tex
update main text

@@
 
-\subsection{Most hierarchical network architecture maximizes robustness}
-Examining this expression noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
+\subsection{Most hierarchical network architecture maximizes dynamical robustness}
+Here we show that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
+This is accomplished by demonstrating that any graph $G$ with fewer than the maximum, $l_{\mathrm{max}}$, of links between the given collection of SCCs must have a lower dynamical robustness than the graph, $G_{\mathrm{max}}$, with $l=l_{\mathrm{max}}$ and that this $G_{\mathrm{max}}$ is indeed the graph $G_{\mathrm{tot}}$ which gives the total ordering on the SCCs.
+Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
 Given two connected components $C_{\alpha}$ and $C_{\beta}$ with $v_{\alpha}$ and $v_{\beta}$ nodes respectively, we have a maximum of $v_{\alpha} v_{\beta}$ links going from $C_{\alpha}$ to $C_{\beta}$.
@@
 been ordered in a way such that, if $(\alpha,\beta) \in \hier(G)$, then $\alpha <
-\beta$.  Hence, $l \le l_{max}$ where
-$$l_{max} = \sum_{\alpha=1}^{n-1}\sum_{\beta=\alpha+1}^{n}v_{\alpha}
+\beta$.  Hence, $l \le l_{\mathrm{max}}$ where
+$$l_{\mathrm{max}} = \sum_{\alpha=1}^{n-1}\sum_{\beta=\alpha+1}^{n}v_{\alpha}
 v_{\beta}~=~\frac{1}{2} \left( \sum_{\alpha=1}^{n} v_{\alpha} \right)^2-\frac{1}{2} \sum_{\alpha=1}^{n}
@@
 Then we have a graph $G_{\mathrm{max}}$ such that $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$, the components of $G_{\mathrm{max}}$ are also $C_1, \ldots, C_n$, but that includes all possible links between each pair of components.
-In this case $R(G)~\le~R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$
+In this case $R(G)~\le~R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{\mathrm{max}}$ substituted for $l$
 \begin{widetext}
@@
 
-This implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs.
-This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the total ordering will be more robust than those associated to any of the other interaction graphs in \reffigscc{}.
-The graph associated to the total ordering is the most hierarchical network architecture for any given number of system components like that of \reffigscc{} top for three component systems where the \emph{highest} component in the hierarchy has directed links to all other nodes in the network, the second highest component has directed links to all other nodes in the network except the highest one, \emph{et cetera} (\refsupp{}).
+This analytical result implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs.
 Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.
+Any network whose associated dynamical system has the interaction graph among its SCCs equivalent to the total ordering is more robust than those associated to any of the other interaction graphs in \reffigscc{}.
+The graph associated to the total ordering is the most hierarchical network architecture for any given number of SCCs like that of \reffigscc{} top for three component systems where the \emph{highest} component in the hierarchy has directed links to all other nodes in the network, the second highest component has directed links to all other nodes in the network except the highest one, \emph{et cetera} (\refsupp{}).
 The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
 
-\subsection{Demonstration of the analytical relationship between hierarchy and robustness}
+\subsection{Examples of the analytical relationship between hierarchy and robustness}
 

### 2015-07-17 tex/methods.tex
add section describing the relationship of dynamical robustness to structural stability

@@
 
-The quality of interest in this paper is robustness, which is related to the concept of structural stability \cite{Smale1967}, whose evaluation requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters, its rate functions, or environmental constraints that restrict it to a subset of its basins of attraction.
+The quality of interest in this paper is robustness, which is related to the concept of structural stability (see \ref{sec:dynrobustandstructstab}), whose evaluation requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters, its rate functions, or environmental constraints that restrict it to a subset of its basins of attraction.
 We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the components and not only to perturbations of the state vector at a given point in time.
@@
 Some of these operations leave dynamical robustness invariant, so we may regard them as a symmetry groupoid of the systems associated to a given graph $G$ with respect to the robustness property.
+
+\section{Relationship between dynamical robustness and structural stability}\label{sec:dynrobustandstructstab}
+
+Dynamical robustness differs from structural stability in several ways: most importantly, dynamical robustness is quantitative whereas structural stability is qualitative.
+The usual definition states that a vector field is structurally stable if there exists an open neighborhood of that vector field so that every element of that neighborhood is topologically conjugate to the original vector field \cite{Smale1967}.
+This definition only requires that there be an open set but makes no reference to the size of that set.
+Thus a system would be considered structurally stable if there were but a tiny open neighborhood of topologically conjugate vector fields.
+In such a situation, even a relatively small perturbation to the system could change its dynamics appreciably.
+To distinguish this from a system which is also structurally stable, but which is surrounded by a large open set of conjugate systems and hence not susceptible to having its dynamics changed by a small perturbation, we introduced our quantitative probabilistic measure of dynamical robustness.
+
+Structural stability is usually formulated in terms of the space of all smooth vector fields endowed with the $C^1$ topology.
+We did not want to do this for two reasons.
+Theoretically, it is not clear how one would impose a probability measure on the set of all vector fields.
+Practically, one does not expect that an environmental fluctuation will often change a dynamical system into any other arbitrary system but that it may introduce a change into a \emph{nearby} system that shares many features with the original.
+Hence we formulate our definition in terms of a finite-dimensional, if still large, parameter space.
+
+Structural stability is defined in terms of topological conjugacy which is a global condition ensuring that all features of the phase space are preserved.
+We, however, are only interested in the portion of the state space near a certain fixed point and are not concerned with perturbations that might change the character of the phase space at points far away from the region of interest as long as they do not affect the neighborhood of the fixed point.
+Thus, rather than being formulated globally, our definition is given locally in terms of the stability of a given fixed point.

### 2015-07-17 tex/figuressupp.tex
update supplemental figure captions

@@
 \noindent\includegraphics[width=0.6\columnwidth]{fig/apstab3x3.pdf}
-\caption{{\bf System stability as a function of connectivity.} The red points represent the average of system stability at each connectivity.}
+\caption{{\bf System stability as a function of connectivity.} The probability of stability is plotted versus connectivity for all three component systems.
+We note that, in contrast to robustness, the stability ranges over less than 10\% of the possible dynamic range and demonstrates no obvious significant correlation with connectivity.
+The red points represent the average of system stability at each connectivity.}
 \label{fig:apstab3x3}
@@
 \noindent\includegraphics[width=0.6\columnwidth]{fig/cycle3x3.pdf}
-\caption{{\bf dynamical robustness as a function of number of cycles.} }
+\caption{{\bf Dynamical robustness as a function of cycle number.} Note that number of cycles alone does not classify networks according to their dynamical robustness.}
 \label{fig:cycle3x3}

### 2015-07-21 tex/methods.tex
update sections on structural stability

@@
 
-The quality of interest in this paper is robustness, which is related to the concept of structural stability (see \ref{sec:dynrobustandstructstab}), whose evaluation requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters, its rate functions, or environmental constraints that restrict it to a subset of its basins of attraction.
+% The quality of interest in this paper is robustness, which is related to the concept of structural stability (see \ref{sec:dynrobustandstructstab}), whose
+The evaluation of \ref{eq:robustnessonjacobians} requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters, its rate functions, or environmental constraints that restrict it to a subset of its basins of attraction.
 We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the components and not only to perturbations of the state vector at a given point in time.
@@
 Theoretically, it is not clear how one would impose a probability measure on the set of all vector fields.
-Practically, one does not expect that an environmental fluctuation will often change a dynamical system into any other arbitrary system but that it may introduce a change into a \emph{nearby} system that shares many features with the original.
-Hence we formulate our definition in terms of a finite-dimensional, if still large, parameter space.
+Practically, we do not consider environmental fluctuations that introduce a smooth simultaneous change to all possible interactions among system variables, but rather those that introduce a change in one or changes in a few interactions among system variables.
+Hence we formulate our definition in terms of a finite-dimensional, if still large, space of possible systems.
 
 Structural stability is defined in terms of topological conjugacy which is a global condition ensuring that all features of the phase space are preserved.
-We, however, are only interested in the portion of the state space near a certain fixed point and are not concerned with perturbations that might change the character of the phase space at points far away from the region of interest as long as they do not affect the neighborhood of the fixed point.
+We, however, are only interested in the portion of the state space near a particular fixed point and are not concerned with perturbations that might change the character of the phase space at points far away from the region of interest as long as they do not affect the neighborhood of the fixed point.
 Thus, rather than being formulated globally, our definition is given locally in terms of the stability of a given fixed point.

### 2015-07-21 tex/overview.tex
motivate the inclusion of symmetries
update sections on structural stability

@@
 \end{equation}
-The space of all $F$ is too large to precisely define probabilities over, so we consider various subspaces $\mathcal{D} \subset \mathcal{D}'$.
-In some cases these may be given by parametrized models with a single fixed form of $F$ and then only $\vec{p}$ is allowed to vary.
+The space of all possible $F$ may be too large to precisely define probabilities over, in which case it is possible to consider various subspaces $\mathcal{D} \subset \mathcal{D}'$.
+In some cases these may even be given by parametrized models with a single fixed form of $F$ and then varying $\vec{p}$ alone would be sufficient to vary within that class of dynamical systems.
 
@@
 We then ask what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
-This quantifies the intuitive statement that, over evolutionary timescales, it is not enough for a system to be stable, but rather that it have a reasonably high probability of remaining stable over continguous timeframes.
+This quantifies the intuitive statement that, over evolutionary timescales, it is not enough for a system to be stable, but rather that it have a reasonably high probability of remaining stable over contiguous timeframes.
 In terms of a history $s^T$ over the states specifying the stability property, $\mathcal{S}$, this is given by
@@
 
+Conceptually, one might consider the quantity defined by \ref{eq:robustnessfromstoch} to be related to structural stability \cite{Smale1967}, but its definition differs in several important ways (see \ref{sec:dynrobustandstructstab}).
+Briefly, structural stability is a qualitative, global property that distinguishes dynamical systems whose associated vector field topologies are invariant to \emph{any} sufficiently small perturbation from those whose topologies undergo some change under any such perturbations.
+Dynamical robustness, on the other hand, is a quantitative, local property that specifies the probability of the stability of a fixed point undergoing a transformation from stable to unstable when the system is subjected to, potentially large, perturbations in one or a few interactions among system variables.
+
 In terms of Jacobian matrices $A$ and $A'$, if the underlying process $P([A]^T)$ has analogous limits
@@
 for $A,A' \in \mathbb{R}^{n \times n}$, \ref{eq:robustnessfromstoch} becomes
-\begin{equation}
+\begin{equation}\label{eq:robustnessonjacobians}
 R(S,\mu,\tau) = \frac{\int_{A,A' \in \mathbb{R}^{n \times n}} d\mu(A) d\tau(A' \mid A) \mathcal{S}(A) \mathcal{S}(A')}{\int_{A \in \mathbb{R}^{n \times n}} d\mu(A) \mathcal{S}(A)}.
@@
 \ref{fig:robustnesssymmetries} shows an example of these latter symmetries applied to a specific interaction graph.
+The collections of networks that have equivalent dynamical robustness characterized by its symmetries allow for a classification of neutral networks with respect to any process in which maximal robustness is selected for.
 

### 2015-07-21 tex/overview.tex
clarify the use of local and global interaction graphs

@@
 \subsection{Interaction graphs encoding network architecture}
-The interactions among variables in a dynamical model for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
+We are interested in the manner in which interactions among system variables are encoded by a directed graph associated to the system or some property of it.
+Strictly speaking, the interactions among variables in a dynamical model for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
 For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$.
-These two conditions on global system interactions are expressed respectively for all $\vec{x}$ in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$.
+These two conditions on global system interactions are expressed locally for all $\vec{x}$ in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$.
 %  \begin{displaymath}
@@
 \end{equation}
-In general, the graph $G_A$ is a subgraph of $G_X$, however, $G_A$ is almost always equivalent to $G_X$ (\refsupp{}).
+In general, the graph $G_A$ is a subgraph of $G_X$, however, $G_A$ is equal to $G_X$, and thus $\adj(G_A) = \adj(G_X)$, for the classes of systems of interest here (see \refsupp{} \ref{sec:reactionnetjacobian}).
 We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
@@
 
-Each distinct directed graph $G$, of which there are $k=2^{n(n-1)}$ that could be associated to the interactions in a model defined on $n$ variables, selects a subset of $\mathcal{F} / {\sim}$ (see~\ref{fig:robustnessprocess}C)
+Each distinct labeled directed graph $G$ (of which there are $k=2^{n(n-1)}$) that could be associated to the interactions in a model defined on $n$ variables, selects a subset of $\mathcal{F} / {\sim}$ (see~\ref{fig:robustnessprocess}C)
 \begin{equation}\label{eq:jacgrapheqs}

### 2015-07-21 tex/abstract.tex
update abstract

@@
 The relationship between network topology and system dynamics has significant implications for unifying our understanding of the interplay among metabolic, gene-regulatory, and ecosystem network architecures.
-Here we analyze the stability and robustness of a large class of dynamics on such networks.
-We determine the probability distribution of robustness as a function of network topology and show that robustness is classified by the number of links between modules of the network.
+Here we analyze the stability and dynamical robustness of a large class of dynamics on such networks, the latter considered to occur over evolutionary timescales.
+We determine the probability distribution of dynamical robustness as a function of network topology and show that dynamical robustness is classified by the number of links between modules of the network.
 We also demonstrate that permutation of these modules is a fundamental symmetry of dynamical robustness.
-Analysis of these findings leads to the conclusion that the most robust systems have the most hierarchical structure.
+Analysis of these findings leads to the conclusion that the most dynamically robust systems correspond to the most hierarchical connectivity of these modules.
 This relationship provides a means by which evolutionary selection for a purely dynamical phenomenon may shape network architectures across scales of the biological hierarchy.

### 2015-07-22 tex/introduction.tex
add section corresponding to introduction
move reference to total ordering from the introduction to the quantification of network hierarchy section

@@
 %!TEX root = ../paper.tex
+\section{Introduction}
+
 The traditional approach taken in the study of chemical reaction, gene-regulatory, population, and ecosystem networks is to derive a system of differential equations to model a particular biological network, attempt to fit that model to data and adjust the modeling assumptions along with parameter values until a good fit is achieved \cite{Meyer2014}.
@@
 
-Here we determine the relationship between network hierarchy, a topological property, and \emph{robustness}, a dynamical property.
+Here we determine the relationship between network hierarchy, a topological property, and \emph{dynamical robustness}, a property of the dynamics.
 Robustness is of interest in biological systems at all scales, and has been previously studied in the context of biochemical networks \cite{Alon1999,Shinar2010}, gene-regulatory networks \cite{VanNimwegen1999,Siegal2002,Draghi2010,Wagner2013}, and ecological networks \cite{Rohr2014}.
@@
 
-We demonstrate that systems exhibiting maximal robustness with respect to this definition have the most hierarchical network topology and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
-For the purpose of formulating this result, the maximally hierarchical network is considered to be the graph associated to the total ordering (\refsupp{} and \cite{Cormen2009}).
-An example of this for three system components is shown in \reffigscc{} top.
-We use a measure of hierarchy based upon the edit distance from this maximally hierarchical network \cite{Axenovich2011}.
+We demonstrate that systems exhibiting maximal robustness with respect to this definition have the most hierarchical network topology relative to a specification of system modules and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
 Our results hold for networks of arbitrary size and are independent of the probability distribution from which the strengths of interaction are sampled.

### 2015-07-22 tex/overview.tex
move reference to total ordering from the introduction to the quantification of network hierarchy section

@@
 The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
-The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \reffigscc{} (top panel).
+The most hierarchical system is considered to be the graph corresponding to the total ordering (\refsupp{} and \cite{Cormen2009}), which for three nodes is given in \reffigscc{} (top panel).
 This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.

### 2015-07-22 tex/overview.tex
fix overflow equation numbering

@@
 For example, in the Lotka-Volterra model in \ref{fig:biomodelexamples}, $\vec{x} = (x_1, \ldots, x_n)$, $\vec{p}=(r_1,\ldots,r_n,b_{11},\ldots,b_{nn})$, and $F_i = r_i x_i + \sum_{j=1}^{n} b_{ij} x_i x_j$.
-The set of all dynamical systems, $\mathcal{D}$, for a given number of state variables $n$ and a given number of parameters $n'$, is then
+The set of all dynamical systems, $\mathcal{D}'$, for a given number of state variables $n$ and a given number of parameters $n'$, is then
 \begin{equation}\label{eq:setdynsys}
-\mathcal{D'} = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^{n'} \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
+\mathcal{D}' = \{ (F(-,\vec{p}),\vec{p}) | F \colon \mathbb{R}^n \times \mathbb{R}^{n'} \rightarrow \mathbb{R}^n, \vec{p} \in \mathbb{R}^s \}.
 \end{equation}
@@
 The Lotka-Volterra model has two fixed points: the trivial one of all zero species $x_i=0$ and the other given implicitly by $r_i + \sum_{j=1}^{n} b_{ij} x_j = 0$.
-The set of all such fixed points, $\mathcal{F}$, is then
+The set of all such fixed points, $\mathcal{F}$, for a collection of dynamical systems $\mathcal{D}$ is then
 \begin{equation}\label{eq:setfixedpoints}
-\mathcal{F} = \{ (d,\vec{x}^0) \; | \; d=(F(-,\vec{p}),\vec{p}) \in \mathcal{D}, \vec{x}^0 \in \mathbb{R}^n,F(\vec{x}^0,\vec{p})=0 \}.
+\mathcal{F} = \left\{ (d,\vec{x}^0) \; \left| \;
+\begin{aligned}
+d = (F(-,\vec{p}),\vec{p}) &\in \mathcal{D}, \\
+\vec{x}^0 &\in \mathbb{R}^n,\\
+F(\vec{x}^0,\vec{p}) &= 0
+\end{aligned}
+\right.
+\right\}.
 \end{equation}

### 2015-07-22 tex/introduction.tex
add relevant citations to introduction

@@
 This is necessary to understand in order to determine whether or not a given dynamical phenomenon should be regarded as unique or generic in the development and investigation of models applied to particular systems \cite{Gunawardena2013,Gunawardena2014}.
-This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems that, in comparison to one another, appear to have components that are randomly interlinked.
+This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems \cite{Bialek2012}.
+For the case considered here, the ensemble of interest is defined by dynamical systems that, in comparison to one another, appear to have components that are randomly interlinked.
 % \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}
@@
 The class of dynamical systems studied by May is, however, not restricted to ecosystem dynamics and encompasses, among others, the dynamics of all of the networks represented in \ref{fig:biomodelexamples}.
-May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
+May conjectured on the basis of results from random matrix theory what eventually came to be referred to as the May-Wigner stability theorem \cite{Cohen1984,May1972a,McMurtrie1975,Radius2014,Majumdar2014}, which implies a relationship between a topological property, system connectivity, and a dynamical property, stability.
 

### 2015-07-22 tex/overview.tex
add relevant citations to introduction

@@
 \subsection{Stability analysis of biological networks}
-To determine stability, we use the Taylor series expansion of the equations of motion \ref{eq:eom} about the fixed point $\vec{x}^0$ where $\vec{y} = \vec{x} - \vec{x}^0$ by
+To determine stability, we use the Taylor series expansion of the equations of motion, \ref{eq:eom}, about the fixed point $\vec{x}^0$ where $\vec{y} = \vec{x} - \vec{x}^0$ given by
 \begin{equation}\label{eq:taylorseries}
@@
 & + \frac{1}{2}\sum_{j,k=1}^{N} \left. \frac{\partial^2 F_i}{\partial x_j \partial x_k} \right|_{\vec{x} = \vec{x}^0} y_j y_k + \cdots
-\end{aligned}
+\end{aligned}.
 \end{equation}

### 2015-07-22 tex/introduction.tex
update introduction

@@
 
-Rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change.
+Rather than evaluating system stability for a particular model of a biological network over physiological timescales, we are interested in evaluating dynamical robustness over evolutionary timescales where the form of the most accurate underlying model is itself subject to change.
 Over evolutionary timescales it is expected for there to be fluctuations, not only in the state, but in any aspect of the model specifying the dynamics of the network itself (i.e.
@@
 However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process subject to environmental constraints.
-This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon2002}.
+This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon1962,Simon2002}.
 What is important over at least moderate evolutionary timescales is then the conditional probability, $R$, given a system is in a stable state, that upon modifications to its structure in the context of environmental fluctuations, it remains stable, regardless of where the stable state is located within the state space \ref{fig:robustnessconcept}.
@@
 
-We demonstrate that systems exhibiting maximal robustness with respect to this definition have the most hierarchical network topology relative to a specification of system modules and explain why this results from an invariance of robustness to particular kinds of transformations of the network topology.
+We demonstrate that systems exhibiting maximal dynamical robustness with respect to this definition have the most hierarchical network topology relative to a specification of system modules and explain why this results from an invariance of dynamical robustness to particular kinds of transformations of the network topology.
 Our results hold for networks of arbitrary size and are independent of the probability distribution from which the strengths of interaction are sampled.

### 2015-07-22 tex/introduction.tex
update introduction

@@
 It is therefore also expected that network architectures managing to persist over such timescales may be required to do so with modifications to the location of their stable states and even to the geometry of their state spaces.
-However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining in a stable state upon modifications to the underlying dynamical process subject to environmental constraints.
+However, what must remain invariant on such evolutionary timescales is the higher-level property that the system possess a relatively high overall probability of remaining near a stable state upon modifications to the underlying dynamical process subject to environmental constraints.
 This is necessary in order for networks of lower-level components to exist, regardless of what state they exist in, long enough to serve as the substrate out of which networks of relatively higher-level components are constructed \cite{Simon1962,Simon2002}.

### 2015-07-22 tex/overview.tex
add intuitive sentences motivating the definition of equivalence classes of fixed points determined by Jacobian matrices

@@
 \subsection{Equivalence classes of systems associated to Jacobian Matrices}
-Jacobian matrices define an equivalence relation, $\sim$, on fixed points given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}\,',\vec{x}^{0'}) $ if and only if
+We would ultimately like to determine the probability of a stable system remaining stable in the context of evolution modeled as a stochastic process operating on a large set of dynamical systems.
+Any given Jacobian matrix, which is what characterizes the stability of a system near a given fixed point, does not refer to a single dynamical system, but rather a set of fundamentally distinct dynamical systems from which that Jacobian matrix could have derived.
+More precisely, Jacobian matrices define an equivalence relation, $\sim$, on fixed points, $\mathcal{F}$, given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}\,',\vec{x}^{0'}) $ if and only if
 \begin{equation}\label{eq:jaceqrel}
@@
 \begin{equation}\label{eq:jaceqs}
-[A] = \left\{ (F,\vec{p},\vec{x}^0) \; | \; \left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} = A \right\}.
+[A] = \left\{ (F,\vec{p},\vec{x}^0) \; \Bigg| \; \left. \frac{\partial F(\vec{x},\vec{p})}{\partial \vec{x}} \right|_{\vec{x} = \vec{x}^0} = A \right\}.
 \end{equation}

### 2015-07-22 tex/discussion.tex
change some usages of components to variables

@@
 In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness as we have defined it here \cite{Zhao2006,Bhardwaj2010,Colm,Prill2005}.
-Indeed, Figure 2 from \cite{Prill2005} demonstrates an enrichment of various hierarchical network architectures consisting of multiple SCCs relative to those containing cycles with single SCCs for networks with three and four components in \emph{E.
+Indeed, Figure 2 from \cite{Prill2005} demonstrates an enrichment of various hierarchical network architectures consisting of multiple SCCs relative to those containing cycles with single SCCs for networks with three and four variables in \emph{E.
 coli}, \emph{S.

### 2015-07-22 tex/figuressupp.tex
change some usages of components to variables

@@
 \noindent\includegraphics[width=0.9\columnwidth]{fig/jacobianvectorfields.pdf}
-\caption{{\bf Vector fields resulting from the random sampling of two component systems.} System parameters of \ref{eq:arbitraryjacobianreactionnetwork} are rescaled to ensure the fixed point is located at $(1,1)$.
+\caption{{\bf Vector fields resulting from the random sampling of two variable systems.} System parameters of \ref{eq:arbitraryjacobianreactionnetwork} are rescaled to ensure the fixed point is located at $(1,1)$.
 The color of the dot located at the fixed point indicates whether it is stable (black) or unstable (gray).}
@@
 \noindent\includegraphics[width=0.6\columnwidth]{fig/apstab3x3.pdf}
-\caption{{\bf System stability as a function of connectivity.} The probability of stability is plotted versus connectivity for all three component systems.
+\caption{{\bf System stability as a function of connectivity.} The probability of stability is plotted versus connectivity for all three variable systems.
 We note that, in contrast to robustness, the stability ranges over less than 10\% of the possible dynamic range and demonstrates no obvious significant correlation with connectivity.

### 2015-07-22 tex/introduction.tex
change some usages of components to variables

@@
 This can be achieved using a method common in statistical physics involving the consideration of an ensemble of systems \cite{Bialek2012}.
-For the case considered here, the ensemble of interest is defined by dynamical systems that, in comparison to one another, appear to have components that are randomly interlinked.
+For the case considered here, the ensemble of interest is defined by dynamical systems that, in comparison to one another, appear to have variables that are randomly interlinked.
 % \cite{Brown2003,Gutenkunst2007,Daniels2008a,Machta2013,Hines2014,Prabakaran2014,Tonsing2014}
@@
 Over physiological timescales, the robustness of a particular network state may be evaluated by determining its linear stability \cite{Davis1962}.
-Network states that are linearly stable, are robust to perturbations in the states of their components.
+Network states that are linearly stable, are robust to perturbations in the states of their variables.
 For example, in the case of gene-regulatory networks, a state vector containing protein molecule counts or concentrations of each gene that is stable with respect to the dynamics of the network will exponentially suppress any relatively small modifications to the state of a gene and return to the initial stable state.

### 2015-07-22 tex/methods.tex
change some usages of components to variables

@@
 The evaluation of \ref{eq:robustnessonjacobians} requires the determination of whether or not a given dynamical system that is determined to be stable remains stable under a perturbation to one or more of its defining parameters, its rate functions, or environmental constraints that restrict it to a subset of its basins of attraction.
-We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the components and not only to perturbations of the state vector at a given point in time.
+We mean to refer to perturbations to the structure of the system itself as determined by the strengths of the couplings between the variables and not only to perturbations of the state vector at a given point in time.
 It is justified to consider resampling elements of $A$ to generate $A'$ as a proxy for resampling elements of $\vec{p}$ to produce $\vec{p}\,'$ if any matrix $A$ can be obtained for some $F_i$, $\vec{p}$ and $\vec{x}^0$.
@@
 
- We refer to the most hierarchical network architecture as the directed graph associated to a total ordering on the set of system components corresponding to the set of nodes, $V$, of the graph \cite{Cormen2009}.
+ We refer to the most hierarchical network architecture as the directed graph associated to a total ordering on the set of system variables corresponding to the set of nodes, $V$, of the graph \cite{Cormen2009}.
 In general, a totally ordered set is a pair $(S,R)$ consisting of a set $S$ together with a total order relation $R$ on it.

### 2015-07-22 tex/overview.tex
change some usages of components to variables

@@
 
-In the general case of a dynamical system with $n$ components, where the components may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of state variables or observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
+In the general case of a dynamical system with $n$ variables, where the variables may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of state variables or observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
 $
@@
 \subsection{Interaction graphs encoding network architecture}
-We are interested in the manner in which interactions among system variables are encoded by a directed graph associated to the system or some property of it.
+We are interested in the manner in which system variables interact with one another, and, ultimately, whether dynamical robustness depends upon the topology of these interactions.
+Interactions among system variables can be encoded in a directed graph associated to the system.
 Strictly speaking, the interactions among variables in a dynamical model for any network can be represented in terms of a global interaction graph (\ref{fig:biomodelexamples} top row).
-For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another is given by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$.
+For a general system $X \in \mathcal{D}$ the directed graph $G_X$ that describes the manner in which each of the variables depends upon one another can be represented by the adjacency matrix $\adj(G_X)$ where $\adj(G_X)_{ij}$ is $1$ if $F_i \hbox{ depends on } x_j$ and $0$ if $F_i \hbox{ does not depend on } x_j$.
 These two conditions on global system interactions are expressed locally for all $\vec{x}$ in terms of elements of the Jacobian matrix as $\frac{\partial F_i}{\partial x_j}(\vec{x}) \neq 0$ and $\frac{\partial F_i}{\partial x_j}(\vec{x}) = 0$.
@@
 % \end{displaymath}
-For large systems, since any given component is only likely to interact with a relatively small proportion of the other components, these matrices may be sparse.
+For large systems, since any given variable is only likely to interact with a relatively small proportion of the other variables, these matrices may be sparse.
 We can also associate a local interaction graph $G_A$ given by an adjacency matrix $\adj(G_A)$ to each dynamical system having Jacobian matrix $A$ at some fixed point $\vec{x}^0$ where
@@
 for each history of length $T$, $s^T = ( s_1, s_2, \ldots, s_T ) \in \{0,1\}^T$ where $s=1 \equiv S,\, s=0 \equiv U$ as visualized in \ref{fig:robustnessprocess}A and B.
-In order to model this we consider a process whereby perturbations applied to a given dynamical system and fixed point associated to a stable Jacobian matrix $A$ lead to another Jacobian matrix $A'$.
-This corresponds to an ensemble of fixed points of dynamical systems where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{}).
+In order to model this, we consider a process whereby perturbations applied to a given dynamical system and fixed point associated to a stable Jacobian matrix $A$ lead to another Jacobian matrix $A'$.
+This corresponds to an ensemble of fixed points of dynamical systems where each model in the ensemble may otherwise be defined in terms of a different collection of rate functions $F'$, vector of parameters ${\vec{p}}\,'$, or environmental conditions restricting access to $\vec{x}^{0'}$ (\refsupp{} \ref{sec:reactionnetjacobian}).
 We then ask what is the probability, given $A$ is a stable matrix, that $A'$ is also a stable matrix.
@@
 
-Conceptually, one might consider the quantity defined by \ref{eq:robustnessfromstoch} to be related to structural stability \cite{Smale1967}, but its definition differs in several important ways (see \ref{sec:dynrobustandstructstab}).
-Briefly, structural stability is a qualitative, global property that distinguishes dynamical systems whose associated vector field topologies are invariant to \emph{any} sufficiently small perturbation from those whose topologies undergo some change under any such perturbations.
+Conceptually, one might consider the quantity defined by \ref{eq:robustnessfromstoch} to be related to structural stability \cite{Smale1967}, but its definition differs in several important ways (see \refsupp{} \ref{sec:dynrobustandstructstab}).
+Briefly, structural stability is a qualitative, global property that distinguishes dynamical systems whose associated vector field topologies are invariant to \emph{any} sufficiently small perturbation from those whose topologies change for \emph{some} perturbation of any size.
 Dynamical robustness, on the other hand, is a quantitative, local property that specifies the probability of the stability of a fixed point undergoing a transformation from stable to unstable when the system is subjected to, potentially large, perturbations in one or a few interactions among system variables.
@@
 \subsection{Simultaneous perturbation of multiple interaction strengths}
-This argument also works when relationships between multiple system components are perturbed simultaneously,
+This argument also works when relationships between multiple system variables are perturbed simultaneously,
 although the notation becomes more complicated.  Suppose that we
@@
 This analytical result implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs.
-Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system.
+Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the variable interaction strengths are sampled or the size of the system.
 Any network whose associated dynamical system has the interaction graph among its SCCs equivalent to the total ordering is more robust than those associated to any of the other interaction graphs in \reffigscc{}.
-The graph associated to the total ordering is the most hierarchical network architecture for any given number of SCCs like that of \reffigscc{} top for three component systems where the \emph{highest} component in the hierarchy has directed links to all other nodes in the network, the second highest component has directed links to all other nodes in the network except the highest one, \emph{et cetera} (\refsupp{}).
+The graph associated to the total ordering is the most hierarchical network architecture for any given number of SCCs like that of \reffigscc{} top for three variable systems where the \emph{highest} variable in the hierarchy has directed links to all other variables in the network, the second highest variable has directed links to all other variables in the network except the highest one, \emph{et cetera} (\refsupp{}).
 The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
@@
 
-To test the prediction of the analytical results in \ref{eq:robschematic} and \ref{eq:robustnessmultiple}, we computed approximations to the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting components (see \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
+To test the prediction of the analytical results in \ref{eq:robschematic} and \ref{eq:robustnessmultiple}, we computed approximations to the probability distribution of stability and dynamical robustness relative to network architecture for ensembles of systems having two or three interacting variables (see \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}).
 For all of these, we found that robustness is correlated with connectivity, but that the most robust systems have intermediate connectivity for a given network size (\reffigrobustconnect).

### 2015-07-22 tex/overview.tex
fix typo

@@
 We would like to determine whether selection for dynamical robustness would result in a preference for the maintenance of one kind of network architecture relative to another.
-In order to then classify the properties of this process according to network architecture, we consider cases where the condition $\adj(G_A) = \adj(G_{A'})$ holds, which results in an analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
+In order to then classify the properties of this process according to network architecture, we consider cases where the condition $\adj(G_A) = \adj(G_{A'})$ holds, which results in analogous processes defined on each $G$-class $[G] \in \mathcal{F}/G$ given by
 $$

### 2015-07-22 tex/overview.tex
state that all networks with certain architectures are unstable

@@
 This partition, $\mathcal{F} / G$, is a coarsening of $\mathcal{F} / {\sim}$.
+In what follows we consider the stability of classes of fixed points associated to each possible graph $G$.
+In most cases some of these are stable and some of them are unstable, which we quantify as the probability of stability.
+However, note that for some $G$, all fixed points respecting the topology of $G$ are unstable.
 

### 2015-07-22 tex/overview.tex
clarify acyclicity of SCC graph

@@
 There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of SCCs, $\hier (G)$ is acyclic for any $G$.
+Note that $\hier (G)$ is acyclic for any $G$ because of the maximality property of SCCs.
 
@@
 
-The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
+The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of nodes suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents.
 The most hierarchical system is considered to be the graph corresponding to the total ordering (\refsupp{} and \cite{Cormen2009}), which for three nodes is given in \reffigscc{} (top panel).
-This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components.
-The graph edit distance (ED) on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
+This graph possesses the maximum number of SCCs for any graph with a given number of nodes, $n$, as well as the maximum number, $\binom{n}{2}$, of links between SCCs.
+The graph edit distance (ED) on a fixed number of nodes from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second \cite{Axenovich2011}.
 This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being maximally hierarchical.
@@
 
-\subsection{Most hierarchical network architecture maximizes dynamical robustness}
+\subsection{Most hierarchical network architectures maximize dynamical robustness}
 Here we show that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.

### 2015-07-22 tex/overview.tex
consistently use SCC abbreviation

@@
 These three symmetries, \reffigscc{}, represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated.
-For instance, we may interchange the positions of the strongly
-connected components relative to each other \reffighiertransformations$a$.
+For instance, we may interchange the positions of the SCCs relative to each other \reffighiertransformations$a$.
 Leaving the components fixed, we may move links between nodes in a component \reffighiertransformations$b$ or between components, or even add or delete links \reffighiertransformations$c$.
@@
 To relate the robustness of a graph to the robustness of its SCCs, we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessbygraph}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
-Let $L$ denote the set of edges of $G$ that connect distinct strongly connected components.
+Let $L$ denote the set of edges of $G$ that connect distinct SCCs.
 Then, if $i \in C_\alpha$ for some SCC $C_\alpha$, we have
@@
 m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
-$m_0$ links between strongly connected components.  Hence, our
+$m_0$ links between SCCs.  Hence, our
 weighted average becomes
@@
 increase $R(G, S, \mu, \tau^{(m)})$ by increasing the maximum
-possible value of $m_0$ while keeping the strongly connected
-components the same.  Again, if we fix $\hier(G)$, the maximum
+possible value of $m_0$ while keeping the SCCs the same.  Again, if we fix $\hier(G)$, the maximum
 possible value of $m_0$ is $\sum_{(\alpha,\beta) \in \hier(G)} v_\alpha v_\beta$

### 2015-07-22 tex/methods.tex
use abbreviation for SCCs
clarify acyclicity of SCC graph
consistently use node rather than vertex

@@
 Corresponding to the SCC decomposition of $G$ we can construct a directed acyclic graph $\hier (G)$ or the condensed graph \cite{Corominas-Murtra2013}.
-Each node of $\hier (G)$ corresponds to a strongly connected component of $G$.
-There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
-Because of the maximality of strongly connected components, $\hier (G)$ is acyclic.
+Each node of $\hier (G)$ corresponds to a SCC of $G$.
+There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some node in $C$ to some node in $C'$ in $G$.
 

### 2015-07-22 tex/overview.tex
consistently use node rather than vertex

@@
 The network architecture represented in terms of the adjacency matrix, \ref{eq:lininterdepadj}, can be abstracted into modules by decomposing the interaction graph into its associated network of strongly connected components (SCCs).
-A SCC of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other in some number (not necessarily single) number of steps \cite{Cormen2009}.
+A SCC of a graph is a maximal subset of nodes where each node within the subset can be reached from any other in some number of (not necessarily single) steps \cite{Cormen2009}.
 The SCCs of some examples of three variable systems are outlined in \reffigscc{} along with their adjacency matrices.
@@
 Each node of $\hier (G)$ corresponds to a SCC of $G$.
-There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some vertex in $C$ to some vertex in $C'$ in $G$.
+There is an edge from the node corresponding to component $C$ to the node corresponding to component $C'$ if and only if there exists a link from some node in $C$ to some node in $C'$ in $G$.
 Note that $\hier (G)$ is acyclic for any $G$ because of the maximality property of SCCs.
@@
 a strongly connected graph $C_n$.
-To each link $(i,j)$ of $H$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.
+To each link $(i,j)$ of $H$ associate a non-empty subset of $\Node(C_i) \times \Node(C_j)$.
 The result will be a graph $G$ such that $\hier(G) = H$ and furthermore, every graph $G$ such that $\hier(G) = H$ can be obtained in this manner.
@@
 The decomposition of a digraph into SCCs corresponds to a block triangular decomposition of its adjacency matrix.
-Say that the graph $G$ has SCCs $C_1, C_2, \ldots C_n$, where we now use $n$ with no index to specify the number of SCCs, which have been labelled in such a way that there are no links from vertices in component $C_i$ to component $C_j$ when $i < j$.
-Label the vertices in such a way that $V_1, \ldots, V_{n_1}$ belong to $C_1$, $V_{n_1 + 1}, \ldots, V_{n_2}$ belong to $C_2$, etc.
-Then, if we choose basis vectors corresponding to this labelling of the vertices, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
+Say that the graph $G$ has SCCs $C_1, C_2, \ldots C_n$, where we now use $n$ with no index to specify the number of SCCs, which have been labelled in such a way that there are no links from nodes in component $C_i$ to component $C_j$ when $i < j$.
+Label the nodes in such a way that $V_1, \ldots, V_{n_1}$ belong to $C_1$, $V_{n_1 + 1}, \ldots, V_{n_2}$ belong to $C_2$, etc.
+Then, if we choose basis vectors corresponding to this labelling of the nodes, we will have $a_{ij} = 0$ whenever $i$ and $j$ correspond to different components and $i > j$.
 This condition is equivalent to stating that the matrix is block triangular with blocks of size $n_1, n_2, \ldots$.

### 2015-07-22 tex/overview.tex
clarify introduction to derivation section

@@
 
+To begin our analysis we will define dynamical systems on biological networks, review the manner in which their linear stability can be evaluated and describe how each one can be associated to a graph encoding interactions among system variables.
 In the general case of a dynamical system with $n$ variables, where the variables may be concentrations of chemical species, genes, or biological species, we have an $n$-dimensional vector of state variables or observables $(x_1(t), \ldots x_n(t)) = \vec{x}(t)
@@
 Any given Jacobian matrix, which is what characterizes the stability of a system near a given fixed point, does not refer to a single dynamical system, but rather a set of fundamentally distinct dynamical systems from which that Jacobian matrix could have derived.
-More precisely, Jacobian matrices define an equivalence relation, $\sim$, on fixed points, $\mathcal{F}$, given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}\,',\vec{x}^{0'}) $ if and only if
+More precisely, matrices of numbers corresponding to some Jacobian matrix evaluated at some fixed point define an equivalence relation, $\sim$, on the set of fixed points, $\mathcal{F}$, given by $(F,\vec{p},\vec{x}^0) \sim (F',\vec{p}\,',\vec{x}^{0'}) $ if and only if
 \begin{equation}\label{eq:jaceqrel}
@@
 \end{equation}
-In general, the graph $G_A$ is a subgraph of $G_X$, however, $G_A$ is equal to $G_X$, and thus $\adj(G_A) = \adj(G_X)$, for the classes of systems of interest here (see \refsupp{} \ref{sec:reactionnetjacobian}).
+In general, the graph $G_A$ is a subgraph of $G_X$, however, for the classes of systems of interest here, $G_A$ is equal to $G_X$, and thus $\adj(G_A) = \adj(G_X)$ (see \refsupp{} \ref{sec:reactionnetjacobian}).
 We define the connectivity to be equal to the number of edges in $G_A$, which is equivalent to summing up the number of non-zero entries of $\adj(G_A)$.
@@
 In the course of biological evolution, the parameter values $\vec{p}$, form of the functions $F$, and environmental conditions restricting access to the basins associated to different fixed points $\vec{x}^0$ corresponding to all different types of networks considered in \ref{fig:biomodelexamples} are subject to, potentially drastic, modifications due to environmental fluctuations.
-The stochastic process by which these modifications occur induces one on the set of fixed points that results in the assignment of a probability $P(f^T)$ to each history of length $T$, $f^T = ( f_1,f_2,\ldots,f_T ) \in \mathcal{F}^T$ \cite{RobertM.Gray130}.
+The stochastic process by which these modifications occur induces a process on the set of fixed points that results in the assignment of a probability $P(f^T)$ to each history of length $T$, $f^T = ( f_1,f_2,\ldots,f_T ) \in \mathcal{F}^T$ \cite{RobertM.Gray130}.
 % $$
@@
 This map $\hier$ is many-to-one and so there is a large class of
-operations which leaves $\hier(G)$ invariant for a given graph $G$ \reffighiertransformations{}.
+operations which leaves $\hier(G)$ invariant for a given graph $G$, \reffighiertransformations{}.
 These three symmetries, \reffigscc{}, represent transformations that can be performed on the interaction graph that do not change the network of SCCs to which it is associated.
@@
 \ref{fig:robustnesssymmetries} shows an example of these latter symmetries applied to a specific interaction graph.
-The collections of networks that have equivalent dynamical robustness characterized by its symmetries allow for a classification of neutral networks with respect to any process in which maximal robustness is selected for.
+The collections of networks that have equivalent dynamical robustness characterized by its symmetries allow for a classification of neutral networks with respect to any evolutionary process selecting for dynamical robustness.
 
@@
 Here we show that dynamical robustness is maximized by the most hierarchical of network architectures.
-In doing so, we restrict to a set of graphs which has the same SCCs but may have different links between the components, consider a stochastic process as a simple model for evolution that independently resamples individual elements of Jacobian matrices, and maximize robustness in that context.
+In doing so, we restrict to a set of networks that has the same SCCs but may have different number and connectivity of links between the components, consider a stochastic process as a simple model for evolution that independently resamples individual elements of Jacobian matrices, and maximize robustness in that context.
 
@@
 We derive an analytical expression for dynamical robustness, $R$, of a network in terms of its interaction graph, $G$, as a weighted average of the dynamical robustness values, $R_{\alpha}$, of the SCCs, $C_{\alpha}$, the corresponding number of links within each SCC, $d_{\alpha}$, and the number of links between the SCCs, $l$ to give a total connectivity $d = l + \sum_{\alpha} d_{\alpha}$.
-The network architecture having $l=l_{max}$ is the most hierarchical one.
-Comparing the robustness of a network architecture having $l = l_{max}$ to any other network architecture having any other value of $l$, we show that the most hierarchical network architecture maximizes dynamical robustness.
-The result ultimately holds for those processes involving simultaneous perturbation of any number of elements of the Jacobian.
-To anchor the intuition before stating the more general result, we derive the expression for the case of perturbing a single element at each timestep of the process.
+For a given collection of SCCs, we show that the network architectures maximizing dynamical robustness must maximize the number of links between those SCCs.
+The latter corresponds to the most hierarchical network architecture that can be constructed from that set of SCCs.
+%The network architecture having $l=l_{max}$ is the most hierarchical one.
+Comparing the robustness of a network architecture having $l = l_{max}$ to any other network architecture having any other value of $l$, we show that the network architecture having $l = l_{max}$ maximizes dynamical robustness.
+These results ultimately hold for those processes involving simultaneous perturbation of any number of elements of the Jacobian.
+To anchor the intuition before stating the more general result, we derive the expression for the case of perturbing a single element at each timestep of the stochastic process being used as a simplistic model of evolution.
+
 \subsection{Decomposition of stability over SCCs}

### 2015-07-22 tex/overview.tex
add introductory sentence to stability decomposition subsection

@@
 \subsection{Decomposition of stability over SCCs}
+We begin by expressing the stability of a network in terms of the stabilities of its SCCs.
 Let the index $i$ range over the non-zero entries $a_i$ of $A$.

### 2015-07-23 tex/overview.tex
add example demonstrating maximization of robustness by maximizing number of links between sccs

@@
 
-\subsection{General computation of dynamical robustness}
+\subsection{Decompositon of dynamical robustness over SCCs}
 
-To relate the robustness of a graph to the robustness of its SCCs, we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessbygraph}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
+We use the fact that stability decomposes over the SCCs of a network to relate the robustness of a network to the robustness of its SCCs.
+To derive this relationship we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessbygraph}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
 Let $L$ denote the set of edges of $G$ that connect distinct SCCs.
@@
 
-\subsection{Example computation of robustness}
-Examples of vector fields that correspond to the sampling of Jacobian matrices used in the computation of robustness for particular examples are shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
-\ref{eq:robschematic} shows a schematized version of \ref{eq:robustnessbygraph} for the case of perturbing a single element at a time  (see \reffigscc$\,$ for examples demonstrating this expression).
+\subsection{Example computation of dynamical robustness}
+Before proceeding to demonstrate which network architectures maximize dynamical robustness we illustrate how the result of the previous section applies to simple examples.
+Dynamical robustness can be computed for systems like those that derive from the sampling of Jacobian matrices shown in \ref{fig:robustnessconcept} and \ref{fig:jacobianvectorfields}.
+\ref{eq:robschematic} is a schematized version of \ref{eq:robustnessbygraph} for the case of perturbing a single element at a time.
+\reffigscc$\,$ shows examples demonstrating the usage of this expression.
 For instance, if our graph is the one in \reffigscc{} (middle panels), then we have two connected components, one with two nodes, and one with one node.
@@
 The graph with one node corresponds to a $1 \times 1$ matrix, so we have probability $0.5$ of stability and robustness $0.5$.
-Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc{}, which agrees with the value computed in \ref{tab:structstabmat} up to sampling error.
+Thus, the probability of our 3-node graph being stable is $0.5 \times 0.25 = 0.125$ and its robustness is computed from \ref{eq:robschematic} in \reffigscc{}, which agrees with the value computed in \ref{tab:structstabmat3} up to sampling error.
+
+Due to the symmetries of robustness, there are two more network architectures having the same SCCs but different dynamical robustness values.
+Their values for dynamical robustness are given for $l=0,1,2$ in
+$$
+R = \frac{l+1 \times 0.5 + 4 \times 0.62}{l+1+4}.
+$$
+Inspection of this formula demonstrates that dynamical robustness takes on higher values for $l=2$ than $l=1$ and also for $l=1$ than $l=0$.
+These correspond to network architectures specified by the following adjacency matrices
+\begin{equation}
+\begin{aligned}
+\begin{pmatrix}
+1 & 0 & 0\\
+0 & 1 & 1\\
+0 & 1 & 1\\
+\end{pmatrix}, & \;\; l = 0, \;\; & R = 0.596,\\
+\begin{pmatrix}
+1 & 1 & 0\\
+0 & 1 & 1\\
+0 & 1 & 1\\
+\end{pmatrix}, & \;\; l = 1, \;\; & R = 0.663\\
+\begin{pmatrix}
+1 & 1 & 1\\
+0 & 1 & 1\\
+0 & 1 & 1\\
+\end{pmatrix}, & \;\; l = 2, \;\; & R = 0.711,
+\end{aligned},
+\end{equation}
+where the SCCs correspond to the upper $1 \times 1$ and lower $2 \times 2$ blocks and the other entries correspond to $l$, the links between the SCCs.
+Thus, in this simple case, we see that maximizing the number of links between SCCs maximizes dynamical robustness.
 
 \subsection{Most hierarchical network architectures maximize dynamical robustness}
-Here we show that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
+Following the examples of the previous section, here we show, in general, that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
 This is accomplished by demonstrating that any graph $G$ with fewer than the maximum, $l_{\mathrm{max}}$, of links between the given collection of SCCs must have a lower dynamical robustness than the graph, $G_{\mathrm{max}}$, with $l=l_{\mathrm{max}}$ and that this $G_{\mathrm{max}}$ is indeed the graph $G_{\mathrm{tot}}$ which gives the total ordering on the SCCs.

### 2015-07-23 tex/overview.tex
clarify meaning of gmax and gtot

@@
 Following the examples of the previous section, here we show, in general, that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
-This is accomplished by demonstrating that any graph $G$ with fewer than the maximum, $l_{\mathrm{max}}$, of links between the given collection of SCCs must have a lower dynamical robustness than the graph, $G_{\mathrm{max}}$, with $l=l_{\mathrm{max}}$ and that this $G_{\mathrm{max}}$ is indeed the graph $G_{\mathrm{tot}}$ which gives the total ordering on the SCCs.
+This is accomplished by demonstrating that any graph $G$ with $l < l_{\mathrm{max}}$, that is with fewer than the maximum possible number of links between the given collection of SCCs, must have a lower dynamical robustness than one of the graphs having $l=l_{\mathrm{max}}$, $G_{\mathrm{max}}$, and that this $G_{\mathrm{max}}$ has SCCs related by the graph $G_{\mathrm{tot}}$ which gives the total ordering on those SCCs.
+
 Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
@@
 \subsection{Simultaneous perturbation of multiple interaction strengths}
-This argument also works when relationships between multiple system variables are perturbed simultaneously,
+The results of the previous sections all work when relationships between multiple system variables are perturbed simultaneously,
 although the notation becomes more complicated.  Suppose that we
-resample $m$ interactions.  Then the analogue of \ref{eq:singleresamplemutau} is
+resample $m$ interactions.  Then the analog of \ref{eq:singleresamplemutau} is
 \begin{equation}\label{eq:multipleresamplemutau}

### 2015-07-23 tex/overview.tex
minor edits

@@
 Following the examples of the previous section, here we show, in general, that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
-This is accomplished by demonstrating that any graph $G$ with $l < l_{\mathrm{max}}$, that is with fewer than the maximum possible number of links between the given collection of SCCs, must have a lower dynamical robustness than one of the graphs having $l=l_{\mathrm{max}}$, $G_{\mathrm{max}}$, and that this $G_{\mathrm{max}}$ has SCCs related by the graph $G_{\mathrm{tot}}$ which gives the total ordering on those SCCs.
+This is accomplished by demonstrating that any graph $G$ with $l < l_{\mathrm{max}}$, where $l_{\mathrm{max}}$ refers to the maximum possible number of links between the given collection of SCCs, must have a lower dynamical robustness than one of the graphs having $l=l_{\mathrm{max}}$, referred to as $G_{\mathrm{max}}$, and that these $G_{\mathrm{max}}$ have SCCs related by the graph $G_{\mathrm{tot}}$ which gives the total ordering on those SCCs.
 
@@
 \subsection{Simultaneous perturbation of multiple interaction strengths}
-The results of the previous sections all work when relationships between multiple system variables are perturbed simultaneously,
+The results of the previous sections also hold when relationships between multiple system variables are perturbed simultaneously,
 although the notation becomes more complicated.  Suppose that we

### 2015-07-23 tex/discussion.tex
minor edits

@@
 In the short term, this prediction may be further evaluated at the levels of both metabolic and transcription factor networks, which have already been shown to display hierarchical structure, but whose dynamics have not been sufficiently well characterized to ascertain their dynamical robustness as we have defined it here \cite{Zhao2006,Bhardwaj2010,Colm,Prill2005}.
+
 Indeed, Figure 2 from \cite{Prill2005} demonstrates an enrichment of various hierarchical network architectures consisting of multiple SCCs relative to those containing cycles with single SCCs for networks with three and four variables in \emph{E.
@@
 elegans} neural network.
-The network architecture that we predict to be enriched by selection for robustness for three variable networks ranks 4th out of 13 in overall abundance across these networks and within each network individually with the exception of the \emph{C.
-elegans} neural network, and the three more abundant network architectures are all proper subgraphs of this one.
-These results are essentially consistent with our prediction, though, of course, we would not claim that there could not be various alternative explanations for these observations.
+The network architecture that we predict to be most enriched by selection for robustness for three variable networks ranks fourth out of thirteen in overall abundance across these empirically observed classes of networks and within each class of network individually.
+In each case, the three most abundant network architectures are all proper subgraphs of the fourth-ranked most hierarchical one.
+These observations are consistent with both our analytical and numerical predictions, though, of course, we do not claim that there could not be alternative explanations for the relative enrichment of more hierarchical network architectures in these systems.
+
 At the ecological level, a system subjected to the environmental stress of overfishing, which may imply selection for robustness, has been observed to exhibit such a bias toward more hierarchical network architectures \cite{Bascompte2005}.
@@
 Of course, in this case, factors other than selection for robustness would not be ruled out by this analysis alone.
-In the long term, this prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that impose differential selection strengths for dynamical robustness \cite{Leroi1994}.
+In the long term, our prediction may be evaluated using experimental evolution by comparing the degree of hierarchy that emerges in the evolution of gene regulatory network topology in the context of both static and fluctuating environments that impose differential selection strengths for dynamical robustness \cite{Leroi1994,Hindre2012}.
 

### 2015-07-23 tex/figures.tex
minor edits

@@
 (B) SCCs highlighted in gray for each of the four graphs representing the interdependencies relevant to four different three variable systems.
-The most hierarchical network, top panel, is the one that maximizes the number of SCCs and the number of links between them.
+The most hierarchical network, top panel, maximizes the number of links between SCCs.
 % We therefore define hierarchy as $max(\hbox{ED}) - \hbox{ED}$ where ED is the edit distance representing the number of link addition/deletion operations necessary to transform a given graph into the most hierarchical one.
@@
 SCCs with more than one variable) and hierarchy.
-(C) Symmetries of the $\hier$ transformation between graphs and SCCs.
+(C) Symmetries of the $\hier$ transformation that maps a graph representing interactions among system variables to a graph representing the relationships among the SCCs of the system.
 The transformation $a$ represents an interchange of SCCs, $b$ moving a link between nodes in a component and $c$ adding a link.
-All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries of robustness.}
+All three transformations represent symmetries of the $\hier$ transformation from graphs to SCCs while only $a$ and $b$ are symmetries that preserve dynamical robustness.}
 \label{fig:modsccsym}

### 2015-07-23 tex/overview.tex
minor edits

--- 2015-07-23 tex/overview.tex
@@
 Following the examples of the previous section, here we show, in general, that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
-This is accomplished by demonstrating that any graph $G$ with fewer than the maximum, $l_{\mathrm{max}}$, of links between the given collection of SCCs must have a lower dynamical robustness than the graph, $G_{\mathrm{max}}$, with $l=l_{\mathrm{max}}$ and that this $G_{\mathrm{max}}$ is indeed the graph $G_{\mathrm{tot}}$ which gives the total ordering on the SCCs.
-Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
+This is accomplished by demonstrating that any graph $G$ with $l < l_{\mathrm{max}}$ ($l_{\mathrm{max}}$ refers to the maximum possible number of links between the given collection of SCCs) must have a lower dynamical robustness than one of the graphs $G_{\mathrm{max}}$ having $l=l_{\mathrm{max}}$, and that these $G_{\mathrm{max}}$ have SCCs related by the graph $G_{\mathrm{tot}}$ which gives the total ordering on the given set of SCCs.
+
+Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$ for a given set of SCCs, will also maximize $R$.
 Given two connected components $C_{\alpha}$ and $C_{\beta}$ with $v_{\alpha}$ and $v_{\beta}$ nodes respectively, we have a maximum of $v_{\alpha} v_{\beta}$ links going from $C_{\alpha}$ to $C_{\beta}$.
@@
 \subsection{Simultaneous perturbation of multiple interaction strengths}
-This argument also works when relationships between multiple system variables are perturbed simultaneously,
+The results of the previous sections also hold when relationships between multiple system variables are perturbed simultaneously,
 although the notation becomes more complicated.  Suppose that we
-resample $m$ interactions.  Then the analogue of \ref{eq:singleresamplemutau} is
+resample $m$ interactions, as a model for cases in which multiple interaction strengths among system variables are modified in a single timestep of the evolutionary process.
+Then the analog of \ref{eq:singleresamplemutau} is
 \begin{equation}\label{eq:multipleresamplemutau}
@@
 Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
-m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
+m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from within $C_i$ and
 $m_0$ links between SCCs.  Hence, our
@@
 This analytical result implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs.
-Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the variable interaction strengths are sampled or the size of the system.
+Because this result is purely topological in nature, it does not depend at all upon any particular details such as the size of the system or the probability distribution from which the variable interaction strengths are sampled.
 Any network whose associated dynamical system has the interaction graph among its SCCs equivalent to the total ordering is more robust than those associated to any of the other interaction graphs in \reffigscc{}.
-The graph associated to the total ordering is the most hierarchical network architecture for any given number of SCCs like that of \reffigscc{} top for three variable systems where the \emph{highest} variable in the hierarchy has directed links to all other variables in the network, the second highest variable has directed links to all other variables in the network except the highest one, \emph{et cetera} (\refsupp{}).
+The graph associated to the total ordering is the most hierarchical network architecture for any given number of SCCs like that of \reffigscc{} top for three variable systems where the \emph{highest} variable in the hierarchy has directed links to all other variables in the network, the second highest variable has directed links to all other variables in the network except the highest one, \emph{et cetera} (\refsupp{} \ref{sec:totalordering}).
 The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.
 
-\subsection{Examples of the analytical relationship between hierarchy and robustness}
+\subsection{Examples of the relationship between hierarchy and robustness}
 
@@
 While the most hierarchical network architecture will always lack cycles altogether, cycle number alone is clearly insufficient to account for robustness as the members of each class span nearly the entire range of possible robustness values.
-Consistent with our analysis of the symmetries of robustness, we found that the most hierarchical network architecture is the most robust (\reffigrobusthierarchy).
+Consistent with our analysis, we found that the most hierarchical network architecture is the most robust (\reffigrobusthierarchy).
 Moreover, if we consider hierarchy partitioned by connectivity, we find that there is a monotonic increase in robustness following any line of increasing hierarchy in \reffigconnectdist3D3x3.

### 2015-07-24 tex/overview.tex
refactor most hierarchical network architectures section

--- 2015-07-23 tex/overview.tex
@@
 \subsection{Most hierarchical network architectures maximize dynamical robustness}
-Following the examples of the previous section, here we show, in general, that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
-This is accomplished by demonstrating that any graph $G$ with $l < l_{\mathrm{max}}$, where $l_{\mathrm{max}}$ refers to the maximum possible number of links between the given collection of SCCs, must have a lower dynamical robustness than one of the graphs having $l=l_{\mathrm{max}}$, referred to as $G_{\mathrm{max}}$, and that these $G_{\mathrm{max}}$ have SCCs related by the graph $G_{\mathrm{tot}}$ which gives the total ordering on those SCCs.
+Following the examples of the previous section, we now show, in general, that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them.
+The proof consists of two stages; deriving an upper bound on the robustness of a network in terms of its SCCs and then showing that this bound is attained by a graph if and only if it maximizes the number of links between its SCCs.
 
-Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.
+Suppose we have a directed graph $G$ with SCCs $C_1,\ldots,C_n$.
+Since every acyclic digraph can be embedded into a totally ordered set, we may assume without loss of generality that our components have been ordered in a way such that, if $(\alpha,\beta) \in \hier(G)$, then $\alpha < \beta$.
 Given two connected components $C_{\alpha}$ and $C_{\beta}$ with $v_{\alpha}$ and $v_{\beta}$ nodes respectively, we have a maximum of $v_{\alpha} v_{\beta}$ links going from $C_{\alpha}$ to $C_{\beta}$.
-Hence, $l \le \sum_{(\alpha,\beta) \in \hier(G)} v_{\alpha} v_{\beta}$.
-Since every acyclic digraph can be embedded into a totally ordered
-set, we may assume without loss of generality that our components have
-been ordered in a way such that, if $(\alpha,\beta) \in \hier(G)$, then $\alpha <
-\beta$.  Hence, $l \le l_{\mathrm{max}}$ where
+Hence, $l \le l_{\mathrm{max}}$ where
 $$l_{\mathrm{max}} = \sum_{\alpha=1}^{n-1}\sum_{\beta=\alpha+1}^{n}v_{\alpha}
@@
 v_{\alpha}^2.$$
-Suppose we have a graph $G$ with SCCs $C_1,\ldots,C_n$ and that $G_{\mathrm{tot}}$
-is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.
-Then we have a graph $G_{\mathrm{max}}$ such that $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$, the components of $G_{\mathrm{max}}$ are also $C_1, \ldots, C_n$, but that includes all possible links between each pair of components.
-In this case $R(G)~\le~R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{\mathrm{max}}$ substituted for $l$
+Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one leads to the conclusion that networks maximizing $l$ for a given set of SCCs, will also maximize $R$.
+Hence, we have $R(G,S,\mu,\tau) \le R_{\mathrm{max}}$ where $R_{\mathrm{max}}$ is equivalent to \ref{eq:robschematic} with $l_{\mathrm{max}}$ substituted for $l$:
 \begin{widetext}
 \begin{align} \label{eq:sccmaxrobustness}
-R(G_{\mathrm{max}}) =
+R_{\mathrm{max}} =
 \frac{\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
@@
 \end{widetext}
-% where $R(G_{\mathrm{max}},S,\mu,\tau) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
-%This bound is attained by $G_{\mathrm{max}}$.
-% $\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
+
+Let $G_{\mathrm{tot}}$ refer to the directed graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$ and let $G_{\mathrm{max}}$ refer to the directed graph whose SCCs are $C_1, \ldots, C_n$, but which includes all possible links between each pair of components $C_\alpha, C_\beta$ when $\alpha < \beta$.
+Then $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$ and $R(G_{\mathrm{max}}) = R_{\mathrm{max}}$.
+Thus, we see that our bound $R_{\mathrm{max}}$ is attained by $G_{\mathrm{max}}$ so conclude that $R_{\mathrm{max}}$ is indeed the maximum value of dynamical robustness for any G-class of fixed points.
+Furthermore, we note that for any other graph $G$, which has the maximum number of links between its SCCs, we also have $R(G) = R_{\mathrm{max}}$.
+By contrast, for any graph $G$, which has less than the maximum number of links between its SCCs, we have $R(G) < R_{\mathrm{max}}$.
+
 \subsection{Simultaneous perturbation of multiple interaction strengths}
@@
 although the notation becomes more complicated.  Suppose that we
-resample $m$ interactions.  Then the analog of \ref{eq:singleresamplemutau} is
+resample $m$ interactions, as a model for cases in which multiple interaction strengths among system variables are modified in a single timestep of the evolutionary process.
+Then the analog of \ref{eq:singleresamplemutau} is
 \begin{equation}\label{eq:multipleresamplemutau}

### 2015-07-24 tex/overview.tex
add section demonstrating that the most dynamically robust network architectures have a single type of SCC

@@
 
+\subsection{The most dynamically robust network architectures have a single type of SCC}
+The previously considered optimization problem maximized dynamical robustness subject to the constraint that the network have a particular set of SCCs.
+We will now generalize this to situations in which we fix the types of SCCs which may appear in our network, but allow arbitrary numbers of each type.
+We will study the constrained optimization problem where we fix the total number of nodes to constant $N$.
+This analysis will show that the most dynamically robust networks will be homogeneous in the types of their SCCs thereby having only a single type.
+This result drastically limits the number of different networks on $N$ nodes that are candidates for being the most dynamically robust.
+
+Suppose that ${\hat C}_1, \ldots, {\hat C}_n$ is a collection of non-isomorphic strongly connected directed graphs and that $G$ is a directed graph such that every SCC of $G$ is isomorphic to some ${\hat C}_i$ for some $i$.
+For all $i \in \{1, \ldots, n\}$, let $k_i$ be the number of SCCs of $G$ which are isomorphic to ${\hat C}_i$.
+Then, following algebraic manipulation, \ref{eq:robschematic} can be rewritten as follows:
+\begin{equation}\label{eq:robtype}
+1-R(G) = \frac{\sum_{i=1}^n k_i d_i (1-R_i)}{l + \sum_{i=1}^n k_i d_i}.
+\end{equation}
+Maximizing $R$ is equivalent to minimizing $1-R$.
+Even though the $k_i$'s are integer variables, we will approximate them as positive real variables.
+The accuracy of this approximation increases in the limit of large networks.
+
+When maximizing robustness subject to a fixed number $N$ of nodes, we can make use of our previous result in \ref{eq:sccmaxrobustness} to restrict attention to directed graphs with the maximal number of links between SCCs.
+Hence, we can replace $l$ with $l_{\mathrm{max}}$ in \ref{eq:robtype} to obtain
+\begin{equation}
+1-R(G) = \frac{\sum_{i=1}^n k_i d_i (1-R_i)}
+  {\frac{1}{2} N^2 - \frac{1}{2} \sum_{i=1}^n k_i v_i^2 + \sum_{i=1}^n k_i d_i} .
+\end{equation}
+The condition on the total number of nodes means that the $k_i$'s cannot be varied independently but are subject to the constraint $\sum_{i=1}^n k_i v_i = N$.
+Geometrically, this constraint defines a hyperplane and the interesection of this hyperplane with the positive orthant is the simplex whose vertices are the points $(N/v_1, 0, \ldots, 0), (0, N/v_2, \ldots, 0), \ldots, (0, 0, \ldots, N/v_n)$.
+
+Since both numerator and denominator are linear in the variables $k_i$ over which we are optimizing, an elementary argument shows the quantity will assume its minimum value at one of the vertices of the simplex.
+The value at the $i$-th vertex equals
+\begin{equation}
+ q_i =
+ \frac{2 d_i (1 - R_i)}
+      {(k_i - 1) v_i^2 + 2 d_i},
+\end{equation}
+hence the network which maximizes robustness will have as many SCCs as possible be of the type which have the smallest value of $q_i$.
+
 \subsection{Examples of the relationship between hierarchy and robustness}

### 2015-07-24 tex/table2x2stab.tex
modify tables and imported packages to work with revtex
fix reference alignment and footnoting
update tables with scc numbers

@@
 %!TEX root = ../paper.tex
+% \onecolumn
 \newcommand{\specialcell}[2][c]{%
@@
 
-\begin{table}[h]
+\begin{table*}[h]
 \begin{center}
-\begin{tabular}{ c || c | c | c }
+\begin{tabular}{ c || c | c | c | c | c }
 \hline
-matrix & connectivity & \specialcell{probability of stability\\to perturbation} & \specialcell{probability\\of stability}\\
+matrix & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{q \\ (\ref{eq:robqual})}\\
 \hline
@@
 d & c
-\end{pmatrix}$ & 4 & 0.62 & 0.25 \\
+\end{pmatrix}$ & 1 & 4 & 0.62 & 0.25 & 1.52\\
   $\begin{pmatrix}
@@
 d & c
-\end{pmatrix}$ & 3 & 0.5 & 0.25 \\
+\end{pmatrix}$ & 1 & 3 & 0.5 & 0.25 & 1.5\\
   $\begin{pmatrix}
@@
 0 & c
-\end{pmatrix}$ & 3 & 0.67 & 0.25 \\
+\end{pmatrix}$ & 2 & 3 & 0.67 & 0.25 & \\
 $\begin{pmatrix}
@@
 0 & c
-\end{pmatrix}$ & 2 & 0.5 & 0.25 \\
+\end{pmatrix}$ & 2 & 2 & 0.5 & 0.25 & \\
 \end{tabular}
 \end{center}
-\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for two variable systems derived analytically}.
-All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
-\end{table}
+\caption{{\bf Dynamical robustness and probability of stability for two variable systems derived analytically}. All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
+\end{table*}
+% \twocolumn

### 2015-07-24 tex/overview.tex
add equation reference
update transition

@@
 
-We use the fact that stability decomposes over the SCCs of a network to relate the robustness of a network to the robustness of its SCCs.
-To derive this relationship we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessbygraph}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
+In this section we derive \ref{eq:robschematic}, which relates the dynamical robustness of a network to the dynamical robustness' of its SCCs and the number of links that connect distinct SCCs.
+We use the fact that stability decomposes over the SCCs of a network.
+More specifically, to derive this relationship we substitute \ref{eq:singleresamplemutau} and \ref{eq:stabfactor} into \ref{eq:robustnessbygraph}, collect factors corresponding to components, decompose integrals into their respective products, collapse integrals over delta distributions, and cancel common factors between numerator and denominator.
 Let $L$ denote the set of edges of $G$ that connect distinct SCCs.
@@
 Here $R_\alpha$ is shorthand for $R(C_\alpha,S,\mu_\alpha,\tau^{(1)}_\alpha)$ and $R$ is shorthand for $R(G,S,\mu,\tau^{(1)})$.
+\ref{eq:robschematic} demonstrates just as stability decomposes over the SCCs of a network architecture, which was shown in the previous section, so does dynamical robustness.
 
@@
 The value at the $i$-th vertex equals
-\begin{equation}
+\begin{equation}\label{eq:robqual}
  q_i =

### 2015-07-26 tex/overview.tex
explain quantity that gives relative contribution of SCCs to robustness

@@
 \end{equation}
-hence the network which maximizes robustness will have as many SCCs as possible be of the type which have the smallest value of $q_i$.
+hence the network that maximizes robustness will have as many SCCs as possible be of the type which have the smallest value of $q_i$.
+For the SCC consisting of a single node that has a link connecting to itself $\lim_{N \rightarrow \infty} Nq = 1$.
+For other SCCs of two and three nodes this quantity is computed in \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}.
 

### 2015-07-26 tex/table2x2stab.tex
explain quantity that gives relative contribution of SCCs to robustness

@@
 \hline
-matrix & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{q \\ (\ref{eq:robqual})}\\
+matrix & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{$\lim_{N \rightarrow \infty} Nq$ \\ (\ref{eq:robqual})}\\
 \hline

### 2015-07-27 tex/overview.tex
update explanation of relative contribution of SCCs to robustness

@@
 Geometrically, this constraint defines a hyperplane and the interesection of this hyperplane with the positive orthant is the simplex whose vertices are the points $(N/v_1, 0, \ldots, 0), (0, N/v_2, \ldots, 0), \ldots, (0, 0, \ldots, N/v_n)$.
+When $v_i = 1$ this corresponds to the dynamical robustness of the SCC with a single node and link to itself.
 
-Since both numerator and denominator are linear in the variables $k_i$ over which we are optimizing, an elementary argument shows the quantity will assume its minimum value at one of the vertices of the simplex.
+Since both numerator and denominator are linear in the variables $k_i$ over which we are optimizing, the quantity will assume its minimum value at one of the vertices of the simplex.
 The value at the $i$-th vertex equals
@@
 hence the network that maximizes robustness will have as many SCCs as possible be of the type which have the smallest value of $q_i$.
-For the SCC consisting of a single node that has a link connecting to itself $\lim_{N \rightarrow \infty} Nq = 1$.
-For other SCCs of two and three nodes this quantity is computed in \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}.
+
+% $\lim_{N \rightarrow \infty} Nq = 1$
 
@@
 Moreover, if we consider hierarchy partitioned by connectivity, we find that there is a monotonic increase in robustness following any line of increasing hierarchy in \reffigconnectdist3D3x3.
+
+For the SCC consisting of a single node that has a link connecting to itself, which is the only SCC in the network corresponding to the total ordering on the network variables, \ref{eq:robqual} gives $q= \frac{1}{N+1}$.
+Assuming this value needs to be smaller than the corresponding value of $q_i$ for any other SCC, $\frac{1}{N+1} < q_i$, and simplifying the algebra, results in a sufficient condition for networks that contain only the single node SCC to have a higher value of dynamical robustness than networks containing any combination of alternative SCCs:
+\begin{equation}\label{eq:maxrobcond}
+\frac{2 d_i}{v_i} (1-R_i) > 1.
+\end{equation}
+For SCCs of two and three nodes this quantity is computed in \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}, where all of these SCCs satisfy this condition.
+On this basis, we conjecture that, in addition to the fact that the most dynamically robust network architectures correspond to the total ordering on a given collection of SCCs demonstrated by the analysis leading up to \ref{eq:robschematic}, that the most dynamically robust network architectures also correspond to the graph associated to the total ordering on all network variables.

### 2015-07-27 tex/table2x2stab.tex
update explanation of relative contribution of SCCs to robustness

--- 2015-07-24 tex/table2x2stab.tex
@@
 \hline
-matrix & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{q \\ (\ref{eq:robqual})}\\
+\specialcell{adjacency \\ matrix} & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{$\frac{2d_i}{v_i}(1-R_i)$ \\ (\ref{eq:maxrobcond})}\\
 \hline
   $\begin{pmatrix}
-a & b \\
-d & c
+1 & 1 \\
+1 & 1
 \end{pmatrix}$ & 1 & 4 & 0.62 & 0.25 & 1.52\\
   $\begin{pmatrix}
-a & b \\
-d & 0
+1 & 1 \\
+1 & 0
 \end{pmatrix}$, $\begin{pmatrix}
-0 & b \\
-d & c
+0 & 1 \\
+1 & 1
 \end{pmatrix}$ & 1 & 3 & 0.5 & 0.25 & 1.5\\
   $\begin{pmatrix}
-a & 0 \\
-d & c
+1 & 0 \\
+1 & 1
 \end{pmatrix}$, $\begin{pmatrix}
-a & b \\
-0 & c
+1 & 1 \\
+0 & 1
 \end{pmatrix}$ & 2 & 3 & 0.67 & 0.25 & \\
 $\begin{pmatrix}
-a & 0 \\
-0 & c
+1 & 0 \\
+0 & 1
 \end{pmatrix}$ & 2 & 2 & 0.5 & 0.25 & \\
